---
title: "Lai Study 1 replication with VH estimators"
subtitle: "Technical review of historical comparability, reporting scale, and slope-SD diagnostics"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    html-math-method: mathjax
execute:
  echo: false
  warning: false
  message: false
params:
  results_dir: "outputs/lai_study1_vh_production_20260731"
  figure_dir: null
  historical_figure_reference: null
---

## Technical summary

The replication succeeds on its primary like-for-like point-estimate check.
On Lai's historical reporting scale, our dual-BLUP analogue has negative bias
near $-.02$ for every random-effect correlation, closely matching Lai's
original **dual-EB** series.  The previously suspected sign disagreement came
from comparing that dual method with Lai's separate slope-only EB series,
whose bias changes sign with the random-effect correlation.  Our 2S-PA and
MSEM analogues also remain close to zero, as in the original results.

That conclusion is scale-specific.  On the raw and common latent-slope-SD
views, the dual-BLUP estimator is centrally near zero (with rare, very large
failures), so much of its historical negative bias is attributable to Lai's
method-specific EB-slope-SD multiplier.  In contrast, closed-form OLS remains
negatively biased on the common scale, while Fuller, Fuller alpha-stepdown,
2S-PA, and MSEM are generally much closer to zero.

The interval results from the seven-method run are encouraging but not exact
replications: our 2S-PA is somewhat conservative and our **VH/Mplus MSEM**
somewhat anti-conservative relative to Lai's original summaries, especially
at $J=30$.  A historical-only eighth method now reproduces Lai's original
OpenMx MSEM and obtains the standardized estimate and delta-method SE directly
from `mxSE(stdbeta)`.  It is added to the historical panels when its separate
array has been completed and aggregated; it never enters the seven-method VH
panels.  Lai's 2S-PAB and slope-only EB variants remain outside this refresh.

```{r setup}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(knitr)
library(scales)

report_input <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) "")
report_dir <- if (nzchar(report_input)) {
  report_input <- normalizePath(report_input, mustWork = FALSE)
  if (dir.exists(report_input)) report_input else dirname(report_input)
} else {
  getwd()
}
repo_root <- normalizePath(file.path(report_dir, ".."), mustWork = FALSE)

resolve_repo_path <- function(path) {
  if (is.null(path) || !nzchar(path)) return(NULL)
  if (grepl("^/", path)) {
    normalizePath(path, mustWork = FALSE)
  } else {
    normalizePath(file.path(repo_root, path), mustWork = FALSE)
  }
}

safe_mean <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) == 0L) NA_real_ else mean(x)
}

safe_median <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) == 0L) NA_real_ else stats::median(x)
}

fmt_num <- function(x, digits = 3L) {
  ifelse(is.finite(x), formatC(x, digits = digits, format = "f"), NA_character_)
}

fmt_pct <- function(x, accuracy = 0.1) {
  ifelse(is.finite(x), scales::percent(x, accuracy = accuracy), NA_character_)
}

method_labels <- c(
  oracle_dual = "Oracle dual",
  naive_dual_blup = "Naive dual BLUP",
  closed_form_dual = "Closed-form dual",
  fuller_closed_form = "Fuller closed-form",
  fuller_alpha_stepdown_closed_form = "Fuller alpha-stepdown",
  lai_2spa = "2S-PA",
  msem = "MSEM (VH/Mplus)",
  lai_original_msem = "MSEM (original OpenMx)"
)

method_palette <- c(
  "Oracle dual" = "#4D4D4D",
  "Naive dual BLUP" = "#00BFC4",
  "Closed-form dual" = "#619CFF",
  "Fuller closed-form" = "#B79F00",
  "Fuller alpha-stepdown" = "#F564E3",
  "2S-PA" = "#F8766D",
  "MSEM (VH/Mplus)" = "#007A3D",
  "MSEM (original OpenMx)" = "#00BA38"
)

results_dir <- resolve_repo_path(params$results_dir)
figure_dir <- resolve_repo_path(params$figure_dir)
if (is.null(figure_dir)) figure_dir <- file.path(results_dir, "figure2_analogue")
summary_path <- file.path(results_dir, "lai_study1_vh_summary.csv")
original_overlap_path <- file.path(
  repo_root,
  "lai_study1_replication_vh/reference/lai_original_study1_overlap_summary.csv"
)

if (!file.exists(summary_path)) {
  stop("Could not find the aggregate summary: ", summary_path, call. = FALSE)
}

summary_df <- readr::read_csv(summary_path, show_col_types = FALSE)
required_columns <- c(
  "condition_id", "reporting_scale", "method", "num_clus", "clus_size",
  "icc", "vr_u1_u0", "cor_u0_u1", "beta_zu1", "var_u1", "sigma2",
  "dgm_population_slope_sd", "dgm_posterior_slope_reliability", "truth",
  "n_rep", "convergence", "vh_analysis_eligibility_rate",
  "lai_original_eligibility_rate", "n_eligible_both", "n_lai_original_only",
  "n_vh_only", "bias", "rmse", "coverage",
  "mean_eb_to_population_slope_sd",
  "mean_corrected_to_population_slope_sd",
  "mean_fitted_to_population_slope_sd",
  "mean_lai_original_method_multiplier",
  "lai_original_reporting_available_rate"
)
missing_columns <- setdiff(required_columns, names(summary_df))
if (length(missing_columns) > 0L) {
  stop(
    "Aggregate summary is missing required columns: ",
    paste(missing_columns, collapse = ", "),
    call. = FALSE
  )
}

summary_df <- summary_df |>
  mutate(
    method_label = factor(method, levels = names(method_labels), labels = unname(method_labels)),
    reporting_scale = factor(
      reporting_scale,
      levels = c("raw", "latent_sd", "lai_original_standardized")
    )
  )

raw_df <- summary_df |> filter(reporting_scale == "raw")
latent_df <- summary_df |> filter(reporting_scale == "latent_sd")
historical_df <- summary_df |> filter(reporting_scale == "lai_original_standardized")

n_conditions <- n_distinct(summary_df$condition_id)
n_methods <- n_distinct(summary_df$method)
n_replications <- raw_df |> distinct(condition_id, n_rep) |> summarise(total = sum(n_rep)) |> pull(total)

read_optional_csv <- function(path) {
  if (!file.exists(path)) return(NULL)
  readr::read_csv(path, show_col_types = FALSE)
}

historical_figure_summary <- read_optional_csv(
  file.path(figure_dir, "figure2_analogue_cell_summary.csv")
)
vh_figure_summary <- read_optional_csv(
  file.path(figure_dir, "vh_primary_companion_cell_summary.csv")
)
vh_primary_inference_summary <- read_optional_csv(
  file.path(figure_dir, "vh_primary_inference_companion_cell_summary.csv")
)
historical_inference_summary <- read_optional_csv(
  file.path(figure_dir, "figure3_4_analogue_cell_summary.csv")
)
paired_engine_audit <- read_optional_csv(
  file.path(results_dir, "mplus_openmx_audit", "paired_engine_comparisons.csv")
)
paired_engine_audit_selection_column <- if (!is.null(paired_engine_audit)) {
  selection_columns <- grep("^selection", names(paired_engine_audit), value = TRUE)
  if (length(selection_columns) == 1L) selection_columns[[1]] else NULL
} else {
  NULL
}
has_paired_engine_audit <- !is.null(paired_engine_audit) &&
  !is.null(paired_engine_audit_selection_column) &&
  all(c("mplus_variant", "openmx_standardized_estimate", "openmx_delta_se") %in%
        names(paired_engine_audit))
original_overlap_summary <- read_optional_csv(original_overlap_path)
has_original_msem <- !is.null(historical_figure_summary) &&
  "lai_original_msem" %in% historical_figure_summary$method
original_msem_status_text <- if (has_original_msem) {
  "complete and included in the historical panels"
} else {
  "implemented but not yet present in the post-estimation aggregate"
}

historical_bias_by_correlation <- if (!is.null(historical_figure_summary)) {
  historical_figure_summary |>
    filter(beta_zu1 == 0.4) |>
    group_by(method, cor_u0_u1) |>
    summarise(
      median_robust_bias = safe_median(robust_bias),
      mean_robust_bias = safe_mean(robust_bias),
      .groups = "drop"
    )
} else {
  NULL
}

format_correlation_medians <- function(method) {
  if (is.null(historical_bias_by_correlation)) return("not available")
  historical_bias_by_correlation |>
    filter(method == !!method) |>
    arrange(cor_u0_u1) |>
    transmute(text = paste0("rho = ", cor_u0_u1, ": ", fmt_num(median_robust_bias))) |>
    pull(text) |>
    paste(collapse = "; ")
}

naive_historical_correlation_text <- format_correlation_medians("naive_dual_blup")
two_spa_historical_correlation_text <- format_correlation_medians("lai_2spa")
msem_historical_correlation_text <- format_correlation_medians("msem")
original_msem_historical_correlation_text <- format_correlation_medians("lai_original_msem")

vh_instability_summary <- if (!is.null(vh_figure_summary)) {
  vh_figure_summary |>
    group_by(method) |>
    summarise(
      median_bias = safe_median(mean_bias),
      median_abs_bias = safe_median(abs(mean_bias)),
      max_abs_bias = max(abs(mean_bias), na.rm = TRUE),
      n_abs_bias_over_01 = sum(abs(mean_bias) > 0.1, na.rm = TRUE),
      n_abs_bias_over_1 = sum(abs(mean_bias) > 1, na.rm = TRUE),
      .groups = "drop"
    )
} else {
  NULL
}

naive_instability <- if (!is.null(vh_instability_summary)) {
  vh_instability_summary |> filter(method == "naive_dual_blup")
} else {
  NULL
}
naive_instability_count <- if (!is.null(naive_instability) && nrow(naive_instability) == 1L) {
  naive_instability$n_abs_bias_over_1[[1]]
} else {
  NA_integer_
}
naive_instability_max <- if (!is.null(naive_instability) && nrow(naive_instability) == 1L) {
  naive_instability$max_abs_bias[[1]]
} else {
  NA_real_
}

historical_inference_by_method <- if (!is.null(historical_inference_summary)) {
  historical_inference_summary |>
    group_by(method) |>
    summarise(
      coverage_conditions = sum(is.finite(coverage)),
      mean_coverage = safe_mean(coverage),
      median_coverage = safe_median(coverage),
      type1_conditions = sum(beta_zu1 == 0 & is.finite(type1_error)),
      mean_type1_error = safe_mean(type1_error[beta_zu1 == 0]),
      median_type1_error = safe_median(type1_error[beta_zu1 == 0]),
      mean_retained = safe_mean(n_retained),
      .groups = "drop"
    )
} else {
  NULL
}

historical_inference_by_method_j <- if (!is.null(historical_inference_summary)) {
  historical_inference_summary |>
    group_by(method, num_clus) |>
    summarise(
      mean_coverage = safe_mean(coverage),
      mean_type1_error = safe_mean(type1_error),
      .groups = "drop"
    )
} else {
  NULL
}

historical_inference_value <- function(method, metric, num_clus = NULL) {
  df <- if (is.null(num_clus)) {
    historical_inference_by_method
  } else {
    historical_inference_by_method_j
  }
  if (is.null(df)) return(NA_real_)
  row <- df |> filter(method == !!method)
  if (!is.null(num_clus)) row <- row |> filter(num_clus == !!num_clus)
  if (nrow(row) != 1L || !metric %in% names(row)) return(NA_real_)
  row[[metric]][[1]]
}

original_inference_by_method <- if (!is.null(original_overlap_summary)) {
  original_overlap_summary |>
    distinct(method, mean_coverage, mean_type1_error)
} else {
  NULL
}

vh_inference_by_method <- if (!is.null(vh_primary_inference_summary)) {
  vh_primary_inference_summary |>
    group_by(method) |>
    summarise(
      mean_coverage = safe_mean(coverage),
      min_coverage = min(coverage, na.rm = TRUE),
      n_coverage_below_70 = sum(coverage < 0.70, na.rm = TRUE),
      mean_type1_error = safe_mean(type1_error[beta_zu1 == 0]),
      .groups = "drop"
    )
} else {
  NULL
}

vh_inference_value <- function(method, metric) {
  if (is.null(vh_inference_by_method)) return(NA_real_)
  row <- vh_inference_by_method |> filter(method == !!method)
  if (nrow(row) != 1L || !metric %in% names(row)) return(NA_real_)
  row[[metric]][[1]]
}
```

The production aggregate contains `r n_conditions` DGM conditions, `r n_methods`
methods, and `r scales::comma(n_replications)` simulated data sets per raw
reporting view.  Its source is
`r normalizePath(summary_path, mustWork = FALSE)`.

### Result at a glance

For the historical comparison at $\beta_{Zu1}=.4$, the naïve dual-BLUP
median trimmed bias is negative at every random-effect correlation
(`r naive_historical_correlation_text`).  This closely reproduces Lai's
original dual-EB $(u_0+u_1)$ series.  The correlation-signed pattern in the
original paper belongs to the two *slope-only* EB specifications, which are
not in the seven-method VH bundle.  2S-PA remains near zero
(`r two_spa_historical_correlation_text`) and the VH/Mplus MSEM is small and
positive (`r msem_historical_correlation_text`) on this historical view.
The source-faithful OpenMx MSEM add-on is `r original_msem_status_text`.
`r if (has_original_msem) paste0("Its correlation-specific medians are ", original_msem_historical_correlation_text, ".") else ""`

On the common latent-SD VH view, the naïve dual-BLUP median is not itself the
main issue; its distribution has `r naive_instability_count` conditions with
$|\mathrm{mean\ bias}| > 1$, including a maximum of
`r fmt_num(naive_instability_max, 1)`.  The report therefore shows its central
behavior and its rare instability in separate panels rather than letting those
failures flatten every other method's plot.

### Replication goals and current status

```{r replication-status}
replication_status <- tibble::tribble(
  ~Goal, ~Status, ~Evidence_or_limit,
  "Preserve Lai Study 1 DGM and 486-cell grid", "Complete", "Audited in Section 1; 5,000 replications per cell",
  "Use the seven VH Study 2 estimators", "Complete", "All seven methods appear in the common-scale bias and inference panels",
  "Add Lai's original OpenMx MSEM without changing the VH bundle", ifelse(has_original_msem, "Complete", "Implemented; run pending"), original_msem_status_text,
  "Retain raw, common latent-SD, and historical scales", "Complete", "All three reporting views are explicit and separately labeled",
  "Track VH and original-Lai eligibility", "Complete", "Both rules and their overlap are reported in Section 6",
  "Diagnose slope SD and posterior reliability", "Complete", "DGM and fitted slope-SD diagnostics are shown in Section 5",
  "Reproduce Lai Figures 2--4 exactly", "Partial", "The original MSEM now uses Lai's p value and delta-method SE; 2S-PAB and two slope-only EB methods remain absent"
)
knitr::kable(replication_status, caption = "Extent to which the refreshed study meets the goals in README.md")
```

The design therefore answers the intended codebase-validation question, but
does not replace a strict rerun of all six original methods.  Numeric
comparisons below use a frozen summary extracted from Lai's original
`results-sim1-revise2.rds` at commit
`9ffe53168f6bb04e13ef977dc19a8d953d0bf29d`.

## Methods and terminology

The target is the effect of the level-2 covariate $Z$ on the random slope
$u_1$.  “Dual” means that the stage-2 regression uses both the random
intercept $u_0$ and random slope $u_1$ as predictors; this is important
when those random effects are correlated.

```{r method-glossary}
method_glossary <- tibble(
  Method = c(
    "Oracle dual", "Naive dual BLUP", "Closed-form dual",
    "Fuller closed-form", "Fuller alpha-stepdown", "2S-PA",
    "MSEM (VH/Mplus)", "MSEM (original OpenMx; historical only)"
  ),
  `Stage-2 predictor or model` = c(
    "True simulated u0 and u1",
    "Empirical-Bayes (BLUP) u0 and u1, treated as observed",
    "Closed-form corrected u0 and u1, treated as observed",
    "Closed-form corrected scores with their measurement-error covariance",
    "Fuller correction with alpha stepdown for a guarded correction",
    "Lai two-stage path-analysis model fitted in OpenMx",
    "Joint multilevel structural-equation model fitted in Mplus",
    "Lai's original cross-level OpenMx MSEM; direct standardized algebra and mxSE delta-method SE"
  ),
  `Interpretive role` = c(
    "Simulation benchmark; not available in applications",
    "Shows attenuation/instability from ignoring first-stage uncertainty",
    "Separates corrected-score construction from stage-2 EIV correction",
    "EIV-style correction using estimated first-stage error covariance",
    "Stabilized Fuller variant in the VH Study 2 bundle",
    "Direct Lai-method overlap for historical comparison",
    "VH Study 2 latent-variable benchmark; historical comparison diagnostic",
    "Source-faithful historical benchmark; excluded from VH-primary seven-method comparisons"
  )
)
knitr::kable(method_glossary, caption = "Estimator terminology used throughout this report")
```

The first three methods use a two-predictor observed-score regression in stage
2.  The distinction is whether those predictors are true random effects,
shrunken EB estimates, or closed-form corrected scores.  The Fuller methods
also propagate the estimated measurement-error covariance into the stage-2
correction.  2S-PA and both MSEM implementations instead fit their own
latent-variable/path models.  The VH/Mplus MSEM is one of the seven primary
methods.  The original OpenMx MSEM is an eighth, historical-only diagnostic:
its standardized effect and SE jointly account for uncertainty in the fitted
latent slope SD through OpenMx's delta method.

## 1. DGM audit: the Lai grid and the estimand are explicit

The review starts by auditing the condition grid, rather than inferring the
design from a plotted result.  The external-outcome effect of the random
intercept is fixed at $\beta_{zu0}=.4$; the focal slope effect is
$\beta_{zu1}$.  The summary rows below use the raw reporting view only, so
reporting-scale rows do not duplicate conditions.

```{r dgm-audit}
dgm_conditions <- raw_df |>
  distinct(
    condition_id, num_clus, clus_size, icc, vr_u1_u0, cor_u0_u1,
    beta_zu1, var_u1, sigma2, dgm_population_slope_sd,
    dgm_posterior_slope_reliability, n_rep
  )

dgm_grid <- tibble(
  factor = c(
    "Clusters", "Cluster size", "ICC", "Slope/intercept variance ratio",
    "Random-effect correlation", "Slope external effect"
  ),
  levels = c(
    paste(sort(unique(dgm_conditions$num_clus)), collapse = ", "),
    paste(sort(unique(dgm_conditions$clus_size)), collapse = ", "),
    paste(sort(unique(dgm_conditions$icc)), collapse = ", "),
    paste(sort(unique(dgm_conditions$vr_u1_u0)), collapse = ", "),
    paste(sort(unique(dgm_conditions$cor_u0_u1)), collapse = ", "),
    paste(sort(unique(dgm_conditions$beta_zu1)), collapse = ", ")
  )
)

knitr::kable(dgm_grid, caption = "Lai Study 1 condition grid represented in the VH replication")
```

The condition count is `r nrow(dgm_conditions)`.  The population slope SD and
posterior slope reliability below are DGM quantities, not post-hoc calibration
targets; they provide the reference for the later slope-SD and reliability
diagnostics.

```{r dgm-reference-table}
dgm_reference <- dgm_conditions |>
  summarise(
    `Population slope SD range` = paste(range(dgm_population_slope_sd), collapse = " to "),
    `Posterior reliability range` = paste(round(range(dgm_posterior_slope_reliability), 3), collapse = " to "),
    `Replications per condition` = paste(sort(unique(n_rep)), collapse = ", ")
  )
knitr::kable(dgm_reference)
```

## 2. Historical Figure 2 analogue: compare like with like

This panel is the primary visual comparison with Lai's Figure 2.  It contains
the three methods already available from the VH run (naive dual BLUP, 2S-PA,
and VH/Mplus MSEM) plus Lai's original OpenMx MSEM when the additive historical
run is complete.  It uses Lai's method-specific historical standardization,
Lai's `status_code == 0` retention rule, and a 20%-trimmed Monte Carlo bias.
Those choices are deliberately different from the VH-primary analysis below.

Lai's published Figure 2 code plots six methods: 2S-PA, 2S-PAB, MSEM,
EB $(u_0+u_1)$, centered EB, and uncentered EB.  This analogue retains only
the directly corresponding 2S-PA and dual-EB series, the current VH/Mplus MSEM
diagnostic, and—when available—a source-faithful rerun of the original OpenMx
MSEM.  It cannot test the other three lines.  The original plotting code is available in the
[Lai supplementary repository](https://github.com/marklhc/2spa-random-slopes-supp/blob/9ffe53168f6bb04e13ef977dc19a8d953d0bf29d/notebooks/_sim1_results_supp.qmd).
The colors preserve that notebook's ggplot2 mapping: 2S-PA is red, the
**original OpenMx MSEM** is the original green, and dual EB is cyan.  The
VH/Mplus MSEM uses a darker green so the two implementations are not conflated.
Colors assigned to VH-only methods in later panels are stable extensions, not
claims that those methods correspond to omitted Lai methods.

```{r historical-figure, fig.cap="Lai Study 1 Figure 2 analogue: historical reporting scale and Lai eligibility rule."}
historical_plot_path <- file.path(figure_dir, "figure2_analogue.png")
if (file.exists(historical_plot_path)) {
  knitr::include_graphics(historical_plot_path)
} else {
  knitr::asis_output(
    "**Figure not yet generated.** Run `make_figure2_analogue.R` against the results directory, then re-render this report."
  )
}
```

```{r historical-availability}
if (!is.null(historical_figure_summary)) {
  historical_availability <- historical_figure_summary |>
    mutate(Method = recode(method, !!!method_labels)) |>
    group_by(Method) |>
    summarise(
      conditions = n(),
      `Mean scale availability` = safe_mean(historical_scale_availability_rate),
      `Mean Lai eligibility` = safe_mean(lai_original_eligibility_rate),
      `Mean retained replications` = safe_mean(n_retained),
      .groups = "drop"
    ) |>
    mutate(
      `Mean scale availability` = fmt_pct(`Mean scale availability`),
      `Mean Lai eligibility` = fmt_pct(`Mean Lai eligibility`),
      `Mean retained replications` = fmt_num(`Mean retained replications`, 1)
    )
  knitr::kable(
    historical_availability,
    caption = "Availability and retention in the historical comparison view"
  )
}
```

```{r historical-correlation-table}
if (!is.null(historical_bias_by_correlation) && !is.null(original_overlap_summary)) {
  historical_comparison_table <- historical_bias_by_correlation |>
    mutate(reference_method = if_else(method == "lai_original_msem", "msem", method)) |>
    left_join(
      original_overlap_summary |>
        select(reference_method = method, cor_u0_u1, original_bias = median_robust_bias_beta04),
      by = c("reference_method", "cor_u0_u1")
    ) |>
    mutate(
      method = recode(
        method,
        naive_dual_blup = "Naive dual BLUP",
        lai_2spa = "2S-PA",
        msem = "MSEM (VH/Mplus)",
        lai_original_msem = "MSEM (original OpenMx)"
      ),
      `Current median bias` = fmt_num(median_robust_bias),
      `Original Lai median bias` = fmt_num(original_bias),
      Difference = fmt_num(median_robust_bias - original_bias)
    ) |>
    select(Method = method, `Correlation` = cor_u0_u1,
           `Current median bias`, `Original Lai median bias`, Difference)
  knitr::kable(
    historical_comparison_table,
    caption = "Current versus original median 20%-trimmed bias at beta[Z.u1] = .4, summarized over the remaining DGM factors"
  )
}
```

The table resolves the central replication question.  Our dual-BLUP medians
are about $-.022$ across the three values of $\rho$, versus roughly
$-.024$ to $-.026$ for Lai's dual-EB method.  Both 2S-PA implementations
are essentially unbiased; the VH/Mplus MSEM differs slightly in sign but
remains small.  The original OpenMx MSEM row, when present, is the decisive
like-for-like implementation check because it uses Lai's original model and
delta-method interval rather than reconstructing a standardized interval from
a raw Mplus path.
The slope-only centered and uncentered EB lines in Lai's figure do change sign
with $\rho$, but they answer a different stage-2 regression question and are
not represented here.  The current Figure 2 analogue is therefore a strong
replication for the overlapping point estimators, not evidence of a
contradiction.

### Historical coverage and Type I error analogues

The next two panels mimic Lai's Figures 3 and 4 using the same historical
methods, historical scale, and `status_code == 0` retention rule.
Coverage is the proportion of retained 95% intervals containing the truth.
The Type I panel is limited to $\beta_{Zu1}=0$ and uses two-sided 95% CI
exclusion as the rejection rule for the three methods reconstructed from the
VH output.  The original OpenMx MSEM add-on instead stores and uses Lai's exact
normal-Wald `p < .05` definition.  These rules ordinarily agree, but keeping
the provenance explicit prevents the reconstructed VH/Mplus interval from
being mistaken for Lai's delta-method interval.

```{r figure3-coverage, fig.cap="Figure 3 analogue: empirical 95% coverage under historical Lai reporting and retention."}
coverage_plot_path <- file.path(figure_dir, "figure3_coverage_analogue.png")
if (file.exists(coverage_plot_path)) {
  knitr::include_graphics(coverage_plot_path)
} else {
  knitr::asis_output("**Coverage analogue not yet generated.** Run the post-estimation figure command, then re-render this report.")
}
```

```{r figure4-type1, fig.cap="Figure 4 analogue: CI-based empirical Type I error under historical Lai reporting and retention."}
type1_plot_path <- file.path(figure_dir, "figure4_type1_analogue.png")
if (file.exists(type1_plot_path)) {
  knitr::include_graphics(type1_plot_path)
} else {
  knitr::asis_output("**Type I error analogue not yet generated.** Run the post-estimation figure command, then re-render this report.")
}
```

```{r historical-inference-table}
if (!is.null(historical_inference_by_method)) {
  historical_inference_table <- historical_inference_by_method |>
    mutate(reference_method = if_else(method == "lai_original_msem", "msem", method)) |>
    left_join(
      original_inference_by_method |>
        rename(reference_method = method) |>
        rename(
          original_mean_coverage = mean_coverage,
          original_mean_type1_error = mean_type1_error
        ),
      by = "reference_method"
    ) |>
    mutate(
      Method = recode(
        method,
        naive_dual_blup = "Naive dual BLUP",
        lai_2spa = "2S-PA",
        msem = "MSEM (VH/Mplus)",
        lai_original_msem = "MSEM (original OpenMx)"
      ),
      `Mean coverage` = fmt_pct(mean_coverage),
      `Original coverage` = fmt_pct(original_mean_coverage),
      `Median coverage` = fmt_pct(median_coverage),
      `Mean Type I error` = fmt_pct(mean_type1_error),
      `Original Type I error` = fmt_pct(original_mean_type1_error),
      `Median Type I error` = fmt_pct(median_type1_error),
      `Mean retained replications` = fmt_num(mean_retained, 1)
    ) |>
    select(
      Method, `Coverage conditions` = coverage_conditions,
      `Mean coverage`, `Original coverage`, `Median coverage`,
      `Type I conditions` = type1_conditions,
      `Mean Type I error`, `Original Type I error`, `Median Type I error`,
      `Mean retained replications`
    )
  knitr::kable(
    historical_inference_table,
    caption = "Current and original condition-weighted interval performance; reconstructed VH methods use CI exclusion, while original OpenMx MSEM uses Lai's stored normal-Wald p value"
  )
}
```

The reference lines are .95 for coverage and .05 for Type I error.  The table
uses condition-level rates (not a pooled replication denominator), so it
describes how performance varies across the Lai design.  Interpret a departure
from either reference together with the matching bias panel and the number of
retained replications: a nominal interval result can change when the
historical reporting multiplier or status-code retention changes.

For these three overlapping methods, 2S-PA is slightly conservative overall:
mean coverage is `r fmt_pct(historical_inference_value("lai_2spa", "mean_coverage"))`
and mean CI-based Type I error is
`r fmt_pct(historical_inference_value("lai_2spa", "mean_type1_error"))`.
Naïve dual BLUP averages
`r fmt_pct(historical_inference_value("naive_dual_blup", "mean_coverage"))`
coverage and `r fmt_pct(historical_inference_value("naive_dual_blup", "mean_type1_error"))`
Type I error; MSEM averages
`r fmt_pct(historical_inference_value("msem", "mean_coverage"))` and
`r fmt_pct(historical_inference_value("msem", "mean_type1_error"))`, respectively.
The strongest MSEM departure is in the $J=30$ cells, where its mean coverage
is `r fmt_pct(historical_inference_value("msem", "mean_coverage", 30))` and
its mean Type I error is
`r fmt_pct(historical_inference_value("msem", "mean_type1_error", 30))`.
`r if (has_original_msem) paste0("The source-faithful OpenMx MSEM averages ", fmt_pct(historical_inference_value("lai_original_msem", "mean_coverage")), " coverage and ", fmt_pct(historical_inference_value("lai_original_msem", "mean_type1_error")), " Type I error overall; at J = 30 the corresponding values are ", fmt_pct(historical_inference_value("lai_original_msem", "mean_coverage", 30)), " and ", fmt_pct(historical_inference_value("lai_original_msem", "mean_type1_error", 30)), ".") else "The original OpenMx MSEM results are not yet available; rerun the historical figure postprocessor after its additive aggregate completes."`
Relative to Lai's original summaries, current dual-BLUP coverage and Type I
error are close.  Current 2S-PA is more conservative (about 96% coverage and
4% Type I error versus about 95% and 4.5% originally), while MSEM is more
anti-conservative (about 94% and 5.7% versus about 96% and 4.2%).  The
`r if (has_original_msem) "historical OpenMx row now separates a model/software/interval-construction difference from a DGM difference" else "implemented OpenMx add-on will separate model/software/interval construction from the DGM once its production array is complete"`.
Those are practical inference differences worth tracing; they do not overturn
the point-estimate replication.

### Paired Mplus–OpenMx audit: two mechanisms, not one

The completed paired-engine audit regenerates the exact data for two close and
two maximally divergent seed/condition pairs, then retains all Mplus input and
output files.  It compares the current Mplus specification (whose retained
output identifies its default estimator as MLR), an explicit-ML version of the
same model, MLR and ML versions with the standardized product declared in
`MODEL CONSTRAINT`, and the original OpenMx model.  This separates a
difference in standard-error construction from a difference in the fitted
numerical solution.

```{r paired-engine-audit-table}
if (has_paired_engine_audit) {
  audit_rows <- paired_engine_audit |>
    rename(selection = all_of(paired_engine_audit_selection_column)) |>
    filter(mplus_variant %in% c("current_default", "default_constrained", "ml_constrained")) |>
    mutate(
      `Seed class` = recode(
        selection,
        closest_saved_solution = "Closest paired solution",
        most_divergent_saved_solution = "Most divergent paired solution"
      ),
      `Mplus fit` = recode(
        mplus_variant,
        current_default = "Current default (MLR; fixed multiplier)",
        default_constrained = "Current default (MLR; product delta method)",
        ml_constrained = "ML constraint (product delta method)"
      ),
      `Mplus standardized estimate` = fmt_num(mplus_fixed_multiplier_estimate),
      `Mplus SE` = if_else(
        mplus_variant %in% c("default_constrained", "ml_constrained"),
        fmt_num(mplus_constraint_delta_se),
        fmt_num(mplus_fixed_multiplier_se)
      ),
      `OpenMx standardized estimate` = fmt_num(openmx_standardized_estimate),
      `OpenMx delta SE` = fmt_num(openmx_delta_se),
      `Mplus − OpenMx estimate` = if_else(
        mplus_variant %in% c("default_constrained", "ml_constrained"),
        fmt_num(constrained_estimate_difference_vs_openmx),
        fmt_num(fixed_multiplier_estimate_difference_vs_openmx)
      )
    ) |>
    select(
      Condition = condition_id, Replication = rep, `Seed class`, `Mplus fit`,
      `Mplus standardized estimate`, `Mplus SE`,
      `OpenMx standardized estimate`, `OpenMx delta SE`,
      `Mplus − OpenMx estimate`
    )
  knitr::kable(
    audit_rows,
    caption = "One-to-one regenerated-data audit. The same current-default Mplus and OpenMx values reproduce their saved production values exactly."
  )
}
```

`r if (has_paired_engine_audit) {
  stable <- paired_engine_audit |> rename(selection = all_of(paired_engine_audit_selection_column)) |> filter(selection == "closest_saved_solution", mplus_variant == "ml_constrained") |> slice(1L)
  divergent <- paired_engine_audit |> rename(selection = all_of(paired_engine_audit_selection_column)) |> filter(selection == "most_divergent_saved_solution", mplus_variant == "ml_constrained") |> slice(1L)
  paste0(
    "In the closest pair, the Mplus ML constrained product and OpenMx agree to ",
    fmt_num(abs(stable$constrained_estimate_difference_vs_openmx[[1]]), 6),
    " on the standardized estimate, and their delta-method SEs are ",
    fmt_num(stable$mplus_constraint_delta_se[[1]]), " and ",
    fmt_num(stable$openmx_delta_se[[1]]), ". This verifies that the two model parameterizations can represent the same fitted solution. ",
    "In the selected divergent pair, the Mplus ML constrained estimate is ",
    fmt_num(divergent$mplus_constraint_estimate[[1]]), " whereas OpenMx returns ",
    fmt_num(divergent$openmx_standardized_estimate[[1]]),
    "; changing Mplus from its default MLR to ML does not remove that discrepancy. Thus this part of the gap is optimizer/weak-identification behavior, not a reporting-scale transformation."
  )
} else {
  "The paired-engine audit has not yet been run for this result directory."
}`

`r if (has_paired_engine_audit) {
  stable_fixed <- paired_engine_audit |> rename(selection = all_of(paired_engine_audit_selection_column)) |> filter(selection == "closest_saved_solution", mplus_variant == "current_default") |> slice(1L)
  stable_mlr_delta <- paired_engine_audit |> rename(selection = all_of(paired_engine_audit_selection_column)) |> filter(selection == "closest_saved_solution", mplus_variant == "default_constrained") |> slice(1L)
  stable_ml_delta <- paired_engine_audit |> rename(selection = all_of(paired_engine_audit_selection_column)) |> filter(selection == "closest_saved_solution", mplus_variant == "ml_constrained") |> slice(1L)
  paste0(
    "For that same stable pair, the current MLR fixed-multiplier SE is ",
    fmt_num(stable_fixed$mplus_fixed_multiplier_se[[1]]), ", the MLR product-delta SE is ",
    fmt_num(stable_mlr_delta$mplus_constraint_delta_se[[1]]), ", and the ML product-delta SE is ",
    fmt_num(stable_ml_delta$mplus_constraint_delta_se[[1]]),
    ". Thus both the post-hoc reporting rule and the MLR-versus-ML choice can move an interval even when the fitted point estimate agrees."
  )
}`

The divergent Mplus outputs nevertheless state that estimation terminated
normally. They also report information-matrix condition ratios of about
$1.22\times10^{-14}$ and $3.61\times10^{-16}$ and explicitly flag possible
model nonidentification. The saved current-run rows mark those fits as
noncritical, because the VH Mplus-warning screen does not currently treat the
`CONDITION NUMBER` / `MODEL NONIDENTIFICATION` wording as a critical warning.
This is therefore a concrete eligibility-screening gap, rather than evidence
that the two models are stably estimating different parameters.

The audit also identifies an inference mismatch in the current VH extraction:
its historical-scale Mplus row takes the SE of the raw $z \leftarrow s$ path
and multiplies it by the fitted slope SD.  That treats the fitted SD as fixed.
Lai's OpenMx result instead obtains `mxSE(stdbeta)` for the product itself,
including uncertainty in the slope variance and its covariance with the path.
The ML `MODEL CONSTRAINT` audit is the direct Mplus analogue of the latter.
This difference can be material, but it is not sufficient to explain every
outlying seed because some paired fits have different point estimates before
standardization.

At the condition-summary level, the historical OpenMx coverage is also based
on a more selective usable-interval population.  At $J=30$, OpenMx retains a
mean 91.3% of replications for interval calculation (with some cells near
50%), while current Mplus retains about 98.5%.  Therefore the high historical
OpenMx coverage should not be read as an engine-only advantage until coverage
is recomputed under a common numerical-eligibility rule and a common
product-delta estimand.

If `params$historical_figure_reference` points to a locally saved image of the
published Figure 2, render it alongside this panel during review.  The visual
comparison should ask whether the sign and correlation pattern agree, not
whether the current all-method VH analysis matches a method-specific Lai scale.

```{r published-reference, fig.cap="Optional local reference image of Lai's published Figure 2."}
published_reference <- resolve_repo_path(params$historical_figure_reference)
if (!is.null(published_reference) && file.exists(published_reference)) {
  knitr::include_graphics(published_reference)
}
```

## 3. VH-primary performance: seven methods on a common latent-SD scale

The VH-primary view uses the `latent_sd` reporting scale and the VH eligibility
rule.  Each point in the next plot is a condition-level Monte Carlo mean bias;
the boxplots summarize the 486-condition distribution rather than pooling
replications across different DGMs.

```{r vh-instability-table}
if (!is.null(vh_instability_summary)) {
  vh_instability_table <- vh_instability_summary |>
    mutate(
      Method = recode(method, !!!method_labels),
      `Median bias` = fmt_num(median_bias),
      `Median |bias|` = fmt_num(median_abs_bias),
      `Maximum |bias|` = fmt_num(max_abs_bias, 2)
    ) |>
    select(Method, `Median bias`, `Median |bias|`,
           `Maximum |bias|`, `Conditions: |bias| > .1` = n_abs_bias_over_01,
           `Conditions: |bias| > 1` = n_abs_bias_over_1)
  knitr::kable(
    vh_instability_table,
    caption = "Central performance and extreme condition-level bias on the VH-primary scale"
  )
}
```

```{r vh-performance-table}
vh_performance <- latent_df |>
  group_by(method_label) |>
  summarise(
    conditions = n(),
    `Mean bias` = safe_mean(bias),
    `Median |bias|` = safe_median(abs(bias)),
    `Mean RMSE` = safe_mean(rmse),
    `Mean coverage` = safe_mean(coverage),
    `Mean VH eligibility` = safe_mean(vh_analysis_eligibility_rate),
    .groups = "drop"
  ) |>
  mutate(
    `Mean bias` = fmt_num(`Mean bias`),
    `Median |bias|` = fmt_num(`Median |bias|`),
    `Mean RMSE` = fmt_num(`Mean RMSE`),
    `Mean coverage` = fmt_pct(`Mean coverage`),
    `Mean VH eligibility` = fmt_pct(`Mean VH eligibility`)
  )
knitr::kable(vh_performance, caption = "Condition-weighted VH-primary performance summary")
```

```{r vh-primary-companion, fig.cap="VH-primary central-range companion. The fixed ±0.2 range is intentional; rare larger failures are shown in the following diagnostic."}
vh_plot_path <- file.path(figure_dir, "vh_primary_companion.png")
if (file.exists(vh_plot_path)) {
  knitr::include_graphics(vh_plot_path)
} else {
  knitr::asis_output(
    "**Companion Figure 2 panel not yet generated.** It will appear after the post-estimation figure command has run."
  )
}
```

```{r vh-primary-extremes, fig.cap="Full absolute-bias distribution on a log scale. This retains, rather than trims, rare unstable conditions."}
vh_extreme_path <- file.path(figure_dir, "vh_primary_extremes.png")
if (file.exists(vh_extreme_path)) {
  knitr::include_graphics(vh_extreme_path)
}
```

The central-range panel shows that the closed-form dual estimator is the main
source of systematic negative central bias when $\beta_{Zu1}=.4$, while the
two Fuller estimators, 2S-PA, MSEM, and oracle are generally much closer to
zero.  The log-scale diagnostic adds the complementary fact that naïve dual
BLUP has rare, severe failures even though its median is near zero.  Neither
claim is visible in the original unbounded all-method boxplot by itself.

### VH-primary coverage and Type I error companions

The following two panels extend the interval diagnostics to all seven VH
methods, including Fuller closed-form and Fuller alpha-stepdown.  They use the
common latent-SD reporting view and VH primary eligibility rule.  Unlike the
three-method historical panels above, these are not Lai Figure 3/4 analogues:
Fuller has no historical Lai method-specific standardization.  A common
constant scale does not alter interval inclusion or CI exclusion, but the
explicit common-scale convention keeps the method comparison and retention rule
aligned with the VH bias panels.

```{r vh-primary-coverage, fig.cap="VH-primary companion: empirical 95% coverage for all seven methods."}
vh_coverage_plot_path <- file.path(figure_dir, "vh_primary_coverage_companion.png")
if (file.exists(vh_coverage_plot_path)) {
  knitr::include_graphics(vh_coverage_plot_path)
} else {
  knitr::asis_output("**VH-primary coverage companion not yet generated.** Run the post-estimation figure command, then re-render this report.")
}
```

The focused coverage panel makes departures around 95% readable.  It
intentionally clips values below 70%, so the full-range diagnostic below is
required for judging catastrophic failure rather than only near-nominal
performance.

```{r vh-primary-coverage-extremes, fig.cap="Full-range VH-primary coverage distribution. This exposes values hidden by the focused 70%--100% panel."}
vh_coverage_extremes_path <- file.path(figure_dir, "vh_primary_coverage_extremes.png")
if (file.exists(vh_coverage_extremes_path)) {
  knitr::include_graphics(vh_coverage_extremes_path)
}
```

```{r vh-primary-type1, fig.cap="VH-primary companion: CI-based Type I error for all seven methods, at beta[Z.u1] = 0."}
vh_type1_plot_path <- file.path(figure_dir, "vh_primary_type1_companion.png")
if (file.exists(vh_type1_plot_path)) {
  knitr::include_graphics(vh_type1_plot_path)
} else {
  knitr::asis_output("**VH-primary Type I-error companion not yet generated.** Run the post-estimation figure command, then re-render this report.")
}
```

```{r vh-primary-inference-table}
if (!is.null(vh_primary_inference_summary)) {
  vh_primary_inference_table <- vh_primary_inference_summary |>
    group_by(method) |>
    summarise(
      `Mean coverage` = safe_mean(coverage),
      `Mean Type I error` = safe_mean(type1_error[beta_zu1 == 0]),
      `Mean retained replications` = safe_mean(n_retained),
      .groups = "drop"
    ) |>
    mutate(
      Method = recode(method, !!!method_labels),
      `Mean coverage` = fmt_pct(`Mean coverage`),
      `Mean Type I error` = fmt_pct(`Mean Type I error`),
      `Mean retained replications` = fmt_num(`Mean retained replications`, 1)
    ) |>
    select(Method, `Mean coverage`, `Mean Type I error`, `Mean retained replications`)
  knitr::kable(
    vh_primary_inference_table,
    caption = "Condition-weighted VH-primary interval performance; Type I error is evaluated only when beta[Z.u1] = 0"
  )
}
```

Closed-form dual is the clear interval failure: its condition-weighted mean
coverage is `r fmt_pct(vh_inference_value("closed_form_dual", "mean_coverage"))`,
with `r vh_inference_value("closed_form_dual", "n_coverage_below_70")`
conditions below 70% and a minimum near
`r fmt_pct(vh_inference_value("closed_form_dual", "min_coverage"))`.
Both Fuller variants recover roughly 96% coverage and Type I error near 4%,
with little practical separation between ordinary Fuller and alpha-stepdown.
Naïve dual BLUP and MSEM are modestly anti-conservative overall; oracle is
close to nominal.  Read these averages with the faceted panels because small
clusters and particular correlations concentrate the departures.

## 4. Reporting-scale sensitivity: separate rescaling from estimator behavior

Raw and latent-SD rows report the same fitted estimator on different common
condition-level scales.  The historical standardized row is different: for
the directly comparable methods it preserves Lai's method-specific multiplier.
Consequently, this section is a diagnostic of reporting conventions, not a
competition between differently scaled numerical values.

```{r scale-sensitivity, fig.width=11, fig.height=6, fig.cap="Central condition-level bias for the three overlapping methods under each reporting view. The fixed range makes shifts between views visible; tail failures are reported separately."}
scale_plot_df <- summary_df |>
  filter(method %in% c("naive_dual_blup", "lai_2spa", "msem")) |>
  mutate(scale_label = recode(
    as.character(reporting_scale),
    raw = "Raw",
    latent_sd = "Common latent-SD",
    lai_original_standardized = "Lai historical standardized"
  ))

ggplot(scale_plot_df, aes(x = method_label, y = bias, fill = method_label)) +
  geom_hline(yintercept = 0, colour = "grey35", linewidth = 0.4) +
  geom_boxplot(colour = "grey25", outlier.alpha = 0.2) +
  scale_fill_manual(values = method_palette) +
  coord_flip(ylim = c(-0.15, 0.15)) +
  facet_wrap(~ scale_label) +
  labs(
    x = NULL,
    y = "Condition-level mean bias (view-specific units)",
    title = "Reporting-scale sensitivity",
    subtitle = "Fixed display range: -0.15 to 0.15; historical colors match Lai's original method mapping",
    fill = NULL
  ) +
  theme_bw(base_size = 10) +
  theme(legend.position = "bottom")
```

```{r historical-multiplier-table}
multiplier_summary <- historical_df |>
  group_by(method_label) |>
  summarise(
    `Mean Lai historical multiplier` = safe_mean(mean_lai_original_method_multiplier),
    `Availability` = safe_mean(lai_original_reporting_available_rate),
    .groups = "drop"
  ) |>
  mutate(
    `Mean Lai historical multiplier` = fmt_num(`Mean Lai historical multiplier`),
    Availability = fmt_pct(Availability)
  )
knitr::kable(
  multiplier_summary,
  caption = "Method-specific historical scaling diagnostics"
)
```

The key finding is method-specific: the dual-BLUP distribution shifts
downward on Lai's historical view but is centrally near zero on both common
views.  Raw-to-latent-SD scaling alone cannot create or remove a bias pattern,
because it is a fixed condition-level multiplier applied to both estimate and
truth.  The extra change on the historical view comes from using the observed
EB-slope SD for the naïve method while retaining the true latent-slope SD in
the benchmark.  This is why a negative historical dual-EB bias can coexist
with little central raw-coefficient bias.

## 5. Slope-SD and reliability diagnostics: test the calibration explanation

The posterior reliability plotted here is calculated from the true full
two-dimensional random-effect covariance and Lai's RMS-scaled within-cluster
design.  It should not be replaced by `lambda22` when random effects are
correlated.  The first plot removes duplicate values across the number of
clusters and $\beta_{Zu1}$, neither of which enters the posterior-reliability
calculation, and shows the remaining determinants explicitly.

```{r reliability-by-size, fig.width=10, fig.height=5.5, fig.cap="True posterior slope reliability by homogeneous cluster size, ICC, slope/intercept variance ratio, and random-effect correlation."}
reliability_df <- raw_df |>
  distinct(
    clus_size, icc, vr_u1_u0, cor_u0_u1,
    dgm_posterior_slope_reliability
  ) |>
  mutate(
    clus_size = factor(clus_size),
    icc = factor(icc),
    vr_u1_u0 = factor(vr_u1_u0),
    cor_u0_u1 = factor(cor_u0_u1)
  )

ggplot(
  reliability_df,
  aes(
    x = clus_size,
    y = dgm_posterior_slope_reliability,
    colour = icc,
    shape = vr_u1_u0
  )
) +
  geom_jitter(width = 0.13, height = 0, alpha = 0.8, size = 2) +
  facet_wrap(~ cor_u0_u1, nrow = 1, labeller = label_both) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = "Homogeneous cluster size",
    y = "True posterior slope reliability",
    colour = "ICC",
    shape = "Slope/intercept\nvariance ratio",
    title = "Reliability is not interchangeable with cluster size",
    subtitle = "Each size retains substantial variation from ICC, variance ratio, and random-effect correlation"
  ) +
  theme_bw(base_size = 10)
```

```{r reliability-by-size-table}
reliability_by_size <- reliability_df |>
  group_by(clus_size) |>
  summarise(
    `Minimum reliability` = min(dgm_posterior_slope_reliability),
    `Median reliability` = median(dgm_posterior_slope_reliability),
    `Maximum reliability` = max(dgm_posterior_slope_reliability),
    .groups = "drop"
  ) |>
  mutate(across(-clus_size, fmt_num))
knitr::kable(
  reliability_by_size,
  caption = "True posterior slope reliability varies substantially within every homogeneous cluster-size level"
)
```

```{r slope-sd-ratios, fig.width=10, fig.height=5, fig.cap="Stage-1 slope-SD diagnostics relative to the DGM population slope SD."}
slope_sd_df <- raw_df |>
  filter(method == "naive_dual_blup") |>
  select(
    condition_id, dgm_posterior_slope_reliability,
    mean_eb_to_population_slope_sd,
    mean_corrected_to_population_slope_sd,
    mean_fitted_to_population_slope_sd
  ) |>
  pivot_longer(
    cols = starts_with("mean_") ,
    names_to = "diagnostic",
    values_to = "slope_sd_ratio"
  ) |>
  mutate(diagnostic = recode(
    diagnostic,
    mean_eb_to_population_slope_sd = "EB / population SD",
    mean_corrected_to_population_slope_sd = "Corrected / population SD",
    mean_fitted_to_population_slope_sd = "Fitted / population SD"
  ))

ggplot(slope_sd_df, aes(x = dgm_posterior_slope_reliability, y = slope_sd_ratio)) +
  geom_hline(yintercept = 1, colour = "grey35", linetype = "dashed") +
  geom_point(alpha = 0.45, colour = "#7a3e9d", size = 1.4) +
  facet_wrap(~ diagnostic, nrow = 1) +
  labs(
    x = "True posterior slope reliability",
    y = "Mean slope SD / population slope SD",
    title = "Stage-1 slope-SD diagnostics versus true reliability"
  ) +
  theme_bw(base_size = 10)
```

The slope-SD ratios use the same Stage-1 fit attached to the naive dual BLUP
raw summary, avoiding duplicate copies of the diagnostics across reporting
views.  A ratio of one is a diagnostic reference, not a convergence criterion.

The reliability result is consistent with cluster size being an important
determinant, but rejects the stronger claim that it is merely a proxy for
cluster size: at size 3 the reliability spans approximately .07--.86, and at
size 25 it still spans approximately .40--.98.  The EB slope SD is most
attenuated at low reliability, whereas the fitted slope SD remains close to
the population slope SD across the reliability range.  This makes the
slope-SD correction a useful diagnostic for the observed bias pattern rather
than a substitute for modelling reliability itself.  Because reliability is
emergent rather than independently randomized here, however, a trend against
reliability in this study cannot be interpreted as the isolated causal effect
of reliability; that is precisely the role of the direct calibration in VH
Study 2.

## 6. Eligibility and convergence: compare the two retention rules directly

VH eligibility and original-Lai eligibility answer different questions.
The historical rule retains runs with `status_code == 0`; the VH-primary rule
also applies estimator-specific finite-estimate, finite-SE, design, and fit
diagnostics.  The following plot displays the average condition-level rate for
each rule; it should accompany any comparison that changes the reporting scale
or estimator bundle.

```{r eligibility-rates, fig.width=10, fig.height=5, fig.cap="Average condition-level retention under the original-Lai and VH-primary rules."}
eligibility_df <- raw_df |>
  select(method_label, vh_analysis_eligibility_rate, lai_original_eligibility_rate) |>
  pivot_longer(
    cols = c(vh_analysis_eligibility_rate, lai_original_eligibility_rate),
    names_to = "rule",
    values_to = "eligibility_rate"
  ) |>
  mutate(rule = recode(
    rule,
    vh_analysis_eligibility_rate = "VH primary",
    lai_original_eligibility_rate = "Original Lai"
  ))

ggplot(eligibility_df, aes(x = method_label, y = eligibility_rate, shape = rule)) +
  stat_summary(fun = mean, geom = "point", position = position_dodge(width = 0.45), size = 2.6, colour = "#1f4e79") +
  coord_flip(ylim = c(0.85, 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(
    x = NULL,
    y = "Mean condition-level eligibility rate",
    shape = NULL,
    title = "Eligibility depends on the reporting rule",
    subtitle = "Focused 85%--100% range; exact rates and overlap counts appear below"
  ) +
  theme_bw(base_size = 11)
```

```{r eligibility-overlap-table}
eligibility_overlap <- raw_df |>
  group_by(method_label) |>
  summarise(
    `VH eligible` = safe_mean(vh_analysis_eligibility_rate),
    `Lai eligible` = safe_mean(lai_original_eligibility_rate),
    `Both eligible (mean count)` = safe_mean(n_eligible_both),
    `Lai-only (mean count)` = safe_mean(n_lai_original_only),
    `VH-only (mean count)` = safe_mean(n_vh_only),
    `Convergence` = safe_mean(convergence),
    .groups = "drop"
  ) |>
  mutate(
    `VH eligible` = fmt_pct(`VH eligible`),
    `Lai eligible` = fmt_pct(`Lai eligible`),
    `Both eligible (mean count)` = fmt_num(`Both eligible (mean count)`, 1),
    `Lai-only (mean count)` = fmt_num(`Lai-only (mean count)`, 1),
    `VH-only (mean count)` = fmt_num(`VH-only (mean count)`, 1),
    Convergence = fmt_pct(Convergence)
  )
knitr::kable(
  eligibility_overlap,
  caption = "Condition-averaged eligibility overlap and convergence"
)
```

For this run, the VH screen is a strict subset of the original-Lai screen:
there are no VH-only replications.  The material differences occur for 2S-PA
and naïve dual BLUP, for which about 2% of replications per condition are
retained by the original status-code rule but excluded from the VH-primary
analysis.  Thus the rare large naïve-BLUP biases are not an artifact of a more
permissive VH retention rule; the stricter VH screen is already in force.

## Next steps and conclusions

The cross-study interpretation here is benchmarked against the current
[VH Study 2 analysis](../vig_hallquist_2026/analysis/vh_2026_study2.Rmd), with
the [VH Study 1 analysis](../vig_hallquist_2026/analysis/vh_2026_study1.Rmd)
used to distinguish the random-effect-as-predictor and
random-effect-as-outcome settings.

### The overlapping Lai point-estimate result replicates

The strongest conclusion is positive: when method, reporting scale, bias
statistic, and eligibility are aligned, the current machinery reproduces
Lai's principal pattern for the estimators that overlap.  Dual BLUP is
negatively biased on Lai's historical scale, while 2S-PA and the VH/Mplus MSEM
are near zero.  The original OpenMx MSEM add-on provides the stricter
implementation-and-interval check without changing the VH seven-method bundle.
The apparent failure to reproduce a correlation-signed pattern
was a terminology mismatch: Lai's sign-changing lines are slope-only EB
models, not the dual-EB model used here.

The remaining replication gap is inferential rather than a broad failure of
the DGM or estimator implementation.  Dual BLUP intervals are close to Lai's
original aggregate behavior, while 2S-PA is modestly more conservative and
VH/Mplus MSEM more anti-conservative in this run.  The historical OpenMx MSEM
now supplies the original standardized delta-method interval and p-value; a
fully strict Figures 3--4 replication still requires all six original methods,
especially 2S-PAB and the two slope-only EB specifications.

### The scale result narrows what can explain VH Study 2

The historical negative dual-EB bias is not a generic property of using BLUPs
as predictors.  It is much weaker centrally on the raw and common latent-SD
views and becomes prominent when the naïve estimate is multiplied by the
shrunken observed EB-slope SD but compared with a truth scaled by the latent
slope SD.  VH Study 2's negative **raw/common-scale** bias in low-reliability
conditions therefore cannot be dismissed as the same historical scaling
artifact.  Its source must lie in estimator behavior, DGM structure,
conditioning, or their interaction.

```{r reliability-method-pattern, fig.width=11, fig.height=6, fig.cap="Descriptive median bias across posterior-reliability bands in the Lai DGM. Reliability is emergent here, so these are associations rather than isolated reliability effects."}
reliability_method_summary <- latent_df |>
  filter(beta_zu1 == 0.4) |>
  mutate(
    reliability_band = cut(
      dgm_posterior_slope_reliability,
      breaks = c(-Inf, 0.25, 0.50, 0.75, Inf),
      labels = c("< .25", ".25--.49", ".50--.74", ">= .75"),
      right = FALSE
    )
  ) |>
  group_by(method_label, reliability_band) |>
  summarise(
    median_bias = safe_median(bias),
    conditions = n(),
    .groups = "drop"
  )

ggplot(
  reliability_method_summary,
  aes(x = reliability_band, y = median_bias, colour = method_label, group = method_label)
) +
  geom_hline(yintercept = 0, colour = "grey35") +
  geom_line(linewidth = 0.7) +
  geom_point(size = 2) +
  scale_colour_manual(values = method_palette) +
  coord_cartesian(ylim = c(-0.10, 0.05)) +
  labs(
    x = "True posterior slope-reliability band",
    y = "Median condition-level bias",
    colour = NULL,
    title = "Low reliability separates uncorrected closed-form OLS from corrected methods",
    subtitle = "Common latent-SD scale; beta[Zu1] = .4; VH eligibility"
  ) +
  theme_bw(base_size = 10) +
  theme(legend.position = "bottom")
```

This descriptive bridge resembles the broader VH pattern in one important
respect: uncorrected closed-form OLS is negatively biased, particularly when
reliability is low, and both Fuller corrections materially reduce that bias.
It differs in another: the central naïve dual-BLUP bias in the Lai DGM is near
zero or slightly positive across reliability bands, rather than showing the
pronounced low-reliability negative raw bias observed in VH Study 2.  Rare
naïve-BLUP failures remain consequential, so medians, trimmed means, and tail
counts should continue to accompany ordinary means.

### Reliability is not merely cluster size, but it is not isolated here

The diagnostic range within each homogeneous cluster size is wide enough to
reject “reliability is just cluster size” as a description of the Lai DGM.
ICC, the slope/intercept variance ratio, and random-effect correlation all
matter.  Nevertheless, Lai reliability is an emergent combination of these
factors rather than a randomized design dimension.  Associations in the plot
above therefore cannot identify a pure reliability effect.  The direct
posterior-reliability calibration in the VH project remains essential and is
the stronger design for testing reliability-specific hypotheses.

### Recommended next analyses

1. Run a bridge experiment that starts from the Lai DGM and introduces the VH
   reliability calibration one component at a time.  Hold the raw/common
   reporting scale and eligibility rule fixed, then change the cluster-size
   design, reliability calibration, and outcome model sequentially.  This will
   localize where the VH low-reliability negative bias enters.
2. Compare matched reliability bands across Lai and VH using median bias,
   20%-trimmed bias, RMSE, and tail-failure counts.  Do not rely on the mean
   alone for naïve BLUP, whose rare failures can dominate it.
3. Use the source-faithful OpenMx MSEM versus VH/Mplus MSEM contrast to check
   whether the $J=30$ coverage gap comes from model parameterization,
   standardization/SE propagation, or software.  Then check 2S-PA interval
   construction and—if exact paper replication is needed—add 2S-PAB plus the
   centered and uncentered slope-only EB methods outside the VH bundle.
4. In VH reporting, keep raw/common latent-SD results primary and retain Lai's
   historical method-specific scale only as a labeled diagnostic.  Continue to
   show both eligibility rules and slope-SD ratios so scale, retention, and
   estimator effects cannot be conflated.

Overall, the replication increases confidence that the shared 2S-PA,
VH/Mplus MSEM, and dual-BLUP machinery is recovering Lai's point-estimate
result, while the eighth OpenMx method is designed to provide a source-faithful
historical control without changing the VH bundle.  It does
not explain away the different VH Study 2 pattern.  Instead, it sharpens the
next question: which aspect of VH's direct reliability calibration or DGM
turns the low-reliability dual-BLUP association negative on a common scale?

## Reproducibility notes

The report is parameterized.  After generating the Figure 2 post-processing
outputs, render it from the repository root with:

```sh
quarto render reports/lai_study1_vh_review.qmd \
  -P results_dir:outputs/lai_study1_vh_production_20260731
```

Optionally supply `figure_dir` when figures live elsewhere and
`historical_figure_reference` for a locally saved copy of Lai's published
Figure 2.  This report does not rerun any models or alter the aggregate output.
