---
title: "Lai Study 1 Replication: Fuller Stepdown, Fuller, EIV, Naive Dual EB, and Lai"
author: "Codex"
date: "May 19, 2026"
format:
  html:
    toc: true
    code-fold: true
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
params:
  output_dir: "/proj/mnhallqlab/projects/blup_unshrinking/lai_replication/outputs/lai_sim_18may2026"
---

## Purpose

This report summarizes the Study 1 replication run in:

`r params$output_dir`

The main comparison is deliberately focused on the methods most relevant to the current question:

- `fuller_stepdown`, the Fuller stepdown variant;
- `fuller`, the original Fuller-style estimator;
- dual EIV (`eiv_dual_corrected`) and its HC3 standard-error variant;
- naive dual empirical-Bayes / BLUP regression (`naive_dual_eb`) and its HC3 standard-error variant;
- Lai's `2S-PA` model, with `2S-PAA` retained as a secondary Lai comparator.

The performance criteria are bias, RMSE, 95% interval coverage, standard-error calibration, and usable-estimate rate. The report treats singular and near-singular Stage 1 random-effect fits as part of method performance because those cases are exactly where dual predictors can become ill-conditioned.

## Setup

```{r setup}
library(dplyr)
library(tidyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(scales)
library(data.table)

safe_mean <- function(x) {
  if (length(x) == 0L || all(is.na(x))) {
    return(NA_real_)
  }
  mean(x, na.rm = TRUE)
}

safe_sd <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) < 2L) {
    return(NA_real_)
  }
  stats::sd(x)
}

safe_median <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) == 0L) {
    return(NA_real_)
  }
  stats::median(x)
}

fmt_num <- function(x, digits = 3) {
  ifelse(is.na(x), NA_character_, formatC(x, digits = digits, format = "f"))
}

fmt_sci <- function(x, digits = 2) {
  ifelse(is.na(x), NA_character_, formatC(x, digits = digits, format = "e"))
}

fmt_mixed <- function(x, digits = 3, sci_cut = 1000) {
  ifelse(
    is.na(x),
    NA_character_,
    ifelse(abs(x) >= sci_cut, fmt_sci(x, 2), fmt_num(x, digits))
  )
}

fmt_pct <- function(x, accuracy = 0.1) {
  ifelse(is.na(x), NA_character_, scales::percent(x, accuracy = accuracy))
}

resolve_run_dir <- function(path) {
  if (grepl("^/", path)) {
    normalizePath(path, mustWork = FALSE)
  } else {
    normalizePath(file.path(getwd(), path), mustWork = FALSE)
  }
}

read_selected <- function(path, select) {
  available <- names(data.table::fread(path, nrows = 0L, showProgress = FALSE))
  data.table::fread(path, select = intersect(select, available), showProgress = FALSE)
}

read_many_selected <- function(paths, select, missing_ok = FALSE) {
  paths <- paths[file.exists(paths)]
  if (length(paths) == 0L) {
    if (missing_ok) {
      return(tibble::tibble())
    }
    stop("No input files found.", call. = FALSE)
  }
  data.table::rbindlist(
    lapply(paths, read_selected, select = select),
    fill = TRUE,
    use.names = TRUE
  )
}

parse_chunk_replication_index <- function(paths) {
  if (length(paths) == 0L) {
    return(tibble::tibble(path = character(), first_condition = integer(), last_condition = integer()))
  }
  ranges <- lapply(paths, function(path) {
    match <- regexec("conditions_([0-9]{4})_([0-9]{4})_replication_results", basename(path))
    parts <- regmatches(basename(path), match)[[1]]
    if (length(parts) != 3L) {
      return(NULL)
    }
    tibble::tibble(
      path = path,
      first_condition = as.integer(parts[[2]]),
      last_condition = as.integer(parts[[3]])
    )
  })
  dplyr::bind_rows(ranges)
}

coerce_common_types <- function(df) {
  numeric_cols <- intersect(
    c(
      "estimate", "se", "ci_low", "ci_high", "truth",
      "convergence", "bias", "coverage", "rmse", "mean_estimate",
      "mx_condition_number", "eiv_measurement_weight_requested",
      "eiv_measurement_weight_used", "eiv_latent_cov_min_eigen",
      "eiv_latent_cov_condition_number", "fuller_lambda1",
      "fuller_lambda2", "fuller_sigma2", "fuller_weight_min",
      "fuller_weight_max", "fuller_correction_c",
      "fuller_measurement_weight_requested", "fuller_measurement_weight_used",
      "fuller_sx_star_condition", "fuller_sx_star_min_eigen",
      "fuller_sx_uncorr_condition", "fuller_sx_uncorr_min_eigen",
      "fuller_sx_observed_max_eigen", "fuller_sx_uncorr_relative_min_eigen",
      "fuller_scaling_condition", "fuller_scaling_min_eigen",
      "fuller_scaling_observed_max_eigen", "fuller_scaling_relative_min_eigen",
      "fuller_reference_se", "fuller_se_ratio", "fuller_auto_guard_score",
      "fuller_auto_full_weight_se_ratio", "fuller_auto_search_evaluations",
      "stage1_re_corr", "stage1_eb_corr", "stage1_design_kappa",
      "mean_stage1_re_corr", "mean_stage1_eb_corr",
      "mean_stage1_design_kappa", "icc", "vr_u1_u0", "cor_u0_u1",
      "beta_zu1", "sigma2", "var_u1", "sigma_z"
    ),
    names(df)
  )
  integer_cols <- intersect(
    c("condition_id", "rep", "num_clus", "clus_size", "status_code", "n_rep", "n_success"),
    names(df)
  )
  logical_cols <- intersect(
    c(
      "stage1_singular_problem", "stage1_lmer_singular", "mx_info_definite",
      "eiv_regularized", "fuller_auto_tempered", "fuller_auto_guard_pass",
      "fuller_auto_full_weight_guard_pass", "fuller_auto_search_nonmonotone"
    ),
    names(df)
  )

  if (length(numeric_cols) > 0L) {
    df <- df %>% mutate(across(all_of(numeric_cols), ~ suppressWarnings(as.numeric(.x))))
  }
  if (length(integer_cols) > 0L) {
    df <- df %>% mutate(across(all_of(integer_cols), ~ suppressWarnings(as.integer(.x))))
  }
  if (length(logical_cols) > 0L) {
    df <- df %>% mutate(across(all_of(logical_cols), as.logical))
  }
  tibble::as_tibble(df)
}

metric_summary <- function(df, group_vars = character()) {
  df %>%
    mutate(
      success = !is.na(estimate),
      covered = ifelse(!is.na(ci_low) & !is.na(ci_high) & !is.na(truth), ci_low <= truth & ci_high >= truth, NA),
      bias = estimate - truth,
      sq_error = bias^2
    ) %>%
    group_by(across(all_of(group_vars))) %>%
    summarise(
      Replications = n(),
      Conditions = n_distinct(condition_id),
      Success_Rate = safe_mean(success),
      Mean_Bias = safe_mean(bias),
      Median_Bias = safe_median(bias),
      RMSE = if (all(is.na(sq_error))) NA_real_ else sqrt(mean(sq_error, na.rm = TRUE)),
      Coverage = safe_mean(covered),
      Mean_SE = safe_mean(se),
      Empirical_SD = safe_sd(estimate),
      SE_to_SD = Mean_SE / Empirical_SD,
      .groups = "drop"
    )
}

format_metric_table <- function(df) {
  df %>%
    mutate(
      Success_Rate = fmt_pct(Success_Rate),
      Mean_Bias = fmt_mixed(Mean_Bias),
      Median_Bias = fmt_mixed(Median_Bias),
      RMSE = fmt_mixed(RMSE),
      Coverage = fmt_pct(Coverage),
      Mean_SE = fmt_mixed(Mean_SE),
      Empirical_SD = fmt_mixed(Empirical_SD),
      SE_to_SD = fmt_num(SE_to_SD, 2)
    )
}

metric_value <- function(tbl, method_name, metric) {
  out <- tbl %>%
    filter(method == method_name) %>%
    pull({{ metric }})
  if (length(out) == 0L) NA_real_ else out[[1]]
}

stage1_metric_value <- function(tbl, method_name, stratum_name, metric) {
  out <- tbl %>%
    filter(method == method_name, stage1_stratum == stratum_name) %>%
    pull({{ metric }})
  if (length(out) == 0L) NA_real_ else out[[1]]
}

run_dir <- resolve_run_dir(params$output_dir)
condition_dir <- file.path(run_dir, "conditions")
summary_path <- file.path(run_dir, "lai_apples_to_apples_summary.csv")
stage1_path <- file.path(run_dir, "lai_apples_to_apples_stage1_problem_summary.csv")
replication_path <- file.path(run_dir, "lai_apples_to_apples_replication_results.csv.gz")

manifest_files <- list.files(
  run_dir,
  pattern = "^lai_apples_to_apples_chunk_.*_manifest[.]csv$",
  full.names = TRUE
)
manifest_df <- if (length(manifest_files) > 0L) {
  coerce_common_types(data.table::rbindlist(lapply(manifest_files, data.table::fread), fill = TRUE))
} else {
  tibble::tibble()
}

summary_files <- list.files(
  condition_dir,
  pattern = "^condition_[0-9]{4}_summary[.]csv$",
  full.names = TRUE
)
stage1_files <- list.files(
  condition_dir,
  pattern = "^condition_[0-9]{4}_stage1_problem_summary[.]csv$",
  full.names = TRUE
)

summary_df <- if (file.exists(summary_path)) {
  data.table::fread(summary_path, showProgress = FALSE)
} else {
  data.table::rbindlist(lapply(summary_files, data.table::fread), fill = TRUE, use.names = TRUE)
}
summary_df <- coerce_common_types(summary_df) %>%
  filter(study == "study1")

stage1_df <- if (file.exists(stage1_path)) {
  data.table::fread(stage1_path, showProgress = FALSE)
} else {
  data.table::rbindlist(lapply(stage1_files, data.table::fread), fill = TRUE, use.names = TRUE)
}
stage1_df <- coerce_common_types(stage1_df) %>%
  filter(study == "study1")

expected_condition_ids <- if (all(c("condition_id", "study") %in% names(manifest_df))) {
  sort(unique(manifest_df$condition_id[manifest_df$study == "study1"]))
} else {
  sort(unique(summary_df$condition_id))
}
available_condition_ids <- sort(unique(summary_df$condition_id))
missing_condition_ids <- setdiff(expected_condition_ids, available_condition_ids)
missing_condition_label <- if (length(missing_condition_ids) > 0L) {
  paste(missing_condition_ids, collapse = ", ")
} else {
  "none"
}
run_completion_note <- if (length(missing_condition_ids) > 0L) {
  paste0(
    "The manifests list ", length(expected_condition_ids),
    " Study 1 conditions; completed output files are present for ",
    length(available_condition_ids), ". Missing condition id(s): ",
    missing_condition_label, "."
  )
} else {
  paste0("Completed output files are present for all ", length(available_condition_ids), " manifest conditions.")
}

replication_select <- c(
  "method", "estimate", "se", "ci_low", "ci_high", "status_code", "se_type",
  "mx_issue_class", "eiv_measurement_weight_used", "eiv_regularized",
  "eiv_latent_cov_min_eigen", "eiv_latent_cov_condition_number",
  "fuller_lambda1", "fuller_lambda2", "fuller_sigma2", "fuller_weight_min",
  "fuller_weight_max", "fuller_correction_c",
  "fuller_measurement_weight_requested", "fuller_measurement_weight_used",
  "fuller_auto_tempered", "fuller_sx_star_condition",
  "fuller_sx_star_min_eigen", "fuller_sx_uncorr_condition",
  "fuller_sx_uncorr_min_eigen", "fuller_sx_observed_max_eigen",
  "fuller_sx_uncorr_relative_min_eigen", "fuller_scaling_condition",
  "fuller_scaling_min_eigen", "fuller_scaling_observed_max_eigen",
  "fuller_scaling_relative_min_eigen", "fuller_reference_se",
  "fuller_se_ratio", "fuller_auto_guard_pass", "fuller_auto_guard_reason",
  "fuller_auto_guard_score", "fuller_auto_full_weight_guard_pass",
  "fuller_auto_full_weight_guard_reason", "fuller_auto_full_weight_se_ratio",
  "fuller_auto_search_evaluations", "fuller_auto_search_nonmonotone",
  "stage1_singular_problem",
  "stage1_lmer_singular", "stage1_re_corr", "stage1_eb_corr",
  "stage1_design_kappa", "study", "truth", "num_clus", "clus_size", "icc",
  "vr_u1_u0", "cor_u0_u1", "beta_zu1", "condition_id", "rep"
)

chunk_replication_files <- list.files(
  run_dir,
  pattern = "^lai_apples_to_apples_chunk_.*_replication_results[.]csv[.]gz$",
  full.names = TRUE
)
chunk_replication_index <- parse_chunk_replication_index(chunk_replication_files)
if (length(available_condition_ids) > 0L && nrow(chunk_replication_index) > 0L) {
  chunk_replication_index <- chunk_replication_index %>%
    filter(
      last_condition >= min(available_condition_ids),
      first_condition <= max(available_condition_ids)
    )
}
chunk_replication_files <- chunk_replication_index$path
chunk_replication_condition_ids <- sort(unique(unlist(Map(
  seq.int,
  chunk_replication_index$first_condition,
  chunk_replication_index$last_condition
))))
condition_replication_ids <- setdiff(available_condition_ids, chunk_replication_condition_ids)
condition_replication_files <- file.path(
  condition_dir,
  sprintf("condition_%04d_replication_results.csv.gz", condition_replication_ids)
)

replication_df <- if (file.exists(replication_path)) {
  read_selected(replication_path, replication_select)
} else {
  read_many_selected(c(chunk_replication_files, condition_replication_files), replication_select)
}
replication_df <- replication_df %>%
  coerce_common_types() %>%
  filter(study == "study1")

method_labels <- c(
  oracle_dual = "Oracle dual",
  naive_slope_only = "Naive BLUP slope only",
  naive_slope_only_hc3 = "Naive BLUP slope only + HC3 SE",
  centered_slope_only = "Centered BLUP slope only",
  centered_slope_only_hc3 = "Centered BLUP slope only + HC3 SE",
  naive_dual_eb = "Naive dual BLUP OLS",
  naive_dual_eb_hc3 = "Naive dual BLUP OLS + HC3 SE",
  ridge_dual_eb = "Ridge dual BLUP",
  corrected_slope_only = "Corrected slope only",
  corrected_slope_only_hc3 = "Corrected slope only + HC3 SE",
  corrected_dual = "Corrected dual",
  corrected_dual_hc3 = "Corrected dual + HC3 SE",
  eiv_dual_corrected = "Dual EIV",
  eiv_dual_corrected_hc0 = "Dual EIV + HC0 SE",
  eiv_dual_corrected_hc3 = "Dual EIV + HC3 SE",
  eiv_dual_corrected_nearpd = "Dual EIV + near-PD",
  eiv_dual_corrected_nearpd_hc0 = "Dual EIV + near-PD + HC0 SE",
  eiv_dual_corrected_nearpd_hc3 = "Dual EIV + near-PD + HC3 SE",
  eiv_dual_corrected_ridge = "Dual EIV + ridge",
  eiv_dual_corrected_ridge_hc0 = "Dual EIV + ridge + HC0 SE",
  eiv_dual_corrected_ridge_hc3 = "Dual EIV + ridge + HC3 SE",
  fuller = "Fuller",
  fuller_stepdown = "Fuller stepdown",
  lai_2spa = "Lai 2S-PA",
  lai_2spaa = "Lai 2S-PAA"
)

method_order <- unname(method_labels)
label_method <- function(x) {
  out <- unname(method_labels[x])
  ifelse(is.na(out), x, out)
}

focus_methods <- c(
  "fuller",
  "fuller_stepdown",
  "eiv_dual_corrected",
  "eiv_dual_corrected_hc3",
  "naive_dual_eb",
  "naive_dual_eb_hc3",
  "lai_2spa",
  "lai_2spaa"
)

primary_methods <- c(
  "fuller",
  "fuller_stepdown",
  "eiv_dual_corrected_hc3",
  "naive_dual_eb_hc3",
  "lai_2spa"
)

fuller_pair_methods <- c("fuller", "fuller_stepdown")

summary_df <- summary_df %>%
  mutate(method_label = factor(label_method(method), levels = method_order))

stage1_df <- stage1_df %>%
  mutate(
    method_label = factor(label_method(method), levels = method_order),
    stage1_stratum = ifelse(stage1_singular_problem, "Stage 1 singular", "Stage 1 regular")
  )

replication_df <- replication_df %>%
  mutate(
    method_label = factor(label_method(method), levels = method_order),
    stage1_stratum = ifelse(stage1_singular_problem, "Stage 1 singular", "Stage 1 regular")
  )

study1_completed_conditions <- n_distinct(summary_df$condition_id)
study1_total_replications <- n_distinct(paste(replication_df$condition_id, replication_df$rep, sep = ":"))
stage1_singular_rate <- replication_df %>%
  filter(method == "fuller") %>%
  summarise(rate = safe_mean(stage1_singular_problem)) %>%
  pull(rate)

focus_overall_raw <- replication_df %>%
  filter(method %in% focus_methods) %>%
  metric_summary("method_label") %>%
  left_join(tibble(method = names(method_labels), method_label = factor(method_order, levels = method_order)), by = "method_label")

focus_by_stage1_raw <- replication_df %>%
  filter(method %in% focus_methods) %>%
  metric_summary(c("method_label", "stage1_stratum")) %>%
  left_join(tibble(method = names(method_labels), method_label = factor(method_order, levels = method_order)), by = "method_label")

primary_converged_keys <- replication_df %>%
  filter(method %in% primary_methods) %>%
  group_by(condition_id, rep) %>%
  summarise(
    n_primary_methods = n_distinct(method),
    all_primary_converged = all(!is.na(estimate)),
    stage1_singular_problem = any(stage1_singular_problem),
    .groups = "drop"
  ) %>%
  filter(n_primary_methods == length(primary_methods), all_primary_converged) %>%
  select(condition_id, rep, stage1_singular_problem)

primary_converged_rate <- nrow(primary_converged_keys) / study1_total_replications

primary_converged_raw <- replication_df %>%
  semi_join(primary_converged_keys, by = c("condition_id", "rep")) %>%
  filter(method %in% primary_methods) %>%
  metric_summary("method_label") %>%
  left_join(tibble(method = names(method_labels), method_label = factor(method_order, levels = method_order)), by = "method_label")

fuller_success <- metric_value(focus_overall_raw, "fuller", Success_Rate)
stepdown_success <- metric_value(focus_overall_raw, "fuller_stepdown", Success_Rate)
lai_success <- metric_value(focus_overall_raw, "lai_2spa", Success_Rate)
lai_singular_success <- stage1_metric_value(focus_by_stage1_raw, "lai_2spa", "Stage 1 singular", Success_Rate)
fuller_singular_success <- stage1_metric_value(focus_by_stage1_raw, "fuller", "Stage 1 singular", Success_Rate)
stepdown_singular_success <- stage1_metric_value(focus_by_stage1_raw, "fuller_stepdown", "Stage 1 singular", Success_Rate)
fuller_singular_rmse <- stage1_metric_value(focus_by_stage1_raw, "fuller", "Stage 1 singular", RMSE)
stepdown_singular_rmse <- stage1_metric_value(focus_by_stage1_raw, "fuller_stepdown", "Stage 1 singular", RMSE)
fuller_regular_rmse <- stage1_metric_value(focus_by_stage1_raw, "fuller", "Stage 1 regular", RMSE)
stepdown_regular_rmse <- stage1_metric_value(focus_by_stage1_raw, "fuller_stepdown", "Stage 1 regular", RMSE)
naive_singular_rmse <- stage1_metric_value(focus_by_stage1_raw, "naive_dual_eb_hc3", "Stage 1 singular", RMSE)
eiv_singular_rmse <- stage1_metric_value(focus_by_stage1_raw, "eiv_dual_corrected_hc3", "Stage 1 singular", RMSE)
shared_fuller_bias <- metric_value(primary_converged_raw, "fuller", Mean_Bias)
shared_fuller_rmse <- metric_value(primary_converged_raw, "fuller", RMSE)
shared_fuller_coverage <- metric_value(primary_converged_raw, "fuller", Coverage)
shared_fuller_mean_se <- metric_value(primary_converged_raw, "fuller", Mean_SE)
shared_stepdown_bias <- metric_value(primary_converged_raw, "fuller_stepdown", Mean_Bias)
shared_stepdown_rmse <- metric_value(primary_converged_raw, "fuller_stepdown", RMSE)
shared_stepdown_coverage <- metric_value(primary_converged_raw, "fuller_stepdown", Coverage)
shared_stepdown_mean_se <- metric_value(primary_converged_raw, "fuller_stepdown", Mean_SE)
shared_lai_rmse <- metric_value(primary_converged_raw, "lai_2spa", RMSE)
shared_lai_coverage <- metric_value(primary_converged_raw, "lai_2spa", Coverage)
shared_lai_mean_se <- metric_value(primary_converged_raw, "lai_2spa", Mean_SE)
shared_eiv_hc3_rmse <- metric_value(primary_converged_raw, "eiv_dual_corrected_hc3", RMSE)
shared_eiv_hc3_coverage <- metric_value(primary_converged_raw, "eiv_dual_corrected_hc3", Coverage)
shared_eiv_hc3_mean_se <- metric_value(primary_converged_raw, "eiv_dual_corrected_hc3", Mean_SE)
shared_naive_hc3_rmse <- metric_value(primary_converged_raw, "naive_dual_eb_hc3", RMSE)
shared_naive_hc3_coverage <- metric_value(primary_converged_raw, "naive_dual_eb_hc3", Coverage)
shared_naive_hc3_mean_se <- metric_value(primary_converged_raw, "naive_dual_eb_hc3", Mean_SE)
```

The May 18 run contributes `r study1_completed_conditions` completed Study 1 conditions and `r format(study1_total_replications, big.mark = ",")` condition-replication pairs to this report. `r run_completion_note` Across the completed output, `r fmt_pct(stage1_singular_rate)` of Stage 1 fits were flagged as singular or near-singular.

## Key Findings

Convergence and accuracy answer different questions in this run. Fuller returns usable estimates in `r fmt_pct(fuller_success)` of replications and Fuller stepdown returns usable estimates in `r fmt_pct(stepdown_success)`, compared with `r fmt_pct(lai_success)` for Lai 2S-PA. Bias, RMSE, standard errors, and coverage should be compared on a shared converged subset: the primary comparators all converged in `r fmt_pct(primary_converged_rate)` of completed condition-replications.

Within that shared converged subset, the Fuller comparison is central. Fuller has mean bias `r fmt_mixed(shared_fuller_bias)`, RMSE `r fmt_mixed(shared_fuller_rmse)`, coverage `r fmt_pct(shared_fuller_coverage)`, and mean SE `r fmt_mixed(shared_fuller_mean_se)`. Fuller stepdown has mean bias `r fmt_mixed(shared_stepdown_bias)`, RMSE `r fmt_mixed(shared_stepdown_rmse)`, coverage `r fmt_pct(shared_stepdown_coverage)`, and mean SE `r fmt_mixed(shared_stepdown_mean_se)`. Lai 2S-PA's shared-subset RMSE is `r fmt_mixed(shared_lai_rmse)`, coverage is `r fmt_pct(shared_lai_coverage)`, and mean SE is `r fmt_mixed(shared_lai_mean_se)`.

The singular and near-singular Stage 1 cases are interpreted separately as robustness failures, not as ordinary accuracy comparisons. Fuller keeps `r fmt_pct(fuller_singular_success)` success in the singular stratum and Fuller stepdown keeps `r fmt_pct(stepdown_singular_success)`, whereas Lai 2S-PA keeps only `r fmt_pct(lai_singular_success)`. When singular cases are retained, naive dual BLUP OLS + HC3 and dual EIV + HC3 show very large singular-stratum RMSEs (`r fmt_mixed(naive_singular_rmse)` and `r fmt_mixed(eiv_singular_rmse)`), which is why their singular behavior is treated as a stability problem rather than as a fair converged-model comparison.

## Focal Overall Results

```{r focal-overall-table}
focus_overall_tbl <- focus_overall_raw %>%
  select(method_label, Replications, Conditions, Success_Rate, Mean_Bias, Median_Bias, RMSE, Coverage, Mean_SE, Empirical_SD, SE_to_SD) %>%
  format_metric_table()

kable(
  focus_overall_tbl,
  col.names = c(
    "Method", "Rows", "Conditions", "Success Rate", "Mean Bias",
    "Median Bias", "RMSE", "Coverage", "Mean SE", "Empirical SD", "SE / SD"
  ),
  align = "lrrrrrrrrrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

The full-grid table is primarily a convergence and robustness summary. The success-rate column should be read as a first-class performance metric: a method with good coverage only after many difficult replications are dropped is not operationally equivalent to a method that returns usable estimates across the full grid.

The bias, RMSE, standard-error, and coverage columns in this full-grid table are descriptive, but they are not the main basis for accuracy comparisons because methods contribute different sets of successful replications. Those comparisons are made below in the shared primary-converged subset, where Fuller, Fuller stepdown, dual EIV + HC3, naive dual BLUP OLS + HC3, and Lai 2S-PA all returned usable estimates for the same condition-replications.

## Fuller Versus Fuller Stepdown

This is the focal within-family comparison. The table reports bias, RMSE, standard errors, and coverage for Fuller and Fuller stepdown in the full completed May 18 output and again in the shared primary-converged subset used for the main accuracy comparisons.

```{r fuller-stepdown-comparison-table}
fuller_stepdown_compare_tbl <- bind_rows(
  focus_overall_raw %>%
    filter(method %in% fuller_pair_methods) %>%
    mutate(Analysis_Set = "All completed output"),
  primary_converged_raw %>%
    filter(method %in% fuller_pair_methods) %>%
    mutate(Analysis_Set = "Shared primary-converged subset")
) %>%
  arrange(Analysis_Set, method_label) %>%
  select(Analysis_Set, method_label, Replications, Conditions, Success_Rate, Mean_Bias, Median_Bias, RMSE, Coverage, Mean_SE, Empirical_SD, SE_to_SD) %>%
  format_metric_table()

kable(
  fuller_stepdown_compare_tbl,
  col.names = c(
    "Analysis Set", "Method", "Rows", "Conditions", "Success Rate",
    "Mean Bias", "Median Bias", "RMSE", "Coverage", "Mean SE",
    "Empirical SD", "SE / SD"
  ),
  align = "llrrrrrrrrrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r fuller-stepdown-design-plot, fig.height=9, fig.width=10}
fuller_stepdown_design_plot <- replication_df %>%
  semi_join(primary_converged_keys, by = c("condition_id", "rep")) %>%
  filter(method %in% fuller_pair_methods) %>%
  metric_summary(c("method_label", "num_clus", "clus_size", "icc")) %>%
  select(method_label, num_clus, clus_size, icc, Mean_Bias, RMSE, Coverage, Mean_SE) %>%
  pivot_longer(cols = c(Mean_Bias, RMSE, Coverage, Mean_SE), names_to = "metric", values_to = "value") %>%
  mutate(
    metric = factor(
      metric,
      levels = c("Mean_Bias", "RMSE", "Mean_SE", "Coverage"),
      labels = c("Mean bias", "RMSE", "Mean SE", "Coverage")
    ),
    clus_size = factor(clus_size),
    icc = factor(icc)
  )

ggplot(fuller_stepdown_design_plot, aes(x = clus_size, y = value, color = icc, group = icc)) +
  geom_hline(
    data = tibble(metric = factor("Coverage", levels = levels(fuller_stepdown_design_plot$metric)), yint = 0.95),
    aes(yintercept = yint),
    inherit.aes = FALSE,
    color = "#555555",
    linetype = "dashed",
    linewidth = 0.35
  ) +
  geom_line(linewidth = 0.55, alpha = 0.85) +
  geom_point(size = 1.6) +
  facet_grid(metric ~ method_label + num_clus, scales = "free_y", labeller = label_both) +
  scale_y_continuous(trans = scales::pseudo_log_trans(sigma = 0.01)) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Fuller and Fuller Stepdown by Design Cell",
    subtitle = "Shared primary-converged subset. Dashed line marks 95% coverage.",
    x = "Cluster size",
    y = NULL,
    color = "ICC"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.45, "lines")
  )
```

## Stage 1 Regular Versus Singular Fits

```{r focal-stage1-table}
focus_by_stage1_tbl <- focus_by_stage1_raw %>%
  arrange(method_label, stage1_stratum) %>%
  select(method_label, stage1_stratum, Replications, Conditions, Success_Rate, Mean_Bias, Median_Bias, RMSE, Coverage, Mean_SE, Empirical_SD, SE_to_SD) %>%
  format_metric_table()

kable(
  focus_by_stage1_tbl,
  col.names = c(
    "Method", "Stage 1 Stratum", "Rows", "Conditions", "Success Rate",
    "Mean Bias", "Median Bias", "RMSE", "Coverage", "Mean SE",
    "Empirical SD", "SE / SD"
  ),
  align = "llrrrrrrrrrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

This split is the main robustness diagnostic, not the main accuracy comparison. Lai can be judged by conditional performance where the SEM returns a usable solution, but the singular stratum shows how much of the design the method effectively refuses. Naive dual BLUP OLS is vulnerable because singular Stage 1 fits can make the EB intercept and slope nearly collinear. HC3 changes the interval calculation for the same point estimates; it does not fix a rank-deficient predictor pair.

```{r stage1-performance-plot, fig.height=7, fig.width=10}
stage1_plot <- focus_by_stage1_raw %>%
  filter(method %in% primary_methods) %>%
  select(method_label, stage1_stratum, Success_Rate, RMSE, Coverage, Mean_SE) %>%
  pivot_longer(cols = c(Success_Rate, RMSE, Coverage, Mean_SE), names_to = "metric", values_to = "value") %>%
  mutate(
    metric = factor(
      metric,
      levels = c("Success_Rate", "RMSE", "Coverage", "Mean_SE"),
      labels = c("Success rate", "RMSE", "Coverage", "Mean SE")
    )
  )

ggplot(stage1_plot, aes(x = method_label, y = value, fill = stage1_stratum)) +
  geom_col(position = position_dodge(width = 0.78), width = 0.68) +
  geom_hline(
    data = tibble(metric = factor("Coverage", levels = levels(stage1_plot$metric)), yint = 0.95),
    aes(yintercept = yint),
    inherit.aes = FALSE,
    color = "#555555",
    linetype = "dashed",
    linewidth = 0.35
  ) +
  facet_wrap(~ metric, scales = "free_y", ncol = 1) +
  scale_y_continuous(trans = scales::pseudo_log_trans(sigma = 0.01)) +
  scale_fill_manual(values = c("Stage 1 regular" = "#4e79a7", "Stage 1 singular" = "#c44e52")) +
  labs(
    title = "Focal Method Performance by Stage 1 Singularity",
    x = NULL,
    y = NULL,
    fill = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

## Convergence and Failure Patterns

```{r failure-condition-table}
failure_focus <- replication_df %>%
  filter(method %in% focus_methods) %>%
  mutate(
    failure = is.na(estimate),
    failure_class = case_when(
      !failure ~ "ok",
      method %in% c("lai_2spa", "lai_2spaa") & !is.na(mx_issue_class) ~ as.character(mx_issue_class),
      method %in% c("lai_2spa", "lai_2spaa") ~ "other_lai_failure",
      method %in% c("naive_dual_eb", "naive_dual_eb_hc3") & stage1_singular_problem ~ "stage1_singular_dual_blup",
      method %in% c("eiv_dual_corrected", "eiv_dual_corrected_hc3") & eiv_regularized %in% TRUE ~ "eiv_regularized",
      method %in% fuller_pair_methods & fuller_auto_guard_pass %in% FALSE ~ "fuller_guard_failure",
      method %in% fuller_pair_methods ~ "fuller_failure",
      TRUE ~ "other_failure"
    )
  )

failure_condition_tbl <- failure_focus %>%
  group_by(method_label, condition_id) %>%
  summarise(
    failure_rate = safe_mean(failure),
    any_failure = any(failure),
    .groups = "drop"
  ) %>%
  group_by(method_label) %>%
  summarise(
    Conditions = n(),
    Mean_Failure_Rate = safe_mean(failure_rate),
    Median_Failure_Rate = safe_median(failure_rate),
    P90_Failure_Rate = unname(stats::quantile(failure_rate, probs = 0.9, na.rm = TRUE)),
    Max_Failure_Rate = max(failure_rate, na.rm = TRUE),
    Conditions_With_Any_Failure = safe_mean(any_failure),
    Conditions_Over_10pct_Failure = safe_mean(failure_rate > 0.10),
    Conditions_Over_50pct_Failure = safe_mean(failure_rate > 0.50),
    .groups = "drop"
  ) %>%
  mutate(across(ends_with("Rate") | starts_with("Conditions_With") | starts_with("Conditions_Over"), ~ fmt_pct(.x))) %>%
  arrange(method_label)

kable(
  failure_condition_tbl,
  col.names = c(
    "Method", "Conditions", "Mean Failure Rate", "Median Failure Rate",
    "90th pct Failure Rate", "Max Failure Rate", "Conditions with Any Failure",
    "Conditions > 10% Failure", "Conditions > 50% Failure"
  ),
  align = "lrrrrrrrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r failure-class-table}
failure_issue_tbl <- failure_focus %>%
  filter(failure) %>%
  group_by(method_label, failure_class) %>%
  summarise(Failed_Reps = n(), .groups = "drop") %>%
  group_by(method_label) %>%
  mutate(Share_Of_Method_Failures = Failed_Reps / sum(Failed_Reps)) %>%
  ungroup() %>%
  arrange(method_label, desc(Failed_Reps)) %>%
  mutate(Share_Of_Method_Failures = fmt_pct(Share_Of_Method_Failures))

kable(
  failure_issue_tbl,
  col.names = c("Method", "Failure Type", "Failed Replications", "Share of Method Failures"),
  align = "llrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r failure-heatmap, fig.height=9, fig.width=11}
failure_design <- failure_focus %>%
  group_by(method_label, num_clus, clus_size, icc) %>%
  summarise(Failure_Rate = safe_mean(failure), .groups = "drop") %>%
  mutate(clus_size = factor(clus_size), icc = factor(icc))

ggplot(failure_design, aes(x = clus_size, y = icc, fill = Failure_Rate)) +
  geom_tile(color = "white") +
  facet_grid(method_label ~ num_clus, labeller = label_both) +
  scale_fill_gradient(low = "#eef3f7", high = "#b33a3a", labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  labs(
    title = "Usable-Estimate Failure Rate by Design Cell",
    subtitle = "Averaged over all Study 1 conditions sharing the displayed cluster count, cluster size, and ICC.",
    x = "Cluster size",
    y = "ICC",
    fill = "Failure rate"
  ) +
  theme_minimal(base_size = 10)
```

The failure heatmap separates robustness from conditional inference. Fuller, Fuller stepdown, and EIV-style algebraic estimators can return estimates in cases where Lai fails, but accuracy comparisons should not mix those retained difficult cases with the subset where all focal models converged.

## Bias, RMSE, SE, and Coverage by Design

This design-cell view is restricted to non-singular Stage 1 replications where all primary comparators converged. It therefore compares bias, RMSE, standard errors, and coverage on the same replications for Fuller, Fuller stepdown, dual EIV + HC3, naive dual BLUP OLS + HC3, and Lai 2S-PA.

```{r design-metric-plot, fig.height=10, fig.width=14}
design_keys <- primary_converged_keys %>%
  filter(stage1_singular_problem %in% FALSE) %>%
  select(condition_id, rep)

design_metric_plot <- replication_df %>%
  filter(method %in% primary_methods) %>%
  semi_join(design_keys, by = c("condition_id", "rep")) %>%
  metric_summary(c("method_label", "num_clus", "clus_size", "icc")) %>%
  select(method_label, num_clus, clus_size, icc, Success_Rate, Mean_Bias, RMSE, Coverage, Mean_SE) %>%
  pivot_longer(cols = c(Success_Rate, Mean_Bias, RMSE, Coverage, Mean_SE), names_to = "metric", values_to = "value") %>%
  mutate(
    metric = factor(
      metric,
      levels = c("Success_Rate", "Mean_Bias", "RMSE", "Coverage", "Mean_SE"),
      labels = c("Success rate", "Mean bias", "RMSE", "Coverage", "Mean SE")
    ),
    clus_size = factor(clus_size),
    icc = factor(icc)
  )

ggplot(design_metric_plot, aes(x = clus_size, y = value, color = icc, group = icc)) +
  geom_hline(
    data = tibble(metric = factor("Coverage", levels = levels(design_metric_plot$metric)), yint = 0.95),
    aes(yintercept = yint),
    inherit.aes = FALSE,
    color = "#555555",
    linetype = "dashed",
    linewidth = 0.35
  ) +
  geom_line(linewidth = 0.55, alpha = 0.85) +
  geom_point(size = 1.6) +
  facet_grid(metric ~ method_label + num_clus, scales = "free_y", labeller = label_both) +
  scale_y_continuous(trans = scales::pseudo_log_trans(sigma = 0.01)) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Design-Cell Performance in Shared Converged, Non-Singular Replications",
    subtitle = "Restricted to non-singular replications where all primary comparators converged. Dashed line marks 95% coverage.",
    x = "Cluster size",
    y = NULL,
    color = "ICC"
  ) +
  theme_minimal(base_size = 8) +
  theme(
    strip.text = element_text(size = 7),
    panel.spacing = unit(0.45, "lines")
  )
```

## HC3 Standard-Error Effects

HC3 variants use the same point estimates as their base estimators but change the standard-error and confidence-interval calculation. This section isolates whether HC3 helps inference for naive dual EB and EIV within the shared primary-converged subset, so coverage and SE comparisons are not driven by different convergence sets.

```{r hc3-comparison-table}
hc3_pairs <- tibble::tribble(
  ~base_method, ~hc3_method, ~family_label,
  "naive_dual_eb", "naive_dual_eb_hc3", "Naive dual BLUP OLS",
  "eiv_dual_corrected", "eiv_dual_corrected_hc3", "Dual EIV"
)

hc3_replication_focus <- replication_df %>%
  filter(method %in% c(hc3_pairs$base_method, hc3_pairs$hc3_method)) %>%
  semi_join(primary_converged_keys, by = c("condition_id", "rep")) %>%
  mutate(
    covered = ifelse(!is.na(ci_low) & !is.na(ci_high), ci_low <= truth & ci_high >= truth, NA),
    se_variant = ifelse(grepl("_hc3$", method), "hc3", "base")
  ) %>%
  left_join(
    hc3_pairs %>%
      tidyr::pivot_longer(cols = c(base_method, hc3_method), values_to = "method") %>%
      select(-name),
    by = "method"
  )

hc3_compare_tbl <- hc3_replication_focus %>%
  group_by(family_label, stage1_stratum, se_variant) %>%
  summarise(
    Coverage = safe_mean(covered),
    Mean_SE = safe_mean(se),
    Empirical_SD = safe_sd(estimate),
    SE_to_SD = Mean_SE / Empirical_SD,
    .groups = "drop"
  ) %>%
  tidyr::pivot_wider(names_from = se_variant, values_from = c(Coverage, Mean_SE, SE_to_SD)) %>%
  mutate(
    `Base Coverage` = fmt_pct(Coverage_base),
    `HC3 Coverage` = fmt_pct(Coverage_hc3),
    `Coverage Gap` = fmt_pct(Coverage_hc3 - Coverage_base),
    `Base Mean SE` = fmt_mixed(Mean_SE_base),
    `HC3 Mean SE` = fmt_mixed(Mean_SE_hc3),
    `Base SE / SD` = fmt_num(SE_to_SD_base, 2),
    `HC3 SE / SD` = fmt_num(SE_to_SD_hc3, 2)
  ) %>%
  select(family_label, stage1_stratum, `Base Coverage`, `HC3 Coverage`, `Coverage Gap`, `Base Mean SE`, `HC3 Mean SE`, `Base SE / SD`, `HC3 SE / SD`)

kable(
  hc3_compare_tbl,
  col.names = c(
    "Method Family", "Stage 1 Stratum", "Base Coverage", "HC3 Coverage",
    "Coverage Gap", "Base Mean SE", "HC3 Mean SE", "Base SE / SD", "HC3 SE / SD"
  ),
  align = "llrrrrrrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r hc3-coverage-gap-plot, fig.height=4.8, fig.width=8}
hc3_gap_plot <- hc3_compare_tbl %>%
  transmute(
    family_label,
    stage1_stratum,
    coverage_gap = suppressWarnings(as.numeric(sub("%$", "", `Coverage Gap`)) / 100)
  )

ggplot(hc3_gap_plot, aes(x = family_label, y = coverage_gap, fill = stage1_stratum)) +
  geom_col(position = position_dodge(width = 0.75), width = 0.65) +
  geom_hline(yintercept = 0, color = "#555555", linewidth = 0.35) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = c("Stage 1 regular" = "#4e79a7", "Stage 1 singular" = "#c44e52")) +
  labs(
    title = "Coverage Change from HC3 Standard Errors",
    x = NULL,
    y = "Coverage gap (HC3 - base)",
    fill = NULL
  ) +
  theme_minimal(base_size = 11)
```

## Fuller Diagnostics

The Fuller estimators add method-specific diagnostics that are useful for understanding when they are behaving like stable corrections versus when they are being driven by the same low-information geometry that challenges Lai and EIV.

```{r fuller-diagnostics-table}
fuller_diag_tbl <- replication_df %>%
  filter(method %in% fuller_pair_methods) %>%
  group_by(method_label, stage1_stratum) %>%
  summarise(
    Replications = n(),
    Success_Rate = safe_mean(!is.na(estimate)),
    Mean_Lambda1 = safe_mean(fuller_lambda1),
    Mean_Lambda2 = safe_mean(fuller_lambda2),
    Mean_Sigma2 = safe_mean(fuller_sigma2),
    Mean_Weight_Min = safe_mean(fuller_weight_min),
    Mean_Weight_Max = safe_mean(fuller_weight_max),
    Mean_Correction_C = safe_mean(fuller_correction_c),
    Mean_SE_Ratio = safe_mean(fuller_se_ratio),
    Auto_Tempered_Rate = safe_mean(fuller_auto_tempered %in% TRUE),
    Guard_Pass_Rate = safe_mean(fuller_auto_guard_pass %in% TRUE),
    Median_Design_Kappa = safe_median(stage1_design_kappa),
    .groups = "drop"
  ) %>%
  mutate(
    Success_Rate = fmt_pct(Success_Rate),
    Auto_Tempered_Rate = fmt_pct(Auto_Tempered_Rate),
    Guard_Pass_Rate = fmt_pct(Guard_Pass_Rate),
    across(
      c(Mean_Lambda1, Mean_Lambda2, Mean_Sigma2, Mean_Weight_Min, Mean_Weight_Max, Mean_Correction_C, Mean_SE_Ratio, Median_Design_Kappa),
      ~ fmt_mixed(.x)
    )
  )

kable(
  fuller_diag_tbl,
  col.names = c(
    "Method", "Stage 1 Stratum", "Replications", "Success Rate", "Mean lambda1",
    "Mean lambda2", "Mean sigma2", "Mean min weight", "Mean max weight",
    "Mean correction c", "Mean SE ratio", "Auto-tempered rate",
    "Guard pass rate", "Median design kappa"
  ),
  align = "llrrrrrrrrrrrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r fuller-kappa-plot, fig.height=6, fig.width=10}
fuller_kappa_plot <- replication_df %>%
  filter(method %in% fuller_pair_methods, !is.na(estimate), is.finite(stage1_design_kappa)) %>%
  mutate(
    abs_error = abs(estimate - truth),
    log10_kappa = log10(pmax(stage1_design_kappa, 1)),
    kappa_bin = cut(
      log10_kappa,
      breaks = seq(floor(min(log10_kappa, na.rm = TRUE)), ceiling(max(log10_kappa, na.rm = TRUE)), length.out = 36),
      include.lowest = TRUE
    )
  ) %>%
  group_by(method_label, stage1_stratum, kappa_bin) %>%
  summarise(
    log10_kappa = safe_mean(log10_kappa),
    Median_Abs_Error = safe_median(abs_error),
    P25_Abs_Error = unname(stats::quantile(abs_error, probs = 0.25, na.rm = TRUE)),
    P75_Abs_Error = unname(stats::quantile(abs_error, probs = 0.75, na.rm = TRUE)),
    Replications = n(),
    .groups = "drop"
  )

ggplot(fuller_kappa_plot, aes(x = log10_kappa, y = Median_Abs_Error, color = stage1_stratum, fill = stage1_stratum)) +
  geom_ribbon(aes(ymin = P25_Abs_Error, ymax = P75_Abs_Error), alpha = 0.16, color = NA) +
  geom_line(linewidth = 0.8) +
  geom_point(aes(size = Replications), alpha = 0.75) +
  scale_y_continuous(trans = scales::pseudo_log_trans(sigma = 0.01)) +
  scale_color_manual(values = c("Stage 1 regular" = "#4e79a7", "Stage 1 singular" = "#c44e52")) +
  scale_fill_manual(values = c("Stage 1 regular" = "#4e79a7", "Stage 1 singular" = "#c44e52")) +
  scale_size_continuous(range = c(0.8, 2.4), guide = "none") +
  facet_wrap(~ method_label, ncol = 1) +
  labs(
    title = "Fuller Absolute Error as Stage 1 Design Conditioning Worsens",
    subtitle = "Points show median absolute error within log-condition-number bins; bands show the interquartile range.",
    x = "log10(Stage 1 dual-BLUP design condition number)",
    y = "Absolute error",
    color = NULL
  ) +
  guides(fill = "none") +
  theme_minimal(base_size = 11)
```

## Shared Primary-Converged Subset

Convergence differences create survivorship bias. To compare bias, RMSE, standard errors, and coverage on the same replications, the next table restricts the primary comparators to condition-replication pairs where all primary methods returned usable estimates: Fuller, Fuller stepdown, dual EIV + HC3, naive dual BLUP OLS + HC3, and Lai 2S-PA.

```{r primary-converged-subset}
primary_converged_subset_tbl <- replication_df %>%
  semi_join(primary_converged_keys, by = c("condition_id", "rep")) %>%
  filter(method %in% primary_methods) %>%
  metric_summary("method_label") %>%
  arrange(method_label) %>%
  select(method_label, Replications, Conditions, Success_Rate, Mean_Bias, Median_Bias, RMSE, Coverage, Mean_SE, Empirical_SD, SE_to_SD) %>%
  format_metric_table()

kable(
  primary_converged_subset_tbl,
  col.names = c(
    "Method", "Rows", "Conditions", "Success Rate", "Mean Bias",
    "Median Bias", "RMSE", "Coverage", "Mean SE", "Empirical SD", "SE / SD"
  ),
  align = "lrrrrrrrrrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

This table is the main accuracy comparison. The full-grid and Stage 1 tables answer how often each method survives; this table answers how the methods compare once they have all survived on the same data.

On this shared converged subset, the Fuller and Fuller stepdown rows are the main within-family accuracy comparison. Lai remains the SEM reference method, while EIV + HC3 and naive dual BLUP + HC3 are included to show whether their standard-error adjustments or measurement-error correction are competitive once convergence is held fixed.

## Appendix: Broader Method Context

```{r all-methods-table}
all_methods_tbl <- replication_df %>%
  metric_summary("method_label") %>%
  arrange(method_label) %>%
  select(method_label, Replications, Conditions, Success_Rate, Mean_Bias, Median_Bias, RMSE, Coverage, Mean_SE, Empirical_SD, SE_to_SD) %>%
  format_metric_table()

kable(
  all_methods_tbl,
  col.names = c(
    "Method", "Rows", "Conditions", "Success Rate", "Mean Bias",
    "Median Bias", "RMSE", "Coverage", "Mean SE", "Empirical SD", "SE / SD"
  ),
  align = "lrrrrrrrrrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

```{r stage1-summary-appendix}
stage1_appendix_tbl <- stage1_df %>%
  filter(method %in% c(focus_methods, "eiv_dual_corrected_hc0", "eiv_dual_corrected_nearpd_hc3", "eiv_dual_corrected_ridge_hc3")) %>%
  group_by(method_label, stage1_stratum) %>%
  summarise(
    Conditions = n_distinct(condition_id),
    Mean_Convergence = safe_mean(convergence),
    Mean_Bias = safe_mean(bias),
    Mean_RMSE = safe_mean(rmse),
    Mean_Coverage = safe_mean(coverage),
    Mean_Stage1_Design_Kappa = safe_mean(mean_stage1_design_kappa),
    .groups = "drop"
  ) %>%
  mutate(
    Mean_Convergence = fmt_pct(Mean_Convergence),
    Mean_Bias = fmt_mixed(Mean_Bias),
    Mean_RMSE = fmt_mixed(Mean_RMSE),
    Mean_Coverage = fmt_pct(Mean_Coverage),
    Mean_Stage1_Design_Kappa = fmt_mixed(Mean_Stage1_Design_Kappa)
  ) %>%
  arrange(method_label, stage1_stratum)

kable(
  stage1_appendix_tbl,
  col.names = c(
    "Method", "Stage 1 Stratum", "Conditions", "Mean Convergence",
    "Mean Bias", "Mean RMSE", "Mean Coverage", "Mean Stage 1 Design Kappa"
  ),
  align = "llrrrrrr"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

## Interpretation

The central accuracy result should be read from the shared primary-converged subset. In that subset, the Fuller and Fuller stepdown rows give the cleanest within-family comparison of bias, RMSE, mean SE, and coverage because they are evaluated on exactly the same condition-replication pairs as Lai 2S-PA, dual EIV + HC3, and naive dual BLUP + HC3.

The HC3 results should also be interpreted only within this shared converged frame. HC3 can alter coverage by widening or reshaping intervals, but it does not change point estimates. Thus, when naive dual BLUP OLS + HC3 has worse RMSE than Lai or Fuller in the shared converged subset, that is not an SE problem; it is the same estimator with a different variance calculation.

Dual EIV remains useful as a conceptual bridge to Lai because it applies an analytical measurement-error correction. But once convergence is held fixed, its RMSE and standard-error behavior should be read against both Fuller variants rather than against Fuller alone.

Singular and near-singular Stage 1 fits are a separate robustness story. They should not be used to make ordinary bias/RMSE/coverage comparisons because Lai often fails there and the competing algebraic methods retain cases that Lai drops. In that separate robustness analysis, the two Fuller variants should be judged first on retained usable estimates and then on whether their singular-stratum RMSE avoids the catastrophic behavior seen for naive dual BLUP and EIV.

The practical conclusion is therefore two-part. Conditional on all primary models converging, use the dedicated Fuller-versus-stepdown table as the focal comparison. Across the full completed design, especially when singular Stage 1 fits matter, the success-rate and Stage 1 tables show how much robustness each Fuller variant buys relative to Lai, naive dual BLUP, and EIV.
