---
title: "Lai Study 1 replication with VH estimators"
subtitle: "Technical review of historical comparability, reporting scale, and slope-SD diagnostics"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
execute:
  echo: false
  warning: false
  message: false
params:
  results_dir: "outputs/lai_study1_vh_production_20260731"
  figure_dir: null
  historical_figure_reference: null
---

## Technical summary

This report reviews the Lai Study 1 data-generating mechanisms using the
seven-method Vig--Hallquist Study 2 estimator bundle.  It is designed to keep
two questions separate:

1. Can the overlapping methods reproduce the *pattern* in Lai's Figure 2
   under the historical scale and retention rule?
2. What do all seven VH methods show when estimates are put on a common
   latent-slope-SD reporting scale and screened using the VH eligibility rule?

The historical Figure 2 panel is therefore a **Figure 2 analogue**, not an
exact reproduction.  The VH bundle does not contain Lai's historical
u1-only-EB and 2S-PAB variants, and Lai's original standardization can vary
by method.  The diagnostic sections make that distinction inspectable rather
than treating a scale change as an estimator effect.

```{r setup}
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(knitr)
library(scales)

report_input <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) "")
report_dir <- if (nzchar(report_input)) {
  report_input <- normalizePath(report_input, mustWork = FALSE)
  if (dir.exists(report_input)) report_input else dirname(report_input)
} else {
  getwd()
}
repo_root <- normalizePath(file.path(report_dir, ".."), mustWork = FALSE)

resolve_repo_path <- function(path) {
  if (is.null(path) || !nzchar(path)) return(NULL)
  if (grepl("^/", path)) {
    normalizePath(path, mustWork = FALSE)
  } else {
    normalizePath(file.path(repo_root, path), mustWork = FALSE)
  }
}

safe_mean <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) == 0L) NA_real_ else mean(x)
}

safe_median <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) == 0L) NA_real_ else stats::median(x)
}

fmt_num <- function(x, digits = 3L) {
  ifelse(is.finite(x), formatC(x, digits = digits, format = "f"), NA_character_)
}

fmt_pct <- function(x, accuracy = 0.1) {
  ifelse(is.finite(x), scales::percent(x, accuracy = accuracy), NA_character_)
}

method_labels <- c(
  oracle_dual = "Oracle dual",
  naive_dual_blup = "Naive dual BLUP",
  closed_form_dual = "Closed-form dual",
  fuller_closed_form = "Fuller closed-form",
  fuller_alpha_stepdown_closed_form = "Fuller alpha-stepdown",
  lai_2spa = "2S-PA",
  msem = "MSEM"
)

results_dir <- resolve_repo_path(params$results_dir)
figure_dir <- resolve_repo_path(params$figure_dir)
if (is.null(figure_dir)) figure_dir <- file.path(results_dir, "figure2_analogue")
summary_path <- file.path(results_dir, "lai_study1_vh_summary.csv")

if (!file.exists(summary_path)) {
  stop("Could not find the aggregate summary: ", summary_path, call. = FALSE)
}

summary_df <- readr::read_csv(summary_path, show_col_types = FALSE)
required_columns <- c(
  "condition_id", "reporting_scale", "method", "num_clus", "clus_size",
  "icc", "vr_u1_u0", "cor_u0_u1", "beta_zu1", "var_u1", "sigma2",
  "dgm_population_slope_sd", "dgm_posterior_slope_reliability", "truth",
  "n_rep", "convergence", "vh_analysis_eligibility_rate",
  "lai_original_eligibility_rate", "n_eligible_both", "n_lai_original_only",
  "n_vh_only", "bias", "rmse", "coverage",
  "mean_eb_to_population_slope_sd",
  "mean_corrected_to_population_slope_sd",
  "mean_fitted_to_population_slope_sd",
  "mean_lai_original_method_multiplier",
  "lai_original_reporting_available_rate"
)
missing_columns <- setdiff(required_columns, names(summary_df))
if (length(missing_columns) > 0L) {
  stop(
    "Aggregate summary is missing required columns: ",
    paste(missing_columns, collapse = ", "),
    call. = FALSE
  )
}

summary_df <- summary_df |>
  mutate(
    method_label = factor(method, levels = names(method_labels), labels = unname(method_labels)),
    reporting_scale = factor(
      reporting_scale,
      levels = c("raw", "latent_sd", "lai_original_standardized")
    )
  )

raw_df <- summary_df |> filter(reporting_scale == "raw")
latent_df <- summary_df |> filter(reporting_scale == "latent_sd")
historical_df <- summary_df |> filter(reporting_scale == "lai_original_standardized")

n_conditions <- n_distinct(summary_df$condition_id)
n_methods <- n_distinct(summary_df$method)
n_replications <- raw_df |> distinct(condition_id, n_rep) |> summarise(total = sum(n_rep)) |> pull(total)

read_optional_csv <- function(path) {
  if (!file.exists(path)) return(NULL)
  readr::read_csv(path, show_col_types = FALSE)
}

historical_figure_summary <- read_optional_csv(
  file.path(figure_dir, "figure2_analogue_cell_summary.csv")
)
vh_figure_summary <- read_optional_csv(
  file.path(figure_dir, "vh_primary_companion_cell_summary.csv")
)
historical_inference_summary <- read_optional_csv(
  file.path(figure_dir, "figure3_4_analogue_cell_summary.csv")
)

historical_bias_by_correlation <- if (!is.null(historical_figure_summary)) {
  historical_figure_summary |>
    filter(beta_zu1 == 0.4) |>
    group_by(method, cor_u0_u1) |>
    summarise(
      median_robust_bias = safe_median(robust_bias),
      mean_robust_bias = safe_mean(robust_bias),
      .groups = "drop"
    )
} else {
  NULL
}

format_correlation_medians <- function(method) {
  if (is.null(historical_bias_by_correlation)) return("not available")
  historical_bias_by_correlation |>
    filter(method == !!method) |>
    arrange(cor_u0_u1) |>
    transmute(text = paste0("rho = ", cor_u0_u1, ": ", fmt_num(median_robust_bias))) |>
    pull(text) |>
    paste(collapse = "; ")
}

naive_historical_correlation_text <- format_correlation_medians("naive_dual_blup")
two_spa_historical_correlation_text <- format_correlation_medians("lai_2spa")
msem_historical_correlation_text <- format_correlation_medians("msem")

vh_instability_summary <- if (!is.null(vh_figure_summary)) {
  vh_figure_summary |>
    group_by(method) |>
    summarise(
      median_bias = safe_median(mean_bias),
      median_abs_bias = safe_median(abs(mean_bias)),
      max_abs_bias = max(abs(mean_bias), na.rm = TRUE),
      n_abs_bias_over_01 = sum(abs(mean_bias) > 0.1, na.rm = TRUE),
      n_abs_bias_over_1 = sum(abs(mean_bias) > 1, na.rm = TRUE),
      .groups = "drop"
    )
} else {
  NULL
}

naive_instability <- if (!is.null(vh_instability_summary)) {
  vh_instability_summary |> filter(method == "naive_dual_blup")
} else {
  NULL
}
naive_instability_count <- if (!is.null(naive_instability) && nrow(naive_instability) == 1L) {
  naive_instability$n_abs_bias_over_1[[1]]
} else {
  NA_integer_
}
naive_instability_max <- if (!is.null(naive_instability) && nrow(naive_instability) == 1L) {
  naive_instability$max_abs_bias[[1]]
} else {
  NA_real_
}

historical_inference_by_method <- if (!is.null(historical_inference_summary)) {
  historical_inference_summary |>
    group_by(method) |>
    summarise(
      coverage_conditions = sum(is.finite(coverage)),
      mean_coverage = safe_mean(coverage),
      median_coverage = safe_median(coverage),
      type1_conditions = sum(beta_zu1 == 0 & is.finite(type1_error)),
      mean_type1_error = safe_mean(type1_error[beta_zu1 == 0]),
      median_type1_error = safe_median(type1_error[beta_zu1 == 0]),
      mean_retained = safe_mean(n_retained),
      .groups = "drop"
    )
} else {
  NULL
}
```

The production aggregate contains `r n_conditions` DGM conditions, `r n_methods`
methods, and `r scales::comma(n_replications)` simulated data sets per raw
reporting view.  Its source is
`r normalizePath(summary_path, mustWork = FALSE)`.

### Result at a glance

For the historical comparison at \(\beta_{Zu1}=.4\), the naïve dual-BLUP
median trimmed bias is negative at every random-effect correlation
(`r naive_historical_correlation_text`).  Thus this VH implementation does
**not** reproduce the correlation-signed naïve-EB pattern described for Lai's
Figure 2.  In contrast, 2S-PA remains near zero
(`r two_spa_historical_correlation_text`) and MSEM is small and positive
(`r msem_historical_correlation_text`) on this historical view.

On the common latent-SD VH view, the naïve dual-BLUP median is not itself the
main issue; its distribution has `r naive_instability_count` conditions with
\(|\mathrm{mean\ bias}| > 1\), including a maximum of
`r fmt_num(naive_instability_max, 1)`.  The report therefore shows its central
behavior and its rare instability in separate panels rather than letting those
failures flatten every other method's plot.

## Methods and terminology

The target is the effect of the level-2 covariate \(Z\) on the random slope
\(u_1\).  “Dual” means that the stage-2 regression uses both the random
intercept \(u_0\) and random slope \(u_1\) as predictors; this is important
when those random effects are correlated.

```{r method-glossary}
method_glossary <- tibble(
  Method = c(
    "Oracle dual", "Naive dual BLUP", "Closed-form dual",
    "Fuller closed-form", "Fuller alpha-stepdown", "2S-PA", "MSEM"
  ),
  `Stage-2 predictor or model` = c(
    "True simulated u0 and u1",
    "Empirical-Bayes (BLUP) u0 and u1, treated as observed",
    "Closed-form corrected u0 and u1, treated as observed",
    "Closed-form corrected scores with their measurement-error covariance",
    "Fuller correction with alpha stepdown for a guarded correction",
    "Lai two-stage path-analysis model fitted in OpenMx",
    "Joint multilevel structural-equation model fitted in Mplus"
  ),
  `Interpretive role` = c(
    "Simulation benchmark; not available in applications",
    "Shows attenuation/instability from ignoring first-stage uncertainty",
    "Separates corrected-score construction from stage-2 EIV correction",
    "EIV-style correction using estimated first-stage error covariance",
    "Stabilized Fuller variant in the VH Study 2 bundle",
    "Direct Lai-method overlap for historical comparison",
    "Latent-variable benchmark and direct Lai-method overlap"
  )
)
knitr::kable(method_glossary, caption = "Estimator terminology used throughout this report")
```

The first three methods use a two-predictor observed-score regression in stage
2.  The distinction is whether those predictors are true random effects,
shrunken EB estimates, or closed-form corrected scores.  The Fuller methods
also propagate the estimated measurement-error covariance into the stage-2
correction.  2S-PA and MSEM instead fit their own latent-variable/path models,
so their reported slope-SD scale must be interpreted separately in the
historical comparison.

## 1. DGM audit: the Lai grid and the estimand are explicit

The review starts by auditing the condition grid, rather than inferring the
design from a plotted result.  The external-outcome effect of the random
intercept is fixed at \(\beta_{zu0}=.4\); the focal slope effect is
\(\beta_{zu1}\).  The summary rows below use the raw reporting view only, so
reporting-scale rows do not duplicate conditions.

```{r dgm-audit}
dgm_conditions <- raw_df |>
  distinct(
    condition_id, num_clus, clus_size, icc, vr_u1_u0, cor_u0_u1,
    beta_zu1, var_u1, sigma2, dgm_population_slope_sd,
    dgm_posterior_slope_reliability, n_rep
  )

dgm_grid <- tibble(
  factor = c(
    "Clusters", "Cluster size", "ICC", "Slope/intercept variance ratio",
    "Random-effect correlation", "Slope external effect"
  ),
  levels = c(
    paste(sort(unique(dgm_conditions$num_clus)), collapse = ", "),
    paste(sort(unique(dgm_conditions$clus_size)), collapse = ", "),
    paste(sort(unique(dgm_conditions$icc)), collapse = ", "),
    paste(sort(unique(dgm_conditions$vr_u1_u0)), collapse = ", "),
    paste(sort(unique(dgm_conditions$cor_u0_u1)), collapse = ", "),
    paste(sort(unique(dgm_conditions$beta_zu1)), collapse = ", ")
  )
)

knitr::kable(dgm_grid, caption = "Lai Study 1 condition grid represented in the VH replication")
```

The condition count is `r nrow(dgm_conditions)`.  The population slope SD and
posterior slope reliability below are DGM quantities, not post-hoc calibration
targets; they provide the reference for the later slope-SD and reliability
diagnostics.

```{r dgm-reference-table}
dgm_reference <- dgm_conditions |>
  summarise(
    `Population slope SD range` = paste(range(dgm_population_slope_sd), collapse = " to "),
    `Posterior reliability range` = paste(round(range(dgm_posterior_slope_reliability), 3), collapse = " to "),
    `Replications per condition` = paste(sort(unique(n_rep)), collapse = ", ")
  )
knitr::kable(dgm_reference)
```

## 2. Historical Figure 2 analogue: compare like with like

This panel is the primary visual comparison with Lai's Figure 2.  It uses only
the three overlapping methods (naive dual BLUP, 2S-PA, and MSEM), Lai's
method-specific historical standardization, Lai's `status_code == 0` retention
rule, and a 20%-trimmed Monte Carlo bias.  Those choices are deliberately
different from the VH-primary analysis below.

Lai's published Figure 2 code plots six methods: 2S-PA, 2S-PAB, MSEM,
EB \((u_0+u_1)\), centered EB, and uncentered EB.  This analogue retains only
the directly corresponding 2S-PA, MSEM, and dual-EB series; it cannot test the
other three lines.  The original plotting code is available in the
[Lai supplementary repository](https://github.com/marklhc/2spa-random-slopes-supp/blob/main/notebooks/_sim1_results_supp.qmd).

```{r historical-figure, fig.cap="Lai Study 1 Figure 2 analogue: historical reporting scale and Lai eligibility rule."}
historical_plot_path <- file.path(figure_dir, "figure2_analogue.png")
if (file.exists(historical_plot_path)) {
  knitr::include_graphics(historical_plot_path)
} else {
  knitr::asis_output(
    "**Figure not yet generated.** Run `make_figure2_analogue.R` against the results directory, then re-render this report."
  )
}
```

```{r historical-availability}
historical_availability <- historical_df |>
  group_by(method_label) |>
  summarise(
    conditions = n(),
    `Mean scale availability` = safe_mean(lai_original_reporting_available_rate),
    `Mean Lai eligibility` = safe_mean(lai_original_eligibility_rate),
    .groups = "drop"
  ) |>
  mutate(
    `Mean scale availability` = fmt_pct(`Mean scale availability`),
    `Mean Lai eligibility` = fmt_pct(`Mean Lai eligibility`)
  )
knitr::kable(
  historical_availability,
  caption = "Availability and retention in the historical comparison view"
)
```

```{r historical-correlation-table}
if (!is.null(historical_bias_by_correlation)) {
  historical_comparison_table <- historical_bias_by_correlation |>
    mutate(
      method = recode(
        method,
        naive_dual_blup = "Naive dual BLUP",
        lai_2spa = "2S-PA",
        msem = "MSEM"
      ),
      `Median 20%-trimmed bias` = fmt_num(median_robust_bias),
      `Mean 20%-trimmed bias` = fmt_num(mean_robust_bias)
    ) |>
    select(Method = method, `Correlation` = cor_u0_u1,
           `Median 20%-trimmed bias`, `Mean 20%-trimmed bias`)
  knitr::kable(
    historical_comparison_table,
    caption = "Historical-view bias at beta[Z.u1] = .4, summarized over the remaining DGM factors"
  )
}
```

The key discrepancy is concentrated in the naïve dual-BLUP analogue: its
median bias remains roughly \(-.02\) across \(\rho=-.5, 0, .5\), instead of
switching from negative to positive with the random-effect correlation as in
the qualitative pattern reported for Lai's Figure 2.  This is evidence of a
real implementation/DGM or estimand difference worth tracing; it is not
explained away by the 2S-PA and MSEM lines, which remain close to zero on this
historical view.  Because the three omitted Lai series are not part of the VH
bundle, this result narrows the discrepancy but does not establish an exact
paper-level contradiction.

### Historical coverage and Type I error analogues

The next two panels mimic Lai's Figures 3 and 4 using the same three
overlapping methods, historical scale, and `status_code == 0` retention rule.
Coverage is the proportion of retained 95% intervals containing the truth.
The Type I panel is limited to \(\beta_{Zu1}=0\) and uses two-sided 95% CI
exclusion as the rejection rule.  Lai's original code used `p < .05`; these
agree for ordinary two-sided Wald intervals but should not be called an exact
p-value replication when an estimator's interval construction differs.

```{r figure3-coverage, fig.cap="Figure 3 analogue: empirical 95% coverage under historical Lai reporting and retention."}
coverage_plot_path <- file.path(figure_dir, "figure3_coverage_analogue.png")
if (file.exists(coverage_plot_path)) {
  knitr::include_graphics(coverage_plot_path)
} else {
  knitr::asis_output("**Coverage analogue not yet generated.** Run the post-estimation figure command, then re-render this report.")
}
```

```{r figure4-type1, fig.cap="Figure 4 analogue: CI-based empirical Type I error under historical Lai reporting and retention."}
type1_plot_path <- file.path(figure_dir, "figure4_type1_analogue.png")
if (file.exists(type1_plot_path)) {
  knitr::include_graphics(type1_plot_path)
} else {
  knitr::asis_output("**Type I error analogue not yet generated.** Run the post-estimation figure command, then re-render this report.")
}
```

```{r historical-inference-table}
if (!is.null(historical_inference_by_method)) {
  historical_inference_table <- historical_inference_by_method |>
    mutate(
      Method = recode(
        method,
        naive_dual_blup = "Naive dual BLUP",
        lai_2spa = "2S-PA",
        msem = "MSEM"
      ),
      `Mean coverage` = fmt_pct(mean_coverage),
      `Median coverage` = fmt_pct(median_coverage),
      `Mean Type I error` = fmt_pct(mean_type1_error),
      `Median Type I error` = fmt_pct(median_type1_error),
      `Mean retained replications` = fmt_num(mean_retained, 1)
    ) |>
    select(
      Method, `Coverage conditions` = coverage_conditions,
      `Mean coverage`, `Median coverage`,
      `Type I conditions` = type1_conditions,
      `Mean Type I error`, `Median Type I error`,
      `Mean retained replications`
    )
  knitr::kable(
    historical_inference_table,
    caption = "Condition-weighted historical-view interval performance; Type I error is evaluated only when beta[Z.u1] = 0"
  )
}
```

The reference lines are .95 for coverage and .05 for Type I error.  The table
uses condition-level rates (not a pooled replication denominator), so it
describes how performance varies across the Lai design.  Interpret a departure
from either reference together with the matching bias panel and the number of
retained replications: a nominal interval result can change when the
historical reporting multiplier or status-code retention changes.

If `params$historical_figure_reference` points to a locally saved image of the
published Figure 2, render it alongside this panel during review.  The visual
comparison should ask whether the sign and correlation pattern agree, not
whether the current all-method VH analysis matches a method-specific Lai scale.

```{r published-reference, fig.cap="Optional local reference image of Lai's published Figure 2."}
published_reference <- resolve_repo_path(params$historical_figure_reference)
if (!is.null(published_reference) && file.exists(published_reference)) {
  knitr::include_graphics(published_reference)
}
```

## 3. VH-primary performance: seven methods on a common latent-SD scale

The VH-primary view uses the `latent_sd` reporting scale and the VH eligibility
rule.  Each point in the next plot is a condition-level Monte Carlo mean bias;
the boxplots summarize the 486-condition distribution rather than pooling
replications across different DGMs.

```{r vh-instability-table}
if (!is.null(vh_instability_summary)) {
  vh_instability_table <- vh_instability_summary |>
    mutate(
      Method = recode(method, !!!method_labels),
      `Median bias` = fmt_num(median_bias),
      `Median |bias|` = fmt_num(median_abs_bias),
      `Maximum |bias|` = fmt_num(max_abs_bias, 2)
    ) |>
    select(Method, `Median bias`, `Median |bias|`,
           `Maximum |bias|`, `Conditions: |bias| > .1` = n_abs_bias_over_01,
           `Conditions: |bias| > 1` = n_abs_bias_over_1)
  knitr::kable(
    vh_instability_table,
    caption = "Central performance and extreme condition-level bias on the VH-primary scale"
  )
}
```

```{r vh-performance-table}
vh_performance <- latent_df |>
  group_by(method_label) |>
  summarise(
    conditions = n(),
    `Mean bias` = safe_mean(bias),
    `Median |bias|` = safe_median(abs(bias)),
    `Mean RMSE` = safe_mean(rmse),
    `Mean coverage` = safe_mean(coverage),
    `Mean VH eligibility` = safe_mean(vh_analysis_eligibility_rate),
    .groups = "drop"
  ) |>
  mutate(
    `Mean bias` = fmt_num(`Mean bias`),
    `Median |bias|` = fmt_num(`Median |bias|`),
    `Mean RMSE` = fmt_num(`Mean RMSE`),
    `Mean coverage` = fmt_pct(`Mean coverage`),
    `Mean VH eligibility` = fmt_pct(`Mean VH eligibility`)
  )
knitr::kable(vh_performance, caption = "Condition-weighted VH-primary performance summary")
```

```{r vh-primary-companion, fig.cap="VH-primary central-range companion. The fixed ±0.2 range is intentional; rare larger failures are shown in the following diagnostic."}
vh_plot_path <- file.path(figure_dir, "vh_primary_companion.png")
if (file.exists(vh_plot_path)) {
  knitr::include_graphics(vh_plot_path)
} else {
  knitr::asis_output(
    "**Companion Figure 2 panel not yet generated.** It will appear after the post-estimation figure command has run."
  )
}
```

```{r vh-primary-extremes, fig.cap="Full absolute-bias distribution on a log scale. This retains, rather than trims, rare unstable conditions."}
vh_extreme_path <- file.path(figure_dir, "vh_primary_extremes.png")
if (file.exists(vh_extreme_path)) {
  knitr::include_graphics(vh_extreme_path)
}
```

The central-range panel shows that the closed-form dual estimator is the main
source of systematic negative central bias when \(\beta_{Zu1}=.4\), while the
two Fuller estimators, 2S-PA, MSEM, and oracle are generally much closer to
zero.  The log-scale diagnostic adds the complementary fact that naïve dual
BLUP has rare, severe failures even though its median is near zero.  Neither
claim is visible in the original unbounded all-method boxplot by itself.

## 4. Reporting-scale sensitivity: separate rescaling from estimator behavior

Raw and latent-SD rows report the same fitted estimator on different common
condition-level scales.  The historical standardized row is different: for
the directly comparable methods it preserves Lai's method-specific multiplier.
Consequently, this section is a diagnostic of reporting conventions, not a
competition between differently scaled numerical values.

```{r scale-sensitivity, fig.width=11, fig.height=6, fig.cap="Condition-level bias by reporting view. Facets have separate vertical scales because their units differ."}
scale_plot_df <- summary_df |>
  filter(method %in% names(method_labels)) |>
  mutate(scale_label = recode(
    as.character(reporting_scale),
    raw = "Raw",
    latent_sd = "Common latent-SD",
    lai_original_standardized = "Lai historical standardized"
  ))

ggplot(scale_plot_df, aes(x = method_label, y = bias)) +
  geom_hline(yintercept = 0, colour = "grey35", linewidth = 0.4) +
  geom_boxplot(fill = "#fdd0a2", colour = "#9c3d00", outlier.alpha = 0.2) +
  coord_flip() +
  facet_wrap(~ scale_label, scales = "free_x") +
  labs(
    x = NULL,
    y = "Condition-level mean bias (view-specific units)",
    title = "Reporting-scale sensitivity",
    subtitle = "The historical-standardized view is available only for methods with a direct Lai analogue"
  ) +
  theme_bw(base_size = 10)
```

```{r historical-multiplier-table}
multiplier_summary <- historical_df |>
  group_by(method_label) |>
  summarise(
    `Mean Lai historical multiplier` = safe_mean(mean_lai_original_method_multiplier),
    `Availability` = safe_mean(lai_original_reporting_available_rate),
    .groups = "drop"
  ) |>
  mutate(
    `Mean Lai historical multiplier` = fmt_num(`Mean Lai historical multiplier`),
    Availability = fmt_pct(Availability)
  )
knitr::kable(
  multiplier_summary,
  caption = "Method-specific historical scaling diagnostics"
)
```

The key review question is whether an apparent change in bias survives the
common latent-SD view.  If it does not, the historical comparison is likely
being driven partly by reporting scale rather than by a change in the fitted
coefficient.

## 5. Slope-SD and reliability diagnostics: test the calibration explanation

The posterior reliability plotted here is calculated from the true full
two-dimensional random-effect covariance and Lai's RMS-scaled within-cluster
design.  It should not be replaced by `lambda22` when random effects are
correlated.  The first plot makes the dependence on cluster size visible while
retaining the remaining DGM variation as individual points.

```{r reliability-by-size, fig.width=10, fig.height=7, fig.cap="True posterior slope reliability by homogeneous cluster size, stratified by ICC and random-effect correlation."}
reliability_df <- raw_df |>
  distinct(
    condition_id, num_clus, clus_size, icc, cor_u0_u1,
    dgm_posterior_slope_reliability
  ) |>
  mutate(
    clus_size = factor(clus_size),
    num_clus = factor(num_clus),
    icc = factor(icc),
    cor_u0_u1 = factor(cor_u0_u1)
  )

ggplot(reliability_df, aes(x = clus_size, y = dgm_posterior_slope_reliability, shape = num_clus)) +
  geom_jitter(width = 0.16, height = 0, alpha = 0.45, colour = "#1f4e79", size = 1.5) +
  facet_grid(icc ~ cor_u0_u1, labeller = label_both) +
  labs(
    x = "Homogeneous cluster size",
    y = "True posterior slope reliability",
    shape = "Clusters",
    title = "Reliability is an emergent property of the Lai DGM"
  ) +
  theme_bw(base_size = 10)
```

```{r reliability-by-size-table}
reliability_by_size <- reliability_df |>
  group_by(clus_size) |>
  summarise(
    `Minimum reliability` = min(dgm_posterior_slope_reliability),
    `Median reliability` = median(dgm_posterior_slope_reliability),
    `Maximum reliability` = max(dgm_posterior_slope_reliability),
    .groups = "drop"
  ) |>
  mutate(across(-clus_size, fmt_num))
knitr::kable(
  reliability_by_size,
  caption = "True posterior slope reliability varies substantially within every homogeneous cluster-size level"
)
```

```{r slope-sd-ratios, fig.width=10, fig.height=5, fig.cap="Stage-1 slope-SD diagnostics relative to the DGM population slope SD."}
slope_sd_df <- raw_df |>
  filter(method == "naive_dual_blup") |>
  select(
    condition_id, dgm_posterior_slope_reliability,
    mean_eb_to_population_slope_sd,
    mean_corrected_to_population_slope_sd,
    mean_fitted_to_population_slope_sd
  ) |>
  pivot_longer(
    cols = starts_with("mean_") ,
    names_to = "diagnostic",
    values_to = "slope_sd_ratio"
  ) |>
  mutate(diagnostic = recode(
    diagnostic,
    mean_eb_to_population_slope_sd = "EB / population SD",
    mean_corrected_to_population_slope_sd = "Corrected / population SD",
    mean_fitted_to_population_slope_sd = "Fitted / population SD"
  ))

ggplot(slope_sd_df, aes(x = dgm_posterior_slope_reliability, y = slope_sd_ratio)) +
  geom_hline(yintercept = 1, colour = "grey35", linetype = "dashed") +
  geom_point(alpha = 0.45, colour = "#7a3e9d", size = 1.4) +
  facet_wrap(~ diagnostic, nrow = 1) +
  labs(
    x = "True posterior slope reliability",
    y = "Mean slope SD / population slope SD",
    title = "Stage-1 slope-SD diagnostics versus true reliability"
  ) +
  theme_bw(base_size = 10)
```

The slope-SD ratios use the same Stage-1 fit attached to the naive dual BLUP
raw summary, avoiding duplicate copies of the diagnostics across reporting
views.  A ratio of one is a diagnostic reference, not a convergence criterion.

The reliability result is consistent with cluster size being an important
determinant, but rejects the stronger claim that it is merely a proxy for
cluster size: at size 3 the reliability spans approximately .07--.86, and at
size 25 it still spans approximately .40--.98.  The EB slope SD is most
attenuated at low reliability, whereas the fitted slope SD remains close to
the population slope SD across the reliability range.  This makes the
slope-SD correction a useful diagnostic for the observed bias pattern rather
than a substitute for modelling reliability itself.

## 6. Eligibility and convergence: compare the two retention rules directly

VH eligibility and original-Lai eligibility answer different questions.
The historical rule retains runs with `status_code == 0`; the VH-primary rule
also applies estimator-specific finite-estimate, finite-SE, design, and fit
diagnostics.  The following plot displays the average condition-level rate for
each rule; it should accompany any comparison that changes the reporting scale
or estimator bundle.

```{r eligibility-rates, fig.width=10, fig.height=5, fig.cap="Average condition-level retention under the original-Lai and VH-primary rules."}
eligibility_df <- raw_df |>
  select(method_label, vh_analysis_eligibility_rate, lai_original_eligibility_rate) |>
  pivot_longer(
    cols = c(vh_analysis_eligibility_rate, lai_original_eligibility_rate),
    names_to = "rule",
    values_to = "eligibility_rate"
  ) |>
  mutate(rule = recode(
    rule,
    vh_analysis_eligibility_rate = "VH primary",
    lai_original_eligibility_rate = "Original Lai"
  ))

ggplot(eligibility_df, aes(x = method_label, y = eligibility_rate, shape = rule)) +
  stat_summary(fun = mean, geom = "point", position = position_dodge(width = 0.45), size = 2.6, colour = "#1f4e79") +
  coord_flip() +
  scale_y_continuous(labels = percent_format(accuracy = 10), limits = c(0, 1)) +
  labs(
    x = NULL,
    y = "Mean condition-level eligibility rate",
    shape = NULL,
    title = "Eligibility depends on the reporting rule"
  ) +
  theme_bw(base_size = 11)
```

```{r eligibility-overlap-table}
eligibility_overlap <- raw_df |>
  group_by(method_label) |>
  summarise(
    `VH eligible` = safe_mean(vh_analysis_eligibility_rate),
    `Lai eligible` = safe_mean(lai_original_eligibility_rate),
    `Both eligible (mean count)` = safe_mean(n_eligible_both),
    `Lai-only (mean count)` = safe_mean(n_lai_original_only),
    `VH-only (mean count)` = safe_mean(n_vh_only),
    `Convergence` = safe_mean(convergence),
    .groups = "drop"
  ) |>
  mutate(
    `VH eligible` = fmt_pct(`VH eligible`),
    `Lai eligible` = fmt_pct(`Lai eligible`),
    `Both eligible (mean count)` = fmt_num(`Both eligible (mean count)`, 1),
    `Lai-only (mean count)` = fmt_num(`Lai-only (mean count)`, 1),
    `VH-only (mean count)` = fmt_num(`VH-only (mean count)`, 1),
    Convergence = fmt_pct(Convergence)
  )
knitr::kable(
  eligibility_overlap,
  caption = "Condition-averaged eligibility overlap and convergence"
)
```

For this run, the VH screen is a strict subset of the original-Lai screen:
there are no VH-only replications.  The material differences occur for 2S-PA
and naïve dual BLUP, for which about 2% of replications per condition are
retained by the original status-code rule but excluded from the VH-primary
analysis.  Thus the rare large naïve-BLUP biases are not an artifact of a more
permissive VH retention rule; the stricter VH screen is already in force.

### Recommended review sequence

1. Read the historical analogue beside Lai's Figure 2, comparing sign and
   correlation patterns only for the three overlapping methods.
2. Use the VH-primary panel to decide whether the low-reliability negative
   bias persists after common latent-SD reporting and VH eligibility.
3. Consult the reporting-scale and slope-SD sections before attributing a
   cross-method bias difference to an estimator.
4. Treat eligibility changes as part of the result: a discrepancy that appears
   only after a retention-rule change is not evidence of a DGM mismatch alone.

## Reproducibility notes

The report is parameterized.  After generating the Figure 2 post-processing
outputs, render it from the repository root with:

```sh
quarto render reports/lai_study1_vh_review.qmd \
  -P results_dir:outputs/lai_study1_vh_production_20260731
```

Optionally supply `figure_dir` when figures live elsewhere and
`historical_figure_reference` for a locally saved copy of Lai's published
Figure 2.  This report does not rerun any models or alter the aggregate output.
