---
title: "Audit of the Fuller (1987) Algebra in the VH Simulations"
author: "Simulation and implementation audit"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
params:
  output_dir: "outputs/fuller_algebra_audit"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 120
)
suppressPackageStartupMessages({
  library(data.table)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
  library(scales)
})

input_path <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) "")
find_repo_root <- function(start_paths) {
  for (start_path in unique(start_paths[nzchar(start_paths)])) {
    candidate <- normalizePath(start_path, mustWork = FALSE)
    for (level in 0:5) {
      if (
        file.exists(file.path(candidate, "R", "stage2_estimators.R")) &&
        dir.exists(file.path(candidate, "vig_hallquist_2026"))
      ) {
        return(normalizePath(candidate, mustWork = TRUE))
      }
      parent <- dirname(candidate)
      if (identical(parent, candidate)) break
      candidate <- parent
    }
  }
  stop("Could not locate the repository root for the Fuller audit report.")
}
repo_root <- find_repo_root(c(
  getwd(),
  if (nzchar(input_path)) dirname(input_path) else "",
  if (nzchar(input_path)) dirname(file.path(getwd(), input_path)) else ""
))
resolve_repo_path <- function(path) {
  if (grepl("^/", path)) path else file.path(repo_root, path)
}

output_dir <- resolve_repo_path(params$output_dir)
replication_path <- file.path(
  output_dir, "fuller_algebra_replication_results.csv.gz"
)
summary_path <- file.path(output_dir, "fuller_algebra_summary.csv")
validation_path <- file.path(
  output_dir, "fuller_algebra_production_validation.csv"
)
stopifnot(
  file.exists(replication_path),
  file.exists(summary_path),
  file.exists(validation_path)
)

replications <- fread(replication_path) %>% as_tibble()
performance <- fread(summary_path) %>% as_tibble()
validation <- fread(validation_path) %>% as_tibble()

scenario_labels <- c(
  no_measurement_error = "No measurement error",
  outcome_homoskedastic = "Outcome error: constant",
  outcome_heteroskedastic = "Outcome error: heterogeneous",
  predictor_only = "Predictor error only",
  both_homoskedastic = "Both errors: constant",
  both_heteroskedastic = "Both errors: heterogeneous"
)
variant_labels <- c(
  current = "Former production (bug)",
  cross_moment_fixed = "Corrected VH (stabilized)",
  documented_preliminary = "Corrected + Fuller preliminary",
  documented_full = "Corrected + Fuller equations"
)
geometry_labels <- c(
  regular = "Regular G shape",
  weak_shape = "Ill-conditioned G shape"
)
variant_palette <- c(
  "Former production (bug)" = "#E15759",
  "Corrected VH (stabilized)" = "#4E79A7",
  "Corrected + Fuller preliminary" = "#B6992D",
  "Corrected + Fuller equations" = "#59A14F"
)

performance <- performance %>%
  mutate(
    scenario_label = factor(
      unname(scenario_labels[scenario]), levels = unname(scenario_labels)
    ),
    variant_label = factor(
      unname(variant_labels[variant]), levels = unname(variant_labels)
    ),
    geometry_label = factor(
      unname(geometry_labels[geometry]), levels = unname(geometry_labels)
    )
  )

replications <- replications %>%
  mutate(
    eligible = status == "ok" & is.finite(estimate) & is.finite(se) & se > 0,
    covered = eligible & ci_low <= truth & ci_high >= truth,
    scenario_label = factor(
      unname(scenario_labels[scenario]), levels = unname(scenario_labels)
    ),
    variant_label = factor(
      unname(variant_labels[variant]), levels = unname(variant_labels)
    ),
    geometry_label = factor(
      unname(geometry_labels[geometry]), levels = unname(geometry_labels)
    )
  )

theme_audit <- function() {
  theme_bw(base_size = 12) +
    theme(
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      strip.background = element_rect(fill = "grey95")
    )
}

fmt <- function(x, digits = 3) formatC(x, digits = digits, format = "f")

outcome_focal <- performance %>%
  filter(
    scenario == "outcome_homoskedastic", m == 200,
    variant %in% c("current", "cross_moment_fixed")
  )
regular_current <- outcome_focal %>%
  filter(geometry == "regular", variant == "current") %>% slice(1)
regular_fixed <- outcome_focal %>%
  filter(geometry == "regular", variant == "cross_moment_fixed") %>% slice(1)
weak_current <- outcome_focal %>%
  filter(geometry == "weak_shape", variant == "current") %>% slice(1)
weak_fixed <- outcome_focal %>%
  filter(geometry == "weak_shape", variant == "cross_moment_fixed") %>% slice(1)

predictor_m200 <- performance %>%
  filter(
    m == 200,
    scenario %in% c(
      "predictor_only", "both_homoskedastic", "both_heteroskedastic"
    )
  ) %>%
  group_by(variant, variant_label) %>%
  summarise(
    coverage = mean(coverage),
    bias = mean(bias),
    median_cell_mean_se_ratio = median(mean_se_ratio),
    median_cell_median_se_ratio = median(median_se_ratio),
    maximum_cell_mean_se_ratio = max(mean_se_ratio),
    .groups = "drop"
  )

overall <- performance %>%
  group_by(variant, variant_label) %>%
  summarise(
    equal_cell_coverage = mean(coverage),
    equal_cell_bias = mean(bias),
    .groups = "drop"
  )

mc_band <- c(
  qbinom(.025, 1000, .95) / 1000,
  qbinom(.975, 1000, .95) / 1000
)
```

# Executive conclusion

The flagged line was a **real algebraic error**, not a harmless coding style
choice. Fuller (1987, pp. 193--194, Equations 3.1.19 and 3.1.26) subtracts the
predictor--outcome measurement-error covariance from the preliminary
predictor--outcome cross moment. The former implementation instead subtracted
the scalar *outcome measurement-error variance* from every element of that
cross-moment vector. R silently recycled the scalar, so the calculation ran
without warning.

This error is now corrected in `fit_fuller_dual_core()`. The estimator accepts
separate predictor--outcome covariance columns, uses them throughout the
Fuller equations, and defaults omitted columns to exact zeros. VH Studies 1--4
also create and pass named zero columns, making their separate-process
assumption explicit in the saved pipeline diagnostics. The pipeline version
was advanced so pre-correction results cannot be mistaken for corrected runs.

This error affects any pathway with an error-prone Stage-2 outcome---VH Studies
1 and 3---but it is algebraically inactive in Studies 2 and 4 because their
Stage-2 outcome-error variance is zero. It therefore cannot explain the severe
non-null undercoverage in Studies 2 and 4. It is, however, a plausible and now
demonstrated contributor to Fuller's overcoverage in Studies 1 and 3.

In the clean outcome-error-only simulation with 200 clusters:

- Under regular covariance geometry, former-production coverage is
  `r percent(regular_current$coverage, accuracy=.1)` and the mean SE is
  `r fmt(regular_current$mean_se_ratio, 2)` times the empirical SD. Fixing only
  the erroneous cross moment gives `r percent(regular_fixed$coverage, accuracy=.1)`
  coverage and an SE ratio of `r fmt(regular_fixed$mean_se_ratio, 2)`.
- Under the ill-conditioned covariance shape, former-production coverage is
  `r percent(weak_current$coverage, accuracy=.1)` and its SE ratio is
  `r fmt(weak_current$mean_se_ratio, 2)`. The one-line algebra fix gives
  `r percent(weak_fixed$coverage, accuracy=.1)` coverage and an SE ratio of
  `r fmt(weak_fixed$mean_se_ratio, 2)`.

Two additional differences from Fuller (1987) were confirmed. The corrected VH
primary estimator uses
the finite-moment modified estimator as its preliminary estimator, and it uses
the modified final predictor matrix in the variance bread. Fuller uses the
ordinary corrected moment estimator for the preliminary weights and the fully
subtracted weighted matrix for the variance bread. These are asymptotically
equivalent substitutions, not errors of the same kind as the recycled outcome
variance. In the simulations they act as strong finite-sample regularizers. A
wholesale switch to the book variance bread brings equal-cell coverage close to
95%, but produces a small number of enormous SEs when the fully corrected bread
approaches singularity. That version should therefore be studied as an explicit
alternative, not installed as an unexamined replacement. Accordingly, Studies
1--4 now retain a nested three-method Fuller sequence: corrected/stabilized,
book preliminary with stabilized bread, and the book preliminary plus book
bread. This sequence isolates which stabilization controls the SE tail.

# Audit standard and sources

The governing primary source is William A. Fuller, *Measurement Error Models*
(1987), especially:

- Section 2.5, pp. 163--175, on finite-moment modified estimators;
- Section 3.1.1, pp. 186--193, on unequal measurement-error variances;
- Section 3.1.2, pp. 193--195, Equations 3.1.19--3.1.28, on a regression with
  measurement error in both the outcome and predictors.

The audit was checked against a local research copy of Fuller (1987), which is
not versioned with this repository. The project's earlier derivation in
`documentation/BLUP_Project.pdf`, Equations 55--69, is likewise a local project
document rather than a PR artifact. The audit therefore evaluates the code
against both the primary source and the project's stated estimator without
placing either PDF in version control.

This report concerns algebra and finite-sample behavior conditional on supplied
measurement-error covariance matrices. Whether estimated Stage-1 covariance
matrices require additional uncertainty propagation is a separate question.

# Mathematical derivation

## Measurement-error model

For cluster $i=1,\ldots,M$, let the observed outcome and predictor vector be

\[
Y_i = y_i + w_i, \qquad
\mathbf X_i = \mathbf x_i + \mathbf u_i,
\]

with structural relation

\[
y_i = \mathbf x_i^\top\boldsymbol\beta + q_i.
\]

The predictor vector includes the Stage-2 intercept. Its corresponding
measurement-error row and column are zero. Define

\[
\mathbf Z_i=(Y_i,\mathbf X_i^\top)
\]

and the supplied measurement-error covariance

\[
\boldsymbol\Sigma_{aa,i}=
\begin{bmatrix}
\sigma_{ww,i} & \boldsymbol\Sigma_{wu,i}\\
\boldsymbol\Sigma_{uw,i} & \boldsymbol\Sigma_{uu,i}
\end{bmatrix}.
\]

For VH Studies 1 and 3, outcome and predictor scores come from separate
first-stage models, so
$\boldsymbol\Sigma_{uw,i}=\mathbf 0$. This does **not** imply that
$\sigma_{ww,i}=0$.

## Fuller's preliminary estimator

Fuller's Equation 3.1.19 is the corrected method-of-moments estimator

\[
\widetilde{\boldsymbol\beta}
=
\widehat{\mathbf M}_{xx}^{-1}\widehat{\mathbf M}_{xy},
\]

where

\[
\widehat{\mathbf M}_{xx}
=\frac{1}{M}\sum_i
\left(\mathbf X_i\mathbf X_i^\top-
\boldsymbol\Sigma_{uu,i}\right),
\]

and

\[
\widehat{\mathbf M}_{xy}
=\frac{1}{M}\sum_i
\left(\mathbf X_iY_i-\boldsymbol\Sigma_{uw,i}\right).
\]

The outcome-error variance $\sigma_{ww,i}$ does not appear in
$\widehat{\mathbf M}_{xy}$. Outcome measurement error increases uncertainty
but does not bias the predictor--outcome cross moment when it is independent of
predictor measurement error.

The former production code instead computed, in simplified notation,

\[
\mathbf d_{0,\text{code}}
=\sum_i \mathbf X_iY_i
-c_1\left(\sum_i\sigma_{ww,i}\right)\mathbf 1_P,
\]

where $c_1$ is a lambda/alpha correction and $\mathbf 1_P$ results from
scalar recycling. Even when
$\boldsymbol\Sigma_{uw,i}=\mathbf0$, this inserts a nonzero correction into
the intercept and every predictor cross moment.

## Residual variance and weights

Using a preliminary estimate, Fuller estimates the structural residual
variance as

\[
\widehat\sigma_q^2=
\max\left\{0,
\frac{1}{M-P}\sum_i
(Y_i-\mathbf X_i^\top\widetilde{\boldsymbol\beta})^2
-\frac{1}{M}\sum_i
\begin{bmatrix}1\\-\widetilde{\boldsymbol\beta}\end{bmatrix}^{\!\top}
\boldsymbol\Sigma_{aa,i}
\begin{bmatrix}1\\-\widetilde{\boldsymbol\beta}\end{bmatrix}
\right\}.
\]

The case weight is

\[
\widehat v_i=\widehat\sigma_q^2+\sigma_{ww,i}
-2\widetilde{\boldsymbol\beta}^{\top}\boldsymbol\Sigma_{uw,i}
+\widetilde{\boldsymbol\beta}^{\top}
\boldsymbol\Sigma_{uu,i}\widetilde{\boldsymbol\beta}.
\]

Consequently, an erroneous preliminary estimate propagates into
$\widehat\sigma_q^2$, every weight, and the model-based SE. With constant
measurement-error covariance, all weights can remain equal; the final
coefficient may then be unchanged even while its SE is badly distorted.

## Modified final estimator

Fuller's Equation 3.1.20 modifies the fully corrected estimator to avoid the
undefined or extremely heavy moments of an unprotected ratio estimator. Let
$\widehat\lambda$ be the smallest root of

\[
\left|\sum_i
(\mathbf Z_i^\top\mathbf Z_i-\lambda
\boldsymbol\Sigma_{aa,i})\right|=0.
\]

The final corrected matrix subtracts $c\boldsymbol\Sigma_{aa,i}$, where

\[
c=\begin{cases}
1-\alpha/M,&\widehat\lambda>1+M^{-1},\\
\widehat\lambda-M^{-1}-\alpha/M,
&\widehat\lambda\leq1+M^{-1}.
\end{cases}
\]

The corrected implementation retains this final modification, including its weighted
analogue. The choice $\alpha=P+1$ is a finite-sample tuning choice consistent
with using a positive stabilizer, but it is not uniquely dictated by Equation
3.1.20. Section 2.5 motivates lower bounds and alternative choices under more
specific normal/Wishart assumptions.

## Variance formula

For the weighted estimator, Fuller's Equations 3.1.27--3.1.28 use

\[
\widehat{\mathbf M}_{xxx}
=\frac1M\sum_i\widehat v_i^{-1}
(\mathbf X_i\mathbf X_i^\top-\boldsymbol\Sigma_{uu,i})
\]

as the variance bread and

\[
\widetilde{\boldsymbol\Sigma}_{uw,i}
=\boldsymbol\Sigma_{uw,i}
-\boldsymbol\Sigma_{uu,i}\widetilde{\boldsymbol\beta}.
\]

Under normal measurement errors,

\[
\widehat{\operatorname{Var}}(\widehat{\boldsymbol\beta})
=M^{-2}\widehat{\mathbf M}_{xxx}^{-1}
\left[\sum_i\left\{
\widehat v_i^{-1}\mathbf X_i\mathbf X_i^\top
+\widehat v_i^{-2}
\widetilde{\boldsymbol\Sigma}_{uw,i}
\widetilde{\boldsymbol\Sigma}_{uw,i}^{\top}
\right\}\right]
\widehat{\mathbf M}_{xxx}^{-1}.
\]

The corrected implementation constructs the meat using the supplied
$\boldsymbol\Sigma_{uw,i}$ (zero in VH Studies 1--4), but its primary variant
uses the modified final
$\mathbf S_x^*/M$ as the bread. Because that matrix subtracts less error
variance in difficult samples, it acts as a regularizer. It is asymptotically
equivalent to Fuller's bread when $c\rightarrow1$, but it is not Equation
3.1.28 in finite samples.

# Line-by-line algebra disposition

```{r algebra-disposition}
tibble::tribble(
  ~Component, ~Fuller_1987, ~Production_implementation, ~Disposition,
  "Measurement-error block",
  "Outcome variance, predictor covariance, and their cross covariance are distinct",
  "Stores all three blocks; omitted cross-covariance columns become exact zeros",
  "Correct and general; VH explicitly supplies zeros",
  "Preliminary X'Y moment",
  "Subtract sum of predictor-outcome error covariances",
  "Subtracts the supplied predictor-outcome covariance; never substitutes outcome variance",
  "Corrected; former code had a definite algebra error",
  "Preliminary X'X moment",
  "Subtract full predictor-error covariance",
  "Subtracts a lambda/alpha-modified predictor covariance",
  "Undocumented asymptotically equivalent stabilization",
  "Residual variance",
  "Residual mean square minus composite measurement variance; truncate at zero",
  "Same formula with the selected corrected preliminary estimate",
  "Correct",
  "Final point estimator",
  "Weighted lambda/alpha-modified corrected moment estimator",
  "Same block construction and correction",
  "Correct",
  "Variance meat",
  "Weighted X'X plus the squared tilde-covariance term",
  "Uses omega_xy - Omega_x beta for supplied or zero cross covariance",
  "Corrected and general",
  "Variance bread",
  "Fully measurement-error-subtracted weighted predictor moment",
  "Lambda/alpha-modified final predictor moment",
  "Undocumented asymptotically equivalent stabilization",
  "Uncertainty in supplied error covariance",
  "Known covariance in Section 3.1.2; estimated covariance treated under additional conditions elsewhere",
  "Treats fitted Stage-1 covariance as supplied",
  "Separate inferential issue, not this algebra error"
) %>%
  kable(
    col.names = c(
      "Component", "Fuller (1987)", "Corrected production", "Audit conclusion"
    ),
    caption = "Disposition of each material Fuller calculation."
  )
```

# Independent implementation check

The audit implementation was written independently in
`fuller_algebra_audit_simulation.R`. Its corrected/stabilized branch was then
compared with `fit_fuller_dual()` over 100 newly generated datasets containing
regular and ill-conditioned covariance shapes and constant or heterogeneous
measurement-error variances.

```{r validation-table}
validation %>%
  summarise(
    Replications = n(),
    `Max |estimate difference|` = max(abs(estimate_difference), na.rm = TRUE),
    `Max |SE difference|` = max(abs(se_difference), na.rm = TRUE),
    `Max |lambda1 difference|` = max(abs(lambda1_difference), na.rm = TRUE),
    `Max |lambda2 difference|` = max(abs(lambda2_difference), na.rm = TRUE),
    `Max |sigma2 difference|` = max(abs(sigma2_difference), na.rm = TRUE)
  ) %>%
  mutate(across(where(is.numeric) & !matches("Replications"), scientific)) %>%
  kable(caption = "Independent audit code versus the production implementation.")
```

The maximum estimate discrepancy is
`r scientific(max(abs(validation$estimate_difference)))`, and the maximum SE
discrepancy is `r scientific(max(abs(validation$se_difference)))`. The audit
therefore reproduces corrected production to numerical precision.

The regression tests also exercise a positive, nonzero
$\boldsymbol\Sigma_{uw,i}$ in both the dual-predictor and single-predictor
paths. Production agrees with the independent implementation for the final
coefficient, structural variance, both generalized-eigen roots, and SE. Thus
zero is a VH design assumption supplied to a general covariance implementation,
not a hard-coded limitation of the estimator.

# Focused simulation

## Design

The simulation is deliberately a direct additive measurement-error model. It
does not fit a mixed model first; doing so would confound the algebra check with
Stage-1 estimation. The supplied error covariance matrices are the true
matrices, matching Fuller's known-covariance setup.

Each cell has 1,000 replications and either 50 or 200 Stage-2 units. The true
standardized focal coefficient is .40. The two true predictor covariance shapes
are:

- regular: $\tau_1^2/\tau_0^2=1$, $\rho=.50$,
  $\kappa(G)=3$;
- ill-conditioned: $\tau_1^2/\tau_0^2=.005$, $\rho=.50$,
  $\kappa(G)=267.3$.

Predictor error, when present, corresponds marginally to reliability .80 for
the intercept-like predictor and .25 for the slope-like predictor. Outcome
error has variance one. Heterogeneous cells use known case-specific variance
multipliers $(.2,1,1.8)$. Predictor and outcome measurement errors are
independent, as in the VH separate-process construction.

Four nested estimators isolate the code differences. The first is retained only
as a historical bug comparator; the remaining three are now implemented as the
VH Fuller method sequence:

1. **Former production (bug)** reproduces the pre-correction scalar recycling.
2. **Corrected VH (stabilized)** uses the correct cross covariance while
   retaining both finite-sample stabilizations; this is the primary method
   `fuller_closed_form`.
3. **Corrected + Fuller preliminary** also uses Equation 3.1.19 for the
   preliminary estimator while retaining the stabilized variance bread; this
   is `fuller_book_preliminary_closed_form`.
4. **Corrected + Fuller equations** additionally uses the Equation 3.1.28
   variance bread; this is `fuller_book_closed_form`.

All `r nrow(performance)` condition-by-estimator summaries have 100% numerical
eligibility. Thus coverage differences are not generated by excluding failed
fits.

## The flagged cross-moment error

```{r focal-outcome-table}
outcome_focal %>%
  transmute(
    Geometry = unname(geometry_labels[geometry]),
    Version = unname(variant_labels[variant]),
    `Mean preliminary slope` = round(mean_gamma0, 3),
    `Mean structural variance` = round(mean_sigma2, 3),
    `Mean SE / empirical SD` = round(mean_se_ratio, 3),
    Coverage = percent(coverage, accuracy = .1),
    Bias = round(bias, 3)
  ) %>%
  kable(
    caption = paste(
      "Outcome-error-only diagnostic with 200 clusters. The true standardized",
      "preliminary/focal slope is .40."
    )
  )
```

The final point estimates remain centered in this homoskedastic diagnostic
because all case weights are equal and their common scale cancels from the
final estimating equation. The preliminary estimate and estimated uncertainty
do not enjoy that cancellation. Under regular geometry, the production
preliminary slope averages `r fmt(regular_current$mean_gamma0)` rather than .40;
under the weak shape it averages `r fmt(weak_current$mean_gamma0, 2)`. These
values inflate the estimated structural variance from approximately
`r fmt(weak_fixed$mean_sigma2)` to `r fmt(weak_current$mean_sigma2, 1)` in the
weak-shape cell.

With heterogeneous outcome-error variances, weights no longer share a common
scale. The erroneous preliminary estimate can then also change the final point
estimate and empirical sampling variance. This is why the issue cannot be
dismissed merely because the homoskedastic point estimates agree.

```{r coverage-plot, fig.width=11, fig.height=7.4}
ggplot(
  performance,
  aes(x = scenario_label, y = coverage, colour = variant_label, group = variant_label)
) +
  annotate(
    "rect", xmin = -Inf, xmax = Inf, ymin = mc_band[[1]], ymax = mc_band[[2]],
    fill = "grey70", alpha = .25
  ) +
  geom_hline(yintercept = .95, linetype = 2) +
  geom_point(position = position_dodge(width = .35), size = 2) +
  facet_grid(geometry_label ~ m, labeller = label_both) +
  scale_colour_manual(values = variant_palette) +
  scale_y_continuous(labels = label_percent(), limits = c(.90, 1.005)) +
  labs(
    x = NULL,
    y = "95% Wald coverage",
    colour = NULL,
    title = "Coverage after isolating each Fuller algebra choice",
    subtitle = "Grey band is the exact 95% Monte Carlo range for 1,000 replications"
  ) +
  theme_audit() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Point bias

```{r bias-plot, fig.width=11, fig.height=7.4}
ggplot(
  performance,
  aes(x = scenario_label, y = bias, colour = variant_label, group = variant_label)
) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_point(position = position_dodge(width = .35), size = 2) +
  facet_grid(geometry_label ~ m, labeller = label_both) +
  scale_colour_manual(values = variant_palette) +
  labs(
    x = NULL,
    y = "Mean bias on standardized scale",
    colour = NULL,
    title = "The cross-moment bug primarily distorts uncertainty",
    subtitle = "Point-estimate differences emerge when case-specific error variances change the weights"
  ) +
  theme_audit() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

At (M=50), the modified final Fuller estimator still has appreciable
finite-sample point bias in predictor-error cells. All four versions share that
final point estimator, so the algebra fix cannot remove that separate ratio/
weak-information problem. This reinforces the Study-5 conclusion that DGM
geometry and interval algebra are distinct sources of coverage error.

## SE calibration and right tails

```{r median-se-ratio-plot, fig.width=11, fig.height=7.4}
ggplot(
  performance,
  aes(
    x = scenario_label, y = median_se_ratio,
    colour = variant_label, group = variant_label
  )
) +
  geom_hline(yintercept = 1, linetype = 2) +
  geom_point(position = position_dodge(width = .35), size = 2) +
  facet_grid(geometry_label ~ m, labeller = label_both) +
  scale_colour_manual(values = variant_palette) +
  scale_y_log10(breaks = c(.5, .75, 1, 1.5, 2, 5, 10, 25)) +
  labs(
    x = NULL,
    y = "Median model SE / empirical SD (log scale)",
    colour = NULL,
    title = "Typical SE calibration",
    subtitle = "Median SE is shown because book-faithful corrected breads can have extreme right tails"
  ) +
  theme_audit() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

The cross-moment fix removes the dramatic SE inflation in all outcome-error
cells. Using Fuller's unmodified preliminary estimator has little effect on a
typical SE, but creates additional tail variability when predictor information
is weak. The book-faithful variance bread generally raises the typical SE and
moves coverage upward.

However, means tell a different tail-sensitive story:

```{r predictor-tail-table}
predictor_m200 %>%
  transmute(
    Version = as.character(variant_label),
    Coverage = percent(coverage, accuracy = .1),
    Bias = round(bias, 3),
    `Median cell: mean SE/SD` = round(median_cell_mean_se_ratio, 2),
    `Median cell: median SE/SD` = round(median_cell_median_se_ratio, 2),
    `Largest cell mean SE/SD` = round(maximum_cell_mean_se_ratio, 1)
  ) %>%
  kable(
    caption = paste(
      "Equal-cell summaries over all predictor-error scenarios and both",
      "covariance shapes at M=200."
    )
  )
```

The full Fuller version averages
`r percent(predictor_m200$coverage[predictor_m200$variant == "documented_full"], accuracy=.1)`
coverage across these predictor-error cells, close to nominal. But its largest
cell mean-SE ratio is
`r fmt(predictor_m200$maximum_cell_mean_se_ratio[predictor_m200$variant == "documented_full"], 1)`,
while the median-cell median-SE ratio is only
`r fmt(predictor_m200$median_cell_median_se_ratio[predictor_m200$variant == "documented_full"], 2)`.
Rare near-singular fully subtracted breads therefore dominate the mean. This is
the same distinction between typical uncertainty and SE tails that appeared in
VH Study 5.

Fuller explicitly developed the modified estimator because ordinary corrected
ratio estimators can lack finite moments in small samples (Section 2.5). Using
the fully subtracted matrix in the variance bread reintroduces an analogous
operational instability even though it follows Equation 3.1.28. Nominal average
coverage does not by itself make those intervals satisfactory.

# Implications for VH Studies 1--4

```{r study-implication-table}
tibble::tribble(
  ~Study, ~Outcome_error_in_Fuller, ~Predictor_error_in_Fuller, ~Consequence,
  "Study 1: random effect as outcome", "Yes", "No",
  "Directly affected. Point estimates may be unchanged in balanced cells, but preliminary coefficients, residual variance, weights, and SEs must be recomputed.",
  "Study 2: random effect as predictor", "No", "Yes",
  "Not affected by the recycled outcome variance. Legacy non-null undercoverage remains a covariance-geometry/point-bias problem.",
  "Study 3: random effects on both sides", "Yes", "Yes",
  "Directly affected and highest priority for rerunning because the bad preliminary estimate enters both outcome- and predictor-error weighting.",
  "Study 4: heterogeneous predictor information", "No", "Yes",
  "Not affected by the recycled outcome variance. Heterogeneity and corrected-bread instability remain relevant."
) %>%
  kable(
    col.names = c(
      "Study", "Outcome score error?", "Predictor score error?", "Audit implication"
    ),
    caption = "Which historical studies are affected by the confirmed algebra error."
  )
```

The existing Fuller average coverages---about 96.2% in Study 1, 90.5% in
Study 2, 97.3% in Study 3, and 92.8% in Study 4---should therefore not receive a
single explanation. The confirmed algebra error likely contributes to the
overcoverage in Studies 1 and 3. It cannot produce the undercoverage in Studies
2 and 4 because the offending term is identically zero there.

The corrected pipeline will generally narrow Study-1 and Study-3 intervals. It
could bring their conservative coverage closer to nominal, but the direction
and magnitude must be established by rerunning those studies. In Study 3 it can
also alter point estimates when outcome and predictor measurement precision
varies across clusters. Studies 2 and 4 must also be rerun because the pipeline
version and comparison-method set changed, even though their primary corrected
estimate is algebraically unchanged when outcome error and the cross block are
both zero.

Study 1 provides a useful invariant check: its Stage-2 predictor is observed
without error, so both preliminary predictor corrections and both variance
breads reduce to the same matrices. The three corrected Fuller algebra variants
must therefore agree exactly after the outcome/cross-moment fix; the pathway
test now enforces that equality. The variants become substantively distinct in
Studies 2--4, where the Stage-2 predictors are error-prone.

# Recommended disposition

## Production correction now implemented

For the VH separate-process assumption,
$\boldsymbol\Sigma_{uw,i}=\mathbf0$, the preliminary cross moment should be

```r
b0_vec <- as.vector(crossprod(x_mat, y_vec))
```

not

```r
b0_vec <- as.vector(
  crossprod(x_mat, y_vec) - c_correction1 * omega_y_sum
)
```

The implementation now accepts separate covariance columns for the
intercept-like and slope-like predictors, constructs the full covariance block,
and propagates the covariance through the generalized eigenvalues, preliminary
cross moment, composite residual variance, weights, final cross moment, and
variance meat. Omitted inputs default to zero; the VH wrappers nevertheless
pass named zero columns so the assumption is visible rather than implicit.

## Stabilizations that should not be silently changed

The modified preliminary estimator and modified variance bread should be
treated as explicit, named finite-sample stabilizations:

- Retaining the modified preliminary estimator is defensible. It is
  asymptotically equivalent to Fuller's preliminary estimator and substantially
  limits the extreme preliminary coefficients and weight tails in weak samples.
- Retaining the modified variance bread is also operationally defensible, but
  its intervals should not be called the unmodified Fuller (1987) variance
  formula. The simple simulation suggests a modest tendency toward
  undercoverage after the cross-moment fix in predictor-error cells.
- Switching to the Equation 3.1.28 bread improves average coverage here but
  creates extreme SE tails. It should be compared with a full cluster bootstrap
  and a test-inversion approach before being chosen for the paper.

The cleanest terminology would distinguish **modified Fuller point estimation**
from **stabilized Fuller Wald inference**. If both preliminary and bread
stabilizations are retained, the methods section should state them explicitly
and the simulation must validate the resulting procedure rather than relying
on Fuller's asymptotic label alone.

## Completed safeguards and required reruns

The following implementation safeguards are complete:

1. The zero-covariance test verifies that positive outcome error does not enter
   the preliminary cross moment, and a nonzero-covariance test verifies the
   complete general implementation against independent audit code.
2. Exact production-versus-reference tests cover the preliminary estimate,
   structural variance, both lambdas, final coefficient, SE, and all three
   retained algebra variants.
3. VH Studies 1--4 explicitly add and pass zero covariance columns; output
   diagnostics record the covariance source, maximum magnitude, preliminary
   moment, and variance bread.
4. The pipeline version is bumped, and the parallel-worker export list includes
   the new common helper and variant fitter.

The remaining empirical work is to:

1. Rerun compact Study-1v2 and Study-3v2 bridges under the corrected code,
   retaining paired random-number streams and both corrected/stabilized and
   Equation-3.1.28 breads.
2. For Studies 2v2 and 4v2, retain the fixed covariance-shape calibration. The algebra
   fix will not repair their historical non-null bias.
3. Compare stabilized Wald, full-pipeline cluster bootstrap, and test-inverted
   intervals in representative $R=.25$, correlated, and heterogeneous cells.
   Report median and mean SE, upper SE quantiles, coverage, Type I error, bias,
   and eligibility.

# Final conclusion

The audit changes the interpretation of the Fuller results in an important but
bounded way. A genuine outcome/cross-covariance substitution error inflated
Fuller uncertainty whenever the Stage-2 outcome was a measured random effect.
The simple simulation reproduces the mechanism and shows that the one-component
fix can move extreme overcoverage back toward nominal performance. That fix is
now implemented and independently tested, making corrected reruns of Studies 1
and 3 the next mandatory empirical step.

It does **not** overturn the Study-5 diagnosis of Studies 2 and 4. Their
Stage-2 outcomes have no measurement error, so their low non-null coverage
still arises primarily from the legacy slope-only reliability calibration,
ill-conditioned covariance geometry, and finite-sample point-estimator bias.

Finally, the audit does not support blindly replacing every stabilized matrix
with the literal book matrix. Fuller's book-faithful variance bread improves
average coverage in this simple design but produces severe SE tails near weak
identification. The decisive next analysis is therefore the amended Study 1--4
run with the corrected/stabilized Fuller estimator as primary, the two nested
book variants as SE-tail diagnostics, and targeted resampling/test-inversion
checks under the v2 fixed-shape DGM.

# Reproducibility

The simulation can be regenerated from the repository root with:

```bash
Rscript vig_hallquist_2026/analysis/fuller_algebra_audit_simulation.R \
  1000 outputs/fuller_algebra_audit 4
```

```{r session-info}
sessionInfo()
```
