---
  title: "Structural Random Effects"
  author: "Zach Vig & Michael Hallquist"
  date: "`r format(Sys.time(), '%d %B, %Y')`"
  output:
    html_document:
      toc: true
      toc_depth: 2
      toc_float: true
      number_sections: true
    editor_options: 
      chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
pacman::p_load(tidyverse, ggh4x)
source_helpers <- function() {
  analysis_dir <- here::here("vig_hallquist_2026", "analysis")
  source(file.path(analysis_dir, "data_import.R"))
  source(here::here("R", "core_utils.R"))
  source(here::here("R", "sim_diagnostics.R"))
  source(here::here("vig_hallquist_2026", "vh_study3.R"))
}
source_helpers()
```

# Study 3: Random Effects on Both Sides of the Equation

## Load Data
```{r}
# Load data
study3 <- read_study_cache("study3", this_run = "vig_hallquist_slurm") %>%
  add_study3_analysis_eligibility() %>%
  mutate(
      converged = point_eligible,
      bias = estimate - truth,
      rel_bias = case_when(
          truth == 0 ~ NA_real_,
          TRUE ~ (estimate - truth) / truth
      ),
      sq_error = (estimate - truth)^2,
      covered = if_else(interval_eligible, ci_low <= truth & ci_high >= truth, NA),
      num_clus_fac = factor(num_clus, levels = c(30, 50, 100, 150, 300))
  )
```

```{r}
study3_method_labels <- c(
  "oracle_dual" = "Oracle with Both",
  "naive_blup_on_blup" = "BLUP on BLUP",
  "closed_form_on_blup" = "MLE on BLUP",
  "blup_on_closed_form" = "BLUP on MLE",
  "closed_form_on_closed_form" = "MLE on MLE",
  "fuller_closed_form" = "Fuller (stabilized)",
  "fuller_book_preliminary_closed_form" = "Fuller (book preliminary)",
  "fuller_book_closed_form" = "Fuller (book equations)",
  "fuller_alpha_stepdown_closed_form" = "Fuller (alpha stepdown)",
  "lai_2spa" = "Lai (2-SPA)",
  "sem" = "Latent Curve Model"
)
```

## Hypotheses
* H1. Naive BLUP-on-BLUP regressions should be unreliable because outcome-side shrinkage, predictor-side measurement error, and heterogeneous reliability act together.
* H2. Correcting only one side should help only for the component of the problem it targets. For example, unshrinking the outcome addresses outcome attenuation but does not fix predictor measurement error.
* H3. Full two-sided measurement correction, either through Lai 2S-PA or a closed-form-score Fuller moment correction, should best recover point estimates and coverage.
* H4. The method differences should be largest when one first-stage process has many observations and the other has few, because this creates asymmetric score reliability.

## Bias

```{r}
# first look at slope-intercept correlation and effect size
bias_cor_es <- study3 %>%
  filter(converged & !stage1_singular_problem) %>%
  filter(standardized_beta_target > 0) %>%
  ggplot(aes(x = standardized_beta_target, y = rel_bias, fill = method)) +
  facet_nested("Slope-Intercept Correlation" + marginal_rho ~ "Number of Clusters" + num_clus_fac) +
  stat_summary(fun = mean, geom = "col", position = "dodge") +
  scale_fill_discrete(labels = study3_method_labels) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Effect Size", y = "Relative Bias", fill = "Method",
       title = "Study 3: Relative Bias (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
# then look at reliabilities
bias_rel <- study3 %>%
  filter(converged & !stage1_singular_problem) %>%
  filter(standardized_beta_target > 0) %>%
  ggplot(aes(x = num_clus_fac, y = rel_bias, fill = method)) +
  facet_nested("Predictor Reliability" + target_reliability_q ~ "Outcome Reliability" + target_reliability_y) +
  stat_summary(fun = mean, geom = "col", position = "dodge") +
  scale_fill_discrete(labels = study3_method_labels) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Number of Clusters", y = "Relative Bias", fill = "Method",
       title = "Study 3: Relative Bias (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

print(bias_cor_es); print(bias_rel)
```
