---
  title: "Structural Random Effects"
  author: "Zach Vig & Michael Hallquist"
  date: "`r format(Sys.time(), '%d %B, %Y')`"
  output:
    html_document:
      toc: true
      toc_depth: 2
      toc_float: true
      number_sections: true
    editor_options: 
      chunk_output_type: inline 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
pacman::p_load(tidyverse, ggh4x, knitr, kableExtra, effectsize)
source_helpers <- function() {
  analysis_dir <- here::here("vig_hallquist_2026", "analysis")
  source(file.path(analysis_dir, "data_import.R"))
  source(here::here("R", "core_utils.R"))
  source(here::here("R", "sim_diagnostics.R"))
  source(here::here("vig_hallquist_2026", "vh_study2.R"))
}
source_helpers()
```

# Study 1: Random-Effect-as-Outcome

```{r}
# Load data
study1 <- read_study_cache("study1", this_run = "vig_hallquist_slurm_7_23_26") %>%
  add_vh_analysis_eligibility() %>%
  mutate(
      status10_failure = !is.na(status_code) & status_code == 10L,
      converged = point_eligible,
      bias = estimate - truth,
      rel_bias = case_when(
          truth == 0 ~ NA_real_,
          TRUE ~ (estimate - truth) / truth
      ),
      sq_error = (estimate - truth)^2,
      covered = if_else(interval_eligible, ci_low <= truth & ci_high >= truth, NA),
      num_clus_fac = factor(num_clus, levels = c(30, 50, 100, 150, 300)),
      is_two_step = !(method %in% c("direct_mlm"))
  )
```

```{r}  
study1_method_labels <- c(
  "fuller_closed_form" = "Fuller (stabilized)",
  "fuller_book_preliminary_closed_form" = "Fuller (book preliminary)",
  "fuller_book_closed_form" = "Fuller (book equations)",
  "lai_2spa" = "Lai (2-SPA)",
  "naive_blup" = "OLS w/ BLUP",
  "fuller_alpha_stepdown_closed_form" = "Fuller (alpha stepdown)",
  "closed_form" = "OLS w/ CF",
  "single_subject_ols" = "OLS w/ SS-OLS",
  "direct_mlm" = "Direct MLM",
  "oracle" = "Oracle"
)
study1_point_est_methods <- c(
  "fuller_closed_form",
  "fuller_book_preliminary_closed_form",
  "fuller_book_closed_form",
  "lai_2spa",
  "naive_blup",
  "direct_mlm",
  "oracle"
)
```


 ## Hypotheses
* H1. Naive BLUP-as-outcome regressions should be attenuated, especially when cluster size is small, slope reliability is low, and shrinkage is strong.
* H2. Closed-form scores should reduce or remove point-estimate bias under correct first-stage specification.
* H3. The direct MLM should be the best benchmark and should agree with the oracle target asymptotically.

## Bias


### Plots

```{r out.extra="style='max-width: 100%; height: auto;'"}
bias_null_effect <- study1 %>%
  filter(converged) %>%
  filter(standardized_beta_target == 0) %>%
  filter(method %in% study1_point_est_methods) %>%
  ggplot(aes(x = target_reliability, y = bias, fill = method)) +
  facet_nested("Slope-Intercept Correlation" + marginal_rho ~ "Number of Clusters" + as.factor(num_clus)) +
  stat_summary(data = ~ dplyr::filter(.x, method != "oracle"), fun = mean, geom = "col", position = "dodge") +
  stat_summary(data = ~ dplyr::filter(.x, method == "oracle"), fun = mean, geom = "point", color = "black", size = 1) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_discrete(labels = study1_method_labels) +
  labs(x = "Slope Reliability", y = "Bias", fill = "Method",
       title = "Study 1: Bias for Null Effect (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
print(bias_null_effect)
```

For the null effect size, all methods show virtually zero bias. Slope-intercept correlation of zero shows slightly lower bias than non-zero ones.

```{r out.extra="style='max-width: 100%; height: auto;'"}
bias_num_clus <- study1 %>%
  filter(converged) %>%
  filter(standardized_beta_target > 0) %>%
  filter(method %in% study1_point_est_methods) %>%
  ggplot(aes(x = target_reliability, y = rel_bias, fill = method)) +
  facet_nested("Slope-Intercept Correlation" + marginal_rho ~ "Number of Clusters" + as.factor(num_clus)) +
  stat_summary(data = ~ dplyr::filter(.x, method != "oracle"), fun = mean, geom = "col", position = "dodge") +
  stat_summary(data = ~ dplyr::filter(.x, method == "oracle"), fun = mean, geom = "point", color = "black", size = 1) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_discrete(labels = study1_method_labels) +
  labs(x = "Slope Reliability", y = "Relative Bias", fill = "Method",
       title = "Study 1: Relative Bias (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
print(bias_num_clus)
```

Bias is virtually equivalent across number of clusters per slope reliability and slope-intercept correlation. This makes sense given that no matter the number of clusters, the slope estimator is unbiased.

```{r}
# Bias doesn't change much over number of clusters, so average over it.
# Effect size == 0 shows no effect so filter that out.
# Naive BLUP bias overtakes everything else so need a second plot
bias_w_blup <- study1 %>%
  filter(converged) %>%
  filter(standardized_beta_target > 0) %>%
  filter(method %in% study1_point_est_methods) %>%
  ggplot(aes(x = target_reliability, y = rel_bias, fill = method)) +
  facet_nested("Effect Size" + standardized_beta_target ~ "Slope-Intercept Correlation" + marginal_rho) +
  stat_summary(data = ~ dplyr::filter(.x, method != "oracle"), fun = mean, geom = "col", position = "dodge") +
  stat_summary(data = ~ dplyr::filter(.x, method == "oracle"), fun = mean, geom = "point", color = "black", size = 1) +
  scale_fill_discrete(labels = study1_method_labels) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Slope Reliability", y = "Relative Bias", fill = "Method",
       title = "Study 1: Relative Bias (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
print(bias_w_blup)
```

The naive BLUP method evidently produces severe negative bias, i.e., slope attenuation. This verifies hypothesis one. The following plot removes naive BLUP to more clearly see the remaining methods.

```{r}
# remove naive BLUP to see the other methods more clearly
bias_wo_blup <- study1 %>%
  filter(converged) %>%
  filter(standardized_beta_target > 0) %>%
  filter(method %in% study1_point_est_methods) %>%
  filter(method != "naive_blup") %>%
  ggplot(aes(x = target_reliability, y = rel_bias, fill = method)) +
  facet_nested("Effect Size" + standardized_beta_target ~ "Slope-Intercept Correlation" + marginal_rho) +
  stat_summary(data = ~ dplyr::filter(.x, method != "oracle"), fun = mean, geom = "col", position = "dodge") +
  stat_summary(data = ~ dplyr::filter(.x, method == "oracle"), fun = mean, geom = "point", color = "black", size = 1) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_discrete(labels = study1_method_labels) +
  labs(x = "Slope Reliability", y = "Relative Bias", fill = "Method",
       title = "Study 1: Relative Bias (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
print(bias_wo_blup)
```

Among the remaining methods, 2-SPA shows the most *positive* relative bias, nearing 15\% in the low reliability, high slope-intercept correlation conditions. Conversely, direct MLM and Fuller show near-zero bias. As hypothesized, closed-form scores via Fuller remove the bias of the naive BLUP approach.

### ANOVA Model

```{r}
# super slow estimation right now
rel_bias_aov <- study1 %>%
  filter(converged & !stage1_singular_problem) %>%
  filter(standardized_beta_target > 0) %>%
  filter(method %in% study1_point_est_methods & method != "oracle" & method != "naive_blup") %>%
  mutate(
    Method = as.factor(method),
    EffectSize = as.factor(standardized_beta_target),
    Correlation = as.factor(marginal_rho),
    Reliability = as.factor(target_reliability),
    NumClus = as.factor(num_clus),
    ClusSize = as.factor(mean_clus_size)
  ) %>%
  aov(rel_bias ~ (Method + Reliability + EffectSize + Correlation + NumClus + ClusSize)^3, data = .)

#summary(rel_bias_aov)
eta_squared(rel_bias_aov, partial = TRUE)
```

The ANOVA model confirms that among the remaining methods, estimators are unbiased in all conditions (as expected).

## RMSE

```{r}
rmse_eff_size <- study1 %>%
  filter(converged) %>%
  filter(standardized_beta_target > 0) %>%
  filter(method %in% study1_point_est_methods) %>%
  ggplot(aes(x = target_reliability, y = sq_error, fill = method)) +
  facet_nested("Effect Size" + standardized_beta_target ~ "Number of Clusters" + num_clus_fac) +
  stat_summary(data = ~ dplyr::filter(.x, method != "oracle"), fun = function(x) sqrt(mean(x)), geom = "col", position = "dodge") +
  stat_summary(data = ~ dplyr::filter(.x, method == "oracle"), fun = mean, geom = "point", color = "black", size = 1) +
  scale_fill_discrete(labels = study1_method_labels) +
  labs(x = "Slope Reliability", y = "RMSE", fill = "Method",
       title = "Study 1: RMSE by Effect Size (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

print(rmse_eff_size)
```

RMSE decreases with more clusters and greater slope reliability. Lai 2-SPA shows marginally greater RMSE in low reliability-small number of cluster conditions than Fuller/direct MLM.

```{r}
rmse_corr <- study1 %>%
  filter(converged) %>%
  filter(standardized_beta_target > 0) %>%
  filter(method %in% study1_point_est_methods) %>%
  ggplot(aes(x = target_reliability, y = sq_error, fill = method)) +
  facet_nested("Slope-Intercept Correlation" + marginal_rho ~ "Number of Clusters" + num_clus_fac) +
  stat_summary(data = ~ dplyr::filter(.x, method != "oracle"), fun = function(x) sqrt(mean(x)), geom = "col", position = "dodge") +
  stat_summary(data = ~ dplyr::filter(.x, method == "oracle"), fun = function(x) sqrt(mean(x)), geom = "point", color = "black", size = 1) +
  scale_fill_discrete(labels = study1_method_labels) +
  labs(x = "Slope Reliability", y = "RMSE", fill = "Method",
       title = "Study 1: RMSE by Effect Size (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

print(rmse_corr)
```

When faceting by slope-intercept correlation, zero correlation conditions showed lower RMSE on average. Previous trends remainined the same.


## Coverage

```{r}
coverage_blup <- study1 %>%
  filter(interval_eligible) %>%
  mutate(covered = as.numeric(covered)) %>%
  filter(method == "naive_blup") %>%
  mutate(effect_label = "Effect Size", cor_label = "Slope-Intercept Correlation") %>%
  ggplot(aes(x = target_reliability, y = covered)) +
  facet_nested(cor_label + marginal_rho ~ effect_label + standardized_beta_target) +
  stat_summary(fun = mean, geom = "col", position =  "dodge", fill = "#00BA38") +  
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_hline(yintercept = 0.95, linetype = "dashed") +
  labs(x = "Slope Reliability", y = "Coverage",
       title = "Study 1: Coverage in Naive BLUPs (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

print(coverage_blup)
```

Naive BLUP showed severe undercoverage in non-zero effect size conditions. Low slope reliability conditions showed especially small coverage (around 10\%).

```{r coverage}
# naive blups led to severe undercoverage, so remove them to see the other methods more clearly
coverage <- study1 %>%
  filter(interval_eligible) %>%
  mutate(covered = as.numeric(covered)) %>%
  filter(method != "naive_blup" & method != "oracle") %>%
  mutate(num_clus_label = "Number of Clusters", cor_label = "Slope-Intercept Correlation") %>%
  ggplot(aes(x = target_reliability, y = covered, fill = method)) +
  facet_nested(cor_label + marginal_rho ~ num_clus_label + num_clus_fac) +
  stat_summary(fun = mean, geom = "col", position =  "dodge") +
  scale_fill_discrete(labels = study1_method_labels) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(ylim = c(0.9, 1)) +
  geom_hline(yintercept = 0.95, linetype = "dashed") +
  labs(x = "Slope Reliability", y = "Coverage", fill = "Method",
       title = "Study 1: Coverage (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

print(coverage)
```

Fuller (with and without alpha-stepdown) tended to show overcoverage (though this decreased with greater slope reliability), and the remaining methods showed good coverage or slight undercoverage.

## Standard Errors

```{r}
se_table <- study1 %>%
  filter(interval_eligible) %>%
  filter(method != "naive_blup" & method != "oracle") %>%
  group_by(method, num_clus_fac, target_reliability, marginal_rho, mean_clus_size, standardized_beta_target) %>%
  summarize(
    mean_se = mean(se),
    empirical_se = sd(estimate),
    se_rel_bias = (mean_se - empirical_se) / empirical_se,
    .groups = "drop_last"
  )
se <- se_table %>%
  ggplot(aes(x = target_reliability, y = se_rel_bias, fill = method)) +
  facet_nested("Slope-Intercept Correlation" + marginal_rho ~ "Number of Clusters" + num_clus_fac) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_discrete(labels = study1_method_labels) +
  labs(x = "Slope Reliability", y = "SE Bias", fill = "Method",
       title = "Study 1: SE Bias (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

print(se)
```

Fuller variants show generally positively biased standard errors, while Lai 2-SPA shows negatively biased standard errors, especially in low reliability conditions.

## Convergence

```{r}
plot_convergence <- function(df, subtitle) {
  df %>%
  ggplot(
    aes(
      x = target_reliability,
      y = converged,
      color = method,
      group = method
    )
  ) +
  stat_summary(fun = safe_mean, geom = "line", linewidth = 1) +
  stat_summary(fun = safe_mean, geom = "point", size = 2) +
  scale_y_continuous(
    labels = scales::label_percent(),
    limits = c(0, 1)
  ) +
  facet_nested(
    "Slope-Intercept Correlation" + marginal_rho ~ "Number of Clusters" + as.factor(num_clus)
  ) +
  labs(
    x = "Slope Reliability",
    y = "Convergence Rate",
    color = "Method",
    title = "Study 1: Convergence Rate",
    subtitle = subtitle
  ) +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  scale_color_discrete(labels = study1_method_labels)
}

convergence <- plot_convergence(
  study1 %>%
  filter(
    method %in% c("fuller_closed_form", "lai_2spa", "direct_mlm")
  ),
  "*Includes Stage 1 Singular Fits"
)
convergence2 <- plot_convergence(
  study1 %>%
  filter(
    method %in% c("fuller_closed_form", "lai_2spa", "direct_mlm")
  ) %>%
  filter(!stage1_singular_problem),
  "*Does NOT Include Stage 1 Singular Fits"
)
print(convergence); print(convergence2)
```

The vast majority of convergence problems were isolated to the first stage, specifically cases of singularity. Removing these cases leaves nearly 100\% convergence for all conditions. Thus, Fuller/Fuller variants do not show a noticeable convergence advantage in the random-effect-as-outcome scenario.
