---
title: "VH Study 3v2: Random Slopes as Structural Predictor and Outcome"
author: "Michael Hallquist and Zach Vig"
date: "26 August 2026"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    embed-resources: true
    html-math-method: katex
params:
  output_dir: "outputs/vig_hallquist_v2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 120, out.width = "100%"
)

suppressPackageStartupMessages({
  library(data.table)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
  library(scales)
})

input_path <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) "")
analysis_dir <- if (nzchar(input_path)) {
  dirname(normalizePath(input_path, mustWork = FALSE))
} else {
  getwd()
}
repo_root <- normalizePath(file.path(analysis_dir, "..", ".."), mustWork = TRUE)
resolve_repo_path <- function(path) {
  if (grepl("^/", path)) path else file.path(repo_root, path)
}
source(
  file.path(repo_root, "vig_hallquist_2026", "vh_designs.R"),
  local = knitr::knit_global()
)

output_dir <- resolve_repo_path(params$output_dir)
condition_dir <- file.path(output_dir, "conditions")
canonical_counts <- study_condition_counts()
canonical_offsets <- c(0L, cumsum(canonical_counts)[-length(canonical_counts)])
names(canonical_offsets) <- names(canonical_counts)
expected_ids <- canonical_offsets[["study3v2"]] +
  seq_len(canonical_counts[["study3v2"]])
summary_paths <- file.path(
  condition_dir,
  sprintf("condition_%04d_summary.csv", expected_ids)
)
replication_paths <- file.path(
  condition_dir,
  sprintf("condition_%04d_replication_results.csv.gz", expected_ids)
)
missing_summary_paths <- summary_paths[!file.exists(summary_paths)]
missing_replication_paths <- replication_paths[!file.exists(replication_paths)]
if (length(missing_summary_paths) > 0L || length(missing_replication_paths) > 0L) {
  stop(
    "Study 3v2 is incomplete: missing ", length(missing_summary_paths),
    " condition summaries and ", length(missing_replication_paths),
    " replication files."
  )
}

cell_metrics <- data.table::rbindlist(
  lapply(summary_paths, data.table::fread),
  fill = TRUE,
  use.names = TRUE
) %>%
  tibble::as_tibble() %>%
  filter(.data$study == "study3v2") %>%
  mutate(
    sd_standardized_error = suppressWarnings(
      as.numeric(.data$sd_standardized_error)
    )
  )

expected_method_names <- c(
  "oracle_dual", "naive_blup_on_blup", "closed_form_on_blup",
  "blup_on_closed_form", "closed_form_on_closed_form",
  "fuller_closed_form", "fuller_book_preliminary_closed_form",
  "fuller_book_closed_form", "fuller_alpha_stepdown_closed_form",
  "lai_2spa", "sem"
)
expected_method_cells <- tidyr::crossing(
  condition_id = expected_ids,
  method = expected_method_names
)
observed_method_cells <- cell_metrics %>%
  count(.data$condition_id, .data$method, name = "summary_rows")
missing_method_cells <- dplyr::anti_join(
  expected_method_cells,
  observed_method_cells,
  by = c("condition_id", "method")
)
unexpected_method_cells <- dplyr::anti_join(
  observed_method_cells,
  expected_method_cells,
  by = c("condition_id", "method")
)
v2_contract_ok <-
  n_distinct(cell_metrics$condition_id) == length(expected_ids) &&
  nrow(missing_method_cells) == 0L &&
  nrow(unexpected_method_cells) == 0L &&
  all(observed_method_cells$summary_rows == 1L) &&
  all(cell_metrics$n_rep == 1000L) &&
  all(cell_metrics$study_version == "v2") &&
  all(cell_metrics$calibration_version == "shape_preserving_v2") &&
  all(cell_metrics$covariance_shape_fixed %in% TRUE) &&
  all(cell_metrics$slope_intercept_variance_ratio == 1)
if (!v2_contract_ok) {
  stop(
    "Refusing to render: Study 3v2 summaries violate the v2-only ",
    "completion, method, replication, or calibration contract."
  )
}
summary_file_info <- file.info(summary_paths)
replication_file_info <- file.info(replication_paths)
summary_artifact_oldest <- min(summary_file_info$mtime)
summary_artifact_newest <- max(summary_file_info$mtime)
replication_artifact_oldest <- min(replication_file_info$mtime)
replication_artifact_newest <- max(replication_file_info$mtime)

method_labels <- c(
  oracle_dual = "Oracle",
  naive_blup_on_blup = "BLUP outcome / BLUP predictors",
  closed_form_on_blup = "Unshrunk outcome / BLUP predictors",
  blup_on_closed_form = "BLUP outcome / unshrunk predictors",
  closed_form_on_closed_form = "Unshrunk outcome / unshrunk predictors",
  fuller_closed_form = "Fuller (stabilized)",
  fuller_book_preliminary_closed_form = "Fuller (book preliminary)",
  fuller_book_closed_form = "Fuller (book equations)",
  fuller_alpha_stepdown_closed_form = "Fuller (alpha-stepdown)",
  lai_2spa = "Lai 2S-PA",
  sem = "MSEM"
)
stopifnot(setequal(names(method_labels), expected_method_names))

primary_methods <- c("oracle_dual", "fuller_closed_form", "lai_2spa", "sem")
comparison_methods <- c("fuller_closed_form", "lai_2spa", "sem")
negative_control_methods <- c(
  "naive_blup_on_blup", "closed_form_on_blup",
  "blup_on_closed_form", "closed_form_on_closed_form"
)
fuller_methods <- c(
  "fuller_closed_form", "fuller_alpha_stepdown_closed_form",
  "fuller_book_preliminary_closed_form", "fuller_book_closed_form"
)

cell_metrics <- cell_metrics %>%
  mutate(
    method_label = factor(
      unname(method_labels[.data$method]),
      levels = unname(method_labels)
    ),
    effect_label = factor(
      paste0("beta = ", .data$standardized_beta_target),
      levels = paste0("beta = ", c(0, .2, .5))
    ),
    predictor_reliability_label = factor(
      paste0("Predictor R = ", sprintf("%.2f", .data$target_reliability_y)),
      levels = paste0("Predictor R = ", sprintf("%.2f", c(.25, .8)))
    ),
    outcome_reliability_label = factor(
      paste0("Outcome R = ", sprintf("%.2f", .data$target_reliability_q)),
      levels = paste0("Outcome R = ", sprintf("%.2f", c(.25, .8)))
    ),
    reliability_pair = factor(
      paste0(
        "Predictor R = ", sprintf("%.2f", .data$target_reliability_y),
        "; outcome R = ", sprintf("%.2f", .data$target_reliability_q)
      ),
      levels = c(
        "Predictor R = 0.25; outcome R = 0.25",
        "Predictor R = 0.25; outcome R = 0.80",
        "Predictor R = 0.80; outcome R = 0.25",
        "Predictor R = 0.80; outcome R = 0.80"
      )
    ),
    n_label = factor(
      paste0("N = ", .data$num_clus),
      levels = paste0("N = ", c(50, 100, 150, 300))
    )
  )

pooled_mc_se <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) == 0L) return(NA_real_)
  sqrt(sum(x^2)) / length(x)
}
fmt_pct <- function(x, accuracy = .1) percent(x, accuracy = accuracy)
fmt_num <- function(x, digits = 3) formatC(x, digits = digits, format = "f")
theme_vh <- function() {
  theme_bw(base_size = 10) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      strip.background = element_rect(fill = "grey95", colour = "grey75")
    )
}
method_palette <- c(
  "Oracle" = "#767676",
  "Fuller (stabilized)" = "#E15759",
  "Lai 2S-PA" = "#4E79A7",
  "MSEM" = "#59A14F"
)

performance_grid <- cell_metrics %>%
  filter(
    .data$method %in% primary_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$effect_label,
    .data$standardized_beta_target, .data$num_clus, .data$n_label,
    .data$target_reliability_y, .data$target_reliability_q,
    .data$reliability_pair
  ) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    bias_mc_se = pooled_mc_se(.data$mc_se_bias),
    bias_mc_low = .data$bias - qnorm(.975) * .data$bias_mc_se,
    bias_mc_high = .data$bias + qnorm(.975) * .data$bias_mc_se,
    coverage = mean(.data$coverage, na.rm = TRUE),
    coverage_mc_se = pooled_mc_se(.data$coverage_mc_se),
    coverage_mc_low = pmax(0, .data$coverage - qnorm(.975) * .data$coverage_mc_se),
    coverage_mc_high = pmin(1, .data$coverage + qnorm(.975) * .data$coverage_mc_se),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

null_grid <- cell_metrics %>%
  filter(
    .data$method %in% primary_methods,
    .data$standardized_beta_target == 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$num_clus, .data$n_label,
    .data$target_reliability_y, .data$target_reliability_q,
    .data$reliability_pair
  ) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    type1_error = mean(.data$type1_error, na.rm = TRUE),
    type1_mc_se = pooled_mc_se(.data$type1_error_mc_se),
    type1_mc_low = pmax(0, .data$type1_error - qnorm(.975) * .data$type1_mc_se),
    type1_mc_high = pmin(1, .data$type1_error + qnorm(.975) * .data$type1_mc_se),
    operational_type1_error = mean(
      .data$operational_type1_error,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

method_summary <- cell_metrics %>%
  filter(
    .data$method %in% primary_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(.data$method, .data$method_label) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    mean_bias = mean(.data$bias, na.rm = TRUE),
    absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    coverage = mean(.data$coverage, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    cells_below_nominal = sum(.data$coverage_mc_high < .95, na.rm = TRUE),
    cells_above_nominal = sum(.data$coverage_mc_low > .95, na.rm = TRUE),
    .groups = "drop"
  )

null_summary <- cell_metrics %>%
  filter(
    .data$method %in% primary_methods,
    .data$standardized_beta_target == 0
  ) %>%
  group_by(.data$method, .data$method_label) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    type1_error = mean(.data$type1_error, na.rm = TRUE),
    operational_type1_error = mean(
      .data$operational_type1_error,
      na.rm = TRUE
    ),
    coverage = mean(.data$coverage, na.rm = TRUE),
    .groups = "drop"
  )

reliability_summary <- cell_metrics %>%
  filter(
    .data$method %in% primary_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$target_reliability_y,
    .data$target_reliability_q, .data$reliability_pair
  ) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    mean_bias = mean(.data$bias, na.rm = TRUE),
    mean_absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    conditional_coverage = mean(.data$coverage, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

reliability_null_summary <- cell_metrics %>%
  filter(
    .data$method %in% primary_methods,
    .data$standardized_beta_target == 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$target_reliability_y,
    .data$target_reliability_q, .data$reliability_pair
  ) %>%
  summarise(
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    conditional_type1 = mean(.data$type1_error, na.rm = TRUE),
    operational_type1 = mean(
      .data$operational_type1_error,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

metric_for <- function(summary, method, metric) {
  value <- summary[summary$method == method, metric, drop = TRUE]
  if (length(value) != 1L) NA_real_ else value
}

reliability_metric <- function(method_key, predictor_r, outcome_r, metric) {
  value <- reliability_summary %>%
    filter(
      .data$method == .env$method_key,
      .data$target_reliability_y == .env$predictor_r,
      .data$target_reliability_q == .env$outcome_r
    ) %>%
    pull(all_of(metric))
  if (length(value) != 1L) NA_real_ else value
}

calibration_profiles <- bind_rows(
  cell_metrics %>%
    filter(.data$method == "oracle_dual") %>%
    transmute(
      process = "Predictor process (y)",
      target = .data$target_reliability_y,
      mean_cluster_size = .data$mean_clus_size_y,
      marginal_rho = .data$marginal_rho,
      population_reliability = .data$achieved_reliability_y,
      fitted_reliability =
        .data$mean_stage1_y_fitted_marginal_slope_posterior_reliability_mean
    ),
  cell_metrics %>%
    filter(.data$method == "oracle_dual") %>%
    transmute(
      process = "Outcome process (q)",
      target = .data$target_reliability_q,
      mean_cluster_size = .data$mean_clus_size_q,
      marginal_rho = .data$marginal_rho,
      population_reliability = .data$achieved_reliability_q,
      fitted_reliability =
        .data$mean_stage1_q_fitted_marginal_slope_posterior_reliability_mean
    )
) %>%
  group_by(
    .data$process, .data$target, .data$mean_cluster_size,
    .data$marginal_rho
  ) %>%
  summarise(
    population_reliability = mean(.data$population_reliability, na.rm = TRUE),
    fitted_reliability = mean(.data$fitted_reliability, na.rm = TRUE),
    .groups = "drop"
  )

# MSEM eligibility audit. The compact reason strings retain exact replication
# counts (for example, "mplus_critical_warning=594"). Parse them rather than
# inferring failure counts from rounded cell-level rates.
msem_cells <- cell_metrics %>%
  filter(.data$method == "sem")

parse_reason_counts <- function(reason_string, condition_id) {
  if (length(reason_string) == 0L || is.na(reason_string) ||
      !nzchar(reason_string)) {
    return(tibble::tibble(
      condition_id = integer(), reason = character(), count = integer()
    ))
  }
  entries <- strsplit(reason_string, "; ", fixed = TRUE)[[1]]
  tibble::tibble(
    condition_id = as.integer(condition_id),
    reason = sub("=.*$", "", entries),
    count = as.integer(sub("^.*=", "", entries))
  )
}

msem_reason_counts <- bind_rows(lapply(seq_len(nrow(msem_cells)), function(i) {
  parse_reason_counts(
    msem_cells$interval_exclusion_reasons[[i]],
    msem_cells$condition_id[[i]]
  )
}))

msem_reason_by_cell <- msem_reason_counts %>%
  group_by(.data$condition_id) %>%
  summarise(
    critical_warning_n = sum(
      .data$count[.data$reason == "mplus_critical_warning"]
    ),
    unavailable_n = sum(
      .data$count[.data$reason == "estimation_unavailable"]
    ),
    .groups = "drop"
  )

msem_disposition <- msem_cells %>%
  select(
    .data$condition_id, .data$reliability_pair,
    .data$standardized_beta_target, .data$n_rep,
    .data$n_interval_eligible, .data$success_and_cover
  ) %>%
  left_join(msem_reason_by_cell, by = "condition_id") %>%
  mutate(
    critical_warning_n = coalesce(.data$critical_warning_n, 0L),
    unavailable_n = coalesce(.data$unavailable_n, 0L),
    success_and_cover_n = pmin(
      .data$n_interval_eligible,
      as.integer(round(.data$success_and_cover * .data$n_rep))
    ),
    eligible_noncover_n =
      .data$n_interval_eligible - .data$success_and_cover_n
  ) %>%
  select(
    .data$condition_id, .data$reliability_pair,
    .data$standardized_beta_target,
    .data$success_and_cover_n, .data$eligible_noncover_n,
    .data$critical_warning_n, .data$unavailable_n
  ) %>%
  pivot_longer(
    cols = ends_with("_n"),
    names_to = "disposition",
    values_to = "count"
  ) %>%
  mutate(
    disposition = factor(
      .data$disposition,
      levels = c(
        "success_and_cover_n", "eligible_noncover_n",
        "critical_warning_n", "unavailable_n"
      ),
      labels = c(
        "Eligible and covers", "Eligible but misses",
        "Excluded: critical Mplus warning",
        "Excluded: estimation unavailable"
      )
    )
  )

msem_disposition_summary <- msem_disposition %>%
  group_by(
    .data$reliability_pair, .data$standardized_beta_target,
    .data$disposition
  ) %>%
  summarise(count = sum(.data$count), .groups = "drop") %>%
  group_by(.data$reliability_pair, .data$standardized_beta_target) %>%
  mutate(proportion = .data$count / sum(.data$count)) %>%
  ungroup()

msem_disposition_totals <- msem_disposition %>%
  group_by(.data$disposition) %>%
  summarise(count = sum(.data$count), .groups = "drop") %>%
  mutate(proportion = .data$count / sum(.data$count))

msem_total_replications <- sum(msem_cells$n_rep)
msem_total_eligible <- sum(msem_cells$n_interval_eligible)
msem_total_critical <- sum(
  msem_reason_counts$count[
    msem_reason_counts$reason == "mplus_critical_warning"
  ]
)
msem_total_unavailable <- sum(
  msem_reason_counts$count[
    msem_reason_counts$reason == "estimation_unavailable"
  ]
)

msem_by_n_effect_rho <- msem_cells %>%
  group_by(
    .data$reliability_pair, .data$num_clus,
    .data$standardized_beta_target, .data$marginal_rho
  ) %>%
  summarise(
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    critical_warning_rate = mean(
      .data$mean_mplus_critical_warning, na.rm = TRUE
    ),
    outcome_covariance_boundary_rate = mean(
      .data$mean_mplus_outcome_latent_covariance_boundary,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

msem_boundary_correlation <- cor(
  1 - msem_cells$interval_eligibility,
  msem_cells$mean_mplus_outcome_latent_covariance_boundary,
  use = "complete.obs"
)
msem_warning_correlation <- cor(
  1 - msem_cells$interval_eligibility,
  msem_cells$mean_mplus_critical_warning,
  use = "complete.obs"
)
msem_point_interval_difference <- max(
  abs(msem_cells$point_eligibility - msem_cells$interval_eligibility),
  na.rm = TRUE
)

msem_eligibility_for <- function(
    predictor_r, outcome_r, n_clusters = NULL,
    beta = NULL, rho = NULL) {
  keep <-
    msem_cells$target_reliability_y == predictor_r &
    msem_cells$target_reliability_q == outcome_r
  if (!is.null(n_clusters)) keep <- keep & msem_cells$num_clus == n_clusters
  if (!is.null(beta)) {
    keep <- keep & msem_cells$standardized_beta_target == beta
  }
  if (!is.null(rho)) keep <- keep & msem_cells$marginal_rho == rho
  mean(msem_cells$interval_eligibility[keep], na.rm = TRUE)
}

# A single prespecified low-low sentinel is read at replication level to
# identify the warning and boundary pattern hidden by aggregate eligibility
# rates. Resolve it from design columns so a future condition-ID offset change
# cannot silently redirect the audit.
msem_sentinel_condition <- msem_cells %>%
  filter(
    .data$num_clus == 50L,
    .data$mean_clus_size_y == 3L,
    .data$mean_clus_size_q == 3L,
    .data$target_reliability_y == .25,
    .data$target_reliability_q == .25,
    .data$marginal_rho == 0,
    .data$standardized_beta_target == 0
  ) %>%
  pull(.data$condition_id)
if (length(msem_sentinel_condition) != 1L) {
  stop("The Study 3v2 MSEM sentinel did not resolve to exactly one condition.")
}
msem_sentinel_path <- replication_paths[expected_ids == msem_sentinel_condition]
msem_sentinel <- data.table::fread(
  msem_sentinel_path,
  select = c(
    "method", "status_code", "interval_eligible",
    "interval_exclusion_reason", "mplus_critical_warning",
    "mplus_critical_warning_detail", "mplus_min_reported_variance",
    "mplus_nonpositive_variance_count",
    "mplus_predictor_latent_covariance_boundary",
    "mplus_outcome_latent_covariance_boundary",
    "mplus_predictor_latent_covariance_min_eigenvalue",
    "mplus_outcome_latent_covariance_min_eigenvalue"
  )
) %>%
  tibble::as_tibble() %>%
  filter(.data$method == "sem")

msem_sentinel_disposition <- msem_sentinel %>%
  mutate(
    disposition = case_when(
      .data$interval_eligible ~ "Eligible interval",
      .data$interval_exclusion_reason == "mplus_critical_warning" ~
        "Critical Mplus warning",
      .data$interval_exclusion_reason == "estimation_unavailable" ~
        "Estimation unavailable",
      TRUE ~ "Other exclusion"
    )
  ) %>%
  count(.data$disposition, name = "replications")

msem_sentinel_count <- function(disposition_label) {
  value <- msem_sentinel_disposition$replications[
    msem_sentinel_disposition$disposition == disposition_label
  ]
  if (length(value) == 0L) 0L else value[[1L]]
}

msem_sentinel_warning_targets <- msem_sentinel %>%
  filter(.data$mplus_critical_warning %in% TRUE) %>%
  mutate(
    warning_target = case_when(
      grepl("VARIABLE S2", .data$mplus_critical_warning_detail) ~
        "Outcome slope (S2)",
      grepl("VARIABLE I2", .data$mplus_critical_warning_detail) ~
        "Outcome intercept (I2)",
      grepl("VARIABLE S1", .data$mplus_critical_warning_detail) ~
        "Predictor slope (S1)",
      grepl("VARIABLE I1", .data$mplus_critical_warning_detail) ~
        "Predictor intercept (I1)",
      TRUE ~ "Other"
    )
  ) %>%
  count(.data$warning_target, name = "replications")

msem_warning_target_count <- function(target_labels) {
  sum(msem_sentinel_warning_targets$replications[
    msem_sentinel_warning_targets$warning_target %in% target_labels
  ])
}

msem_low_low_summary <- msem_cells %>%
  filter(
    .data$target_reliability_y == .25,
    .data$target_reliability_q == .25
  ) %>%
  group_by(.data$standardized_beta_target) %>%
  summarise(
    conditional_coverage = mean(.data$coverage, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    .groups = "drop"
  )

msem_low_low_metric <- function(beta, metric) {
  value <- msem_low_low_summary %>%
    filter(.data$standardized_beta_target == .env$beta) %>%
    pull(all_of(metric))
  if (length(value) != 1L) NA_real_ else value
}

msem_sentinel_boundary <- msem_sentinel %>%
  mutate(
    eligibility = if_else(
      .data$interval_eligible, "Eligible", "Ineligible"
    )
  ) %>%
  group_by(.data$eligibility) %>%
  summarise(
    Replications = n(),
    `Mean minimum variance` = mean(
      .data$mplus_min_reported_variance, na.rm = TRUE
    ),
    `Mean nonpositive-variance count` = mean(
      .data$mplus_nonpositive_variance_count, na.rm = TRUE
    ),
    `Predictor covariance boundary` = mean(
      .data$mplus_predictor_latent_covariance_boundary, na.rm = TRUE
    ),
    `Outcome covariance boundary` = mean(
      .data$mplus_outcome_latent_covariance_boundary, na.rm = TRUE
    ),
    `Predictor minimum eigenvalue` = mean(
      .data$mplus_predictor_latent_covariance_min_eigenvalue, na.rm = TRUE
    ),
    `Outcome minimum eigenvalue` = mean(
      .data$mplus_outcome_latent_covariance_min_eigenvalue, na.rm = TRUE
    ),
    .groups = "drop"
  )
```

# Executive summary

Study 3v2 is the most demanding of the primary v2 studies because estimated
random slopes appear on both sides of the structural equation. The predictor
process is denoted $y$ in the simulation code and the outcome process is
denoted $q$. The amended DGM uses the shared shape-preserving calibration for
both processes: posterior reliability changes while the intercept variance,
slope variance, and covariance retain a common shape
($\tau_1^2/\tau_0^2=1$).

All 864 planned conditions are complete, with all 11 methods and 1,000
replications per condition. This report reads only Study 3v2 condition files
and stops if any legacy or incomplete output is encountered. The original
Study 3 Rmd is retained unchanged as the archive of the v1 analysis.

The principal findings are:

1. Stabilized Fuller has complete interval eligibility, small average bias
   (`r fmt_num(metric_for(method_summary, "fuller_closed_form", "mean_bias"))`),
   and `r fmt_pct(metric_for(method_summary, "fuller_closed_form", "coverage"))`
   non-null conditional coverage. Its performance is close to nominal when
   predictor reliability is .80, but conservative when predictor reliability
   is .25.
2. Low predictor reliability, more than low outcome reliability, defines the
   difficult regime. At predictor reliability .25, all three feasible latent-
   uncertainty methods have inflated typical SEs, conservative null tests, and
   method-specific bias or fit problems. When predictor reliability is .80,
   stabilized Fuller is close to centered and nominal even if outcome
   reliability remains .25.
3. Lai 2S-PA conditionally covers well but has positive bias and incomplete
   interval eligibility. Its non-null eligibility is
   `r fmt_pct(metric_for(method_summary, "lai_2spa", "eligibility"))`, so its
   `r fmt_pct(metric_for(method_summary, "lai_2spa", "coverage"))` conditional
   coverage corresponds to only
   `r fmt_pct(metric_for(method_summary, "lai_2spa", "success_and_cover"))`
   probability of both producing an eligible interval and covering.
4. MSEM also has serious availability costs in this two-process design:
   non-null interval eligibility is
   `r fmt_pct(metric_for(method_summary, "sem", "eligibility"))`. Its
   conditional intervals are conservative on average, particularly when the
   predictor process has reliability .25.
5. Fuller's low-reliability protection is not exact calibration. Its average
   null Type I error is
   `r fmt_pct(metric_for(null_summary, "fuller_closed_form", "type1_error"))`;
   at predictor reliability .25 it is near 1%, whereas it is near 5% at
   predictor reliability .80. This is substantive overcoverage.
6. The one-sided and naive score regressions confirm why a two-sided correction
   is needed. Some are attenuated; regressions that retain low-reliability BLUP
   predictors also develop catastrophic coefficient tails. Neither unshrinking
   the outcome alone nor unshrinking both sides without correcting predictor
   measurement error is a general solution.

# Scientific motivation

Study 3 represents research questions that relate two person-specific dynamic
features. Examples include asking whether an individual's affective-reactivity
slope predicts that same individual's symptom-reactivity slope, or whether a
learning-rate parameter predicts a later change parameter. Both scientific
variables are latent random slopes recovered from repeated measurements, so
uncertainty enters on both sides of the structural equation.

A compact representation is

$$
\begin{aligned}
y_{it} &= \gamma_{0y}+\gamma_{1y}x_{it}+y_{0i}+y_{1i}x_{it}+e_{y,it},\\
q_{is} &= \gamma_{0q}+\gamma_{1q}x_{is}+q_{0i}+q_{1i}x_{is}+e_{q,is},\\
q_{1i} &= \theta_0 y_{0i}+\theta_1 y_{1i}+\zeta_i.
\end{aligned}
$$

The focal estimand is the standardized $\theta_1$. This construction makes the
predictor/outcome asymmetry explicit: noise in $y_{1i}$ contaminates a
predictor moment and can attenuate or destabilize the coefficient, whereas
noise in $q_{1i}$ primarily increases outcome-side uncertainty unless its
shrinkage is left uncorrected.

# Design and v2-only completion contract

The design crosses four cluster counts (50, 100, 150, and 300), three mean
cluster sizes for each process (3, 5, and 10), two posterior-reliability targets
for each process (.25 and .80), two intercept--slope correlations (0 and .50),
and three standardized structural effects (0, .20, and .50). This yields 864
conditions. The null cells support direct Type I error evaluation; non-null
cells evaluate bias, coverage, SE calibration, and power-related operational
performance.

The methods have distinct roles:

- **Oracle** uses the simulated true $y$ intercept and slope and true $q$
  slope.
- Four **score-regression negative controls** cross BLUP versus unshrunk
  predictor scores with BLUP versus unshrunk outcome scores. They identify
  which failures arise from outcome shrinkage, predictor shrinkage, or
  untreated predictor measurement error.
- **Fuller (stabilized)** corrects the predictor moments and propagates both
  processes' Stage-1 uncertainty; its algebra variants audit SE-tail
  stabilization.
- **Lai 2S-PA** fits a two-process latent-variable path model in OpenMx using
  the Stage-1 scores and error matrices.
- **MSEM** jointly estimates both growth processes and the structural path in
  Mplus. It avoids generated scores, but asks each sample to estimate a larger
  latent covariance system.

```{r completion-table}
tibble::tibble(
  item = c(
    "Expected conditions", "Observed conditions", "Methods per condition",
    "Replications per condition", "Study version", "Calibration version",
    "Covariance shape fixed", "Slope/intercept variance ratio",
    "Summary artifacts read", "Replication artifacts read",
    "Oldest summary artifact (local time)",
    "Newest summary artifact (local time)",
    "Newest replication artifact (local time)"
  ),
  value = c(
    length(expected_ids), n_distinct(cell_metrics$condition_id),
    length(expected_method_names), unique(cell_metrics$n_rep),
    unique(cell_metrics$study_version),
    unique(cell_metrics$calibration_version),
    all(cell_metrics$covariance_shape_fixed),
    unique(cell_metrics$slope_intercept_variance_ratio),
    length(summary_paths), length(replication_paths),
    format(summary_artifact_oldest, tz = "America/New_York"),
    format(summary_artifact_newest, tz = "America/New_York"),
    format(replication_artifact_newest, tz = "America/New_York")
  )
) %>%
  kable(align = c("l", "r"))
```

Condition summaries drive the main figures and tables. The exact matching
replication artifacts drive the MSEM disposition and sentinel diagnostics;
the report stops if either artifact set is incomplete.

The focal structural coefficient is the standardized effect of the true $y$
slope on the true $q$ slope, conditional on the $y$ intercept. Thus, in the
labels below, **predictor reliability** is `target_reliability_y` and
**outcome reliability** is `target_reliability_q`. Keeping this direction clear
is important: the two reliabilities do not play symmetric statistical roles.

{{< include _vh_v2_metric_guide.qmd >}}

# Reliability manipulation checks

```{r calibration-plot, fig.width=11, fig.height=7}
calibration_profiles %>%
  ggplot(aes(
    x = .data$target,
    y = .data$fitted_reliability,
    colour = factor(.data$mean_cluster_size),
    group = factor(.data$mean_cluster_size)
  )) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey45") +
  geom_line(alpha = .55) +
  geom_point(size = 2) +
  facet_grid(.data$process ~ paste0("rho = ", .data$marginal_rho)) +
  scale_x_continuous(breaks = c(.25, .8), limits = c(.18, .87)) +
  scale_y_continuous(
    breaks = c(.25, .5, .8), limits = c(.18, .87),
    labels = percent_format(accuracy = 1)
  ) +
  labs(
    x = "Target population posterior slope reliability",
    y = "Mean fitted posterior slope reliability",
    colour = "Mean observations\nper cluster",
    title = "Both Stage-1 processes reproduce the intended reliability contrast",
    subtitle = "Finite-sample fitted reliabilities need not equal the population target exactly"
  ) +
  theme_vh()
```

The population calibrator attains the target by construction. The fitted
reliabilities summarize the Stage-1 covariance estimates and posterior
covariances obtained within the Monte Carlo samples. Their modest deviations
from the diagonal are finite-sample estimation effects, not a return to the v1
one-coordinate calibration. Crucially, both processes use the same covariance
shape and calibration rule.

# Bias

```{r primary-bias-plot, fig.width=13, fig.height=9.5}
performance_grid %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$bias,
    ymin = .data$bias_mc_low,
    ymax = .data$bias_mc_high,
    colour = .data$method_label,
    group = .data$method_label
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey40") +
  geom_errorbar(width = 6, alpha = .25, linewidth = .35) +
  geom_line(alpha = .65) +
  geom_point(size = 1.7) +
  facet_grid(.data$effect_label ~ .data$reliability_pair) +
  scale_x_continuous(breaks = c(50, 100, 150, 300)) +
  scale_colour_manual(values = method_palette) +
  labs(
    x = "Number of clusters",
    y = "Mean bias with 95% Monte Carlo CI",
    colour = NULL,
    title = "Low predictor reliability separates the feasible methods",
    subtitle = "Points average cluster-size and rho cells within each design combination"
  ) +
  theme_vh()
```

Stabilized Fuller is essentially centered whenever predictor reliability is
.80. At low predictor reliability, its bias is modestly negative when both
processes have low reliability and close to zero on average when only the
predictor is low, although individual cells remain variable. Lai 2S-PA shows
substantial positive bias in both low-predictor-reliability panels. MSEM is
negatively biased in the low--low panel but much closer to centered when
outcome reliability is .80.

The asymmetry is substantively useful. Low outcome reliability alone mainly
increases sampling uncertainty. Low predictor reliability contaminates the
structural moment matrix and is therefore the condition in which measurement-
error correction, numerical stability, and interval construction are most
challenging.

## One-sided and naive score regressions

```{r negative-control-summary, include=FALSE}
negative_control_summary <- cell_metrics %>%
  filter(
    .data$method %in% negative_control_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$effect_label,
    .data$target_reliability_y, .data$target_reliability_q,
    .data$reliability_pair
  ) %>%
  summarise(
    median_cell_bias = median(.data$bias, na.rm = TRUE),
    mean_absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    max_absolute_cell_bias = max(abs(.data$bias), na.rm = TRUE),
    interval_eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    conditional_coverage = mean(.data$coverage, na.rm = TRUE),
    fraction_cells_abs_bias_over_half = mean(abs(.data$bias) > .5, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r negative-control-bias-plot, fig.width=12, fig.height=8}
negative_control_summary %>%
  ggplot(aes(
    x = .data$reliability_pair,
    y = .data$median_cell_bias,
    colour = .data$method_label,
    group = .data$method_label
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey40") +
  geom_line(alpha = .55) +
  geom_point(size = 2) +
  facet_wrap(~ effect_label, nrow = 1) +
  scale_y_continuous(
    trans = pseudo_log_trans(sigma = .02, base = 10),
    breaks = c(-.5, -.2, -.1, 0, .1, .5, 10, 1e3, 1e5),
    labels = label_number(accuracy = .01, scale_cut = cut_short_scale())
  ) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = NULL,
    y = "Median condition-level mean bias (signed pseudo-log scale)",
    colour = NULL,
    title = "Partial score corrections fail for different reasons",
    subtitle = "The pseudo-log axis retains both attenuation and catastrophic low-predictor-reliability tails"
  ) +
  theme_vh() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

The low-predictor-reliability BLUP regressions have enormous condition-level
mean biases because a small set of replication-level coefficient explosions
dominate the means. This is why the plot uses a signed pseudo-log scale rather
than silently deleting or clipping those cells: the transformation compresses
large positive and negative magnitudes while retaining zero and sign.
Unshrinking the predictor
scores eliminates the catastrophic tail but does not eliminate attenuation
when the noisy outcome score is used. Unshrinking both sides improves the
problem but still treats corrected predictor scores as measured without error,
leaving appreciable negative bias and undercoverage.

# Coverage and standard-error calibration

```{r primary-coverage-plot, fig.width=13, fig.height=9.5}
performance_grid %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$coverage,
    ymin = .data$coverage_mc_low,
    ymax = .data$coverage_mc_high,
    colour = .data$method_label,
    group = .data$method_label
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey35") +
  geom_errorbar(width = 6, alpha = .25, linewidth = .35) +
  geom_line(alpha = .65) +
  geom_point(size = 1.7) +
  facet_grid(.data$effect_label ~ .data$reliability_pair) +
  scale_x_continuous(breaks = c(50, 100, 150, 300)) +
  scale_y_continuous(limits = c(.70, 1), labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = method_palette) +
  labs(
    x = "Number of clusters",
    y = "Conditional 95% CI coverage",
    colour = NULL,
    title = "Conditional coverage is high in the low-predictor-reliability cells",
    subtitle = "High coverage there reflects wide intervals and excludes ineligible 2S-PA and MSEM fits"
  ) +
  theme_vh()
```

The broad v1 low-reliability undercoverage pattern does not recur. However,
high conditional coverage is not automatically evidence of ideal inference.
With predictor reliability .25, Fuller's typical SE is about 45%--61% larger
than the empirical SD, depending on outcome reliability. Lai 2S-PA and MSEM
are at least as conservative among eligible fits. When predictor reliability
is .80, all four methods have typical SE ratios near one and coverage much
closer to 95%.

```{r se-calibration-plot, fig.width=13, fig.height=8.5}
performance_grid %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$se_ratio,
    colour = .data$method_label,
    group = .data$method_label
  )) +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey35") +
  geom_line(alpha = .65) +
  geom_point(size = 1.7) +
  facet_grid(.data$effect_label ~ .data$reliability_pair, scales = "free_y") +
  scale_x_continuous(breaks = c(50, 100, 150, 300)) +
  scale_colour_manual(values = method_palette) +
  labs(
    x = "Number of clusters",
    y = "Median condition-level SE ratio",
    colour = NULL,
    title = "Predictor reliability governs the typical SE inflation",
    subtitle = paste(
      "Each ratio is mean reported SE / empirical sampling SD;",
      "lines connect cluster counts within a method"
    )
  ) +
  theme_vh()
```

# Conditional versus operational performance

```{r coverage-eligibility-plot, fig.width=12, fig.height=7.5}
reliability_summary %>%
  filter(.data$method %in% comparison_methods) %>%
  ggplot(aes(
    x = .data$eligibility,
    y = .data$conditional_coverage,
    colour = .data$method_label,
    size = .data$mean_absolute_bias
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey50") +
  geom_vline(xintercept = .95, linetype = 3, colour = "grey65") +
  geom_point(alpha = .9) +
  facet_wrap(~ reliability_pair, nrow = 1) +
  scale_x_continuous(limits = c(.25, 1.01), labels = percent_format(accuracy = 1)) +
  scale_y_continuous(limits = c(.90, 1), labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = method_palette) +
  scale_size_continuous(range = c(2.5, 8)) +
  labs(
    x = "Interval eligibility",
    y = "Conditional 95% CI coverage",
    colour = NULL,
    size = "Mean |bias|",
    title = "Similar conditional coverage can conceal very different usability",
    subtitle = "Fuller returns an interval in every replication; 2S-PA and MSEM often do not"
  ) +
  theme_vh()
```

At low reliability for both processes, stabilized Fuller has
`r fmt_pct(reliability_metric("fuller_closed_form", .25, .25, "conditional_coverage"))`
conditional coverage and complete eligibility. Lai 2S-PA has
`r fmt_pct(reliability_metric("lai_2spa", .25, .25, "conditional_coverage"))`
coverage among eligible intervals, but only
`r fmt_pct(reliability_metric("lai_2spa", .25, .25, "eligibility"))`
eligibility. MSEM eligibility falls to
`r fmt_pct(reliability_metric("sem", .25, .25, "eligibility"))`.
The probability of obtaining an interval that covers is consequently only
`r fmt_pct(reliability_metric("lai_2spa", .25, .25, "success_and_cover"))`
for 2S-PA and
`r fmt_pct(reliability_metric("sem", .25, .25, "success_and_cover"))`
for MSEM, compared with
`r fmt_pct(reliability_metric("fuller_closed_form", .25, .25, "success_and_cover"))`
for Fuller.

This operational comparison should not be interpreted as granting failed fits
zero coverage in a conventional conditional-coverage estimand. Rather, it
answers the practical question a user faces: what proportion of repeated
samples yield a usable interval that contains the truth? Reporting both
quantities prevents a method from appearing well calibrated merely because
difficult replications are excluded.

# Why MSEM loses eligibility and success-and-cover

MSEM's poor operational performance is primarily a **model-solution problem**,
not an interval-construction problem. Under the prespecified eligibility
contract, an MSEM point estimate is excluded when Mplus does not return an
estimable result, when estimation has a nonzero status, when the focal path is
not unique, or when Mplus reports a critical warning such as a non-positive-
definite latent covariance matrix. An interval then additionally requires a
finite positive SE and ordered confidence limits. In Study 3v2, point and
interval eligibility are identical in every MSEM cell (maximum absolute
difference = `r fmt_num(msem_point_interval_difference, 3)`). Thus, no material
attrition occurs because an otherwise admissible point estimate merely lacks
a usable SE or confidence limit.

Across all `r comma(msem_total_replications)` MSEM replications,
`r comma(msem_total_eligible)`
(`r fmt_pct(msem_total_eligible / msem_total_replications)`) produce an
eligible interval. The two exact exclusion categories are critical Mplus
warnings in `r comma(msem_total_critical)` replications
(`r fmt_pct(msem_total_critical / msem_total_replications)`) and estimation
being unavailable in `r comma(msem_total_unavailable)` replications
(`r fmt_pct(msem_total_unavailable / msem_total_replications)`). The first
category often contains a numerical solution and focal coefficient, but the
solution is not accepted because the fitted latent covariance structure is
inadmissible. Treating those coefficients as successful estimates would make
MSEM look more available only by relaxing the definition of a valid fitted
model.

```{r msem-success-disposition, fig.width=12.5, fig.height=8}
ggplot(
  msem_disposition_summary,
  aes(
    x = .data$reliability_pair,
    y = .data$proportion,
    fill = .data$disposition
  )
) +
  geom_col(width = .78) +
  geom_text(
    aes(label = if_else(
      .data$proportion >= .035,
      percent(.data$proportion, accuracy = 1),
      ""
    )),
    position = position_stack(vjust = .5),
    colour = "white",
    size = 3
  ) +
  facet_wrap(
    ~ factor(
      .data$standardized_beta_target,
      levels = c(0, .2, .5),
      labels = c("beta = 0", "beta = .2", "beta = .5")
    ),
    ncol = 1
  ) +
  scale_y_continuous(
    labels = percent_format(accuracy = 1),
    expand = expansion(mult = c(0, .01))
  ) +
  scale_fill_manual(values = c(
    "Eligible and covers" = "#59A14F",
    "Eligible but misses" = "#F28E2B",
    "Excluded: critical Mplus warning" = "#E15759",
    "Excluded: estimation unavailable" = "#9D9D9D"
  )) +
  labs(
    x = NULL,
    y = "Proportion of all replications",
    fill = NULL,
    title = "MSEM success-and-cover is lost mainly before coverage is evaluated",
    subtitle = paste(
      "Critical Mplus warnings and unavailable estimates account for",
      "nearly all of the shortfall"
    )
  ) +
  theme_vh() +
  theme(axis.text.x = element_text(angle = 18, hjust = 1))
```

The narrow orange portions are the replications that yield an eligible
interval but miss the true value. In the low-low reliability cells,
conditional coverage among eligible fits is very high: it averages
`r fmt_pct(msem_low_low_metric(0, "conditional_coverage"))`
for $\beta=0$,
`r fmt_pct(msem_low_low_metric(.2, "conditional_coverage"))`
for $\beta=.2$, and
`r fmt_pct(msem_low_low_metric(.5, "conditional_coverage"))`
for $\beta=.5$. The corresponding success-and-cover rates are only
`r fmt_pct(msem_low_low_metric(0, "success_and_cover"))`,
`r fmt_pct(msem_low_low_metric(.2, "success_and_cover"))`, and
`r fmt_pct(msem_low_low_metric(.5, "success_and_cover"))`, because most samples
never reach the conditional coverage calculation. This is the key reason that
MSEM can display reassuring conditional coverage and poor success-and-cover at
the same time.

## What is going wrong in the fitted MSEM?

The MSEM jointly estimates two intercept--slope growth processes and regresses
the outcome slope (`S2`) on the predictor intercept and slope (`I1` and `S1`).
This asks each sample to support two latent covariance blocks plus the
structural relation. When the slopes are weakly measured, the level-2 sample
contains too little information to keep all of those variance and covariance
estimates well inside the admissible parameter space.

A replication-level audit of condition `r msem_sentinel_condition` provides a
concrete example. This is a deliberately difficult null condition with 50
clusters, three observations per process, zero intercept--slope correlation,
and posterior reliability .25 for both slopes. Of its
`r comma(nrow(msem_sentinel))` MSEM fits, only
`r comma(msem_sentinel_count("Eligible interval"))` produce eligible intervals;
`r comma(msem_sentinel_count("Critical Mplus warning"))` end with a critical
warning despite a zero estimation status, and
`r comma(msem_sentinel_count("Estimation unavailable"))` do not return an
estimate. Every critical warning
states that the latent covariance matrix (`PSI`) is not positive definite.
The variable named in the warning is:

```{r msem-sentinel-warning-table}
msem_sentinel_warning_targets %>%
  arrange(desc(.data$replications)) %>%
  transmute(
    `Latent variable named by Mplus` = .data$warning_target,
    Replications = .data$replications,
    `Percent of 1,000 samples` = percent(
      .data$replications / 1000, accuracy = .1
    )
  ) %>%
  kable(align = c("l", "r", "r"))
```

The outcome block is especially prominent: `S2` or `I2` is named in
`r comma(msem_warning_target_count(c("Outcome slope (S2)", "Outcome intercept (I2)")))`
of the
`r comma(msem_sentinel_count("Critical Mplus warning"))` critical warnings.
The fitted covariance diagnostics tell the same story.
Among the eligible fits, neither latent covariance block is flagged as being
on a boundary and the average minimum eigenvalues are positive. Among the
ineligible fits for which diagnostics can be parsed, nonpositive variance
estimates and an inadmissible outcome covariance block are common:

```{r msem-sentinel-boundary-table}
msem_sentinel_boundary %>%
  mutate(
    across(
      c(
        `Mean minimum variance`, `Mean nonpositive-variance count`,
        `Predictor covariance boundary`, `Outcome covariance boundary`,
        `Predictor minimum eigenvalue`, `Outcome minimum eigenvalue`
      ),
      ~ round(.x, 3)
    )
  ) %>%
  kable(align = c("l", rep("r", 7)))
```

This sentinel is not an isolated anomaly. Across the 864 MSEM cells, the
cell-level correlation between interval ineligibility and the rate of a
critical Mplus warning is `r fmt_num(msem_warning_correlation, 3)`; the
correlation between ineligibility and an outcome latent-covariance boundary is
`r fmt_num(msem_boundary_correlation, 3)`. Boundary diagnostics are only
available when enough Mplus output can be parsed, so they do not themselves
count the fully unavailable fits. The exact exclusion counts above remain the
primary attrition analysis.

```{r msem-boundary-association, fig.width=12, fig.height=7}
msem_cells %>%
  ggplot(aes(
    x = .data$mean_mplus_outcome_latent_covariance_boundary,
    y = 1 - .data$interval_eligibility,
    colour = .data$reliability_pair
  )) +
  geom_point(alpha = .28, size = 1.6) +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, linewidth = .8) +
  facet_wrap(~ effect_label, nrow = 1) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = c(
    "Predictor R = 0.25; outcome R = 0.25" = "#B07AA1",
    "Predictor R = 0.25; outcome R = 0.80" = "#E15759",
    "Predictor R = 0.80; outcome R = 0.25" = "#4E79A7",
    "Predictor R = 0.80; outcome R = 0.80" = "#59A14F"
  )) +
  labs(
    x = "Rate of an outcome latent-covariance boundary",
    y = "Interval ineligibility",
    colour = NULL,
    title = "MSEM attrition tracks inadmissible outcome covariance estimates",
    subtitle = "Each point is one simulation condition; lines summarize the within-effect association"
  ) +
  theme_vh()
```

## Which design features amplify the problem?

The number of level-2 units is the clearest protection against MSEM
ineligibility. Averaged over the other factors, eligibility in the low-low
reliability regime rises from
`r fmt_pct(msem_eligibility_for(.25, .25, n_clusters = 50))` with 50 clusters
to `r fmt_pct(msem_eligibility_for(.25, .25, n_clusters = 300))` with 300
clusters. In the high-high regime it rises from
`r fmt_pct(msem_eligibility_for(.80, .80, n_clusters = 50))` to
`r fmt_pct(msem_eligibility_for(.80, .80, n_clusters = 300))`.

```{r msem-eligibility-design-factors, fig.width=12.5, fig.height=8}
msem_by_n_effect_rho %>%
  mutate(
    beta_label = factor(
      .data$standardized_beta_target,
      levels = c(0, .2, .5),
      labels = c("beta = 0", "beta = .2", "beta = .5")
    ),
    rho_label = factor(
      .data$marginal_rho,
      levels = c(0, .5),
      labels = c("rho = 0", "rho = .5")
    )
  ) %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$eligibility,
    colour = .data$beta_label,
    linetype = .data$rho_label,
    group = interaction(.data$beta_label, .data$rho_label)
  )) +
  geom_line(linewidth = .75) +
  geom_point(size = 1.8) +
  facet_wrap(~ reliability_pair, nrow = 2) +
  scale_x_continuous(breaks = c(50, 100, 150, 300)) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = percent_format(accuracy = 1)
  ) +
  scale_colour_manual(values = c(
    "beta = 0" = "#4E79A7",
    "beta = .2" = "#F28E2B",
    "beta = .5" = "#E15759"
  )) +
  labs(
    x = "Number of clusters",
    y = "MSEM interval eligibility",
    colour = NULL,
    linetype = NULL,
    title = "More clusters help, but weakly measured slopes remain difficult",
    subtitle = "Lines average over the crossed within-cluster observation counts"
  ) +
  theme_vh()
```

Three secondary patterns are informative. First, low outcome reliability is
directly associated with instability in the `I2`--`S2` covariance block.
Second, low predictor reliability also matters because `S1` is an uncertain
latent predictor in the structural regression. Third, the $ρ=.5$ covariance
shape and the larger structural effect both reduce eligibility, especially at
low reliability. In this parameterization, `S2` is endogenous, so the variance
reported for it by Mplus is its residual variance conditional on `I1` and
`S1`, not its total marginal variance. As the structural path strengthens,
that residual component can lie closer to a boundary in finite samples. This
helps explain why eligibility at low-low reliability falls from
`r fmt_pct(msem_eligibility_for(.25, .25, beta = 0))` under the null to
`r fmt_pct(msem_eligibility_for(.25, .25, beta = .5))` at the largest effect.

The within-cluster observation count has much less independent leverage here
than the number of clusters. That result is consistent with the v2 DGM: the
measurement-error scale is recalibrated as the observation count changes so
that posterior slope reliability remains at its target. Consequently, adding
observations while holding the reliability target fixed is not the same as
giving MSEM a better-measured slope.

In substantive terms, MSEM's low success-and-cover is therefore not evidence
that its eligible intervals systematically miss the truth. It is evidence
that, under weak level-2 information, a relatively elaborate joint latent
model often cannot produce an admissible covariance solution. Conditional
coverage describes the selected subset that survives this process;
success-and-cover records the operational consequence for all samples. Both
are needed, and neither should be substituted for the other.

# Type I error

At the null, CI noncoverage and two-sided rejection are the same event among
eligible intervals. Type I error is therefore the clearest interpretation of
null-condition coverage.

```{r type1-plot, fig.width=12.5, fig.height=7.5}
null_grid %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$type1_error,
    ymin = .data$type1_mc_low,
    ymax = .data$type1_mc_high,
    colour = .data$method_label,
    group = .data$method_label
  )) +
  geom_hline(yintercept = .05, linetype = 2, colour = "grey35") +
  geom_errorbar(width = 6, alpha = .25, linewidth = .35) +
  geom_line(alpha = .65) +
  geom_point(size = 1.7) +
  facet_wrap(~ reliability_pair, nrow = 1) +
  scale_x_continuous(breaks = c(50, 100, 150, 300)) +
  scale_y_continuous(limits = c(0, .10), labels = percent_format(accuracy = .1)) +
  scale_colour_manual(values = method_palette) +
  labs(
    x = "Number of clusters",
    y = "Conditional Type I error",
    colour = NULL,
    title = "Low predictor reliability produces marked overcoverage",
    subtitle = "The same methods are approximately nominal when predictor reliability is .80"
  ) +
  theme_vh()
```

This plot resolves an ambiguity that coverage alone cannot. The very high
low-predictor-reliability coverage is inferential conservatism, not merely
better centering. Fuller's null rejection rate is approximately 1% when both
reliabilities are .25 and below 1% when predictor reliability is .25 and
outcome reliability is .80. It returns to approximately 5% in both panels with
predictor reliability .80. Lai 2S-PA and MSEM show the same broad transition,
though their conditional rates apply only to eligible fits.

# Direct comparison: stabilized Fuller, Lai 2S-PA, and MSEM

The three methods address the same two-sided uncertainty problem differently.
Fuller applies an explicit moment correction using the replication-specific
Stage-1 error-covariance matrices; 2S-PA represents the Stage-1 scores and
uncertainties in an OpenMx latent-variable model; MSEM estimates both growth
processes and the structural relation jointly in Mplus. The matched-cell
comparisons below use the same simulated samples and do not rank methods from
coverage alone.

```{r paired-comparison-setup, include=FALSE}
comparison_wide <- cell_metrics %>%
  filter(
    .data$method %in% comparison_methods,
    .data$standardized_beta_target > 0
  ) %>%
  select(
    .data$condition_id, .data$target_reliability_y,
    .data$target_reliability_q, .data$reliability_pair,
    .data$standardized_beta_target, .data$method, .data$bias,
    .data$coverage, .data$interval_eligibility,
    .data$success_and_cover, .data$mean_se_to_interval_empirical_sd
  ) %>%
  pivot_wider(
    names_from = .data$method,
    values_from = c(
      bias, coverage, interval_eligibility, success_and_cover,
      mean_se_to_interval_empirical_sd
    )
  )

make_paired_comparison <- function(wide, comparator_key, comparator_label) {
  tibble::tibble(
    comparator = comparator_label,
    target_reliability_y = wide$target_reliability_y,
    reliability_pair = wide$reliability_pair,
    absolute_bias_difference =
      abs(wide$bias_fuller_closed_form) -
      abs(wide[[paste0("bias_", comparator_key)]]),
    absolute_coverage_error_difference =
      abs(wide$coverage_fuller_closed_form - .95) -
      abs(wide[[paste0("coverage_", comparator_key)]] - .95),
    interval_failure_difference =
      (1 - wide$interval_eligibility_fuller_closed_form) -
      (1 - wide[[paste0("interval_eligibility_", comparator_key)]]),
    absolute_log_se_error_difference =
      abs(log(wide$mean_se_to_interval_empirical_sd_fuller_closed_form)) -
      abs(log(wide[[paste0(
        "mean_se_to_interval_empirical_sd_", comparator_key
      )]]))
  )
}

paired_comparisons <- bind_rows(
  make_paired_comparison(comparison_wide, "lai_2spa", "Lai 2S-PA"),
  make_paired_comparison(comparison_wide, "sem", "MSEM")
) %>%
  pivot_longer(
    cols = c(
      .data$absolute_bias_difference,
      .data$absolute_coverage_error_difference,
      .data$interval_failure_difference,
      .data$absolute_log_se_error_difference
    ),
    names_to = "metric",
    values_to = "Fuller_minus_comparator"
  ) %>%
  mutate(
    metric = factor(
      .data$metric,
      levels = c(
        "absolute_bias_difference", "absolute_coverage_error_difference",
        "interval_failure_difference", "absolute_log_se_error_difference"
      ),
      labels = c(
        "Absolute bias", "Absolute conditional-coverage error",
        "Interval-failure rate", "Absolute log(SE / empirical SD)"
      )
    )
  )
```

```{r paired-comparison-plot, fig.width=12, fig.height=9}
paired_comparisons %>%
  ggplot(aes(
    x = factor(.data$target_reliability_y),
    y = .data$Fuller_minus_comparator,
    fill = .data$comparator
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey35") +
  geom_boxplot(outlier.alpha = .18, width = .65) +
  facet_grid(.data$metric ~ .data$comparator, scales = "free_y") +
  scale_fill_manual(values = c("Lai 2S-PA" = "#4E79A7", "MSEM" = "#59A14F")) +
  labs(
    x = "Predictor posterior reliability",
    y = "Fuller minus comparator (negative favors Fuller)",
    fill = NULL,
    title = "Fuller's clearest advantages are bias and interval availability",
    subtitle = "Conditional-coverage and SE-calibration advantages are less uniform in the hardest cells"
  ) +
  theme_vh() +
  theme(legend.position = "none")
```

**Why Fuller remains attractive.** It returns finite intervals in all
replications, has much smaller bias than 2S-PA when predictor reliability is
low, and avoids MSEM's severe low-reliability fit attrition. When predictor
reliability is .80, its point and interval performance are close to oracle
performance even if the outcome process is measured poorly.

**Where 2S-PA is competitive.** Among eligible solutions, 2S-PA often has
coverage as close to or closer to 95% than Fuller. At high predictor
reliability its bias is small. Its liabilities are positive bias and
substantial selection of eligible fits when predictor reliability is .25.
Conditional coverage therefore answers a narrower question for 2S-PA than for
Fuller.

**Where MSEM is competitive.** In the high--high reliability condition MSEM is
very similar to Fuller and 2S-PA. Joint estimation is also theoretically
appealing because it avoids treating estimated scores as observed. In this
simulation, however, MSEM has the lowest eligibility in the difficult cells,
and the eligible low-reliability fits remain conservative. Its average null
Type I error is not closer to 5% than Fuller's in Study 3v2, unlike the clearer
MSEM advantage observed in some one-process settings.

```{r numerical-comparison-table}
reliability_summary %>%
  filter(.data$method %in% comparison_methods) %>%
  left_join(
    reliability_null_summary %>%
      filter(.data$method %in% comparison_methods),
    by = c(
      "method", "method_label", "target_reliability_y",
      "target_reliability_q", "reliability_pair"
    )
  ) %>%
  transmute(
    Method = as.character(.data$method_label),
    `Predictor R` = .data$target_reliability_y,
    `Outcome R` = .data$target_reliability_q,
    `Mean bias` = round(.data$mean_bias, 3),
    `Mean |bias|` = round(.data$mean_absolute_bias, 3),
    `Conditional coverage` = percent(
      .data$conditional_coverage, accuracy = .1
    ),
    `Interval eligible` = percent(.data$eligibility.x, accuracy = .1),
    `Success-and-cover` = percent(.data$success_and_cover, accuracy = .1),
    `Median condition-level SE ratio` = round(.data$median_se_ratio, 3),
    `Null Type I error` = percent(.data$conditional_type1, accuracy = .1),
    `Null interval eligible` = percent(.data$eligibility.y, accuracy = .1)
  ) %>%
  arrange(.data$`Predictor R`, .data$`Outcome R`, .data$Method) %>%
  kable(align = c("l", rep("r", 10)))
```

# Fuller variants and SE tails

The corrected Fuller implementation retains several algebra/regularization
variants as diagnostics. They share the corrected zero predictor--outcome
measurement-error covariance for this DGM. The variants differ in the
preliminary moment and variance-bread choices used to regularize the sampling-
variance tail.

```{r fuller-variant-plot, fig.width=11.5, fig.height=7.5}
cell_metrics %>%
  filter(
    .data$method %in% fuller_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$reliability_pair,
    .data$target_reliability_y, .data$target_reliability_q
  ) %>%
  summarise(
    conditional_coverage = mean(.data$coverage, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    p90_cell_se_ratio = quantile(
      .data$mean_se_to_interval_empirical_sd,
      .90, na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = .data$median_se_ratio,
    y = .data$conditional_coverage,
    colour = .data$method_label,
    size = .data$p90_cell_se_ratio
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey50") +
  geom_vline(xintercept = 1, linetype = 2, colour = "grey50") +
  geom_point(alpha = .85) +
  facet_wrap(~ reliability_pair, nrow = 1) +
  scale_x_log10() +
  scale_size_continuous(range = c(2.5, 9), trans = "log10") +
  scale_colour_brewer(palette = "Set1") +
  labs(
    x = "Median condition-level SE ratio (log scale)",
    y = "Conditional coverage",
    colour = NULL,
    size = "90th percentile\ncell SE/SD",
    title = "The Fuller book variants expose the SE-tail tradeoff",
    subtitle = "Stabilized Fuller and alpha-stepdown avoid the enormous variance tails of the equation-faithful paths"
  ) +
  theme_vh()
```

The variants have similar point bias because they use closely related
coefficient corrections. Their interval behavior differs sharply. The book-
equation path has a much larger typical SE ratio and an extreme upper tail;
the book-preliminary path is less extreme but remains heavy-tailed. Stabilized
Fuller and alpha-stepdown keep the typical SE scale near one and control the
tail far more effectively. These results support the stabilized path as the
proposed estimator while retaining the other variants as transparency and
sensitivity checks.

# Conclusions

Study 3v2 changes the interpretation of the earlier two-process results in
several important ways.

- The shape-preserving reliability calibration removes the apparent general
  low-reliability coverage failure seen under the v1 one-coordinate DGM. The
  remaining difficulty is structured rather than global: it is concentrated
  in low **predictor** reliability.
- Stabilized Fuller offers the strongest overall compromise for the proposed
  use case. It has complete availability, low average bias, and approximately
  nominal performance when predictor reliability is adequate. It also avoids
  the catastrophic coefficient tails of naive or partially corrected score
  regressions.
- Fuller is not fully calibrated in the hardest cells. With predictor
  reliability .25, its SEs are too large relative to empirical sampling
  variability and its null rejection rate is far below 5%. The paper should
  describe this as conservative inference, not as unqualified 95% coverage
  success.
- Lai 2S-PA and MSEM do not provide a simple remedy. Both can have attractive
  conditional coverage, but both lose many eligible intervals in the same
  cells where Fuller's intervals become conservative. 2S-PA also has marked
  positive bias at low predictor reliability. MSEM's attrition occurs before
  interval construction: most excluded fits either have a non-positive-
  definite latent covariance matrix or do not return an estimate. It is not
  principally a case of valid point estimates acquiring unusable SEs.
- Outcome-side reliability still matters, but primarily by changing uncertainty
  and fit difficulty. A poorly measured outcome does not cause the same
  attenuation mechanism as a poorly measured predictor. This asymmetry should
  be made explicit when explaining Study 3 to a psychological methods audience.
- The null cells are indispensable. They show that high coverage for Fuller,
  2S-PA, and MSEM at low predictor reliability corresponds to reduced Type I
  error, whereas the same methods are near nominal when predictor reliability
  is .80. Coverage, Type I error, bias, fit eligibility, and SE calibration
  should therefore be reported together.

For the manuscript, the balanced conclusion is that stabilized Fuller is an
attractive and unusually available estimator for two-sided random-slope
relations, especially once predictor reliability is moderate or high. Its
practical advantage in the hardest cells is that it continues to return usable
results with much less bias than 2S-PA, not that its intervals are exactly
nominal there. A reasonable applied recommendation is to report the Stage-1
posterior reliability of the predictor slopes, treat very low values as a
warning for conservative Fuller intervals and unstable latent-variable fits,
and accompany the structural result with the reliability and fit diagnostics
used in this report.

# Appendix

## Overall method summaries

```{r method-summary-table}
method_summary %>%
  transmute(
    Method = as.character(.data$method_label),
    Cells = .data$cells,
    `Interval eligible` = percent(.data$eligibility, accuracy = .1),
    `Mean bias` = round(.data$mean_bias, 3),
    `Mean |bias|` = round(.data$absolute_bias, 3),
    `Conditional coverage` = percent(.data$coverage, accuracy = .1),
    `Success-and-cover` = percent(.data$success_and_cover, accuracy = .1),
    `Median condition-level SE ratio` = round(.data$median_se_ratio, 3),
    `MC interval below .95` = .data$cells_below_nominal,
    `MC interval above .95` = .data$cells_above_nominal
  ) %>%
  kable(
    align = c("l", rep("r", 9)),
    caption = paste(
      "Cell counts require the condition-level 95% Monte Carlo interval",
      "to lie wholly below or above .95."
    )
  )
```

## Null summaries

```{r null-summary-table}
null_summary %>%
  transmute(
    Method = as.character(.data$method_label),
    Cells = .data$cells,
    `Interval eligible` = percent(.data$eligibility, accuracy = .1),
    `Mean bias` = round(.data$bias, 3),
    `Conditional Type I error` = percent(.data$type1_error, accuracy = .1),
    `Operational Type I error` = percent(
      .data$operational_type1_error,
      accuracy = .1
    ),
    `Conditional coverage` = percent(.data$coverage, accuracy = .1)
  ) %>%
  kable(align = c("l", rep("r", 6)))
```

## Negative-control tail summary

```{r negative-control-table}
cell_metrics %>%
  filter(
    .data$method %in% negative_control_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(.data$method_label, .data$target_reliability_y) %>%
  summarise(
    `Median cell bias` = median(.data$bias, na.rm = TRUE),
    `Mean |cell bias|` = mean(abs(.data$bias), na.rm = TRUE),
    `Maximum |cell bias|` = max(abs(.data$bias), na.rm = TRUE),
    `Cells with |bias| > .5` = mean(abs(.data$bias) > .5, na.rm = TRUE),
    `Interval eligible` = mean(.data$interval_eligibility, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  transmute(
    Method = as.character(.data$method_label),
    `Predictor R` = .data$target_reliability_y,
    `Median cell bias` = round(.data$`Median cell bias`, 3),
    `Mean |cell bias|` = formatC(
      .data$`Mean |cell bias|`, digits = 3, format = "g"
    ),
    `Maximum |cell bias|` = formatC(
      .data$`Maximum |cell bias|`, digits = 3, format = "g"
    ),
    `Cells with |bias| > .5` = percent(
      .data$`Cells with |bias| > .5`, accuracy = .1
    ),
    `Interval eligible` = percent(.data$`Interval eligible`, accuracy = .1)
  ) %>%
  kable(align = c("l", rep("r", 6)))
```

{{< include _vh_v2_provenance.qmd >}}

## Session information

```{r session-information}
sessionInfo()
```
