---
title: "VH Study 5: Matched Posterior-Reliability Calibration Bridge"
author: "Simulation analysis"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
params:
  output_dir: "outputs/vig_hallquist_study5"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 120
)
suppressPackageStartupMessages({
  library(dplyr)
  library(tidyr)
  library(readr)
  library(ggplot2)
  library(knitr)
  library(scales)
})

input_path <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) "")
analysis_dir <- if (nzchar(input_path)) dirname(normalizePath(input_path)) else getwd()
repo_root <- normalizePath(file.path(analysis_dir, "..", ".."), mustWork = TRUE)
resolve_repo_path <- function(path) {
  if (grepl("^/", path)) path else file.path(repo_root, path)
}

source(file.path(
  repo_root,
  "vig_hallquist_2026",
  "random_effects_structural_simulation.R"
), local = knitr::knit_global())

design <- select_design("5")
output_dir <- resolve_repo_path(params$output_dir)
replication_candidates <- file.path(
  output_dir,
  c(
    "vig_hallquist_5_full_selection_replication_results.csv.gz",
    "vig_hallquist_study5_full_selection_replication_results.csv.gz"
  )
)
replication_path <- if (any(file.exists(replication_candidates))) {
  replication_candidates[which(file.exists(replication_candidates))[[1L]]]
} else {
  replication_candidates[[1L]]
}
summary_path <- file.path(
  output_dir,
  "vig_hallquist_5_full_selection_summary.csv"
)
results_available <- file.exists(replication_path)

arm_labels <- c(
  current_g22 = "Legacy solve-G_22 (= τ₁²)",
  shape_preserving_marginal = "Fixed G shape: marginal R",
  shape_preserving_partial = "Fixed G shape: partial R"
)
method_labels <- c(
  oracle_dual = "Oracle dual",
  naive_dual_blup = "Naive dual BLUP",
  closed_form_dual = "Uncorrected closed-form OLS",
  fuller_closed_form = "Fuller",
  fuller_alpha_stepdown_closed_form = "Fuller alpha-stepdown",
  lai_2spa = "Lai 2S-PA",
  msem = "MSEM"
)
confirmatory_methods <- c(
  "oracle_dual", "naive_dual_blup", "fuller_closed_form", "lai_2spa", "msem"
)
corrected_methods <- c("fuller_closed_form", "lai_2spa", "msem")
bias_plot_methods <- c("oracle_dual", corrected_methods)
method_palette <- c(
  "Oracle dual" = "#4D4D4D",
  "Naive dual BLUP" = "#E15759",
  "Fuller" = "#4E79A7",
  "Lai 2S-PA" = "#B07AA1",
  "MSEM" = "#59A14F"
)
arm_palette <- c(
  "Legacy solve-G_22 (= τ₁²)" = "#E15759",
  "Fixed G shape: marginal R" = "#4E79A7",
  "Fixed G shape: partial R" = "#B6992D"
)

if (results_available) {
  replications <- data.table::fread(replication_path) %>%
    tibble::as_tibble() %>%
    mutate(
      analysis_eligible = if ("analysis_eligible" %in% names(.)) {
        as.logical(analysis_eligible)
      } else {
        !is.na(estimate)
      },
      covered = ci_low <= truth & ci_high >= truth,
      reject_positive = analysis_eligible & ci_low > 0,
      reject_negative = analysis_eligible & ci_high < 0,
      reject_two_sided = reject_positive | reject_negative,
      calibration_arm_label = factor(
        unname(arm_labels[calibration_arm]),
        levels = unname(arm_labels)
      ),
      method_label = factor(
        unname(method_labels[method]),
        levels = unname(method_labels)
      )
    )

  performance <- replications %>%
    group_by(
      condition_id, calibration_arm, calibration_arm_label,
      mean_clus_size, marginal_rho, standardized_beta_target,
      method, method_label
    ) %>%
    summarise(
      n_rep = n(),
      n_eligible = sum(analysis_eligible, na.rm = TRUE),
      convergence = mean(analysis_eligible, na.rm = TRUE),
      mean_estimate = mean(estimate[analysis_eligible], na.rm = TRUE),
      bias = mean((estimate - truth)[analysis_eligible], na.rm = TRUE),
      empirical_sd = sd(estimate[analysis_eligible], na.rm = TRUE),
      mean_se = mean(se[analysis_eligible], na.rm = TRUE),
      median_se = median(se[analysis_eligible], na.rm = TRUE),
      mean_se_ratio = mean_se / empirical_sd,
      typical_se_ratio = median_se / empirical_sd,
      bias_mc_se = empirical_sd / sqrt(n_eligible),
      bias_mc_low = bias - qnorm(.975) * bias_mc_se,
      bias_mc_high = bias + qnorm(.975) * bias_mc_se,
      coverage = mean(covered[analysis_eligible], na.rm = TRUE),
      success_and_cover = sum(analysis_eligible & covered, na.rm = TRUE) / n_rep,
      mc_low = qbinom(.025, n_eligible, .95) / n_eligible,
      mc_high = qbinom(.975, n_eligible, .95) / n_eligible,
      truth = first(truth),
      .groups = "drop"
    )

  arm_summary <- performance %>%
    filter(method %in% confirmatory_methods) %>%
    group_by(calibration_arm_label, method_label) %>%
    summarise(
      cells = n(),
      mean_eligibility = mean(convergence, na.rm = TRUE),
      mean_coverage = mean(coverage, na.rm = TRUE),
      minimum_coverage = min(coverage, na.rm = TRUE),
      cells_below_mc_band = sum(coverage < mc_low, na.rm = TRUE),
      cells_in_mc_band = sum(coverage >= mc_low & coverage <= mc_high, na.rm = TRUE),
      cells_above_mc_band = sum(coverage > mc_high, na.rm = TRUE),
      mean_success_and_cover = mean(success_and_cover, na.rm = TRUE),
      mean_absolute_bias = mean(abs(bias), na.rm = TRUE),
      median_cell_mean_se_ratio = median(mean_se_ratio, na.rm = TRUE),
      .groups = "drop"
    )

  arm_stat <- function(arm, method, variable) {
    arm_summary %>%
      filter(calibration_arm_label == arm, method_label == method) %>%
      pull(all_of(variable)) %>%
      first()
  }

  bias_arm_summary <- performance %>%
    filter(method %in% bias_plot_methods) %>%
    group_by(calibration_arm_label, method_label) %>%
    summarise(
      cells = n(),
      cells_bias_ci_excludes_zero = sum(
        bias_mc_low > 0 | bias_mc_high < 0
      ),
      mean_bias = mean(bias),
      bias_mc_se = sqrt(sum(bias_mc_se^2)) / cells,
      bias_mc_low = mean_bias - qnorm(.975) * bias_mc_se,
      bias_mc_high = mean_bias + qnorm(.975) * bias_mc_se,
      mean_absolute_bias = mean(abs(bias)),
      .groups = "drop"
    )

  null_performance <- replications %>%
    filter(standardized_beta_target == 0) %>%
    group_by(
      condition_id, calibration_arm, calibration_arm_label,
      mean_clus_size, marginal_rho, method, method_label
    ) %>%
    summarise(
      n_rep = n(),
      n_eligible = sum(analysis_eligible, na.rm = TRUE),
      eligibility = n_eligible / n_rep,
      n_type1 = sum(reject_two_sided, na.rm = TRUE),
      n_positive = sum(reject_positive, na.rm = TRUE),
      n_negative = sum(reject_negative, na.rm = TRUE),
      type1_error = n_type1 / n_eligible,
      positive_rejection = n_positive / n_eligible,
      negative_rejection = n_negative / n_eligible,
      operational_type1 = n_type1 / n_rep,
      coverage = mean(covered[analysis_eligible], na.rm = TRUE),
      bias = mean((estimate - truth)[analysis_eligible], na.rm = TRUE),
      empirical_sd = sd((estimate - truth)[analysis_eligible], na.rm = TRUE),
      bias_mc_se = empirical_sd / sqrt(n_eligible),
      .groups = "drop"
    ) %>%
    mutate(
      wilson_denominator = 1 + qnorm(.975)^2 / n_eligible,
      wilson_center = (
        type1_error + qnorm(.975)^2 / (2 * n_eligible)
      ) / wilson_denominator,
      wilson_half_width = qnorm(.975) * sqrt(
        type1_error * (1 - type1_error) / n_eligible +
          qnorm(.975)^2 / (4 * n_eligible^2)
      ) / wilson_denominator,
      type1_ci_low = pmax(0, wilson_center - wilson_half_width),
      type1_ci_high = pmin(1, wilson_center + wilson_half_width),
      nominal_mc_low = qbinom(.025, n_eligible, .05) / n_eligible,
      nominal_mc_high = qbinom(.975, n_eligible, .05) / n_eligible,
      coverage_type1_identity_error = abs(type1_error - (1 - coverage))
    )

  null_arm_summary <- null_performance %>%
    filter(method %in% confirmatory_methods) %>%
    group_by(calibration_arm_label, method_label) %>%
    summarise(
      cells = n(),
      cells_below_nominal_range = sum(type1_error < nominal_mc_low),
      cells_in_nominal_range = sum(
        type1_error >= nominal_mc_low & type1_error <= nominal_mc_high
      ),
      cells_above_nominal_range = sum(type1_error > nominal_mc_high),
      mean_eligibility = mean(eligibility),
      mean_bias = mean(bias),
      bias_mc_se = sqrt(sum(bias_mc_se^2)) / cells,
      mean_type1_error = mean(type1_error),
      mean_positive_rejection = mean(positive_rejection),
      mean_negative_rejection = mean(negative_rejection),
      mean_operational_type1 = mean(operational_type1),
      minimum_type1_error = min(type1_error),
      maximum_type1_error = max(type1_error),
      .groups = "drop"
    )

  null_direction_summary <- null_arm_summary %>%
    dplyr::select(
      calibration_arm_label, method_label,
      mean_positive_rejection, mean_negative_rejection
    ) %>%
    pivot_longer(
      cols = starts_with("mean_"),
      names_to = "direction",
      values_to = "rejection_rate"
    ) %>%
    mutate(
      direction = recode(
        direction,
        mean_positive_rejection = "CI entirely above zero",
        mean_negative_rejection = "CI entirely below zero"
      ),
      signed_rejection_rate = if_else(
        direction == "CI entirely below zero",
        -rejection_rate,
        rejection_rate
      )
    )

  null_stat <- function(arm, method, variable) {
    null_arm_summary %>%
      filter(calibration_arm_label == arm, method_label == method) %>%
      pull(all_of(variable)) %>%
      first()
  }

  worst_naive_cell <- performance %>%
    filter(method == "naive_dual_blup") %>%
    slice_max(abs(bias), n = 1, with_ties = FALSE)
  largest_msem_se <- replications %>%
    filter(method == "msem", analysis_eligible, is.finite(se)) %>%
    slice_max(se, n = 1, with_ties = FALSE)
  largest_msem_se_cell <- performance %>%
    filter(
      method == "msem",
      condition_id == largest_msem_se$condition_id[[1]]
    )
  largest_msem_se_profile <- replications %>%
    filter(
      method == "msem", analysis_eligible,
      condition_id == largest_msem_se$condition_id[[1]]
    ) %>%
    summarise(
      eligible_n = n(),
      median_se = median(se, na.rm = TRUE),
      p99_se = quantile(se, .99, na.rm = TRUE),
      maximum_se = max(se, na.rm = TRUE)
    )
} else {
  replications <- tibble()
  performance <- tibble()
  arm_summary <- tibble()
  bias_arm_summary <- tibble()
  null_performance <- tibble()
  null_arm_summary <- tibble()
  null_direction_summary <- tibble()
}

theme_study5 <- function() {
  theme_bw(base_size = 10) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      strip.background = element_rect(fill = "grey95", colour = "grey75")
    )
}
```

# Technical summary

Study 5 resolves the main design question more clearly than it resolves the
choice of estimator. Preserving a regular random-effects covariance shape
removes much of the severe undercoverage seen for the corrected methods, while
the oracle remains close to 95% in all three arms. It does not make every
method well calibrated: the naive BLUP analysis, for example, has inflated
null rejection rates in both fixed-shape arms. The severe Study 2 pattern is
therefore not evidence that posterior reliability is intrinsically unsuitable.
It is primarily evidence against attaining low reliability by collapsing only
the random-slope variance, $\tau_1^2$, especially when the random intercept and
slope are correlated. The old code calls this the `solve-G_22` arm because the
slope variance occupies `G[2,2]` in R. In the mathematical notation used in this
report, however,

$$
G =
\begin{pmatrix}
\tau_0^2 & \tau_{01} \\
\tau_{01} & \tau_1^2
\end{pmatrix},
$$

so $\tau_0^2$ is the random-intercept variance, $\tau_1^2$ is the random-slope
variance, and $\tau_{01}$ is their covariance. The label `solve-G_22` is
retained only when referring to that legacy code terminology; it means solving
for $\tau_1^2$, not adopting a mixed mathematical subscript convention.

The repaired DGM does **not** make all methods inferentially equivalent. Under
fixed-shape marginal and partial reliability, respectively, mean conditional
coverage is `r percent(arm_stat("Fixed G shape: marginal R", "Fuller", "mean_coverage"), accuracy = .1)`
and `r percent(arm_stat("Fixed G shape: partial R", "Fuller", "mean_coverage"), accuracy = .1)`
for Fuller, `r percent(arm_stat("Fixed G shape: marginal R", "Lai 2S-PA", "mean_coverage"), accuracy = .1)`
and `r percent(arm_stat("Fixed G shape: partial R", "Lai 2S-PA", "mean_coverage"), accuracy = .1)`
for 2S-PA, and `r percent(arm_stat("Fixed G shape: marginal R", "MSEM", "mean_coverage"), accuracy = .1)`
and `r percent(arm_stat("Fixed G shape: partial R", "MSEM", "mean_coverage"), accuracy = .1)`
for MSEM. Thus Fuller and 2S-PA generally overcover after the geometry repair,
whereas MSEM is much closer to nominal and retains some non-null undercoverage
under marginal-reliability calibration.

At $\beta=0$, CI coverage and Type I error are complements for the
CI-inversion test used here: conditional Type I error is exactly
$1-\text{coverage}$. The oracle is nominal in every arm. Fuller and 2S-PA are
extremely conservative (roughly .2%--.6% arm-average Type I error), whereas
naive BLUP is inflated in both fixed-shape arms (about 8%--9%). Legacy-arm
MSEM averages 5.2%, but this hides a sharp split: its two $\rho=.5$ cells have
8.1% and 8.7% Type I error, while its $\rho=0$ cells are conservative. Thus
poor null coverage is direct Type I error inflation, but mean bias is not by
itself the mechanism; scale estimation, tail behavior, and fit selection also
determine rejection.

The 2S-PA result needs a strong qualification: its coverage is conditional on
an analyzable fit. Its mean eligibility is only
`r percent(arm_stat("Fixed G shape: marginal R", "Lai 2S-PA", "mean_eligibility"), accuracy = .1)`
under fixed-shape marginal calibration and
`r percent(arm_stat("Fixed G shape: partial R", "Lai 2S-PA", "mean_eligibility"), accuracy = .1)`
under fixed-shape partial calibration. The corresponding probabilities of
both obtaining an eligible result and covering are only
`r percent(arm_stat("Fixed G shape: marginal R", "Lai 2S-PA", "mean_success_and_cover"), accuracy = .1)`
and `r percent(arm_stat("Fixed G shape: partial R", "Lai 2S-PA", "mean_success_and_cover"), accuracy = .1)`.
High conditional coverage is therefore not evidence that 2S-PA is the best
operational procedure.

The bias results sharpen the method comparison. The oracle is centered in all
three arms. In the legacy arm, the equal-cell-weighted mean biases are about
-.058 for Fuller and -.065 for 2S-PA. Shape preservation removes that negative
centering problem, but the fixed-shape marginal/partial mean biases become
.028/.024 for Fuller, .114/.081 for 2S-PA, and .072/.050 for MSEM. Thus MSEM is
less biased than 2S-PA after repair, while Fuller remains the least biased
proxy method. Narrow Monte Carlo intervals show that most of these mean biases
are estimated precisely; precision does not make them substantively small.

The decision from this bridge is to retain posterior-reliability calibration,
replace the one-coordinate $\tau_1^2$ manipulation with a shape-preserving one,
and include residualized-slope reliability ($R_\eta$) as a primary design
quantity. The partial-reliability arm improves MSEM specifically when
$\rho=.5$, but it does not cure the Fuller/2S-PA overcoverage. The oracle is the
interval-calibration benchmark. Among the feasible estimators, MSEM gives the
best overall balance in this grid; Fuller is a stable, low-bias point-estimation
comparator whose intervals are conservative; and 2S-PA should be reported with
both conditional coverage and eligibility.

# Question and decision

Study 5 asks whether the low coverage in VH Study 2 is caused by posterior
reliability itself, by the legacy practice of attaining low reliability by
collapsing only the slope variance, or by using marginal slope reliability for
a dual-predictor estimand. It changes only the calibration route and reuses the
Study 2 data generator and estimator code.

The prespecified interpretation is:

- fixed-shape marginal coverage improves over legacy calibration: retain
  posterior reliability and replace the one-coordinate calibration;
- only fixed-shape partial-reliability coverage improves: use reliability of
  the residualized slope as the primary information factor;
- both fixed-shape arms improve similarly: covariance regularity is decisive,
  and either reliability definition is usable with geometry guards;
- proxy methods fail for non-null effects but oracle coverage is nominal:
  estimator or interval limitations remain after DGM repair; and
- oracle coverage fails: audit truth and interval construction before deciding
  between information factors.

`r if (results_available) paste0("The production replication file was found at ", replication_path, " and the results below are live.") else paste0("No production replication file was found at ", replication_path, ". This report therefore validates and displays the design, and will populate its results sections after the simulation is run.")`

# Design and manipulation checks

The 36 conditions fix 100 clusters and target reliability .25. They cross
three calibration arms, cluster sizes 10 and 25, correlations 0 and .5, and
standardized focal effects 0, .2, and .4. Each condition is intended to receive
1,000 replications.

```{r design-table}
design %>%
  distinct(
    calibration_arm, calibration_metric, mean_clus_size, marginal_rho,
    sigma, achieved_reliability, achieved_partial_reliability,
    slope_intercept_variance_ratio, G_condition_number,
    conditional_slope_icc
  ) %>%
  mutate(
    calibration_arm = unname(arm_labels[calibration_arm]),
    across(where(is.numeric), ~ round(.x, 4))
  ) %>%
  arrange(calibration_arm, mean_clus_size, marginal_rho) %>%
  kable(
    col.names = c(
      "Calibration arm", "Target metric", "m", "rho", "sigma",
      "Marginal R", "Partial R", "τ₁²/τ₀²", "kappa(G)",
      "Conditional slope ICC"
    ),
    caption = "Population manipulation checks for the 12 unique first-stage designs."
  )
```

The arm comparison is intentionally sharp. At $\rho=.5$, the legacy
calibration attains marginal reliability .25 while residualized-slope
reliability falls below .04; at $m=25$ it is about .015. The fixed-shape arms
retain $\tau_1^2/\tau_0^2=1$ and a well-conditioned $G$, then vary the level-1
residual scale. At $\rho=0$, marginal and residualized reliability coincide,
so the two fixed-shape arms are exact duplicate controls.

# Monte Carlo results

## Regular covariance geometry restores coverage, but methods diverge

```{r coverage-plot, eval=results_available, fig.width=11, fig.height=7}
performance %>%
  filter(method %in% confirmatory_methods) %>%
  ggplot(aes(
    x = calibration_arm_label,
    y = coverage,
    colour = method_label,
    shape = method_label,
    group = method_label
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey35") +
  geom_point(position = position_dodge(width = .35), size = 1.8) +
  geom_line(position = position_dodge(width = .35), linewidth = .45) +
  facet_grid(
    mean_clus_size + marginal_rho ~ standardized_beta_target,
    labeller = labeller(
      mean_clus_size = function(x) paste0("m = ", x),
      marginal_rho = function(x) paste0("rho = ", x),
      standardized_beta_target = function(x) paste0("beta = ", x)
    )
  ) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_colour_manual(values = method_palette, drop = TRUE) +
  scale_shape_manual(values = c(16, 17, 15, 18, 8), drop = TRUE) +
  labs(
    x = NULL,
    y = "95% CI coverage",
    colour = NULL,
    shape = NULL,
    title = "Coverage across matched calibration arms",
    subtitle = "Coverage is conditional on an eligible estimate; dashed line is the 95% target"
  ) +
  theme_study5() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1))
```

```{r arm-summary, eval=results_available}
arm_summary %>%
  mutate(
    across(
      c(mean_eligibility, mean_coverage, minimum_coverage, mean_success_and_cover),
      percent, accuracy = .1
    ),
    mean_absolute_bias = formatC(mean_absolute_bias, digits = 3, format = "g"),
    median_cell_mean_se_ratio = round(median_cell_mean_se_ratio, 3)
  ) %>%
  kable(
    col.names = c(
      "Calibration arm", "Method", "Cells", "Eligibility", "Mean coverage",
      "Minimum coverage", "Below MC band", "In MC band", "Above MC band",
      "Success and cover", "Mean |bias|", "Median cell mean-SE / SD"
    ),
    caption = paste(
      "Arm-level summary. Coverage is conditional on eligibility; success and",
      "cover counts ineligible replications as unsuccessful. Monte Carlo bands",
      "use each cell's eligible denominator."
    )
  )
```

The oracle provides the critical calibration check: mean coverage ranges from
`r percent(min(arm_summary$mean_coverage[arm_summary$method_label == "Oracle dual"]), accuracy = .1)`
to `r percent(max(arm_summary$mean_coverage[arm_summary$method_label == "Oracle dual"]), accuracy = .1)`,
with complete eligibility. This makes a truth-coding or generic confidence-
interval bug an implausible explanation for the proxy-method patterns.

The fixed-shape arms move all three proxy methods away from the severe
legacy-arm undercoverage. However, “improvement” should not be read as
“calibration.”
Fuller is above its cell-specific Monte Carlo band in 11 of 12 marginal-
reliability cells and 10 of 12 partial-reliability cells. 2S-PA is above the
band in 11 of 12 cells in each fixed-shape arm. MSEM is more balanced: under
marginal calibration it is below the band in 5 cells, within it in 5, and above
it in 2; under partial calibration the corresponding counts are 2, 4, and 6.

## Overcoverage depends on method, effect size, and selection

```{r fixed-shape-effect-summary, eval=results_available}
performance %>%
  filter(
    calibration_arm != "current_g22",
    method %in% corrected_methods
  ) %>%
  group_by(calibration_arm_label, method_label, standardized_beta_target) %>%
  summarise(
    eligibility = mean(convergence),
    mean_coverage = mean(coverage),
    mean_bias = mean(bias),
    below = sum(coverage < mc_low),
    within = sum(coverage >= mc_low & coverage <= mc_high),
    above = sum(coverage > mc_high),
    .groups = "drop"
  ) %>%
  mutate(
    eligibility = percent(eligibility, accuracy = .1),
    mean_coverage = percent(mean_coverage, accuracy = .1),
    mean_bias = round(mean_bias, 3)
  ) %>%
  kable(
    col.names = c(
      "Calibration arm", "Method", "Beta", "Eligibility", "Coverage",
      "Mean bias", "Below", "Within", "Above"
    ),
    caption = "Fixed-shape results by effect size; band counts are across the four m-by-rho cells."
  )
```

For Fuller, overcoverage is most pronounced at the null and diminishes as the
effect grows: coverage under fixed-shape marginal calibration moves from
99.8% at $\beta=0$ to 97.4% at $\beta=.4$, and under partial calibration from
99.4% to 97.2%. Fuller has complete eligibility and relatively small bias, so
this is genuine interval conservatism rather than a fit-selection artifact.

2S-PA shows the same conditional-coverage gradient, but it also has appreciable
positive non-null bias and many ineligible replications. Its fixed-shape
conditional coverage near 99% is therefore a property of the selected,
analyzable fits and their intervals, not an unconditional guarantee. By
contrast, MSEM coverage decreases with effect size: under marginal calibration
it moves from 96.1% at the null to 92.9% at $\beta=.4$; partial calibration
raises the $\beta=.4$ result to 94.5% but modestly overcovers the null and
$\beta=.2$ conditions. This is why averaging coverage alone would make the
partial MSEM result look more uniformly successful than it is.

The marginal-versus-partial MSEM difference is concentrated exactly where the
two reliability definitions diverge. Averaged across $m$ and $\beta$ at
$\rho=.5$, MSEM coverage is 93.4% under fixed-shape marginal calibration and
96.4% under fixed-shape partial calibration. At $\rho=0$, where the population
DGMs are duplicates, the corresponding values are 95.6% and 95.8%. Fuller and
2S-PA are already near 99% in both correlated fixed-shape arms. This supports
$R_\eta$ as a consequential information factor for MSEM, while also showing
that it cannot explain the conservative Fuller/2S-PA intervals.

## Null undercoverage is Type I error inflation, not merely a warning sign

The null conditions have a more specific interpretation than the non-null
conditions. For the two-sided CI-inversion rule

$$
\text{reject }H_0:\beta=0 \quad\Longleftrightarrow\quad
0 \notin [L,U],
$$

the conditional Type I error estimate in a null cell is

$$
\widehat{\alpha}_{\mathrm{conditional}}
= \frac{\sum I(\text{eligible})I(0\notin[L,U])}
        {\sum I(\text{eligible})}
= 1-\widehat{\operatorname{coverage}}.
$$

This is an identity, not an approximation. It applies because the same 95%
interval and eligible-fit denominator define both quantities. It would not
necessarily apply to a separately implemented Wald or likelihood-ratio test
whose rejection rule differs from CI inversion.

Accordingly, a separate null-coverage-versus-Type-I scatter would be
algebraically redundant: every point would lie on
$\widehat{\alpha}=1-\widehat{C}$. The report plots the null results on the Type
I error scale because that is the decision-relevant interpretation. This
equivalence is specific to $\beta=0$. In the non-null cells, $1-$coverage is
interval miscoverage, not Type I error; power would instead be the probability
that the interval excludes zero.

```{r null-type1-plot, eval=results_available, fig.width=11, fig.height=6.8}
type1_position <- position_dodge(width = .62)

null_performance %>%
  filter(method %in% confirmatory_methods) %>%
  ggplot(aes(
    x = type1_error,
    y = method_label,
    colour = calibration_arm_label,
    shape = calibration_arm_label,
    group = calibration_arm_label
  )) +
  geom_vline(xintercept = .05, linetype = 2, colour = "grey30") +
  geom_errorbarh(
    aes(xmin = type1_ci_low, xmax = type1_ci_high),
    height = 0,
    linewidth = .65,
    position = type1_position
  ) +
  geom_point(size = 2, position = type1_position) +
  facet_grid(
    mean_clus_size ~ marginal_rho,
    labeller = labeller(
      mean_clus_size = function(x) paste0("m = ", x),
      marginal_rho = function(x) paste0("rho = ", x)
    )
  ) +
  scale_colour_manual(values = arm_palette) +
  scale_shape_manual(values = c(16, 17, 15)) +
  scale_x_continuous(
    labels = percent_format(accuracy = 1),
    breaks = seq(0, .14, .02)
  ) +
  coord_cartesian(xlim = c(0, .14)) +
  labs(
    x = "Conditional Type I error rate",
    y = NULL,
    colour = "Calibration arm",
    shape = "Calibration arm",
    title = "Null-condition Type I error by design cell",
    subtitle = "Whiskers are 95% Wilson intervals using eligible fits; dashed line is the nominal 5% rate"
  ) +
  theme_study5()
```

The cell-level view prevents arm averages from hiding opposite failures.
Oracle Type I error is within the cell-specific nominal Monte Carlo range in
all 12 cells. Legacy-arm MSEM averages 5.2%, but its $\rho=.5$ rates are 8.1%
at $m=10$ and 8.7% at $m=25$; its $\rho=0$ rates are only 2.4% and 1.6%.
Consequently, a nominal-looking arm mean is not sufficient evidence of test
calibration. Fixed-shape MSEM is conservative on average, at 3.9% under
marginal reliability and 3.1% under partial reliability.

Naive dual BLUP shows the reverse arm pattern: its legacy-arm null Type I
error is approximately nominal, but all four null cells are inflated under
each fixed-shape calibration. The arm-average conditional rates are 8.7% and
8.4%. Fuller and 2S-PA are conservative in every null cell, with conditional
rates below about 1.1%. Their null overcoverage therefore corresponds to very
low Type I error, not successful nominal calibration.

```{r null-type1-summary, eval=results_available}
null_arm_summary %>%
  transmute(
    calibration_arm_label,
    method_label,
    eligibility = percent(mean_eligibility, accuracy = .1),
    mean_null_bias = sprintf("%+.3f", mean_bias),
    conditional_type1 = percent(mean_type1_error, accuracy = .1),
    cell_type1_range = paste0(
      percent(minimum_type1_error, accuracy = .1), "--",
      percent(maximum_type1_error, accuracy = .1)
    ),
    positive_rejection = percent(mean_positive_rejection, accuracy = .1),
    negative_rejection = percent(mean_negative_rejection, accuracy = .1),
    operational_type1 = percent(mean_operational_type1, accuracy = .1),
    cells_below_within_above = paste(
      cells_below_nominal_range,
      cells_in_nominal_range,
      cells_above_nominal_range,
      sep = " / "
    )
  ) %>%
  kable(
    col.names = c(
      "Calibration arm", "Method", "Eligibility", "Mean null bias",
      "Conditional Type I", "Cell range", "Positive rejection",
      "Negative rejection", "Success and reject", "Cells below / within / above"
    ),
    caption = paste(
      "Equal-cell-weighted null summaries. 'Success and reject' uses all",
      "replications as the denominator; nominal-range counts use each cell's",
      "eligible denominator and a two-sided 95% Monte Carlo range around .05."
    )
  )
```

“Success and reject” is an operational complement to conditional Type I error,
not a preferred error-rate definition. It counts an ineligible fit as no
rejection, so it is mechanically smaller for methods such as 2S-PA and naive
BLUP. Unless the actual analysis workflow treats every failed fit as a final
non-rejection, a low success-and-reject rate does not demonstrate Type I error
control; eligibility must be reported separately.

## Rejection direction helps distinguish bias from interval failure

```{r null-direction-plot, eval=results_available, fig.width=10, fig.height=7.2}
null_direction_summary %>%
  ggplot(aes(
    x = signed_rejection_rate,
    y = method_label,
    fill = direction
  )) +
  geom_vline(xintercept = 0, colour = "grey30") +
  geom_col(width = .65) +
  facet_wrap(~ calibration_arm_label, ncol = 1) +
  scale_fill_manual(values = c(
    "CI entirely below zero" = "#E17C05",
    "CI entirely above zero" = "#4E79A7"
  )) +
  scale_x_continuous(
    limits = c(-.06, .06),
    breaks = seq(-.06, .06, .02),
    labels = function(x) percent(abs(x), accuracy = 1)
  ) +
  labs(
    x = "Cell-weighted conditional rejection rate",
    y = NULL,
    fill = NULL,
    title = "Direction of false-positive rejections",
    subtitle = "Left: CI entirely below zero; right: CI entirely above zero; the two sides sum to total Type I error"
  ) +
  theme_study5()
```

Bias is one possible cause of directional Type I error, but neither zero bias
nor statistically detectable bias is sufficient to characterize test
calibration. Legacy-arm MSEM at $m=10$, $\rho=.5$ illustrates a bias-related
failure: mean null bias is +.046 and positive-direction rejection is 5.4%,
versus 2.8% in the negative direction. Yet legacy-arm MSEM at $m=25$, $\rho=.5$
has near-zero mean bias (+.007) and still rejects 8.7%, almost symmetrically.
That second failure implicates interval scale or tail behavior rather than
centering.

The converse is also visible. Fixed-shape marginal 2S-PA has positive mean
null bias (+.031), but its intervals are so conservative that Type I error is
only .2%. Partial-arm MSEM has appreciable negative mean null bias (-.056), yet
its Type I error is 3.1%, below nominal. Naive BLUP's enormous mean biases are
driven by rare ratio outliers and do not map cleanly onto rejection direction.
The null conditions should therefore be evaluated with a joint diagnostic:
mean bias and its MC interval for centering, directional Type I error for tail
asymmetry, total Type I error for test calibration, and eligibility for the
selection mechanism.

## Shape preservation changes the direction and magnitude of bias

For each condition and method, mean bias is the mean of
$\hat\beta-\beta$ among eligible replications. Its Monte Carlo standard error
is

$$
\operatorname{MCSE}(\widehat{\operatorname{Bias}})
= \frac{SD(\hat\beta-\beta)}{\sqrt{n_{\mathrm{eligible}}}},
$$

and the plotted interval is the normal-approximation mean bias $\pm 1.96$
MCSE. These are intervals for uncertainty in the simulation estimate of mean
bias. They are not the model-based confidence intervals used to calculate
coverage.

The figure includes the oracle and the three corrected/latent methods. Naive
dual BLUP is omitted because its catastrophic ratio estimates produce
condition-level mean biases as large as $10^5$, which would collapse every
other interval onto the zero line. It remains in the coverage table and the
outlier audit below.

```{r bias-ci-plot, eval=results_available, fig.width=11, fig.height=8.5}
bias_position <- position_dodge(width = .65)

performance %>%
  filter(method %in% bias_plot_methods) %>%
  mutate(
    design_cell = factor(
      paste0("m=", mean_clus_size, ", rho=", marginal_rho),
      levels = c("m=10, rho=0", "m=10, rho=0.5", "m=25, rho=0", "m=25, rho=0.5")
    )
  ) %>%
  ggplot(aes(
    x = bias,
    y = design_cell,
    colour = calibration_arm_label,
    shape = calibration_arm_label,
    group = calibration_arm_label
  )) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey35") +
  geom_errorbarh(
    aes(xmin = bias_mc_low, xmax = bias_mc_high),
    height = 0,
    linewidth = .7,
    position = bias_position
  ) +
  geom_point(size = 2, position = bias_position) +
  facet_grid(
    method_label ~ standardized_beta_target,
    labeller = labeller(
      method_label = label_value,
      standardized_beta_target = function(x) paste0("beta = ", x)
    )
  ) +
  scale_colour_manual(values = arm_palette) +
  scale_shape_manual(values = c(16, 17, 15)) +
  scale_x_continuous(labels = number_format(accuracy = .05)) +
  labs(
    x = "Mean bias with 95% Monte Carlo CI",
    y = NULL,
    colour = "Calibration arm",
    shape = "Calibration arm",
    title = "Mean bias and Monte Carlo uncertainty by design cell",
    subtitle = "Each point is one condition-method mean among eligible replications; all panels use the standardized effect scale"
  ) +
  theme_study5()
```

```{r bias-arm-summary, eval=results_available}
bias_arm_summary %>%
  transmute(
    calibration_arm_label,
    method_label,
    cells,
    mean_bias = round(mean_bias, 3),
    bias_mc_se = round(bias_mc_se, 4),
    bias_mc_ci = sprintf("[%.3f, %.3f]", bias_mc_low, bias_mc_high),
    mean_absolute_bias = round(mean_absolute_bias, 3),
    cells_bias_ci_excludes_zero
  ) %>%
  kable(
    col.names = c(
      "Calibration arm", "Method", "Cells", "Mean bias", "MC SE",
      "95% MC CI", "Mean |bias|", "Cell CIs excluding zero"
    ),
    caption = paste(
      "Equal-cell-weighted arm summaries. The arm MC SE propagates independent",
      "cell-level Monte Carlo uncertainty; cell counts are out of 12."
    )
  )
```

The oracle arm-average intervals all include zero, and only zero or one of its
12 cell intervals excludes zero in each arm, as expected under correct
centering. In contrast, legacy calibration produces reliably negative
arm-average bias for Fuller (95% MC CI [-.063, -.054]) and 2S-PA
([-.070, -.060]). Shape preservation reverses the direction: the corrected
methods are generally positively biased for non-null effects. Partial-
reliability calibration reduces the arm-average bias relative to marginal
calibration for 2S-PA (.081 versus .114) and MSEM (.050 versus .072), but does
not eliminate it.

Arm means can conceal cancellation. Legacy-arm MSEM has mean bias near zero
(-.006; 95% MC CI [-.015, .003]) but mean absolute bias .104, and 9 of 12 cell
intervals exclude zero. The cell plot is therefore primary; the arm table is a
compact summary rather than evidence that a near-zero average implies
condition-level centering.

The bias and coverage evidence also rule out a one-mechanism interpretation.
Fuller overcovers under fixed-shape calibration despite modest positive bias,
so its conservative interval behavior dominates its centering error. 2S-PA
combines larger positive bias with overcoverage because uncertainty is highly
right-skewed and coverage is conditional on successful fits. MSEM's positive
non-null bias, lower typical SE ratios, and effect-specific undercoverage are
more mutually aligned. Because 1,000 replications make the Monte Carlo
intervals narrow, exclusion of zero should be interpreted together with bias
magnitude rather than as a standalone practical-significance test.

## Centering bias versus typical interval width

```{r bias-se-plot, eval=results_available, fig.width=10, fig.height=7.2}
performance %>%
  filter(method %in% corrected_methods) %>%
  mutate(
    design_cell = factor(
      paste0("m=", mean_clus_size, ", rho=", marginal_rho),
      levels = c("m=10, rho=0", "m=10, rho=0.5", "m=25, rho=0", "m=25, rho=0.5")
    )
  ) %>%
  ggplot(aes(
    x = bias,
    y = typical_se_ratio,
    colour = calibration_arm_label,
    shape = design_cell
  )) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey45") +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey45") +
  geom_point(alpha = .85, size = 2) +
  facet_grid(
    method_label ~ standardized_beta_target,
    labeller = labeller(
      method_label = label_value,
      standardized_beta_target = function(x) paste0("beta = ", x)
    )
  ) +
  scale_colour_manual(values = arm_palette) +
  scale_shape_manual(values = c(16, 17, 15, 18)) +
  scale_y_continuous(breaks = seq(.25, 2, .25), limits = c(.3, 1.9)) +
  labs(
    x = "Bias on standardized focal-effect scale",
    y = "Median model SE / empirical SD",
    colour = "Calibration arm",
    shape = "Design cell",
    title = "Separate interval-width failure from centering bias",
    subtitle = "Corrected/latent methods only; median SE prevents one pathological fit from controlling the scale"
  ) +
  guides(
    shape = guide_legend(order = 1, nrow = 1, byrow = TRUE),
    colour = guide_legend(order = 2, nrow = 1, byrow = TRUE)
  ) +
  theme_study5() +
  theme(legend.box = "vertical")
```

The robust plot answers a “typical replication” question. It is deliberately
restricted to Fuller, 2S-PA, and MSEM because naive dual BLUP has catastrophic
cell means that make the focal scale unreadable. It uses the median model SE
within each cell rather than the mean because model SEs have a very long right
tail. This change does not erase the tail: the next audit reports it directly.

The distinction between typical and mean interval width is scientifically
important. In the fixed-shape arms, the median cell mean-SE/empirical-SD ratio
is about 1.4--1.9 for Fuller and 2S-PA, consistent with conservative average
width. Yet the median-SE/empirical-SD ratio for 2S-PA is only about .55. Thus
2S-PA overcoverage is not produced by uniformly wide intervals. A small number
of very large standard errors, dependence between the estimated effect and its
estimated uncertainty, and conditioning on successful fits all contribute.
MSEM has the same right-skewed-SE signature but substantially less overcoverage.

## Outlier and instability audit

```{r outlier-audit, eval=results_available}
tibble::tribble(
  ~Diagnostic, ~Observed_result, ~Interpretation,
  "Naive dual BLUP worst cell",
  paste0(
    "Condition ", worst_naive_cell$condition_id,
    "; cell bias = ", formatC(worst_naive_cell$bias, digits = 4, format = "g")
  ),
  "Catastrophic ratio estimates are off-scale; naive BLUP is excluded from the mechanism plot but retained in coverage summaries.",
  "Largest MSEM model SE",
  paste0(
    "Condition ", largest_msem_se$condition_id,
    "; max SE = ", round(largest_msem_se_profile$maximum_se, 2),
    "; cell median SE = ", round(largest_msem_se_profile$median_se, 3),
    "; 99th percentile = ", round(largest_msem_se_profile$p99_se, 2)
  ),
  "One replication inflated the cell mean SE and the original y-axis; median SE is the appropriate robust display statistic.",
  "Affected MSEM cell",
  paste0(
    "Mean-SE/SD = ", round(largest_msem_se_cell$mean_se_ratio, 3),
    "; median-SE/SD = ", round(largest_msem_se_cell$typical_se_ratio, 3),
    "; coverage = ", percent(largest_msem_se_cell$coverage, accuracy = .1),
    " (", largest_msem_se_cell$n_eligible, " eligible)"
  ),
  "The extreme SE is real numerical/statistical instability, but it should be isolated rather than allowed to determine the visual scale."
) %>%
  kable(caption = "Explicit audit of observations omitted or robustified in the mechanism plot.")
```

The affected MSEM cell is the legacy `solve-G_22` ($\tau_1^2$), $m=25$,
$\rho=0$,
$\beta=.4$ condition. Its largest eligible SE is
`r round(largest_msem_se_profile$maximum_se, 2)`, compared with a median of
`r round(largest_msem_se_profile$median_se, 3)`. The original mean-SE ratio of
`r round(largest_msem_se_cell$mean_se_ratio, 3)` was therefore a tail diagnostic,
not a representative interval-width estimate. Keeping it in the audit and
using a median in the plot makes both facts visible.

# Conclusions and decision

## What Study 5 establishes

1. **Low posterior reliability is not, by itself, the explanation for the poor
   coverage in Study 2.** When low reliability was produced by changing only
   the random-slope variance ($\tau_1^2$), the manipulation also created an
   unusual random-effects covariance structure. Holding that structure fixed
   removed much of the severe undercoverage for the corrected methods, while
   oracle coverage remained near 95%. The clearest design lesson is therefore
   to preserve covariance geometry when manipulating reliability. This result
   supports continued use of posterior reliability as an information measure;
   it does not show that every analysis based on that design will be calibrated.

2. **Marginal and residualized posterior reliability describe different
   information problems.** Marginal reliability concerns the slope as a whole.
   Residualized reliability ($R_\eta$) concerns the part of the slope that
   remains after accounting for its association with the random intercept. The
   two definitions generate the same population condition when $\rho=0$. When
   $\rho=.5$, calibration to $R_\eta$ improves MSEM coverage, especially for
   non-null effects. It does not remove the conservative coverage of Fuller or
   2S-PA. Thus $R_\eta$ is a useful design factor and diagnostic, not a single
   criterion that guarantees valid confidence intervals.

3. **The null conditions are tests of false-positive control, not merely
   another set of coverage cells.** With the two-sided CI-inversion rule used
   here, conditional Type I error at $\beta=0$ is exactly one minus coverage.
   Legacy-arm MSEM looks nominal when averaged across conditions (5.2%), but
   its two correlated-intercept/slope cells reject 8.1% and 8.7% of the time;
   conservative cells at $\rho=0$ conceal this inflation. Naive BLUP rejects
   about 8%--9% of the time in both fixed-shape arms. In contrast, Fuller and
   2S-PA usually reject in fewer than 1% of eligible null replications. Those
   methods are not nominal; they are strongly conservative.

4. **Bias and false-positive error answer related but noninterchangeable
   questions.** Bias can make false positives favor one direction, as in one
   legacy-arm MSEM condition. Yet another MSEM condition has little mean bias
   and inflated Type I error in both directions, indicating a problem with
   interval scale or tail behavior. Conversely, 2S-PA can have visible bias and
   still reject too rarely because its intervals are wide. A methods evaluation
   therefore needs both bias and rejection-rate results, with Monte Carlo
   uncertainty around each.

5. **Neither overall coverage nor an arm-level average is an adequate summary.**
   Averaging can make liberal and conservative conditions appear nominal.
   Conditional coverage can also look excellent when many replications never
   produce an eligible estimate. The minimum reporting set is conditional
   coverage, eligibility, the probability of both success and coverage, null
   Type I error in each direction, mean and absolute bias, empirical sampling
   variability, and typical as well as tail-sensitive interval width.

## What the results imply for each estimator

- **Oracle analysis is the calibration anchor.** Its coverage is approximately
  95%, its null Type I error is 4.8%--5.3% across arms, and every replication is
  eligible. It shows that the structural regression itself can be estimated
  properly when the true latent scores are observed. It is not a feasible
  analysis for empirical data, but it is the standard against which the proxy
  methods should be judged.

- **Fuller is stable and comparatively low-bias, but its intervals are too
  conservative.** In the fixed-shape arms it is always eligible, has
  arm-average bias of .024--.028, and covers 98.5%--98.8% rather than 95%.
  Its null Type I error is only .2%--.6%. Fuller is therefore useful as a
  point-estimation comparator, but the present interval procedure should not be
  described as calibrated. The overcoverage suggests a likely cost in power,
  which should be measured directly rather than inferred from coverage alone.

- **2S-PA combines conservative intervals with selective availability.** Its
  fixed-shape conditional coverage is 98.8%--98.9%, but only about 71% of
  marginal-reliability replications and 83% of partial-reliability replications
  are eligible. Its arm-average bias (.114 and .081) also exceeds Fuller's.
  High coverage among the subset of successful fits is therefore not evidence
  that 2S-PA is the safer procedure overall. It remains important for continuity
  with Lai, but eligibility and joint success-and-cover performance must be
  treated as primary results, not footnotes.

- **MSEM gives the best overall balance among the feasible methods in this
  grid, but it is not yet uniformly calibrated.** In the two fixed-shape arms,
  eligibility is 97.4% and 98.7% and conditional coverage is 94.5% and 96.1%.
  These values are closer to the nominal target than Fuller's or 2S-PA's, and
  partial-reliability calibration repairs much of the non-null undercoverage at
  $\rho=.5$. Nevertheless, some cells remain liberal or conservative, its
  fixed-shape arm-average bias is .072 and .050, and a small number of fits have
  extreme standard errors. MSEM should be the primary feasible estimator to
  refine and compare, not be conflated with the oracle calibration benchmark.

- **Naive dual-BLUP regression is not suitable for inferential use here.** Its
  fixed-shape null Type I error is about 8.7% and 8.4%, and some conditions have
  catastrophic bias. It may be retained as a negative control that illustrates
  the consequences of ignoring measurement error, but it should not be treated
  as a candidate analysis procedure.

## Recommended design and analysis decision

The next study should use a shape-preserving DGM rather than solve only for
$\tau_1^2$, and should manipulate or at least report both marginal and
residualized posterior reliability. The oracle should remain the calibration
anchor. MSEM should be the primary feasible estimator, Fuller the stable
point-estimation comparator, and 2S-PA the continuity comparator with Lai.
Naive BLUP should serve only as a negative control.

The next comparison should also be judged cell by cell. For null effects, the
primary criteria are total and directional Type I error, bias, and eligibility.
For non-null effects, they are coverage, bias, interval width, eligibility, and
directly estimated power. On the present evidence, no method is an unqualified
winner: MSEM offers the most promising balance, Fuller trades calibration for
strong conservatism and stability, and 2S-PA adds both conservatism and a
substantial fit-selection problem. That is a sharper and more useful conclusion
than choosing the method with the largest coverage percentage.

# Recommended next analyses

- Diagnose Fuller and 2S-PA interval construction under the fixed-shape DGM by
  retaining replicate-level interval length, studying the joint distribution
  of absolute error and SE, and comparing model-based intervals with a
  bootstrap or empirical calibration. This directly tests whether
  overcoverage comes from finite-sample SE formulas, a few tail fits, or
  error-SE dependence.
- For MSEM, inspect the small subset of extreme-SE fits and the relation between
  latent-slope variance estimates, boundary proximity, bias, and coverage.
  Report robust summaries, but do not automatically discard those fits.
- Extend the design with $m=3$, $\rho=-.5$, and covariance-shape ratios of 0.5
  and 2.0. Include a fixed raw-effect arm only after the information/geometry
  comparison is preserved, so effect scaling is not re-confounded with the
  covariance manipulation.
- Pre-specify effect-specific calibration criteria. A method should not pass
  merely because null overcoverage cancels non-null undercoverage in an arm
  average.
- Estimate power directly in the $\beta=.2$ and $\beta=.4$ conditions. Report
  both rejection among eligible fits and the joint probability of obtaining an
  eligible fit and rejecting the null. Do not use overcoverage alone as a proxy
  for power, especially when methods differ in eligibility.
- For every null cell, pre-specify the CI-inversion Type I error rate,
  positive- and negative-direction rejection rates, eligibility, and the
  cell-specific Monte Carlo acceptance range around .05. Treat an arm average
  as secondary because conservative and anti-conservative cells can cancel.

# Production execution

Run the full Study 5 array from the repository root:

```bash
sbatch --array=1-36%36 \
  --export=ALL,VIG_HALLQUIST_STUDY=5,VIG_HALLQUIST_N_SIM=1000,VIG_HALLQUIST_CHUNK_SIZE=1,VIG_HALLQUIST_OUT_DIR=outputs/vig_hallquist_study5 \
  vig_hallquist_2026/slurm/vig_hallquist_condition_array.sbatch
```

After all 36 tasks finish, rebuild the aggregate replication and summary
files without rerunning complete conditions:

```bash
Rscript vig_hallquist_2026/random_effects_structural_simulation.R \
  1000 5 outputs/vig_hallquist_study5 1 NA NA NA 1
```

The run is small enough to materialize the aggregate replication file used by
this report (36 conditions × 1,000 replications × 7 methods = 252,000 rows).

# Limitations

Coverage is conditional on the report's method-specific eligibility rule.
This matters little for Fuller, which is always eligible, but materially
changes the interpretation of 2S-PA. “Success and cover” is included as an
operational complement; it is not a replacement for conventional conditional
coverage because it treats every ineligible fit as a failure.

The same qualification applies to null Type I error. The primary rate is
conditional on an eligible fit. “Success and reject” is also shown, but it
treats ineligible fits as non-rejections and can make a method with frequent
failures appear safer than it is. Neither quantity alone defines the behavior
of a real analysis pipeline unless the handling of failed fits is specified.

The grid contains 100 clusters, one regular covariance shape, two moderate-to-
large cluster sizes, nonnegative correlation, one reliability target, and
three effect sizes. Conclusions about negative correlation, very small
clusters, different covariance shapes, other reliability targets, or power
are extrapolations. Arm-level means weight the 12 design cells equally and can
hide effect-specific patterns, so cell-level results remain primary.

At $\rho=0$, the marginal- and partial-reliability fixed-shape arms have the
same population DGM but independent Monte Carlo draws. Small differences
between them are therefore expected simulation noise and provide a useful
internal check, not evidence for one calibration metric. Finally, the robust
mechanism plot uses median SE to represent a typical replication; the report
retains mean-SE ratios and the explicit outlier audit so that heavy-tailed
uncertainty is not concealed.

# Returning to Studies 1--4: a revised calibration plan

Study 5 directly evaluated the Study 2 data-generating model, so its strongest
conclusion applies to Study 2. The same calibration mechanism is nevertheless
used throughout Studies 1--4: residual variance is held fixed while the
random-slope variance is solved to attain the target posterior reliability. This means
that reliability, the slope-to-intercept variance ratio, and the conditioning of
the random-effects covariance matrix can change together in every study. We
should therefore revisit the earlier studies with a common shape-preserving
calibration framework, while treating the Study 5 findings as a reason for
redesign rather than assuming that every Study 5 estimator result will
generalize unchanged.

## Common calibration standard

The revised studies should share the following design rules.

1. **Fix the random-effects covariance shape before targeting reliability.**
   Set the intercept and slope variances, their ratio, and the intercept--slope
   correlation first. For the primary replication, use the regular shape tested
   in Study 5, $\tau_1^2/\tau_0^2=1$. Solve for the Level-1 residual scale
   $\sigma$ to attain each reliability target. This varies the amount of
   measurement information without also collapsing the latent slope
   distribution.

2. **Match the reliability definition to the structural estimand.** Marginal
   slope reliability and residualized-slope reliability should both be stored
   and reported. The primary calibration target should depend on whether the
   focal coefficient concerns the full latent slope or the part of the slope
   unique of the random intercept. This is preferable to declaring either
   definition universally correct.

3. **Separate measurement calibration from structural-effect calibration.**
   First construct the desired marginal random-effects covariance and calibrate
   its posterior reliability. Then add the structural regression while
   preserving that marginal covariance. After this decomposition, verify that
   the residual covariance is positive definite and that the final generated
   process still has the intended marginal variances, correlation, reliability,
   standardized effect, and residual outcome variance.

4. **Record geometry and information diagnostics in every condition
   manifest.** At minimum, retain $\tau_0^2$, $\tau_1^2$,
   $\tau_{01}=\operatorname{Cov}(u_0,u_1)$,
   $\tau_1^2/\tau_0^2$, the condition number of $G$, $\sigma$, marginal
   posterior reliability, residualized posterior reliability, and the
   residualized slope variance. For heterogeneous clusters, also retain the
   mean, spread, and range of cluster-specific reliability. These quantities
   should appear in design checks before any performance result is interpreted.

5. **Keep a limited legacy-calibration arm.** The original `solve-G_22`
   conditions, which were calibrated by solving for $\tau_1^2$, should not
   remain the basis for the main conclusions, but a matched subset
   should be retained as a historical bridge. This will show which original
   findings were robust and which depended on the induced covariance geometry.
   Results from legacy and revised calibrations should be labeled explicitly
   and never pooled into a single reliability effect.

6. **Do not treat $\tau_1^2/\tau_0^2=1$ as proven optimal.** Study 5 used one
   regular covariance shape, cluster sizes of 10 and 25, and nonnegative
   correlations. The revised studies should use this shape for the primary
   comparison but include a smaller sensitivity set with variance ratios such
   as 0.5 and 2.0. Negative correlation, cluster sizes of 3 and 5, and the .50
   and .80 reliability targets also require explicit calibration checks because
   Study 5 did not test them.

## Study-specific recommendations

| Study | Structural question | Recommended primary calibration | Revision needed |
|:--|:--|:--|:--|
| **1: BLUP as outcome** | A cluster-level variable predicts the latent slope | **Marginal slope reliability**, with residualized reliability reported as a diagnostic | Fix the marginal $G$ shape and solve for $\sigma$ before decomposing the slope into the part predicted by $w$ and its residual. Verify the marginal $G$ again after the decomposition. |
| **2: BLUP as predictor** | The latent slope predicts an observed outcome while controlling for the latent intercept | **Residualized-slope reliability**, with marginal reliability retained for comparison | Replace the original `solve-G_22` ($\tau_1^2$) grid with the shape-preserving design. Retain matched marginal- and residualized-reliability arms in the low-reliability, correlated conditions because Study 5 showed that their difference matters for MSEM. |
| **3: BLUP as predictor and outcome** | One latent slope predicts another, conditional on the first process's latent intercept | **Residualized reliability for the predictor slope; marginal reliability for the outcome slope** | Calibrate the two first-stage processes separately with fixed covariance shapes. Then add the structural relation and verify that the generated outcome process retains its target marginal covariance. Include a focused sensitivity arm that swaps the reliability definitions because this asymmetric calibration was not tested in Study 5. |
| **4: heterogeneous cluster information** | Methods are compared as cluster-specific information becomes more heterogeneous | **Residualized-slope reliability for the focal predictor**, matched on its weighted mean across information profiles | Hold $G$ fixed across the homogeneous, moderate, severe, and information-matched profiles. Solve for a profile-specific $\sigma$ to equate mean reliability, while allowing the intended within-condition reliability dispersion to differ. This isolates heterogeneity from covariance-shape changes. |

### Study 1: preserve the meaning of a slope outcome

Study 1 differs conceptually from Study 2 because the latent slope is the
outcome of the structural regression rather than one member of a correlated
predictor pair. Marginal slope reliability is therefore the more direct primary
target. The calibration should begin with a fixed marginal $G$ and solve for
$\sigma$; only afterward should the standardized effect of $w$ be used to split
the total slope into explained and residual variation. This ordering preserves
the population slope distribution that the first-stage model is intended to
recover. Residualized reliability remains important as a diagnostic because a
correlated random intercept can still influence posterior slope uncertainty.

The revised report should reassess the original claims about attenuation,
Fuller overcoverage, 2S-PA standard errors, and agreement with the direct MLM.
The direct MLM and oracle should remain distinct benchmarks: the oracle checks
the structural target when the latent slope is observed, whereas the direct MLM
checks whether a correctly specified one-stage analysis recovers it from the
repeated measures.

### Study 2: make the estimand-aligned arm primary

Study 2 is the closest direct beneficiary of Study 5. Because the structural
model includes both the random intercept and random slope as predictors, the
focal slope coefficient concerns slope variation remaining after the intercept
is accounted for. The fixed-shape residualized-reliability arm should therefore
be the primary revised DGM, with the fixed-shape marginal arm retained as a
planned sensitivity analysis. The original arm should appear only as a legacy
comparison.

The estimator hierarchy should follow the Study 5 conclusions: use the oracle
as the calibration anchor, MSEM as the primary feasible estimator to evaluate,
Fuller as the stable point-estimation comparator, 2S-PA as the comparison with
Lai that requires joint reporting of eligibility and performance, and naive
dual-BLUP regression as a negative control. The full Study 2 grid should not be
rerun until a smaller bridge reproduces the Study 5 pattern at cluster sizes 3
and 5, negative $\rho$, and reliability targets .50 and .80.

### Study 3: calibrate the two processes according to their roles

Study 3 has two measurement problems, so applying one reliability definition
to both processes is unnecessarily restrictive. On the predictor side, the
coefficient on the first process's slope is conditional on its intercept;
residualized-slope reliability is consequently the most aligned primary target.
On the outcome side, the total slope of the second process is the dependent
latent variable, making its marginal slope reliability the natural starting
target. Both processes should have fixed covariance shapes, and their residual
scales should be solved separately for the requested information levels.

The structural equation that generates the second slope must then be applied
without changing its intended marginal covariance. This verification is
especially important because its residual slope variance depends on the
structural effect. A focused factorial check should compare the recommended
asymmetric calibration with all-marginal and all-residualized alternatives
before the large Study 3 grid is launched. The comparison will determine
whether the estimand-aligned choice improves calibration or whether joint-model
behavior requires a different definition.

### Study 4: isolate heterogeneity rather than regenerate the Study 2 confound

Study 4 is intended to vary the dispersion of cluster information while
holding mean reliability comparable. Its revised calibration should therefore
hold the entire $G$ matrix fixed across information profiles. For each profile
and target mean reliability, solve one common residual scale for that profile.
The cluster-size pattern then creates the intended distribution of individual
reliabilities, while covariance geometry no longer changes as a by-product of
matching their mean.

The information-matched falsification condition remains valuable and should be
preserved. Its calibration audit should demonstrate both comparable mean
residualized reliability and reduced dispersion of the full measurement
matrices, not only equality of trial-count means. The revised analysis should
also retain row-specific versus average-measurement comparisons, because those
are the study's substantive focus, but add the same eligibility, Type I error,
and tail-sensitive diagnostics used in Study 5.

## A staged rerun rather than an immediate full replacement

The most efficient return to Studies 1--4 is a two-stage process.

1. **Run calibration bridges first.** For each study, select the
   low-reliability stress conditions, at least one null and one non-null effect,
   $\rho=0$ and a correlated condition, and both the smallest and a moderate
   cluster size. Compare the legacy, fixed-shape marginal, and estimand-aligned
   calibration arms. Study 3 should instead compare legacy, all-marginal,
   all-residualized, and the recommended asymmetric calibration. Use common
   random-number streams across matched arms where practical so arm differences have
   less Monte Carlo noise.

2. **Require design checks before estimator comparisons.** Confirm the target
   reliability, fixed covariance geometry, positive-definite structural
   residual covariance, and nominal oracle behavior. If the oracle fails, the
   DGM or target scaling should be corrected before conclusions about proxy
   methods are drawn.

3. **Then rerun the full revised grids.** Once the bridge establishes that the
   revised DGM behaves as intended at the difficult boundaries, extend it to
   all cluster counts, effect sizes, correlations, and reliability targets.
   Save these as versioned revised studies rather than overwriting the original
   results. The original runs remain useful evidence about what happens under
   the one-coordinate calibration, but they should not carry the main
   generalizations about low reliability.

4. **Use one inferential reporting template across studies.** For null effects,
   report cell-specific total and directional Type I error, eligibility, and
   bias with Monte Carlo intervals. For non-null effects, report bias, coverage,
   direct power, interval width, eligibility, and joint success-and-cover or
   success-and-reject probabilities. Include empirical SD and tail summaries so
   a few unstable estimates cannot be mistaken for typical performance.

5. **Interpret changes, not just revised rankings.** Each revised report should
   state whether a conclusion from the original study is preserved, weakened,
   reversed, or shown to be calibration-dependent. This comparison is the main
   scientific benefit of retaining the legacy bridge: it reveals which claims
   concern score reliability and which concern the particular covariance
   geometry previously used to create it.

The immediate priority should be a Study 2 revision, because Study 5 already
provides its bridge and identifies residualized reliability as consequential
when the intercept and slope are correlated. Study 4 should follow because its
central claim requires reliability heterogeneity to be separated cleanly from
changes in $G$. Studies 1 and 3 then require smaller calibration pilots before
full reruns: Study 1 because its slope is an outcome, and Study 3 because the
predictor and outcome processes call for different, not yet directly validated,
reliability targets.
