---
title: "VH Study 1v2: Random Slope as a Structural Outcome"
author: "Michael Hallquist and Zach Vig"
date: "26 August 2026"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    embed-resources: true
    html-math-method: katex
params:
  output_dir: "outputs/vig_hallquist_v2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 120, out.width = "100%"
)

suppressPackageStartupMessages({
  library(data.table)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
  library(scales)
})

input_path <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) "")
analysis_dir <- if (nzchar(input_path)) {
  dirname(normalizePath(input_path, mustWork = FALSE))
} else {
  getwd()
}
repo_root <- normalizePath(file.path(analysis_dir, "..", ".."), mustWork = TRUE)
resolve_repo_path <- function(path) {
  if (grepl("^/", path)) path else file.path(repo_root, path)
}
source(
  file.path(repo_root, "vig_hallquist_2026", "vh_designs.R"),
  local = knitr::knit_global()
)

output_dir <- resolve_repo_path(params$output_dir)
condition_dir <- file.path(output_dir, "conditions")
canonical_counts <- study_condition_counts()
canonical_offsets <- c(0L, cumsum(canonical_counts)[-length(canonical_counts)])
names(canonical_offsets) <- names(canonical_counts)
expected_ids <- canonical_offsets[["study1v2"]] +
  seq_len(canonical_counts[["study1v2"]])
expected_condition_count <- length(expected_ids)
summary_paths <- file.path(
  condition_dir,
  sprintf("condition_%04d_summary.csv", expected_ids)
)
missing_summary_paths <- summary_paths[!file.exists(summary_paths)]
if (length(missing_summary_paths) > 0L) {
  stop(
    "Study 1v2 is incomplete: missing ", length(missing_summary_paths),
    " condition summaries. First missing file: ",
    missing_summary_paths[[1]]
  )
}

cell_metrics <- data.table::rbindlist(
  lapply(summary_paths, data.table::fread),
  fill = TRUE,
  use.names = TRUE
) %>%
  tibble::as_tibble() %>%
  filter(.data$study == "study1v2") %>%
  mutate(
    sd_standardized_error = suppressWarnings(
      as.numeric(.data$sd_standardized_error)
    ),
    mean_mx_latent_covariance_min_eigenvalue = suppressWarnings(
      as.numeric(.data$mean_mx_latent_covariance_min_eigenvalue)
    )
  )

expected_method_names <- c(
  "oracle", "naive_blup", "closed_form", "fuller_closed_form",
  "fuller_book_preliminary_closed_form", "fuller_book_closed_form",
  "fuller_alpha_stepdown_closed_form", "single_subject_ols", "lai_2spa",
  "direct_mlm"
)
completed_ids <- sort(unique(cell_metrics$condition_id))
missing_ids <- setdiff(expected_ids, completed_ids)
unexpected_ids <- setdiff(completed_ids, expected_ids)
expected_method_cells <- tidyr::crossing(
  condition_id = expected_ids,
  method = expected_method_names
)
observed_method_cells <- cell_metrics %>%
  count(.data$condition_id, .data$method, name = "summary_rows")
missing_method_cells <- dplyr::anti_join(
  expected_method_cells,
  observed_method_cells,
  by = c("condition_id", "method")
)
unexpected_method_cells <- dplyr::anti_join(
  observed_method_cells,
  expected_method_cells,
  by = c("condition_id", "method")
)
v2_contract_ok <-
  n_distinct(cell_metrics$condition_id) == expected_condition_count &&
  length(missing_ids) == 0L &&
  length(unexpected_ids) == 0L &&
  nrow(missing_method_cells) == 0L &&
  nrow(unexpected_method_cells) == 0L &&
  all(observed_method_cells$summary_rows == 1L) &&
  all(cell_metrics$n_rep == 1000L) &&
  all(cell_metrics$study_version == "v2") &&
  all(cell_metrics$calibration_version == "shape_preserving_v2") &&
  all(cell_metrics$covariance_shape_fixed %in% TRUE) &&
  all(cell_metrics$slope_intercept_variance_ratio == 1)
if (!v2_contract_ok) {
  stop(
    "Refusing to render: Study 1v2 summaries violate the v2-only ",
    "completion, method, replication, or calibration contract. Details: ",
    n_distinct(cell_metrics$condition_id), " conditions; ",
    length(missing_ids), " missing IDs; ",
    length(unexpected_ids), " unexpected IDs; ",
    nrow(missing_method_cells), " missing method cells; ",
    nrow(unexpected_method_cells), " unexpected method cells; ",
    sum(observed_method_cells$summary_rows != 1L),
    " duplicated method cells; n_rep range ",
    paste(range(cell_metrics$n_rep), collapse = "--")
  )
}
summary_file_info <- file.info(summary_paths)
summary_artifact_oldest <- min(summary_file_info$mtime)
summary_artifact_newest <- max(summary_file_info$mtime)

method_labels <- c(
  oracle = "Oracle",
  naive_blup = "Naive BLUP",
  closed_form = "Unshrunk-score OLS",
  fuller_closed_form = "Fuller (stabilized)",
  fuller_book_preliminary_closed_form = "Fuller (book preliminary)",
  fuller_book_closed_form = "Fuller (book equations)",
  fuller_alpha_stepdown_closed_form = "Fuller (alpha-stepdown)",
  single_subject_ols = "Single-subject OLS",
  lai_2spa = "Lai 2S-PA",
  direct_mlm = "Direct MLM"
)
stopifnot(setequal(names(method_labels), expected_method_names))
focal_methods <- c(
  "oracle", "naive_blup", "closed_form",
  "fuller_closed_form", "lai_2spa", "direct_mlm"
)
corrected_plot_methods <- c(
  "oracle", "closed_form", "fuller_closed_form", "lai_2spa", "direct_mlm"
)

cell_metrics <- cell_metrics %>%
  mutate(
    method_label = factor(
      unname(method_labels[.data$method]),
      levels = unname(method_labels)
    ),
    effect_label = factor(
      paste0("beta = ", .data$standardized_beta_target),
      levels = paste0("beta = ", c(0, .2, .4, .6))
    ),
    reliability_label = factor(
      paste0("R = ", sprintf("%.2f", .data$target_reliability)),
      levels = paste0("R = ", sprintf("%.2f", c(.25, .5, .8)))
    ),
    rho_label = factor(
      paste0("rho = ", .data$marginal_rho),
      levels = paste0("rho = ", c(-.5, 0, .5))
    ),
    n_label = factor(
      paste0("N = ", .data$num_clus),
      levels = paste0("N = ", c(30, 50, 100, 150, 300))
    ),
    m_label = factor(
      paste0("m = ", .data$mean_clus_size),
      levels = paste0("m = ", c(3, 5, 10, 25))
    )
  )

pooled_mc_se <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) == 0L) return(NA_real_)
  sqrt(sum(x^2)) / length(x)
}
fmt_pct <- function(x, accuracy = .1) percent(x, accuracy = accuracy)
fmt_num <- function(x, digits = 3) formatC(x, digits = digits, format = "f")

theme_vh <- function() {
  theme_bw(base_size = 10) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      strip.background = element_rect(fill = "grey95", colour = "grey75")
    )
}
reliability_palette <- c(
  "R = 0.25" = "#E15759",
  "R = 0.50" = "#F28E2B",
  "R = 0.80" = "#4E79A7"
)

method_effect_summary <- cell_metrics %>%
  filter(.data$method %in% focal_methods) %>%
  group_by(.data$method, .data$method_label, .data$standardized_beta_target) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    mean_bias = mean(.data$bias, na.rm = TRUE),
    bias_mc_se = pooled_mc_se(.data$mc_se_bias),
    bias_mc_low = .data$mean_bias - qnorm(.975) * .data$bias_mc_se,
    bias_mc_high = .data$mean_bias + qnorm(.975) * .data$bias_mc_se,
    mean_absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    coverage = mean(.data$coverage, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd, na.rm = TRUE
    ),
    cells_below_nominal = sum(.data$coverage_mc_high < .95, na.rm = TRUE),
    cells_above_nominal = sum(.data$coverage_mc_low > .95, na.rm = TRUE),
    .groups = "drop"
  )

nonnull_summary <- cell_metrics %>%
  filter(
    .data$method %in% focal_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(.data$method, .data$method_label) %>%
  summarise(
    bias = mean(.data$bias, na.rm = TRUE),
    coverage = mean(.data$coverage, na.rm = TRUE),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

metric_for <- function(summary, method, metric) {
  value <- summary[summary$method == method, metric, drop = TRUE]
  if (length(value) != 1L) NA_real_ else value
}

coverage_grid <- cell_metrics %>%
  filter(
    .data$method %in% focal_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$effect_label,
    .data$standardized_beta_target, .data$target_reliability,
    .data$reliability_label, .data$num_clus
  ) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    coverage = mean(.data$coverage, na.rm = TRUE),
    coverage_mc_se = pooled_mc_se(.data$coverage_mc_se),
    coverage_mc_low = .data$coverage - qnorm(.975) * .data$coverage_mc_se,
    coverage_mc_high = .data$coverage + qnorm(.975) * .data$coverage_mc_se,
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    .groups = "drop"
  )

bias_grid <- cell_metrics %>%
  filter(
    .data$method %in% focal_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$effect_label,
    .data$standardized_beta_target, .data$target_reliability,
    .data$reliability_label, .data$num_clus
  ) %>%
  summarise(
    cells = n(),
    mean_bias = mean(.data$bias, na.rm = TRUE),
    bias_mc_se = pooled_mc_se(.data$mc_se_bias),
    bias_mc_low = .data$mean_bias - qnorm(.975) * .data$bias_mc_se,
    bias_mc_high = .data$mean_bias + qnorm(.975) * .data$bias_mc_se,
    .groups = "drop"
  )

null_summary <- cell_metrics %>%
  filter(
    .data$method %in% focal_methods,
    .data$standardized_beta_target == 0
  ) %>%
  group_by(.data$method, .data$method_label) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    mean_null_bias = mean(.data$bias, na.rm = TRUE),
    conditional_type1 = mean(.data$type1_error, na.rm = TRUE),
    operational_type1 = mean(.data$operational_type1_error, na.rm = TRUE),
    positive_type1 = mean(.data$type1_error_positive, na.rm = TRUE),
    negative_type1 = mean(.data$type1_error_negative, na.rm = TRUE),
    cells_liberal = sum(.data$type1_error_mc_low > .05, na.rm = TRUE),
    cells_conservative = sum(.data$type1_error_mc_high < .05, na.rm = TRUE),
    .groups = "drop"
  )

null_grid <- cell_metrics %>%
  filter(
    .data$method %in% focal_methods,
    .data$standardized_beta_target == 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$target_reliability,
    .data$reliability_label, .data$num_clus
  ) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    type1_error = mean(.data$type1_error, na.rm = TRUE),
    type1_mc_se = pooled_mc_se(.data$type1_error_mc_se),
    type1_mc_low = pmax(0, .data$type1_error - qnorm(.975) * .data$type1_mc_se),
    type1_mc_high = pmin(1, .data$type1_error + qnorm(.975) * .data$type1_mc_se),
    operational_type1 = mean(.data$operational_type1_error, na.rm = TRUE),
    .groups = "drop"
  )

se_grid <- cell_metrics %>%
  filter(.data$method %in% focal_methods) %>%
  group_by(
    .data$method, .data$method_label, .data$target_reliability,
    .data$reliability_label, .data$num_clus
  ) %>%
  summarise(
    cells = n(),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd, na.rm = TRUE
    ),
    median_standardized_error_sd = median(
      .data$sd_standardized_error, na.rm = TRUE
    ),
    p95_standardized_error_sd = quantile(
      .data$sd_standardized_error, .95, na.rm = TRUE
    ),
    .groups = "drop"
  )

population_calibration <- cell_metrics %>%
  filter(.data$method == "oracle") %>%
  distinct(
    .data$mean_clus_size, .data$target_reliability, .data$marginal_rho,
    .data$achieved_reliability, .data$achieved_partial_reliability,
    .data$intercept_icc, .data$conditional_slope_icc,
    .data$slope_variance_marginal, .data$residualized_slope_variance,
    .data$slope_intercept_variance_ratio, .data$G_condition_number
  )

stage1_summary <- cell_metrics %>%
  filter(.data$method == "oracle") %>%
  group_by(.data$target_reliability, .data$marginal_rho) %>%
  summarise(
    fitted_marginal_R = mean(
      .data$mean_stage1_fitted_marginal_slope_posterior_reliability_mean,
      na.rm = TRUE
    ),
    fitted_residualized_R = mean(
      .data$mean_stage1_fitted_residualized_slope_posterior_reliability_mean,
      na.rm = TRUE
    ),
    true_blup_r = mean(
      .data$mean_stage1_true_blup_slope_correlation, na.rm = TRUE
    ),
    true_blup_r2 = mean(
      .data$mean_stage1_true_blup_slope_r_squared, na.rm = TRUE
    ),
    blup_bias = mean(.data$mean_stage1_blup_slope_bias, na.rm = TRUE),
    blup_rmse = mean(.data$mean_stage1_blup_slope_rmse, na.rm = TRUE),
    corrected_bias = mean(
      .data$mean_stage1_corrected_score_slope_bias, na.rm = TRUE
    ),
    corrected_rmse = mean(
      .data$mean_stage1_corrected_score_slope_rmse, na.rm = TRUE
    ),
    .groups = "drop"
  )

naive_attenuation <- cell_metrics %>%
  filter(
    .data$method == "naive_blup",
    .data$standardized_beta_target > 0
  ) %>%
  mutate(
    relative_bias = .data$bias / .data$standardized_beta_target,
    relative_bias_mc_se = .data$mc_se_bias /
      .data$standardized_beta_target
  ) %>%
  group_by(
    .data$target_reliability, .data$reliability_label, .data$num_clus
  ) %>%
  summarise(
    mean_relative_bias = mean(.data$relative_bias, na.rm = TRUE),
    relative_bias_mc_se = pooled_mc_se(.data$relative_bias_mc_se),
    relative_bias_mc_low = .data$mean_relative_bias -
      qnorm(.975) * .data$relative_bias_mc_se,
    relative_bias_mc_high = .data$mean_relative_bias +
      qnorm(.975) * .data$relative_bias_mc_se,
    coverage = mean(.data$coverage, na.rm = TRUE),
    .groups = "drop"
  )

naive_reliability_summary <- cell_metrics %>%
  filter(
    .data$method == "naive_blup",
    .data$standardized_beta_target > 0
  ) %>%
  mutate(relative_bias = .data$bias / .data$standardized_beta_target) %>%
  group_by(.data$target_reliability) %>%
  summarise(
    relative_bias = mean(.data$relative_bias, na.rm = TRUE),
    coverage = mean(.data$coverage, na.rm = TRUE),
    .groups = "drop"
  )

oracle_n_summary <- cell_metrics %>%
  filter(
    .data$method == "oracle",
    .data$standardized_beta_target > 0
  ) %>%
  group_by(.data$num_clus) %>%
  summarise(coverage = mean(.data$coverage, na.rm = TRUE), .groups = "drop")

spa_operational <- cell_metrics %>%
  filter(.data$method == "lai_2spa") %>%
  group_by(
    .data$num_clus, .data$target_reliability, .data$reliability_label
  ) %>%
  summarise(
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    conditional_coverage = mean(.data$coverage, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    latent_boundary_rate = mean(
      .data$mean_mx_latent_covariance_boundary, na.rm = TRUE
    ),
    .groups = "drop"
  )

spa_rho_summary <- cell_metrics %>%
  filter(.data$method == "lai_2spa") %>%
  group_by(.data$marginal_rho) %>%
  summarise(
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    .groups = "drop"
  )

type1_identity_error <- cell_metrics %>%
  filter(
    .data$method %in% focal_methods,
    .data$standardized_beta_target == 0
  ) %>%
  summarise(maximum = max(abs(.data$type1_error - (1 - .data$coverage)))) %>%
  pull(.data$maximum)

fuller_variants <- c(
  "fuller_closed_form",
  "fuller_book_preliminary_closed_form",
  "fuller_book_closed_form",
  "fuller_alpha_stepdown_closed_form"
)
fuller_variant_metrics <- cell_metrics %>%
  filter(.data$method %in% fuller_variants) %>%
  select(
    .data$condition_id, .data$method, .data$bias, .data$coverage,
    .data$mean_se, .data$interval_eligibility
  ) %>%
  pivot_wider(
    names_from = .data$method,
    values_from = c(
      .data$bias, .data$coverage, .data$mean_se,
      .data$interval_eligibility
    )
  )
row_spread <- function(data, prefix) {
  selected <- data %>% select(starts_with(prefix))
  max(apply(selected, 1, function(x) max(x) - min(x)), na.rm = TRUE)
}
fuller_variant_max_difference <- max(
  row_spread(fuller_variant_metrics, "bias_"),
  row_spread(fuller_variant_metrics, "coverage_"),
  row_spread(fuller_variant_metrics, "mean_se_"),
  row_spread(fuller_variant_metrics, "interval_eligibility_")
)
```

# Executive summary

Study 1v2 supports the central design decision from the preliminary
slope-calibration investigation and substantially sharpens the original Study
1 conclusions. That investigation was an internal design diagnostic, not a
study intended for publication. The revised simulation holds the
random intercept--slope covariance shape fixed at
$\tau_1^2/\tau_0^2=1$, targets marginal posterior slope reliability by changing
the Level-1 residual scale, and then introduces the structural effect of the
Level-2 predictor $w$. This removes the legacy confounding of low reliability
with a nearly degenerate random-effects covariance matrix.

The repaired design behaves well. Oracle, unshrunk-score OLS, stabilized Fuller,
and the direct MLM are essentially centered. Their non-null conditional
coverage ranges from
`r fmt_pct(min(nonnull_summary$coverage[nonnull_summary$method %in% c("oracle", "closed_form", "fuller_closed_form", "direct_mlm")]))`
to
`r fmt_pct(max(nonnull_summary$coverage[nonnull_summary$method %in% c("oracle", "closed_form", "fuller_closed_form", "direct_mlm")]))`,
their typical model SEs closely match empirical sampling SDs, and their null
Type I error ranges from
`r fmt_pct(min(null_summary$conditional_type1[null_summary$method %in% c("oracle", "closed_form", "fuller_closed_form", "direct_mlm")]))`
to
`r fmt_pct(max(null_summary$conditional_type1[null_summary$method %in% c("oracle", "closed_form", "fuller_closed_form", "direct_mlm")]))`.
Most of the small residual
liberality occurs at $N=30$ and is also present for the oracle, so it is a
finite-cluster benchmark issue rather than evidence that Stage-1 correction
failed.

Three method-specific conclusions are especially clear.

1. **Naive BLUP regression remains a decisive negative control.** Its non-null
   relative bias is
   `r fmt_pct(naive_reliability_summary$relative_bias[naive_reliability_summary$target_reliability == .25])`
   at $R=.25$,
   `r fmt_pct(naive_reliability_summary$relative_bias[naive_reliability_summary$target_reliability == .50])`
   at $R=.50$, and
   `r fmt_pct(naive_reliability_summary$relative_bias[naive_reliability_summary$target_reliability == .80])`
   at $R=.80$. Conditional coverage deteriorates as $N$ grows because the shrinkage bias
   persists while intervals narrow.
2. **The Fuller overcoverage reported in the legacy Study 1 analysis does not
   recur.** Stabilized Fuller has
   `r fmt_pct(metric_for(nonnull_summary, "fuller_closed_form", "coverage"))`
   non-null conditional coverage and
   `r fmt_pct(metric_for(null_summary, "fuller_closed_form", "conditional_type1"))`
   null Type I
   error. In this outcome-error setting, the corrected predictor--outcome
   measurement covariance is zero and all four Fuller variants are numerically
   identical.
3. **Lai 2S-PA is conditionally close to nominal but selectively available.**
   Its non-null conditional coverage is
   `r fmt_pct(metric_for(nonnull_summary, "lai_2spa", "coverage"))`, but
   non-null interval eligibility is only
   `r fmt_pct(metric_for(nonnull_summary, "lai_2spa", "eligibility"))`.
   At $R=.25$, eligibility is much lower, especially with
   30--50 clusters.

The substantive design conclusion is therefore balanced: posterior reliability
is a useful primary information factor once covariance shape is fixed, but
estimator behavior still depends on whether the noisy latent score is used as
an outcome or as a predictor. Study 1 should not inherit the conservative
Fuller and 2S-PA behavior observed in the preliminary investigation's
predictor-error data-generating model.

# Scientific motivation and relation to the calibration repair

Study 1 represents an applied workflow in which a baseline or person-level
covariate is used to explain heterogeneity in a within-person process. Examples
include asking whether a trait predicts individual differences in affective
reactivity, learning rate, or symptom change. The scientific target is not the
noisy fitted slope itself; it is the between-person regression linking the
covariate to the latent slope. The study asks whether a convenient two-stage
analysis recovers that target and its uncertainty.

The preliminary slope-calibration investigation identified two separable
questions:

1. Does the reliability manipulation preserve a regular population covariance
   geometry?
2. Given a regular DGM, does the proposed estimator deliver centered estimates
   and calibrated intervals?

Study 1v2 answers both questions for a random slope used as a structural
**outcome**. Its primary population model can be summarized as

$$
y_{it} = u_{0i} + u_{1i}x_{it} + e_{it},
\qquad
u_{1i} = \gamma w_i + \eta_{1i}.
$$

The reported target is the standardized structural coefficient $\beta$,
equivalently $\gamma$ on the reporting scale used in the simulation. The
measurement process determines how precisely $u_{1i}$ is recovered, while the
structural decomposition preserves its marginal variance.

This role matters. Classical error in a dependent variable increases sampling
variability but does not mechanically attenuate a regression coefficient.
A BLUP is different because it is a systematically shrunken conditional
expectation. Regressing that shrunken outcome on $w$ attenuates the structural
coefficient. Unshrinking restores centering, after which the inferential
question is whether the SE accounts adequately for score uncertainty.

That investigation also emphasized residualized-slope reliability, $R_\eta$,
when the random intercept and slope jointly enter a structural regression. In
Study 1,
only the latent slope is the structural outcome. Marginal slope reliability is
therefore both the cross-study primary factor and the reliability closest to
the focal outcome. We still report $R_\eta$ as a diagnostic, but the design does
not require a separate matched-$R_\eta$ arm unless correlation-specific failure
appears.

# Design, methods, and completeness

The 720 conditions cross:

- $N=30,50,100,150,300$ clusters;
- $m=3,5,10,25$ balanced observations per cluster;
- target marginal slope reliability $R=.25,.50,.80$;
- random intercept--slope correlations $\rho=-.5,0,.5$; and
- standardized structural effects $\beta=0,.2,.4,.6$.

Each condition contains 1,000 Monte Carlo replications. The primary covariance
anchor is $\tau_0^2=\tau_1^2=.81$, so the intercept/slope variance ratio is one.
The covariance condition number is one at $\rho=0$ and three at
$|\rho|=.5$, far from the ill-conditioned legacy low-slope-variance designs.

The methods have distinct diagnostic roles:

- **Oracle:** regress the known true slope on $w$; this checks the structural
  truth and finite-$N$ interval benchmark.
- **Naive BLUP:** regress the Stage-1 empirical Bayes slope on $w$; this is the
  shrinkage negative control.
- **Unshrunk-score OLS and single-subject OLS:** use unshrunken noisy outcomes.
- **Fuller:** use the corrected score and its Stage-1 uncertainty, with zero
  predictor measurement variance and zero predictor--outcome measurement-error
  covariance.
- **Lai 2S-PA:** represent the estimated intercept and slope as latent variables
  in the Stage-2 OpenMx model.
- **Direct MLM:** estimate the $w \times x$ interaction in one mixed model; this
  is the feasible one-stage benchmark.

```{r completion-table}
tibble::tibble(
  check = c(
    "Expected Study 1v2 conditions",
    "Completed condition summaries",
    "Missing condition IDs",
    "Unexpected condition IDs",
    "Missing condition-method summaries",
    "Unexpected condition-method summaries",
    "Methods per condition",
    "Replications per condition",
    "Condition-ID range",
    "Summary artifacts read",
    "Oldest summary artifact (local time)",
    "Newest summary artifact (local time)"
  ),
  result = c(
    expected_condition_count,
    n_distinct(cell_metrics$condition_id),
    length(missing_ids),
    length(unexpected_ids),
    nrow(missing_method_cells),
    nrow(unexpected_method_cells),
    n_distinct(cell_metrics$method),
    paste(range(cell_metrics$n_rep), collapse = "--"),
    paste(range(cell_metrics$condition_id), collapse = "--"),
    length(summary_paths),
    format(summary_artifact_oldest, tz = "America/New_York"),
    format(summary_artifact_newest, tz = "America/New_York")
  )
) %>%
  kable(caption = "Production-output completeness checks.")
```

All subsequent summaries are computed directly from the exact condition files
listed above; no consolidated cache or legacy fallback is used. The render
stops if any file, method, replication count, version tag, calibration tag, or
fixed-shape invariant is wrong.

{{< include _vh_v2_metric_guide.qmd >}}

# Calibration and Stage-1 diagnostics

## Population manipulation check

```{r population-calibration-plot, fig.width=11, fig.height=7.5}
population_calibration %>%
  select(
    .data$mean_clus_size, .data$target_reliability, .data$marginal_rho,
    marginal = .data$achieved_reliability,
    residualized = .data$achieved_partial_reliability
  ) %>%
  pivot_longer(
    cols = c("marginal", "residualized"),
    names_to = "reliability_definition",
    values_to = "achieved_reliability"
  ) %>%
  mutate(
    reliability_definition = recode(
      .data$reliability_definition,
      marginal = "Marginal slope reliability",
      residualized = "Residualized slope reliability"
    ),
    rho_label = factor(
      paste0("rho = ", .data$marginal_rho),
      levels = paste0("rho = ", c(-.5, 0, .5))
    )
  ) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$achieved_reliability,
    colour = .data$reliability_definition,
    shape = .data$rho_label,
    group = interaction(.data$reliability_definition, .data$rho_label)
  )) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey45") +
  geom_line(alpha = .65) +
  geom_point(size = 2.2) +
  facet_wrap(~ mean_clus_size, labeller = label_both) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(
    breaks = c(.1, .25, .5, .8),
    labels = percent_format(accuracy = 1)
  ) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Population reliability achieved",
    colour = NULL,
    shape = NULL,
    title = "The v2 calibrator holds marginal reliability fixed across m",
    subtitle = "Residualized reliability is reported, not silently treated as the same target"
  ) +
  theme_vh()
```

Marginal reliability is attained to numerical precision for every $m$ and
$\rho$. Residualized reliability is lower when the intercept and slope
correlate. For example, at target $R=.25$ it ranges from approximately .156 to
.163 at $|\rho|=.5$, versus .25 at $\rho=0$. This divergence is expected and
does not indicate a failed marginal calibration.

The ICCs also change with $m$ even though posterior reliability does not. At
$R=.25$, the intercept ICC falls from about .11--.14 at $m=3$ to .01 at
$m=25$. This is the same distinction demonstrated by the ICC bridge: posterior
reliability incorporates the observation schedule, whereas a fixed
single-observation ICC does not hold person-specific slope precision constant.

## Fitted reliability and score recovery

The production summaries record quantities that were unavailable in the legacy
Study 1 report: fitted marginal and residualized reliability, empirical
true-slope/BLUP correlation and $R^2$, and BLUP versus corrected-score bias and
RMSE.

```{r stage1-diagnostics-table}
stage1_summary %>%
  mutate(
    target_reliability = fmt_num(.data$target_reliability),
    marginal_rho = fmt_num(.data$marginal_rho, 1),
    fitted_marginal_R = fmt_num(.data$fitted_marginal_R),
    fitted_residualized_R = fmt_num(.data$fitted_residualized_R),
    true_blup_r = fmt_num(.data$true_blup_r),
    true_blup_r2 = fmt_num(.data$true_blup_r2),
    blup_bias = fmt_num(.data$blup_bias),
    blup_rmse = fmt_num(.data$blup_rmse),
    corrected_bias = fmt_num(.data$corrected_bias),
    corrected_rmse = fmt_num(.data$corrected_rmse)
  ) %>%
  kable(
    col.names = c(
      "Target R", "rho", "Fitted marginal R", "Fitted residualized R",
      "Cor(true, BLUP)", "R2(true, BLUP)", "BLUP bias", "BLUP RMSE",
      "Corrected-score bias", "Corrected-score RMSE"
    ),
    caption = paste(
      "Stage-1 recovery diagnostics averaged across m, N, and structural",
      "effect conditions."
    )
  )
```

The fitted marginal reliability is close to its population target at $R=.50$
and $.80$, but ranges from
`r fmt_num(min(stage1_summary$fitted_marginal_R[stage1_summary$target_reliability == .25]), 3)`
to
`r fmt_num(max(stage1_summary$fitted_marginal_R[stage1_summary$target_reliability == .25]), 3)`
across $\rho$ when the population target is $.25$. The empirical
true-slope/BLUP $R^2$ ranges from
`r fmt_num(min(stage1_summary$true_blup_r2[stage1_summary$target_reliability == .25]), 3)`
to
`r fmt_num(max(stage1_summary$true_blup_r2[stage1_summary$target_reliability == .25]), 3)`
in those cells. The two
quantities are related but are not interchangeable: one is computed from the
fitted covariance and posterior-variance model, whereas the other is an
empirical score-recovery diagnostic.

Both BLUP and corrected scores are centered as estimates of the slope. The
cost of unshrinking appears in RMSE: at $R=.25$, corrected-score RMSE ranges
from
`r fmt_num(min(stage1_summary$corrected_rmse[stage1_summary$target_reliability == .25]), 2)`
to
`r fmt_num(max(stage1_summary$corrected_rmse[stage1_summary$target_reliability == .25]), 2)`,
compared with
`r fmt_num(min(stage1_summary$blup_rmse[stage1_summary$target_reliability == .25]), 2)`
to
`r fmt_num(max(stage1_summary$blup_rmse[stage1_summary$target_reliability == .25]), 2)`
for the BLUP. Study 1 therefore illustrates
the familiar bias--variance distinction at two levels. BLUPs predict individual
slopes efficiently, yet regression on them is attenuated; unshrunk scores are
noisier person-level estimates, yet their structural regression is centered.

# Bias

## Naive BLUP attenuation follows the reliability gradient

```{r naive-attenuation-plot, fig.width=9.5, fig.height=6.5}
expected_attenuation <- tibble::tibble(
  target_reliability = c(.25, .5, .8),
  expected_relative_bias = c(.25, .5, .8) - 1
)

ggplot(
  naive_attenuation,
  aes(
    x = .data$target_reliability,
    y = .data$mean_relative_bias,
    colour = factor(.data$num_clus),
    group = .data$num_clus
  )
) +
  geom_errorbar(
    aes(
      ymin = .data$relative_bias_mc_low,
      ymax = .data$relative_bias_mc_high
    ),
    width = .015,
    alpha = .35
  ) +
  geom_line(alpha = .65) +
  geom_point(size = 2) +
  geom_line(
    data = expected_attenuation,
    aes(
      x = .data$target_reliability,
      y = .data$expected_relative_bias
    ),
    inherit.aes = FALSE,
    linetype = 2,
    linewidth = .8,
    colour = "black"
  ) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Relative bias of naive BLUP regression",
    colour = "Clusters",
    title = "Naive BLUP attenuation is approximately the shrinkage predicted by R",
    subtitle = "Dashed line is relative bias R - 1; intervals are Monte Carlo CIs"
  ) +
  theme_vh()
```

The observed attenuation is slightly stronger than the simple $R-1$
approximation at the medium and high targets, but the theoretical pattern is
clear. Increasing $N$ does not remove it. Indeed, the bias becomes easier to
detect as the structural SE shrinks.

## Corrected and benchmark estimators

```{r corrected-bias-plot, fig.width=12, fig.height=10}
bias_grid %>%
  filter(.data$method %in% corrected_plot_methods) %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$mean_bias,
    ymin = .data$bias_mc_low,
    ymax = .data$bias_mc_high,
    colour = .data$reliability_label,
    group = .data$reliability_label
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey45") +
  geom_errorbar(width = 5, alpha = .28, linewidth = .35) +
  geom_line(alpha = .5) +
  geom_point(size = 1.7) +
  facet_grid(method_label ~ effect_label, scales = "free_y") +
  scale_x_continuous(breaks = c(30, 50, 100, 150, 300)) +
  scale_colour_manual(values = reliability_palette) +
  labs(
    x = "Number of clusters",
    y = "Mean bias with 95% Monte Carlo CI",
    colour = NULL,
    title = "Fuller, unshrunk-score OLS, and direct MLM are centered",
    subtitle = "Each point averages 12 cells across m and rho"
  ) +
  theme_vh()
```

The oracle, unshrunk-score OLS, Fuller, and direct MLM biases are practically
zero. Their condition-level mean estimates also agree closely across $\rho$,
including the cells where marginal and residualized reliability differ.

2S-PA has modest positive bias concentrated at low reliability and small $N$.
Its equal-cell mean bias rises from
`r fmt_num(method_effect_summary$mean_bias[method_effect_summary$method == "lai_2spa" & method_effect_summary$standardized_beta_target == .2])`
at $\beta=.2$ to
`r fmt_num(method_effect_summary$mean_bias[method_effect_summary$method == "lai_2spa" & method_effect_summary$standardized_beta_target == .6])`
at $\beta=.6$.
That is much smaller than naive BLUP attenuation, but it is systematic enough
that conditional coverage should be interpreted jointly with fit eligibility.

# Coverage

## Corrected methods and benchmarks

```{r corrected-coverage-plot, fig.width=12, fig.height=10}
coverage_grid %>%
  filter(.data$method %in% corrected_plot_methods) %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$coverage,
    ymin = .data$coverage_mc_low,
    ymax = .data$coverage_mc_high,
    colour = .data$reliability_label,
    group = .data$reliability_label
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey35") +
  geom_errorbar(width = 5, alpha = .25, linewidth = .35) +
  geom_line(alpha = .5) +
  geom_point(size = 1.7) +
  facet_grid(method_label ~ effect_label) +
  scale_x_continuous(breaks = c(30, 50, 100, 150, 300)) +
  scale_y_continuous(
    limits = c(.88, .985),
    labels = percent_format(accuracy = 1)
  ) +
  scale_colour_manual(values = reliability_palette) +
  labs(
    x = "Number of clusters",
    y = "Conditional 95% CI coverage",
    colour = NULL,
    title = "Study 1v2 coverage is close to nominal after the shape repair",
    subtitle = "Points average 12 m-by-rho cells; bars are pooled Monte Carlo CIs"
  ) +
  theme_vh()
```

The most important result is negative in a useful sense: there is no broad
low-reliability coverage collapse for the corrected methods. Across non-null
conditions, unshrunk-score OLS covers
`r fmt_pct(metric_for(nonnull_summary, "closed_form", "coverage"))`, Fuller
`r fmt_pct(metric_for(nonnull_summary, "fuller_closed_form", "coverage"))`,
the direct MLM
`r fmt_pct(metric_for(nonnull_summary, "direct_mlm", "coverage"))`, and the
oracle `r fmt_pct(metric_for(nonnull_summary, "oracle", "coverage"))`.
Conditional coverage for these methods also remains near nominal
within the $R=.25$ conditions.

The small average shortfall from 95% is visible for the oracle as well. Oracle
conditional coverage is
`r fmt_pct(oracle_n_summary$coverage[oracle_n_summary$num_clus == 30])`
at $N=30$ and
`r fmt_pct(oracle_n_summary$coverage[oracle_n_summary$num_clus == 300])`
at $N=300$. That pattern
sets the proper benchmark: a feasible method should be compared with the oracle
and direct MLM in the same finite-$N$ cell, not only with an abstract horizontal
line.

2S-PA's conditional coverage is also close to 95%, but this statement applies
only to successful intervals. Its success-and-cover probability is substantially
lower in the low-reliability cells, as shown later.

## Naive BLUP coverage

```{r naive-coverage-plot, fig.width=9.5, fig.height=6.5}
coverage_grid %>%
  filter(.data$method == "naive_blup") %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$coverage,
    colour = .data$reliability_label,
    group = .data$reliability_label
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey35") +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ effect_label, nrow = 1) +
  scale_x_continuous(breaks = c(30, 50, 100, 150, 300)) +
  scale_y_continuous(
    limits = c(0, 1),
    labels = percent_format(accuracy = 1)
  ) +
  scale_colour_manual(values = reliability_palette) +
  labs(
    x = "Number of clusters",
    y = "Conditional 95% CI coverage",
    colour = NULL,
    title = "More clusters expose rather than repair naive BLUP bias",
    subtitle = "Intervals narrow around an attenuated structural coefficient"
  ) +
  theme_vh()
```

At non-null effects, naive BLUP conditional coverage averages only
`r fmt_pct(metric_for(nonnull_summary, "naive_blup", "coverage"))`.
At $R=.25$ it is
`r fmt_pct(naive_reliability_summary$coverage[naive_reliability_summary$target_reliability == .25])`,
compared with
`r fmt_pct(naive_reliability_summary$coverage[naive_reliability_summary$target_reliability == .80])`
at $R=.80$. The worsening with larger $N$ is a
standard consistency warning: a biased estimator can have decreasing coverage
as its nominal precision increases.

# Standard errors and interval mechanism

## Reported SE scale versus empirical sampling variability

```{r se-calibration-plot, fig.width=11, fig.height=8}
se_grid %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$median_se_ratio,
    colour = factor(.data$num_clus),
    group = .data$num_clus
  )) +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey35") +
  geom_line(alpha = .55) +
  geom_point(size = 1.8) +
  facet_wrap(~ method_label, ncol = 3) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(
    limits = c(.4, 1.1),
    breaks = c(.5, .75, 1)
  ) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Median condition-level SE ratio",
    colour = "Clusters",
    title = "Typical SE calibration is excellent for the corrected Study 1 methods",
    subtitle = paste(
      "Each ratio is mean reported SE / empirical sampling SD;",
      "lines connect reliability levels within a fixed N"
    )
  ) +
  theme_vh()
```

Unshrunk-score OLS, Fuller, direct MLM, and oracle ratios cluster tightly around
one. This agrees with their near-nominal coverage. Fuller is not producing the
inflated typical SEs that generated overcoverage in the preliminary
slope-calibration investigation's predictor-error setting.

For naive BLUP, the ratio falls as the effect grows and reliability falls.
That ratio is not a complete explanation of its coverage failure because the
dominant problem is centering: even a perfectly estimated sampling SD would
produce an interval centered on the attenuated coefficient.

2S-PA's condition-level SE ratio is usually only slightly below one. However, a typical
ratio can conceal a non-normal standardized-error distribution and a selected
fit sample.

## Tail-sensitive standardized-error diagnostic

```{r standardized-error-tail-plot, fig.width=11, fig.height=7.5}
cell_metrics %>%
  filter(
    .data$method %in% corrected_plot_methods,
    is.finite(.data$sd_standardized_error)
  ) %>%
  ggplot(aes(
    x = factor(.data$num_clus),
    y = .data$sd_standardized_error,
    colour = .data$reliability_label
  )) +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey35") +
  geom_jitter(width = .18, height = 0, alpha = .35, size = 1) +
  facet_wrap(~ method_label, ncol = 3) +
  scale_y_log10(breaks = c(1, 1.5, 2, 3, 5)) +
  scale_colour_manual(values = reliability_palette) +
  labs(
    x = "Number of clusters",
    y = "SD of standardized errors (log scale)",
    colour = NULL,
    title = "2S-PA retains heavy standardized-error tails in difficult cells",
    subtitle = "Each point is a condition; a well-calibrated normal pivot has SD near one"
  ) +
  theme_vh()
```

The oracle, unshrunk-score OLS, Fuller, and direct MLM standardized-error SDs
remain close to one. For 2S-PA they reach
`r fmt_num(max(cell_metrics$sd_standardized_error[cell_metrics$method == "lai_2spa"], na.rm = TRUE), 2)`
in the most difficult cell, and
the largest values occur mainly at $R=.25$. These cells do not have enormous
99th-percentile SEs. The problem is instead a combination of selected fits,
bias, and a heavy or irregular studentized-error distribution. This is why mean
SE, SE tails, standardized errors, and eligibility all belong in the analysis.

## Separate centering from interval scale

```{r bias-se-mechanism-plot, fig.width=11, fig.height=7.5}
cell_metrics %>%
  filter(
    .data$method %in% focal_methods,
    .data$standardized_beta_target > 0
  ) %>%
  ggplot(aes(
    x = .data$bias,
    y = .data$mean_se_to_interval_empirical_sd,
    colour = .data$coverage,
    shape = .data$reliability_label
  )) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey45") +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey45") +
  geom_point(alpha = .55, size = 1.25) +
  facet_wrap(~ method_label, scales = "free_x", ncol = 3) +
  scale_colour_viridis_c(
    option = "C",
    limits = c(0, 1),
    labels = percent_format(accuracy = 1)
  ) +
  labs(
    x = "Condition-level mean bias",
    y = "Condition-level mean-SE / empirical-SD ratio",
    colour = "Conditional coverage",
    shape = NULL,
    title = "Bias and interval scale identify different failure mechanisms",
    subtitle = "Free x-scales keep naive BLUP attenuation from compressing the corrected methods"
  ) +
  theme_vh()
```

This plot makes the calibration investigation's diagnostic distinction
concrete for Study 1. Naive BLUP lies
far from zero bias and often below an SE ratio of one. The corrected methods
remain close to both reference lines. 2S-PA is only modestly off center and
modestly low in typical SE scale, yet its eligibility and standardized-error
tails remain problematic. Coverage alone would not separate these mechanisms.

# Type I error

At $\beta=0$, the null hypothesis is true. Under the two-sided CI-inversion rule,
Type I error equals one minus coverage among eligible intervals. The maximum
numerical discrepancy in the production summaries is zero.

```{r type1-plot, fig.width=11, fig.height=8}
null_grid %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$type1_error,
    ymin = .data$type1_mc_low,
    ymax = .data$type1_mc_high,
    colour = .data$reliability_label,
    group = .data$reliability_label
  )) +
  geom_hline(yintercept = .05, linetype = 2, colour = "grey35") +
  geom_errorbar(width = 5, alpha = .28, linewidth = .35) +
  geom_line(alpha = .55) +
  geom_point(size = 1.7) +
  facet_wrap(~ method_label, ncol = 3) +
  scale_x_continuous(breaks = c(30, 50, 100, 150, 300)) +
  scale_y_continuous(
    limits = c(.02, .09),
    labels = percent_format(accuracy = .1)
  ) +
  scale_colour_manual(values = reliability_palette) +
  labs(
    x = "Number of clusters",
    y = "Conditional Type I error",
    colour = NULL,
    title = "Null rejection is close to 5%, with mild small-N liberality",
    subtitle = "Points average 12 m-by-rho cells; bars are pooled Monte Carlo CIs"
  ) +
  theme_vh()
```

```{r null-summary-table}
null_summary %>%
  transmute(
    method = as.character(.data$method_label),
    eligibility = fmt_pct(.data$eligibility),
    mean_null_bias = fmt_num(.data$mean_null_bias),
    conditional_type1 = fmt_pct(.data$conditional_type1),
    operational_type1 = fmt_pct(.data$operational_type1),
    positive = fmt_pct(.data$positive_type1),
    negative = fmt_pct(.data$negative_type1),
    liberal_cells = .data$cells_liberal,
    conservative_cells = .data$cells_conservative
  ) %>%
  kable(
    col.names = c(
      "Method", "Eligibility", "Mean null bias", "Conditional Type I",
      "Operational Type I", "Positive", "Negative",
      "Cells above .05", "Cells below .05"
    ),
    caption = paste(
      "Null diagnostics across 180 cells per method. Cell counts use each",
      "cell's 95% Monte Carlo interval relative to .05."
    )
  )
```

The null results are reassuring but should not be summarized as simply
"coverage is about 95%." Mean null bias is negligible and positive and negative
rejections are balanced. Fuller is neither strongly conservative nor liberal:
its Type I error is
`r fmt_pct(metric_for(null_summary, "fuller_closed_form", "conditional_type1"))`.
The oracle, unshrunk-score OLS, direct MLM, and naive
BLUP are similarly close to nominal at the null.

Naive BLUP's acceptable null test does not rescue it. Shrinkage toward zero
cannot attenuate a coefficient whose truth is already zero, so the null cells
do not reveal its severe non-null bias. This is an important reason to report
Type I error and non-null coverage separately.

2S-PA has conditional Type I error of
`r fmt_pct(metric_for(null_summary, "lai_2spa", "conditional_type1"))`, but
operational Type I error of only
`r fmt_pct(metric_for(null_summary, "lai_2spa", "operational_type1"))`.
The difference is not conservatism of the returned intervals; it
reflects replications that never return an eligible interval.

# Eligibility and 2S-PA fit selection

```{r spa-operational-plot, fig.width=10.5, fig.height=6.5}
spa_operational %>%
  select(
    .data$num_clus, .data$reliability_label,
    eligibility = .data$eligibility,
    conditional_coverage = .data$conditional_coverage,
    success_and_cover = .data$success_and_cover
  ) %>%
  pivot_longer(
    cols = c("eligibility", "conditional_coverage", "success_and_cover"),
    names_to = "metric",
    values_to = "rate"
  ) %>%
  mutate(
    metric = recode(
      .data$metric,
      eligibility = "Interval eligibility",
      conditional_coverage = "Conditional coverage",
      success_and_cover = "Success-and-cover"
    )
  ) %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$rate,
    colour = .data$metric,
    group = .data$metric
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey55") +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ reliability_label, nrow = 1) +
  scale_x_continuous(breaks = c(30, 50, 100, 150, 300)) +
  scale_y_continuous(
    limits = c(.35, 1),
    labels = percent_format(accuracy = 1)
  ) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = "Number of clusters",
    y = "Rate",
    colour = NULL,
    title = "2S-PA's nominal conditional coverage conceals low-R fit selection",
    subtitle = "Metrics average across m, rho, and effect conditions"
  ) +
  theme_vh()
```

At $N=30$ and $R=.25$, 2S-PA eligibility is only
`r fmt_pct(spa_operational$eligibility[spa_operational$num_clus == 30 & spa_operational$target_reliability == .25])`.
It rises to
`r fmt_pct(spa_operational$eligibility[spa_operational$num_clus == 300 & spa_operational$target_reliability == .25])`
at $N=300$, and is essentially complete at $R=.80$. Correlated intercepts and
slopes reduce eligibility further: averaged across the full grid, eligibility
is about
`r fmt_pct(mean(spa_rho_summary$eligibility[abs(spa_rho_summary$marginal_rho) == .5]))`
at $|\rho|=.5$ versus
`r fmt_pct(spa_rho_summary$eligibility[spa_rho_summary$marginal_rho == 0])`
at $\rho=0$.

The OpenMx latent-covariance boundary diagnostic follows the same gradient.
Here, $\Lambda$ is the Stage-1 linear map from the latent intercept--slope
vector to its BLUP score vector; its row/column orientation must match that
ordering. Correcting that orientation solved an implementation error, but it
does not eliminate the finite-sample identification problem of estimating a
latent slope variance from weak Stage-1 information. Conditional coverage
should remain paired with eligibility and success-and-cover in the paper.

# Direct comparison: stabilized Fuller, Lai 2S-PA, and Direct MLM

The preceding diagnostics establish how each estimator behaves on its own. A
publication comparison also needs to ask how the feasible alternatives perform
in the **same design conditions**. Stabilized Fuller and Lai 2S-PA are competing
two-stage corrections. The direct MLM fits the repeated-measures and structural
parts jointly and is therefore the most relevant feasible one-stage benchmark
for this outcome-error setting. The comparisons below match methods by
condition and judge bias, conditional coverage, interval availability, and SE
calibration separately.

```{r direct-comparison-setup, include=FALSE}
comparison_methods <- c("fuller_closed_form", "lai_2spa", "direct_mlm")
comparison_labels <- c(
  fuller_closed_form = "Fuller (stabilized)",
  lai_2spa = "Lai 2S-PA",
  direct_mlm = "Direct MLM"
)

comparison_nonnull <- cell_metrics %>%
  filter(
    .data$method %in% comparison_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(.data$method, .data$target_reliability) %>%
  summarise(
    cells = n(),
    mean_bias = mean(.data$bias, na.rm = TRUE),
    mean_absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    conditional_coverage = mean(.data$coverage, na.rm = TRUE),
    interval_eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

comparison_null <- cell_metrics %>%
  filter(
    .data$method %in% comparison_methods,
    .data$standardized_beta_target == 0
  ) %>%
  group_by(.data$method, .data$target_reliability) %>%
  summarise(
    conditional_type1 = mean(.data$type1_error, na.rm = TRUE),
    operational_type1 = mean(.data$operational_type1_error, na.rm = TRUE),
    .groups = "drop"
  )

comparison_summary <- comparison_nonnull %>%
  left_join(
    comparison_null,
    by = c("method", "target_reliability")
  ) %>%
  mutate(
    method_label = factor(
      unname(comparison_labels[.data$method]),
      levels = unname(comparison_labels)
    )
  )

comparison_metric <- function(method, reliability, metric) {
  value <- comparison_summary[
    comparison_summary$method == method &
      comparison_summary$target_reliability == reliability,
    metric,
    drop = TRUE
  ]
  if (length(value) != 1L) NA_real_ else value
}

comparison_wide <- cell_metrics %>%
  filter(
    .data$method %in% comparison_methods,
    .data$standardized_beta_target > 0
  ) %>%
  select(
    .data$condition_id, .data$target_reliability,
    .data$standardized_beta_target, .data$method, .data$bias,
    .data$coverage, .data$interval_eligibility,
    .data$mean_se_to_interval_empirical_sd
  ) %>%
  pivot_wider(
    names_from = .data$method,
    values_from = c(
      bias, coverage, interval_eligibility,
      mean_se_to_interval_empirical_sd
    )
  )

make_paired_comparison <- function(wide, comparator_key, comparator_label) {
  tibble::tibble(
    comparator = comparator_label,
    target_reliability = wide$target_reliability,
    absolute_bias_difference =
      abs(wide$bias_fuller_closed_form) -
      abs(wide[[paste0("bias_", comparator_key)]]),
    absolute_coverage_error_difference =
      abs(wide$coverage_fuller_closed_form - .95) -
      abs(wide[[paste0("coverage_", comparator_key)]] - .95),
    interval_failure_difference =
      (1 - wide$interval_eligibility_fuller_closed_form) -
      (1 - wide[[paste0("interval_eligibility_", comparator_key)]]),
    absolute_log_se_error_difference =
      abs(log(wide$mean_se_to_interval_empirical_sd_fuller_closed_form)) -
      abs(log(wide[[paste0(
        "mean_se_to_interval_empirical_sd_", comparator_key
      )]]))
  )
}

paired_comparisons <- bind_rows(
  make_paired_comparison(comparison_wide, "lai_2spa", "Lai 2S-PA"),
  make_paired_comparison(comparison_wide, "direct_mlm", "Direct MLM")
) %>%
  pivot_longer(
    cols = c(
      absolute_bias_difference,
      absolute_coverage_error_difference,
      interval_failure_difference,
      absolute_log_se_error_difference
    ),
    names_to = "criterion",
    values_to = "fuller_minus_comparator"
  ) %>%
  mutate(
    criterion = factor(
      .data$criterion,
      levels = c(
        "absolute_bias_difference",
        "absolute_coverage_error_difference",
        "interval_failure_difference",
        "absolute_log_se_error_difference"
      ),
      labels = c(
        "Absolute bias",
        "Absolute conditional-coverage error",
        "Interval-failure probability",
        "Absolute log(SE / empirical SD)"
      )
    )
  ) %>%
  filter(is.finite(.data$fuller_minus_comparator))
```

```{r direct-comparison-plot, fig.width=12, fig.height=9}
paired_comparisons %>%
  ggplot(aes(
    x = factor(.data$target_reliability),
    y = .data$fuller_minus_comparator,
    fill = .data$comparator
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey35") +
  geom_boxplot(outlier.alpha = .18, width = .65) +
  facet_grid(.data$criterion ~ .data$comparator, scales = "free_y") +
  scale_fill_manual(values = c(
    "Lai 2S-PA" = "#4E79A7",
    "Direct MLM" = "#59A14F"
  )) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Fuller minus comparator (negative favors Fuller)",
    fill = NULL,
    title = "Fuller closely tracks the direct MLM and is more available than 2S-PA",
    subtitle = paste(
      "Differences are computed within the same non-null design cell;",
      "zero denotes equal performance"
    )
  ) +
  theme_vh() +
  theme(legend.position = "none")
```

The comparison separates two conclusions that an overall coverage average
would blur. Fuller and the direct MLM are fully interval eligible and both are
close to centered, so their small matched-cell differences mainly describe
finite-sample implementation differences rather than different missing-fit
populations. Lai 2S-PA can be conditionally competitive, but at $R=.25$ its
eligibility is only
`r fmt_pct(comparison_metric("lai_2spa", .25, "interval_eligibility"))`.
Its conditional coverage therefore describes a selected subset, whereas
Fuller's success-and-cover probability at the same reliability is
`r fmt_pct(comparison_metric("fuller_closed_form", .25, "success_and_cover"))`.

```{r direct-comparison-table}
comparison_summary %>%
  transmute(
    Method = .data$method_label,
    Reliability = sprintf("%.2f", .data$target_reliability),
    `Mean bias` = round(.data$mean_bias, 3),
    `Mean |bias|` = round(.data$mean_absolute_bias, 3),
    `Conditional coverage` = percent(
      .data$conditional_coverage,
      accuracy = .1
    ),
    `Interval eligible` = percent(.data$interval_eligibility, accuracy = .1),
    `Success-and-cover` = percent(.data$success_and_cover, accuracy = .1),
    `Median condition-level SE ratio` = round(.data$median_se_ratio, 3),
    `Null Type I error` = percent(.data$conditional_type1, accuracy = .1),
    `Operational null rejection` = percent(
      .data$operational_type1,
      accuracy = .1
    )
  ) %>%
  arrange(.data$Reliability, .data$Method) %>%
  kable(align = c("l", "r", rep("r", 8)))
```

The appropriate comparative claim is not that Fuller uniformly dominates the
joint model. Rather, in Study 1 it reproduces the direct MLM's favorable bias,
coverage, and SE-calibration pattern while retaining a simple two-stage
workflow. Its clearest operational advantage over 2S-PA is complete interval
availability, especially at low reliability.

# Method-specific conclusions

## Oracle and direct MLM

The oracle is centered and close to nominal, validating the structural target.
Its mild $N=30$ liberality establishes the finite-sample baseline. The direct
MLM is also fully eligible and has nearly the same condition-level mean bias as
unshrunk-score OLS. Its SE differs modestly because it estimates the structural
interaction and random-effects model jointly. It is the most useful feasible
benchmark for Study 1.

## Unshrunk-score OLS and single-subject OLS

Unshrunk-score OLS is centered and has typical SE ratios near one. In the
balanced Study 1 design, its condition-level coverage, bias, and mean SE are
numerically identical to single-subject OLS. The additional closed-form
machinery therefore does not change the substantive result here; its importance
emerges in studies with more complicated score use or heterogeneous information.

## Fuller

Fuller is fully eligible, centered, and near nominal. This is a materially
different conclusion from the old Study 1 report and from the preliminary
slope-calibration investigation's predictor-error results. In Study 1 the
error-prone score is the outcome, $w$ is observed
without measurement error, and the predictor--outcome measurement-error
covariance is zero. The Fuller point correction therefore does not subtract an
outcome error variance from the predictor covariance, and the variance
regularization paths do not activate differently.

All four retained Fuller variants agree exactly across condition-level bias,
coverage, mean SE, and eligibility. For the primary Study 1 grid, retaining the
variants is useful as an algebra audit but not as four substantively distinct
estimators.

## Lai 2S-PA

2S-PA has near-nominal conditional coverage and only modest positive bias, but
its low-reliability intervals are available for a selected subset of
replications. Its standardized-error tails are also heavier than its mean
SE/SD ratio suggests. It remains a useful comparator to Lai and a check on the
corrected lambda ordering, but operational performance—not conditional coverage
alone—should govern its interpretation.

## Naive BLUP

Naive BLUP regression behaves exactly as the negative-control logic predicts.
It is approximately centered under the null and therefore has acceptable Type I
error, but it is attenuated whenever $\beta\ne0$. More clusters make its wrong
target more precisely estimated. It should not be presented as a candidate
structural estimator.

# Conclusions for the VH study series

Study 1v2 confirms the common calibration framework motivated by the
preliminary slope-calibration investigation.

1. **The shape-preserving posterior-reliability calibrator works.** The
   population marginal target is attained across $m$, the covariance condition
   number remains one or three, and the corrected estimators no longer show the
   broad coverage distortions attributed to the legacy design.
2. **Study role matters after calibration.** Fuller's strong overcoverage in
   the preliminary slope-calibration investigation's predictor-error setting
   is not a universal consequence of Fuller correction. It does not
   occur when the measured latent slope is an outcome and the structural
   predictor is error free. The Study 2 predictor-error problem cannot be
   generalized mechanically to Study 1.
3. **Marginal reliability remains the appropriate shared primary factor.**
   $R_\eta$ diverges when $\rho\ne0$, but oracle, Fuller, unshrunk-score OLS,
   and direct-MLM performance changes little across $\rho$. Study 1 supplies no
   reason to add a full residualized-reliability arm.
4. **The original attenuation conclusion survives; the original Fuller
   overcoverage conclusion does not.** Naive BLUP attenuation is strong and
   closely follows reliability. The corrected v2 Fuller intervals are close to
   nominal.
5. **The minimum reporting set motivated by the calibration investigation
   remains necessary.** Null Type I error, non-null coverage, bias with Monte
   Carlo uncertainty, model SE versus
   empirical SD, standardized-error tails, eligibility, and success-and-cover
   reveal different aspects of performance.

For a psychological methods audience, the cleanest summary is: posterior
reliability describes how well the latent outcome slope can be recovered under
the repeated-measures design; shape preservation ensures that its levels are
comparable; and the estimator must still be judged according to the role that
the recovered slope plays in the structural model. In Study 1, unshrinking plus
Fuller performs well, agrees closely with the direct MLM, and removes the
attenuation of naive BLUP regression. Lai 2S-PA reaches similar conditional
coverage but at a meaningful cost in low-information fit availability.

# Appendix

## Compact method-by-effect summary

```{r method-effect-summary-table}
method_effect_summary %>%
  mutate(
    effect = sprintf("%.1f", .data$standardized_beta_target),
    eligibility = fmt_pct(.data$eligibility),
    bias = fmt_num(.data$mean_bias),
    bias_mc_ci = paste0(
      "[", fmt_num(.data$bias_mc_low), ", ", fmt_num(.data$bias_mc_high), "]"
    ),
    coverage = fmt_pct(.data$coverage),
    success_and_cover = fmt_pct(.data$success_and_cover),
    se_ratio = fmt_num(.data$median_se_ratio),
    coverage_flags = paste0(
      .data$cells_below_nominal, " low / ",
      .data$cells_above_nominal, " high"
    )
  ) %>%
  select(
    method = .data$method_label, .data$effect, .data$eligibility,
    .data$bias, .data$bias_mc_ci, .data$coverage,
    .data$success_and_cover, .data$se_ratio, .data$coverage_flags
  ) %>%
  kable(
    col.names = c(
      "Method", "Beta", "Eligibility", "Mean bias", "95% MC CI",
      "Conditional coverage", "Success-and-cover",
      "Median condition-level SE ratio",
      "Cells differing from .95"
    ),
    caption = paste(
      "Equal-cell summaries across 180 conditions per effect and method.",
      "Coverage flags use condition-specific 95% Monte Carlo intervals."
    )
  )
```

## Correlation sensitivity

```{r rho-summary-table}
cell_metrics %>%
  filter(.data$method %in% corrected_plot_methods) %>%
  group_by(.data$method_label, .data$marginal_rho) %>%
  summarise(
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    nonnull_coverage = mean(
      .data$coverage[.data$standardized_beta_target > 0], na.rm = TRUE
    ),
    nonnull_bias = mean(
      .data$bias[.data$standardized_beta_target > 0], na.rm = TRUE
    ),
    null_type1 = mean(
      .data$type1_error[.data$standardized_beta_target == 0], na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  mutate(
    marginal_rho = fmt_num(.data$marginal_rho, 1),
    eligibility = fmt_pct(.data$eligibility),
    nonnull_coverage = fmt_pct(.data$nonnull_coverage),
    nonnull_bias = fmt_num(.data$nonnull_bias),
    null_type1 = fmt_pct(.data$null_type1)
  ) %>%
  kable(
    col.names = c(
      "Method", "rho", "Eligibility", "Non-null conditional coverage",
      "Non-null bias", "Null Type I"
    ),
    caption = "Sensitivity to the random intercept--slope correlation."
  )
```

## Fuller variants

```{r fuller-variant-table}
cell_metrics %>%
  filter(.data$method %in% fuller_variants) %>%
  group_by(.data$method_label) %>%
  summarise(
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    null_type1 = mean(
      .data$type1_error[.data$standardized_beta_target == 0], na.rm = TRUE
    ),
    nonnull_coverage = mean(
      .data$coverage[.data$standardized_beta_target > 0], na.rm = TRUE
    ),
    nonnull_bias = mean(
      .data$bias[.data$standardized_beta_target > 0], na.rm = TRUE
    ),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd, na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  mutate(
    eligibility = fmt_pct(.data$eligibility),
    null_type1 = fmt_pct(.data$null_type1),
    nonnull_coverage = fmt_pct(.data$nonnull_coverage),
    nonnull_bias = fmt_num(.data$nonnull_bias),
    median_se_ratio = fmt_num(.data$median_se_ratio)
  ) %>%
  kable(
    col.names = c(
      "Fuller variant", "Eligibility", "Null Type I",
      "Non-null conditional coverage", "Non-null bias",
      "Median condition-level SE ratio"
    ),
    caption = "Retained Fuller algebra variants; all are identical in Study 1v2."
  )
```

{{< include _vh_v2_provenance.qmd >}}
