---
title: "VH Study 3v2 Diagnostic: The Same Joint-Growth Model in Mplus and OpenMx"
author: "Simulation analysis"
date: "25 August 2026"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    embed-resources: true
    html-math-method: katex
params:
  output_dir: "outputs/vig_hallquist_v2/study3_joint_growth_openmx_audit"
  expected_n_rep: 100
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 120, out.width = "100%"
)
suppressPackageStartupMessages({
  library(data.table)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
  library(scales)
})

input_path <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) "")
analysis_dir <- if (nzchar(input_path)) {
  dirname(normalizePath(input_path, mustWork = FALSE))
} else {
  getwd()
}
repo_root <- normalizePath(file.path(analysis_dir, "..", ".."), mustWork = TRUE)
output_dir <- if (grepl("^/", params$output_dir)) {
  params$output_dir
} else {
  file.path(repo_root, params$output_dir)
}
paired_path <- file.path(output_dir, "paired_results.csv.gz")
required_paths <- c(
  paired_path,
  file.path(output_dir, "method_summary.csv.gz"),
  file.path(output_dir, "paired_summary.csv.gz"),
  file.path(output_dir, "exclusion_reason_summary.csv.gz")
)
if (any(!file.exists(required_paths))) {
  stop(
    "The paired diagnostic is incomplete; missing: ",
    paste(basename(required_paths[!file.exists(required_paths)]), collapse = ", ")
  )
}

paired <- data.table::fread(paired_path) %>% tibble::as_tibble()
method_summary <- data.table::fread(required_paths[[2L]]) %>% tibble::as_tibble()
paired_summary <- data.table::fread(required_paths[[3L]]) %>% tibble::as_tibble()
reason_summary <- data.table::fread(required_paths[[4L]]) %>% tibble::as_tibble()

expected_version <- "study3_joint_growth_mplus_openmx_audit_v1_20260825"
expected_reps <- seq_len(as.integer(params$expected_n_rep))
completion <- paired %>%
  group_by(.data$sentinel_index, .data$condition_id) %>%
  summarise(
    rows = n(),
    unique_reps = n_distinct(.data$rep),
    complete_rep_set = setequal(.data$rep, expected_reps),
    .groups = "drop"
  )
audit_complete <-
  nrow(completion) == 6L &&
  all(completion$rows == length(expected_reps)) &&
  all(completion$unique_reps == length(expected_reps)) &&
  all(completion$complete_rep_set) &&
  identical(unique(paired$audit_version), expected_version) &&
  all(paired$replication_seed == paired$mplus_replication_seed) &&
  all(abs(paired$truth - paired$mplus_truth) < 1e-12)
if (!audit_complete) {
  stop("Refusing to render: paired audit completion or seed contract failed.")
}

sentinel_levels <- paired %>%
  distinct(.data$sentinel_index, .data$sentinel_label) %>%
  arrange(.data$sentinel_index)
sentinel_short <- c(
  "1. Low-low, null, rho=0, N=50",
  "2. Low-low, beta=.5, rho=.5, N=50",
  "3. Low-high, beta=.5, rho=.5, N=50",
  "4. High-low, beta=.5, rho=.5, N=50",
  "5. High-high, beta=.5, rho=.5, N=50",
  "6. Low-low, beta=.5, rho=.5, N=300"
)
names(sentinel_short) <- sentinel_levels$sentinel_label
add_labels <- function(data) {
  data %>%
    mutate(
      sentinel = factor(
        unname(sentinel_short[.data$sentinel_label]),
        levels = unname(sentinel_short)
      )
    )
}
paired <- add_labels(paired)
method_summary <- add_labels(method_summary) %>%
  mutate(method = factor(
    .data$method,
    levels = c("Mplus MSEM", "OpenMx matched joint growth")
  ))
paired_summary <- add_labels(paired_summary)
reason_summary <- add_labels(reason_summary)

method_palette <- c(
  "Mplus MSEM" = "#59A14F",
  "OpenMx matched joint growth" = "#4E79A7"
)
theme_vh <- function() {
  theme_bw(base_size = 10.5) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      strip.background = element_rect(fill = "grey95", colour = "grey75")
    )
}
fmt_pct <- function(x, accuracy = .1) percent(x, accuracy = accuracy)
fmt_num <- function(x, digits = 3) formatC(x, digits = digits, format = "f")

method_long <- bind_rows(
  paired %>% transmute(
    .data$sentinel_index, .data$sentinel, .data$truth, .data$rep,
    method = "Mplus MSEM", estimate = .data$mplus_estimate,
    se = .data$mplus_se, point_eligible = .data$mplus_point_eligible,
    interval_eligible = .data$mplus_interval_eligible,
    covered = .data$mplus_covered,
    success_and_cover = .data$mplus_success_and_cover,
    exclusion_reason = .data$mplus_exclusion_reason
  ),
  paired %>% transmute(
    .data$sentinel_index, .data$sentinel, .data$truth, .data$rep,
    method = "OpenMx matched joint growth", estimate = .data$openmx_estimate,
    se = .data$openmx_se, point_eligible = .data$openmx_point_eligible,
    interval_eligible = .data$openmx_interval_eligible,
    covered = .data$openmx_covered,
    success_and_cover = .data$openmx_success_and_cover,
    exclusion_reason = .data$openmx_exclusion_reason
  )
) %>%
  mutate(method = factor(.data$method, levels = names(method_palette)))

metric_plot_data <- method_summary %>%
  select(
    .data$sentinel, .data$method,
    `Interval eligibility` = .data$interval_eligibility,
    `Conditional coverage` = .data$conditional_coverage,
    `Success and cover` = .data$success_and_cover
  ) %>%
  pivot_longer(
    cols = c("Interval eligibility", "Conditional coverage", "Success and cover"),
    names_to = "metric", values_to = "value"
  ) %>%
  mutate(metric = factor(
    .data$metric,
    levels = c("Interval eligibility", "Conditional coverage", "Success and cover")
  ))

disposition <- paired %>%
  count(.data$sentinel, .data$paired_interval_disposition, name = "replications") %>%
  group_by(.data$sentinel) %>%
  mutate(proportion = .data$replications / sum(.data$replications)) %>%
  ungroup() %>%
  mutate(paired_interval_disposition = factor(
    .data$paired_interval_disposition,
    levels = c("both eligible", "Mplus only", "OpenMx only", "neither eligible")
  ))

covariance_disposition <- paired %>%
  count(.data$sentinel, .data$paired_psi_problem, name = "replications") %>%
  group_by(.data$sentinel) %>%
  mutate(proportion = .data$replications / sum(.data$replications)) %>%
  ungroup() %>%
  mutate(paired_psi_problem = factor(
    .data$paired_psi_problem,
    levels = c(
      "both flag covariance problem", "Mplus warning only",
      "OpenMx PSI problem only", "neither flags covariance problem"
    )
  ))

both_eligible <- paired %>%
  filter(.data$mplus_interval_eligible, .data$openmx_interval_eligible)
overall <- tibble::tibble(
  mplus_eligibility = mean(paired$mplus_interval_eligible),
  openmx_eligibility = mean(paired$openmx_interval_eligible),
  both_eligible_rate = mean(
    paired$mplus_interval_eligible & paired$openmx_interval_eligible
  ),
  mplus_only = mean(
    paired$mplus_interval_eligible & !paired$openmx_interval_eligible
  ),
  openmx_only = mean(
    !paired$mplus_interval_eligible & paired$openmx_interval_eligible
  ),
  neither = mean(
    !paired$mplus_interval_eligible & !paired$openmx_interval_eligible
  ),
  both_covariance_problem = mean(
    paired$mplus_critical_warning & !paired$openmx_latent_psi_admissible
  ),
  mplus_warning_openmx_admissible = mean(
    paired$mplus_critical_warning & paired$openmx_latent_psi_admissible
  ),
  mean_abs_estimate_difference = mean(
    abs(both_eligible$openmx_estimate - both_eligible$mplus_estimate),
    na.rm = TRUE
  ),
  max_abs_estimate_difference = max(
    abs(both_eligible$openmx_estimate - both_eligible$mplus_estimate),
    na.rm = TRUE
  ),
  mean_abs_se_difference = mean(
    abs(both_eligible$openmx_se - both_eligible$mplus_se),
    na.rm = TRUE
  ),
  median_abs_se_difference = median(
    abs(both_eligible$openmx_se - both_eligible$mplus_se),
    na.rm = TRUE
  ),
  p99_abs_se_difference = as.numeric(stats::quantile(
    abs(both_eligible$openmx_se - both_eligible$mplus_se),
    .99, na.rm = TRUE
  )),
  max_abs_se_difference = max(
    abs(both_eligible$openmx_se - both_eligible$mplus_se),
    na.rm = TRUE
  ),
  estimate_correlation = cor(
    both_eligible$openmx_estimate, both_eligible$mplus_estimate,
    use = "complete.obs"
  ),
  se_correlation = cor(
    both_eligible$openmx_se, both_eligible$mplus_se,
    use = "complete.obs"
  ),
  mplus_critical_n = sum(paired$mplus_critical_warning),
  mplus_critical_openmx_psi_bad_n = sum(
    paired$mplus_critical_warning & !paired$openmx_latent_psi_admissible
  ),
  mplus_unavailable_n = sum(
    paired$mplus_interval_exclusion_reason == "estimation_unavailable"
  ),
  mplus_unavailable_openmx_ineligible_n = sum(
    paired$mplus_interval_exclusion_reason == "estimation_unavailable" &
      !paired$openmx_interval_eligible
  )
)
se_outlier <- both_eligible %>%
  mutate(abs_se_difference = abs(.data$openmx_se - .data$mplus_se)) %>%
  slice_max(.data$abs_se_difference, n = 1L, with_ties = FALSE)
```

# Executive summary

This diagnostic asks a narrower question than the main Study 3v2 comparison:
does MSEM lose eligibility because the joint two-process growth model is weakly
identified in these samples, or because of a software- or parameterization-
specific fitting problem? The audit fits the **same 16-parameter model** in
OpenMx and pairs it replication by replication with the completed Mplus MSEM
results.

The six sentinels contribute `r comma(nrow(paired))` paired replications. The
overall interval-eligibility rates are
`r fmt_pct(overall$mplus_eligibility)` for Mplus and
`r fmt_pct(overall$openmx_eligibility)` for OpenMx. Both programs are eligible
in `r fmt_pct(overall$both_eligible_rate)` of samples; Mplus alone is eligible in
`r fmt_pct(overall$mplus_only)`, OpenMx alone in
`r fmt_pct(overall$openmx_only)`, and neither in
`r fmt_pct(overall$neither)`.

Among samples that are eligible in both programs, the coefficient correlation
is `r fmt_num(overall$estimate_correlation, 4)` and the mean absolute
coefficient difference is `r fmt_num(overall$mean_abs_estimate_difference, 5)`.
This agreement check is essential: eligibility comparisons are interpretable
as software diagnostics only if the two implementations converge to the same
solution when both succeed.

The result is decisive at the scale this audit was designed to detect. All
`r comma(overall$mplus_critical_n)` Mplus critical-warning samples also have an
inadmissible OpenMx latent `PSI`. All
`r comma(overall$mplus_unavailable_n)` samples in which Mplus estimation is
unavailable are also ineligible in OpenMx. OpenMx rescues **none** of the Mplus
failures; conversely, two Mplus-eligible samples receive an OpenMx status-6
indefinite-Hessian result. Thus, switching software does not remedy the low
MSEM eligibility in these difficult cells.

# What was matched

The current Mplus helper reshapes each process to one row per cluster and fits
two linear latent growth curves. Inspection of the Mplus parameter table shows
that `I1`, `S1`, and `I2` are exogenous and therefore have a freely estimated
3-by-3 covariance matrix; `S2` is regressed on `I1` and `S1`; and `I2 WITH S2`
is a residual covariance. The OpenMx RAM model reproduces those defaults
explicitly.

```{r parameter-table}
bind_rows(
  tibble::tribble(
    ~`Parameter block`, ~Free_parameters, ~Specification,
    "Structural regression", 2L, "S2 ON I1 S1",
    "Exogenous latent covariance", 6L, "Unrestricted covariance of I1, S1, I2",
    "Outcome-slope residual block", 2L, "Residual variance of S2; I2 WITH S2",
    "Manifest residual variances", 2L, "One equality-constrained variance per process",
    "Latent means/intercepts", 4L, "Means of I1, S1, I2 and intercept of S2"
  ),
  tibble::tibble(
    `Parameter block` = "Total",
    Free_parameters = 16L,
    Specification = "Matched in both programs"
  )
) %>%
  kable(align = c("l", "r", "l"))
```

The automated test inspects the actual OpenMx `A`, `S`, and mean matrices,
including the implicit Mplus `I2 WITH I1` and `I2 WITH S1` paths. It also checks
OpenMx's column-major covariance ordering and verifies the total of 16 free
parameters. For eligible smoke-test samples, Mplus and OpenMx coefficients and
SEs agree numerically.

# Sentinel conditions

The diagnostic is intentionally compact rather than another crossed simulation
factor. Every sentinel uses three balanced observations per process. Conditions
1--5 use 50 clusters and span the four reliability combinations; the low-low
null distinguishes weak measurement from pressure caused by the structural
path. Condition 6 repeats the hardest low-low large-effect covariance shape at
300 clusters.

```{r sentinel-table}
paired %>%
  distinct(
    .data$sentinel_index, .data$sentinel, .data$condition_id,
    .data$num_clus, .data$target_reliability_y,
    .data$target_reliability_q, .data$marginal_rho,
    .data$standardized_beta_target
  ) %>%
  transmute(
    Sentinel = as.character(.data$sentinel),
    `Production condition` = .data$condition_id,
    N = .data$num_clus,
    `Predictor reliability` = .data$target_reliability_y,
    `Outcome reliability` = .data$target_reliability_q,
    rho = .data$marginal_rho,
    beta = .data$standardized_beta_target
  ) %>%
  kable(align = c("l", rep("r", 6)))
```

# Eligibility, coverage, and operational success

```{r primary-performance-plot, fig.width=12, fig.height=8}
ggplot(metric_plot_data, aes(
  x = .data$value,
  y = .data$sentinel,
  colour = .data$method,
  group = .data$method
)) +
  geom_vline(
    data = tibble::tibble(
      metric = factor(
        c("Conditional coverage", "Interval eligibility"),
        levels = levels(metric_plot_data$metric)
      ),
      reference = c(.95, 1)
    ),
    aes(xintercept = .data$reference),
    inherit.aes = FALSE, linetype = 2, colour = "grey55"
  ) +
  geom_point(size = 2.8, position = position_dodge(width = .35)) +
  facet_wrap(~ metric, ncol = 1) +
  scale_x_continuous(limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = method_palette) +
  labs(
    x = "Proportion",
    y = NULL,
    colour = NULL,
    title = "The paired audit separates eligibility from conditional coverage",
    subtitle = "Each point uses the same 100 simulated samples in both programs"
  ) +
  theme_vh()
```

Conditional coverage is calculated only among eligible intervals. Success-and-
cover instead counts the probability that a repeated sample produces an
eligible interval containing the truth. The paired design makes any difference
in those operational quantities considerably more precise than an independent
software comparison.

```{r paired-eligibility-disposition, fig.width=12, fig.height=6.5}
ggplot(disposition, aes(
  x = .data$proportion,
  y = .data$sentinel,
  fill = .data$paired_interval_disposition
)) +
  geom_col(width = .72) +
  geom_text(
    aes(label = if_else(
      .data$proportion >= .04,
      percent(.data$proportion, accuracy = 1), ""
    )),
    position = position_stack(vjust = .5), colour = "white", size = 3
  ) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = c(
    "both eligible" = "#59A14F",
    "Mplus only" = "#F28E2B",
    "OpenMx only" = "#4E79A7",
    "neither eligible" = "#9D9D9D"
  )) +
  labs(
    x = "Proportion of paired replications",
    y = NULL,
    fill = NULL,
    title = "Eligibility concordance identifies model-wide versus software-specific failures"
  ) +
  theme_vh()
```

# Do the programs flag the same covariance problem?

The primary OpenMx eligibility definition is deliberately strict. It requires
status zero, a finite positive SE, a definite information matrix, and a
positive-definite full 4-by-4 latent residual covariance matrix (`PSI`). Mplus
eligibility applies the existing production rule: a returned solution with a
critical Mplus warning, including a non-positive-definite `PSI`, is excluded.

```{r covariance-concordance, fig.width=12, fig.height=6.5}
ggplot(covariance_disposition, aes(
  x = .data$proportion,
  y = .data$sentinel,
  fill = .data$paired_psi_problem
)) +
  geom_col(width = .72) +
  geom_text(
    aes(label = if_else(
      .data$proportion >= .04,
      percent(.data$proportion, accuracy = 1), ""
    )),
    position = position_stack(vjust = .5), colour = "white", size = 3
  ) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_fill_manual(values = c(
    "both flag covariance problem" = "#B07AA1",
    "Mplus warning only" = "#E15759",
    "OpenMx PSI problem only" = "#4E79A7",
    "neither flags covariance problem" = "#59A14F"
  )) +
  labs(
    x = "Proportion of paired replications",
    y = NULL,
    fill = NULL,
    title = "Latent-covariance diagnostics localize the source of attrition",
    subtitle = "Fully unavailable Mplus fits need not contain a critical-warning flag"
  ) +
  theme_vh()
```

Across the six sentinels, both programs flag a covariance problem in
`r fmt_pct(overall$both_covariance_problem)` of samples. Mplus gives a critical
warning while OpenMx reaches an admissible `PSI` solution in
`r fmt_pct(overall$mplus_warning_openmx_admissible)`. These categories should
be read with the eligibility disposition: agreement on the covariance problem
supports a data/model-information explanation, whereas a large Mplus-warning/
OpenMx-admissible category points to optimizer, starting-value, or
parameterization sensitivity.

# Agreement when both programs are eligible

```{r estimate-agreement, fig.width=11.5, fig.height=8}
ggplot(both_eligible, aes(
  x = .data$mplus_estimate,
  y = .data$openmx_estimate,
  colour = .data$sentinel
)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey45") +
  geom_point(alpha = .65, size = 1.8) +
  facet_wrap(~ sentinel, scales = "free", ncol = 3) +
  guides(colour = "none") +
  labs(
    x = "Mplus focal estimate",
    y = "OpenMx focal estimate",
    title = "Eligible solutions target the same coefficient",
    subtitle = paste0(
      "Overall r = ", fmt_num(overall$estimate_correlation, 4),
      "; mean absolute difference = ",
      fmt_num(overall$mean_abs_estimate_difference, 5)
    )
  ) +
  theme_vh()
```

```{r se-agreement, fig.width=11.5, fig.height=8}
ggplot(both_eligible, aes(
  x = .data$mplus_se,
  y = .data$openmx_se,
  colour = .data$sentinel
)) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey45") +
  geom_point(alpha = .65, size = 1.8) +
  facet_wrap(~ sentinel, scales = "free", ncol = 3) +
  guides(colour = "none") +
  labs(
    x = "Mplus focal SE",
    y = "OpenMx focal SE",
    title = "Eligible solutions also agree on uncertainty",
    subtitle = paste0(
      "Overall r = ", fmt_num(overall$se_correlation, 4),
      "; mean absolute difference = ", fmt_num(overall$mean_abs_se_difference, 5)
    )
  ) +
  theme_vh()
```

The fixed reporting multiplier is identical in the two programs, so these SEs
do not differ because one program applies a delta method to a fitted scaling
variance. Agreement therefore tests the likelihood specification, parameter
mapping, and information matrix much more directly than comparing MSEM with
2S-PA.

SE agreement is also strong but has one informative tail exception. The median
absolute SE difference is
`r fmt_num(overall$median_abs_se_difference, 5)` and the 99th percentile is
`r fmt_num(overall$p99_abs_se_difference, 4)`. The maximum difference,
`r fmt_num(overall$max_abs_se_difference, 3)`, occurs in sentinel
`r se_outlier$sentinel_index`, replication `r se_outlier$rep`, where both SEs
are extremely large (Mplus `r fmt_num(se_outlier$mplus_se, 3)`; OpenMx
`r fmt_num(se_outlier$openmx_se, 3)`). This is consistent with a very flat,
boundary-adjacent likelihood rather than a systematic SE scaling mismatch.

# Numerical results

```{r method-summary-table}
method_summary %>%
  transmute(
    Sentinel = as.character(.data$sentinel),
    Method = as.character(.data$method),
    `Interval eligible` = percent(.data$interval_eligibility, accuracy = .1),
    Bias = round(.data$bias_among_point_eligible, 3),
    `Mean SE` = round(.data$mean_se_among_interval_eligible, 3),
    `Conditional coverage` = percent(.data$conditional_coverage, accuracy = .1),
    `Success and cover` = percent(.data$success_and_cover, accuracy = .1)
  ) %>%
  kable(align = c("l", "l", rep("r", 5)))
```

```{r paired-summary-table}
paired_summary %>%
  transmute(
    Sentinel = as.character(.data$sentinel),
    Both = percent(.data$both_eligible, accuracy = .1),
    `Mplus only` = percent(.data$mplus_only_eligible, accuracy = .1),
    `OpenMx only` = percent(.data$openmx_only_eligible, accuracy = .1),
    Neither = percent(.data$neither_eligible, accuracy = .1),
    `OpenMx - Mplus eligibility` = percent(
      .data$paired_eligibility_difference, accuracy = .1
    ),
    `Paired MCSE` = percent(
      as.numeric(.data$paired_eligibility_difference_mc_se), accuracy = .1
    ),
    `Eligible estimate r` = round(
      as.numeric(.data$both_eligible_estimate_correlation), 4
    )
  ) %>%
  kable(align = c("l", rep("r", 7)))
```

# Conclusions

The audit was constructed to distinguish three possibilities:

1. **Same estimates and same failures.** This is strong evidence that weak
   information for the joint growth model, rather than a program-specific bug,
   causes the attrition.
2. **Same estimates when eligible, but OpenMx is eligible more often.** The
   estimand and likelihood agree, but numerical strategy matters. Starting
   values, retry behavior, optimizer choice, or covariance parameterization can
   determine whether a sample reaches an admissible solution.
3. **Different estimates among mutually eligible fits.** This would indicate a
   remaining model-specification mismatch and would invalidate the eligibility
   comparison until resolved.

The observed results match the **first** possibility. The mutually eligible
fits agree almost exactly on the focal path, and the eligibility classifications
agree in 598 of 600 paired samples. The two disagreements favor Mplus, not
OpenMx. Most importantly, every Mplus critical covariance warning is reproduced
as an inadmissible latent covariance matrix in OpenMx, and no Mplus-unavailable
sample becomes OpenMx-eligible.

Accordingly, the poor MSEM eligibility in Study 3v2 should be interpreted as a
property of fitting this relatively elaborate joint growth model with weak
level-2 information. It is not explained by a Mplus parsing bug, an overly
aggressive warning screen, or a Mplus-specific optimizer failure. The existing
success-and-cover analysis therefore remains substantively meaningful: many
samples genuinely do not support an admissible joint-model interval.

This finding also changes the priority of possible remedies. Merely porting
MSEM to OpenMx is not promising. Remedies would need to change the information
or estimator, for example by increasing the number of clusters, improving
slope reliability, using a positive-definite covariance parameterization,
regularizing weak variance/covariance components, or introducing principled
Bayesian priors. Those changes may be useful, but they define modified
estimators and must be evaluated for bias and interval calibration; they are
not neutral software bug fixes.

## Limitations and remaining questions

The audit uses 100 paired replications per sentinel, so a single-cell
proportion near 50% has an ordinary Monte Carlo SE of approximately five
percentage points. It is designed to discriminate large mechanisms, not to
provide final publication-grade performance estimates. The OpenMx path also
uses the repository's `mxTryHard` retry wrapper (up to five attempts), whereas
the Mplus rows come from the production fitting path. Any OpenMx advantage
therefore establishes **numerical remediability**, but does not by itself prove
that the software brand is causal. In fact, OpenMx shows no advantage here even
with those retries, which makes optimizer tuning a lower-priority follow-up.

Finally, Mplus prints the saved coefficient and SE to three decimal places,
whereas the OpenMx results retain full numerical precision. Small paired
differences therefore partly reflect reporting precision. The single large SE
difference noted above warrants retaining both programs in any future audit of
SE tails, but it does not affect the eligibility conclusion.

# Reproducibility

The audit regenerates every sample using the saved production replication seed
and stops if the regenerated and saved seeds disagree. It reads, but never
modifies, the Study 3v2 production files. Per-sentinel results are resumable and
are stored separately under the diagnostic output directory.

```{r completion-table}
completion %>%
  transmute(
    Sentinel = .data$sentinel_index,
    `Production condition` = .data$condition_id,
    Rows = .data$rows,
    `Unique replications` = .data$unique_reps,
    Complete = .data$complete_rep_set
  ) %>%
  kable(align = c("r", "r", "r", "r", "c"))
```

```{r session-information}
sessionInfo()
```
