---
  title: "Structural Random Effects"
  author: "Zach Vig & Michael Hallquist"
  date: "`r format(Sys.time(), '%d %B, %Y')`"
  output:
    html_document:
      toc: true
      toc_depth: 2
      toc_float: true
      number_sections: true
    editor_options: 
      chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
pacman::p_load(tidyverse, ggh4x)
source_helpers <- function() {
  analysis_dir <- here::here("vig_hallquist_2026", "analysis")
  source(file.path(analysis_dir, "data_import.R"))
  source(here::here("R", "core_utils.R"))
  source(here::here("R", "sim_diagnostics.R"))
  source(here::here("vig_hallquist_2026", "vh_study2.R"))
}
source_helpers()
```

# Study 2: Random-Effect-as-Predictor

```{r}
# Load data
study2 <- read_study_cache("study2", this_run = "vig_hallquist_slurm") %>%
  mutate(
      status10_failure = !is.na(status_code) & status_code == 10L,
      bias = estimate - truth,
      rel_bias = case_when(
          truth == 0 ~ NA_real_,
          TRUE ~ (estimate - truth) / truth
      ),
      sq_error = (estimate - truth)^2,
      num_clus_fac = factor(num_clus, levels = c(30, 50, 100, 150, 300))
  ) %>%
  add_study2_legacy_eligibility() %>% # for older cached runs
  add_study2_mplus_legacy_threshold(threshold = 5) %>% # for older cached runs
  add_study2_analysis_eligibility() %>% # for all cached runs
  mutate(
    converged = point_eligible,
    analysis_ready = point_eligible,
    covered = if_else(interval_eligible, ci_low <= truth & ci_high >= truth, NA)
  )

manifest <- get_manifest(study_arg = "all", this_run = "vig_hallquist_slurm")

study2_wide <- study2 %>%
  left_join(manifest, by = c("study", "condition_id"))

# failure reasons
# study2 %>%
#   select(method, analysis_exclusion_reason) %>%
#   table
```

```{r}
study2_method_labels <- c(
  "fuller_closed_form" = "Fuller (stabilized)",
  "fuller_book_preliminary_closed_form" = "Fuller (book preliminary)",
  "fuller_book_closed_form" = "Fuller (book equations)",
  "lai_2spa" = "Lai (2-SPA)",
  "naive_dual_blup" = "OLS with Both BLUPs",
  "fuller_alpha_stepdown_closed_form" = "Fuller (alpha stepdown)",
  "closed_form_dual" = "OLS with Both MLEs",
  "msem" = "MSEM",
  "oracle_dual" = "Oracle with Both"
)
```


## Hypotheses
* H1. Naive BLUPs as predictors will have minimally biased point estimates  (esp. compared to Study 1), but their standard errors and coverage will be wrong.
* H1a. Closed-form scores used as predictors require Fuller or Lai 2S-PA to obtain 95\% coverage.
* H2. The largest differences among methods will appear in low-reliability, small-cluster, and dual-predictor conditions.
* H3. Null-effect conditions will show that naive OLS with BLUPs leads to inflated Type I errors.


## Bias
```{r bias_study2}
# first look at slope-intercept correlation (versus effect size)
make_bias_cor <- function(df, clus_size) {
  df %>%
  filter(analysis_ready) %>%
  filter(standardized_beta_target > 0, mean_clus_size == clus_size) %>%
  ggplot(aes(x = target_reliability, y = bias, fill = method)) +
  facet_nested("Slope-Intercept Correlation" + marginal_rho ~ "Number of Clusters" + num_clus_fac) +
  stat_summary(fun = mean, geom = "col", position = "dodge") +
  scale_fill_discrete(labels = study2_method_labels) +
  # scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Slope Reliability", y = "Bias", fill = "Method",
       title = sprintf("Study 2: Bias, Cluster Size = %s (Converged Replications)", clus_size)) +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
}

bias_cor3 <- make_bias_cor(study2, 3)
bias_cor5 <- make_bias_cor(study2, 5)
bias_cor10 <- make_bias_cor(study2, 10)
bias_cor25 <- make_bias_cor(study2, 25)

library(patchwork)
p <- bias_cor3 + bias_cor5 + bias_cor10 + bias_cor25

print(bias_cor3)
print(bias_cor5)
print(bias_cor10)
print(bias_cor25)
print(p)
```

```{r}
# null effect size shows no patterns
bias_eff_size0 <- study2 %>%
  filter(analysis_ready) %>%
  #filter(standardized_beta_target > 0) %>%
  ggplot(aes(x = target_reliability, y = bias, fill = method)) +
  facet_nested("Effect Size" + standardized_beta_target ~ "Number of Clusters" + num_clus_fac) +
  stat_summary(fun = mean, geom = "col", position = "dodge") +
  scale_fill_discrete(labels = study2_method_labels) +
  #scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Slope Reliability", y = "Bias", fill = "Method",
       title = "Study 2: Bias (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

# null effect size shows no patterns
bias_eff_size <- study2 %>%
  filter(analysis_ready) %>%
  filter(standardized_beta_target > 0) %>%
  ggplot(aes(x = target_reliability, y = rel_bias, fill = method)) +
  facet_nested("Effect Size" + standardized_beta_target ~ "Number of Clusters" + num_clus_fac) +
  stat_summary(fun = mean, geom = "col", position = "dodge") +
  scale_fill_discrete(labels = study2_method_labels) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Slope Reliability", y = "Bias", fill = "Method",
       title = "Study 2: Bias (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

print(bias_eff_size0); print(bias_eff_size)
```

OLS using closed-form scores expectedly showed negative bias (i.e., slope attenuation) across the board. The null effect size showed zero bias across all conditions and methods. For nonzero slope-intercept correlation conditions, Fuller showed decreasing (negative) bias with increased cluster size and a faster rate than 2-SPA. Additionally, Fuller showed roughly equal or smaller bias than 2-SPA for all conditions where slope reliability was medium to high. 2-SPA showed roughly equal or smaller bias than Fuller in low reliability conditions.

### Investigating Mismatch Between Our Study and Lai & Liu (2026)
```{r}
bias_clus_size25 <- study2 %>%
  filter(analysis_ready) %>%
  filter(standardized_beta_target > 0, mean_clus_size == 25) %>%
  ggplot(aes(x = target_reliability, y = bias, fill = method)) +
  facet_nested("Effect Size" + standardized_beta_target ~ "Number of Clusters" + num_clus_fac) +
  stat_summary(fun = mean, geom = "col", position = "dodge") +
  scale_fill_discrete(labels = study2_method_labels) +
  #scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Slope Reliability", y = "Bias", fill = "Method",
       title = "Study 2: Bias (Converged Replications, Cluster Size = 25)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
print(bias_clus_size25)
```

```{r}
# heatmap of ICC by other parameters
icc_heat <- study2_wide %>%
  filter(analysis_ready) %>%
  mutate(icc = slope_variance_marginal.x / (slope_variance_marginal.x + outcome_residual_variance)) %>%
  ggplot(aes(x = target_reliability.x, y = as.factor(mean_clus_size.x), fill = icc)) +
  facet_nested("Slope-Intercept Correlation" + marginal_rho.x ~ "Number of Clusters" + num_clus_fac) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(x = "Slope Reliability", y = "Mean Cluster Size", fill = "ICC",
       title = "Study 2: ICC by Condition (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
print(icc_heat)
```

The patterns persist across cluster size. In fact, the mean ICC increases with *smaller* cluster size.

```{r}
# heatmap of stage1_re_corr (since it's one of two parameters we currently export from stage 1)
stage1_re_corr_heat <- study2_wide %>%
  filter(analysis_ready) %>%
  ggplot(aes(x = target_reliability.x, y = as.factor(mean_clus_size.x), fill = stage1_re_corr)) +
  facet_nested("Slope-Intercept Correlation" + marginal_rho.x ~ "Number of Clusters" + num_clus_fac) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(x = "Slope Reliability", y = "Mean Cluster Size", fill = "Stage 1 RE Correlation",
       title = "Study 2: Stage 1 RE Correlation by Condition (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
print(stage1_re_corr_heat)

# heatmap of stage1_eb_corr (since it's one of two parameters we currently export from stage 1)
stage1_eb_corr_heat <- study2_wide %>%
  filter(analysis_ready) %>%
  ggplot(aes(x = target_reliability.x, y = as.factor(mean_clus_size.x), fill = stage1_eb_corr)) +
  facet_nested("Slope-Intercept Correlation" + marginal_rho.x ~ "Number of Clusters" + num_clus_fac) +
  geom_tile() +
  scale_fill_viridis_c() +
  labs(x = "Slope Reliability", y = "Mean Cluster Size", fill = "Stage 1 EB Correlation",
       title = "Study 2: Stage 1 EB Correlation by Condition (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
print(stage1_eb_corr_heat)
```

## RMSE
```{r rmse_study2}
rmse <- study2 %>%
  filter(analysis_ready) %>%
  filter(method != "closed_form_dual") %>%
  ggplot(aes(x = target_reliability, y = sq_error, fill = method)) +
  facet_nested("Slope-Intercept Correlation" + marginal_rho ~ "Number of Clusters" + num_clus_fac) +
  stat_summary(data = ~ dplyr::filter(.x, method != "oracle_dual"), fun = function(x) sqrt(mean(x)), geom = "col", position = "dodge") +
  stat_summary(data = ~ dplyr::filter(.x, method == "oracle_dual"), fun = function(x) sqrt(mean(x)), geom = "point", color = "black", size = 1) +
  scale_fill_discrete(labels = study2_method_labels) +
  labs(x = "Slope Reliability", y = "RMSE", fill = "Method",
       title = "Study 2: RMSE (Converged Replications)") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
print(rmse)
```

OLS with BLUPs showed generally high RMSE, especially in the zero slope-intercept correlation condition. The other methods showed decreasing RMSE with reliability and number of clusters. Fuller with alpha stepdown either perfromed equally or better than the standard Fuller approach.

## Coverage

```{r}
coverage <- study2 %>%
  filter(interval_eligible) %>%
  mutate(covered = as.numeric(covered)) %>%
  filter(method != "closed_form_dual" & method != "oracle_dual") %>%
  mutate(num_clus_label = "Number of Clusters", cor_label = "Slope-Intercept Correlation") %>%
  ggplot(aes(x = target_reliability, y = covered, fill = method)) +
  facet_nested(cor_label + marginal_rho ~ num_clus_label + num_clus_fac) +
  stat_summary(fun = mean, geom = "col", position =  "dodge") +  
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  coord_cartesian(ylim = c(0.4, 1)) +
  scale_fill_discrete(labels = study2_method_labels) +
  geom_hline(yintercept = 0.95, linetype = "dashed") +
  labs(x = "Slope Reliability", y = "Coverage",
       title = "Study 2: Coverage (Converged Replications)", fill = "Method") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))
print(coverage)
```

All methods show undercoverage, sometimes noticeably, for low reliability, nonzero slope-intercept correlation conditions. In medium to high reliability conditions, Fuller, 2-SPA, and MSEM showed nominal coverage, while OLS with BLUPs often showed undercoverage.

## Standard Errors

```{r}
se_table <- study2 %>%
  filter(interval_eligible) %>%
  filter(method != "closed_form_dual" & method != "oracle") %>%
  group_by(method, num_clus_fac, target_reliability, marginal_rho, mean_clus_size, standardized_beta_target) %>%
  summarize(
    mean_se = mean(se),
    empirical_se = sd(estimate),
    se_ratio = mean_se / empirical_se,
    .groups = "drop_last"
  )
se <- se_table %>%
  mutate(cor_label = "Slope-Intercept Correlation",
         num_clus_label = "Number of Clusters") %>%
  ggplot(aes(x = target_reliability, y = se_ratio, fill = method)) +
  facet_nested(cor_label + marginal_rho ~ num_clus_label + num_clus_fac) +
  geom_bar(stat = "identity", position = "dodge") +
  #scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_fill_discrete(labels = study2_method_labels) +
  labs(x = "Slope Reliability", y = "Standard Error Ratio", fill = "Method",
       title = "Study 2: Standard Errors (Converged Replications)") +
  geom_hline(yintercept = 1, linetype = "dashed") +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5))

print(se)
```

## Convergence

```{r}
plot_convergence <- function(df, subtitle) {
  df %>%
  ggplot(
    aes(
      x = target_reliability,
      y = converged,
      color = method,
      group = method
    )
  ) +
  stat_summary(fun = safe_mean, geom = "line", linewidth = 1) +
  stat_summary(fun = safe_mean, geom = "point", size = 2) +
  scale_y_continuous(
    labels = scales::label_percent(),
    limits = c(0, 1)
  ) +
  facet_nested(
    "Slope-Intercept Correlation" + marginal_rho ~ "Number of Clusters" + as.factor(num_clus)
  ) +
  labs(
    x = "Slope Reliability",
    y = "Convergence Rate",
    color = "Method",
    title = "Study 2: Convergence Rate",
    subtitle = subtitle
  ) +
  theme_bw(base_size = 14, base_family = "sans") +
  theme(legend.position = "bottom", plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  scale_color_discrete(labels = study2_method_labels)
}

convergence <- plot_convergence(
  study2 %>%
  filter(
    method %in% c("fuller_closed_form", "lai_2spa", "msem")
  ),
  "*Includes Stage 1 Singular Fits"
)
convergence2 <- plot_convergence(
  study2 %>%
  filter(
    method %in% c("fuller_closed_form", "lai_2spa", "msem")
  ) %>%
  filter(!stage1_singular_problem),
  "*Does NOT Include Stage 1 Singular Fits"
)
print(convergence); print(convergence2)
```
