---
title: "VH ICC Bridge: Posterior Reliability Versus an ICC Anchor"
author: "Simulation analysis"
date: "18 August 2026"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    embed-resources: true
    html-math-method: katex
params:
  output_dir: "outputs/vig_hallquist_icc_bridge"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dpi = 120,
  out.width = "100%"
)

suppressPackageStartupMessages({
  library(data.table)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
  library(scales)
})

input_path <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) "")
analysis_dir <- if (nzchar(input_path)) {
  dirname(normalizePath(input_path, mustWork = FALSE))
} else {
  getwd()
}
repo_root <- normalizePath(file.path(analysis_dir, "..", ".."), mustWork = TRUE)
resolve_repo_path <- function(path) {
  if (grepl("^/", path)) path else file.path(repo_root, path)
}

source(file.path(
  repo_root,
  "vig_hallquist_2026",
  "random_effects_structural_simulation.R"
), local = knitr::knit_global())

output_dir <- resolve_repo_path(params$output_dir)
condition_dir <- file.path(output_dir, "conditions")
summary_files <- list.files(
  condition_dir,
  pattern = "^condition_[0-9]+_summary[.]csv$",
  full.names = TRUE
)
if (length(summary_files) == 0L) {
  stop("No ICC-bridge condition summaries found in: ", condition_dir)
}

cell_metrics <- data.table::rbindlist(
  lapply(summary_files, data.table::fread),
  fill = TRUE,
  use.names = TRUE
) %>%
  tibble::as_tibble()

expected_design <- select_design("iccbridge")
completed_ids <- sort(unique(cell_metrics$condition_id))
missing_ids <- setdiff(expected_design$condition_id, completed_ids)
unexpected_ids <- setdiff(completed_ids, expected_design$condition_id)
expected_method_cells <- tidyr::crossing(
  condition_id = expected_design$condition_id,
  method = study2_methods()
)
observed_method_cells <- cell_metrics %>%
  count(.data$condition_id, .data$method, name = "summary_rows")
missing_method_cells <- dplyr::anti_join(
  expected_method_cells,
  observed_method_cells,
  by = c("condition_id", "method")
)
unexpected_method_cells <- dplyr::anti_join(
  observed_method_cells,
  expected_method_cells,
  by = c("condition_id", "method")
)
if (
  length(missing_ids) > 0L ||
  length(unexpected_ids) > 0L ||
  nrow(missing_method_cells) > 0L ||
  nrow(unexpected_method_cells) > 0L ||
  any(observed_method_cells$summary_rows != 1L) ||
  any(cell_metrics$n_rep != 1000L)
) {
  stop(
    "ICC-bridge output is incomplete or internally inconsistent: ",
    length(missing_ids), " missing condition IDs; ",
    length(unexpected_ids), " unexpected condition IDs; ",
    nrow(missing_method_cells), " missing condition-method cells; ",
    nrow(unexpected_method_cells), " unexpected condition-method cells; ",
    sum(observed_method_cells$summary_rows != 1L),
    " duplicated condition-method cells; n_rep range ",
    paste(range(cell_metrics$n_rep), collapse = "--")
  )
}

arm_labels <- c(
  posterior_reliability_targeted = "Posterior slope reliability held fixed",
  icc_anchored_at_m10 = "ICC anchored at m = 10"
)
method_labels <- c(
  oracle_dual = "Oracle dual",
  naive_dual_blup = "Naive dual BLUP",
  closed_form_dual = "Uncorrected closed-form OLS",
  fuller_closed_form = "Fuller (stabilized)",
  fuller_book_preliminary_closed_form = "Fuller (book preliminary)",
  fuller_book_closed_form = "Fuller (book equations)",
  fuller_alpha_stepdown_closed_form = "Fuller (alpha-stepdown)",
  lai_2spa = "Lai 2S-PA",
  msem = "MSEM"
)
focal_methods <- c("oracle_dual", "fuller_closed_form", "lai_2spa", "msem")
feasible_methods <- c("fuller_closed_form", "lai_2spa", "msem")

cell_metrics <- cell_metrics %>%
  mutate(
    arm_label = factor(
      unname(arm_labels[calibration_arm]),
      levels = unname(arm_labels)
    ),
    method_label = factor(
      unname(method_labels[method]),
      levels = unname(method_labels)
    ),
    m_label = factor(
      paste0("m = ", mean_clus_size),
      levels = paste0("m = ", c(3, 10, 25))
    ),
    rho_label = factor(
      paste0("rho = ", marginal_rho),
      levels = paste0("rho = ", c(-0.5, 0, 0.5))
    ),
    target_label = factor(
      paste0("Target R = ", posterior_reliability_anchor),
      levels = paste0("Target R = ", c(0.25, 0.5, 0.8))
    ),
    effect_label = factor(
      paste0("beta = ", standardized_beta_target),
      levels = paste0("beta = ", c(0, 0.4))
    )
  )

focal <- cell_metrics %>% filter(method %in% focal_methods)

manipulation <- cell_metrics %>%
  distinct(
    calibration_arm, arm_label, mean_clus_size, m_label,
    marginal_rho, rho_label, posterior_reliability_anchor, target_label,
    achieved_reliability, achieved_partial_reliability,
    intercept_icc, marginal_slope_icc, conditional_slope_icc, sigma
  )

arm_effect_summary <- focal %>%
  group_by(arm_label, method_label, standardized_beta_target) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    mean_coverage = mean(.data$coverage, na.rm = TRUE),
    minimum_coverage = min(.data$coverage, na.rm = TRUE),
    maximum_coverage = max(.data$coverage, na.rm = TRUE),
    mean_bias = mean(.data$bias, na.rm = TRUE),
    mean_absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    bias_mc_se = sqrt(sum(.data$mc_se_bias^2, na.rm = TRUE)) / n(),
    mean_model_se = mean(.data$mean_se, na.rm = TRUE),
    median_se_to_sd = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    type1_error = if (first(.data$standardized_beta_target) == 0) {
      mean(.data$type1_error, na.rm = TRUE)
    } else {
      NA_real_
    },
    operational_type1 = if (first(.data$standardized_beta_target) == 0) {
      mean(.data$operational_type1_error, na.rm = TRUE)
    } else {
      NA_real_
    },
    .groups = "drop"
  ) %>%
  mutate(
    bias_mc_low = mean_bias - qnorm(.975) * bias_mc_se,
    bias_mc_high = mean_bias + qnorm(.975) * bias_mc_se
  )

arm_m_summary <- focal %>%
  group_by(arm_label, mean_clus_size, method_label) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    mean_coverage = mean(.data$coverage, na.rm = TRUE),
    mean_bias = mean(.data$bias, na.rm = TRUE),
    mean_absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    mean_model_se = mean(.data$mean_se, na.rm = TRUE),
    median_se_to_sd = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

null_summary <- focal %>%
  filter(standardized_beta_target == 0) %>%
  group_by(arm_label, mean_clus_size, method_label) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    mean_null_bias = mean(.data$bias, na.rm = TRUE),
    conditional_type1 = mean(.data$type1_error, na.rm = TRUE),
    operational_type1 = mean(.data$operational_type1_error, na.rm = TRUE),
    positive_rejection = mean(.data$type1_error_positive, na.rm = TRUE),
    negative_rejection = mean(.data$type1_error_negative, na.rm = TRUE),
    minimum_type1 = min(.data$type1_error, na.rm = TRUE),
    maximum_type1 = max(.data$type1_error, na.rm = TRUE),
    .groups = "drop"
  )

pair_metrics <- cell_metrics %>%
  dplyr::select(
    bridge_pair_id, mean_clus_size, marginal_rho,
    posterior_reliability_anchor, standardized_beta_target,
    method, method_label, calibration_arm,
    coverage, bias, mean_se, interval_eligibility, type1_error
  ) %>%
  pivot_wider(
    names_from = calibration_arm,
    values_from = c(
      coverage, bias, mean_se, interval_eligibility, type1_error
    ),
    names_sep = "__"
  ) %>%
  mutate(
    coverage_difference =
      coverage__posterior_reliability_targeted -
      coverage__icc_anchored_at_m10,
    bias_difference =
      bias__posterior_reliability_targeted -
      bias__icc_anchored_at_m10,
    se_difference =
      mean_se__posterior_reliability_targeted -
      mean_se__icc_anchored_at_m10,
    eligibility_difference =
      interval_eligibility__posterior_reliability_targeted -
      interval_eligibility__icc_anchored_at_m10
  )

paired_summary <- pair_metrics %>%
  filter(method %in% focal_methods) %>%
  group_by(mean_clus_size, method_label) %>%
  summarise(
    paired_cells = n(),
    mean_coverage_difference = mean(coverage_difference, na.rm = TRUE),
    mean_se_difference = mean(se_difference, na.rm = TRUE),
    mean_eligibility_difference = mean(eligibility_difference, na.rm = TRUE),
    .groups = "drop"
  )

anchor_metrics <- pair_metrics %>% filter(mean_clus_size == 10)
anchor_max_difference <- max(abs(c(
  anchor_metrics$coverage_difference,
  anchor_metrics$bias_difference,
  anchor_metrics$se_difference,
  anchor_metrics$eligibility_difference
)), na.rm = TRUE)

type1_identity_error <- focal %>%
  filter(standardized_beta_target == 0) %>%
  summarise(maximum = max(abs(type1_error - (1 - coverage)), na.rm = TRUE)) %>%
  pull(maximum)

metric_mean <- function(
    arm, method, effect, variable, mean_cluster_size = NULL,
    reliability_anchor = NULL) {
  selected <- focal %>%
    filter(
      .data$calibration_arm == .env$arm,
      .data$method == .env$method,
      .data$standardized_beta_target == .env$effect
    )
  if (!is.null(mean_cluster_size)) {
    selected <- selected %>%
      filter(.data$mean_clus_size == .env$mean_cluster_size)
  }
  if (!is.null(reliability_anchor)) {
    selected <- selected %>%
      filter(
        .data$posterior_reliability_anchor == .env$reliability_anchor
      )
  }
  mean(selected[[variable]], na.rm = TRUE)
}

paired_mean <- function(mean_cluster_size, method, variable) {
  selected <- pair_metrics %>%
    filter(
      .data$mean_clus_size == .env$mean_cluster_size,
      .data$method == .env$method
    )
  mean(selected[[variable]], na.rm = TRUE)
}

fmt_pct <- function(x, accuracy = 0.1) percent(x, accuracy = accuracy)
fmt_pp <- function(x, digits = 1) sprintf(paste0("%+.", digits, "f pp"), 100 * x)
fmt_num <- function(x, digits = 3) formatC(x, digits = digits, format = "f")

arm_palette <- c(
  "Posterior slope reliability held fixed" = "#4E79A7",
  "ICC anchored at m = 10" = "#E15759"
)
method_palette <- c(
  "Oracle dual" = "#4D4D4D",
  "Fuller (stabilized)" = "#4E79A7",
  "Lai 2S-PA" = "#B07AA1",
  "MSEM" = "#59A14F"
)

theme_bridge <- function() {
  theme_bw(base_size = 10) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      strip.background = element_rect(fill = "grey95", colour = "grey75")
    )
}
```

# Executive summary

The ICC bridge strengthens the design conclusion from VH Study 5: preserve the
random-effects covariance shape and use posterior slope reliability as the
primary information manipulation. The reason is not that ICC is uninformative.
With the covariance shape fixed and the observation schedule specified, ICC
and posterior reliability are monotone summaries of the same signal-to-noise
system. The problem is that a fixed ICC does **not** hold the precision of an
individual slope estimate constant when the number and placement of repeated
observations change.

This bridge makes that distinction unusually clean. The two calibration arms
are identical at the reference cluster size, $m=10$, and use common random
numbers. Their condition-level coverage, bias, mean SE, and eligibility
summaries agree to a maximum absolute difference of
`r format(anchor_max_difference, scientific = TRUE)`. Away from that anchor,
the posterior-reliability arm recalibrates the level-1 residual scale to keep
slope reliability at .25, .50, or .80. The ICC arm holds the $m=10$ ICC
anchor fixed. Consequently, its achieved posterior slope reliability is only
about .07--.50 at $m=3$, but about .46--.91 at $m=25$.

That information drift has material inferential consequences. Averaging across
both effect conditions, correlations, and reliability anchors, posterior
targeting improves coverage at $m=3$, relative to ICC anchoring, by
`r fmt_pp(paired_mean(3, "fuller_closed_form", "coverage_difference"))` for
Fuller, `r fmt_pp(paired_mean(3, "lai_2spa", "coverage_difference"))` for
2S-PA, and `r fmt_pp(paired_mean(3, "msem", "coverage_difference"))` for
MSEM. The oracle difference is exactly zero because the oracle does not depend
on the noisy Stage-1 measurements. At $m=25$, where the ICC arm gives **more**
posterior information than its $m=10$ anchor, differences are much smaller
and are not uniformly favorable to posterior targeting. This is the intended
result: the bridge shows that holding ICC fixed changes the effective slope
information as $m$ changes, not that one calibration must improve every
estimator in every cell.

The estimator conclusions remain qualified. Fuller is always eligible and has
the smallest feasible-method bias, but its null intervals remain conservative.
2S-PA is also conservative at the null and its coverage is conditional on a
selected set of successful fits. MSEM has Type I error closest to 5%, but it
retains positive non-null bias and some low-reliability undercoverage. Thus the
bridge supports posterior reliability as the **design factor**; it does not
establish that any one feasible interval procedure is uniformly calibrated.

# Relation to the Study 5 argument

VH Study 5 separated two issues that had been confounded in the original
Studies 1--4:

1. attaining low reliability by shrinking only the random-slope variance,
   $\tau_1^2$, can create an irregular covariance geometry; and
2. even after fixing that geometry, marginal slope reliability, residualized
   slope reliability, and ICC are different descriptions of information.

Study 5 decisively rejected the one-coordinate `solve-G_22` calibration (the
old code name for solving only for $\tau_1^2$). It recommended a shared
shape-preserving DGM, marginal posterior slope reliability as the primary
factor, and residualized-slope reliability as a required diagnostic and
sensitivity factor. The present bridge addresses the remaining reviewer-facing
question: once covariance geometry is regular, could an ICC manipulation give
an equally coherent account of the design?

The answer is nuanced. At a fixed measurement design, either ICC or posterior
reliability can index the signal-to-noise balance. Across different cluster
sizes, posterior reliability is more proximal to the Stage-1 quantity used by
Fuller, 2S-PA, and MSEM because it incorporates the full random-effects
covariance, residual variance, and time-design matrix. A single ICC omits part
of that design dependence.

The ICC arm here specifically holds the **intercept ICC** implied by the
$m=10$ posterior-reliability solution. Because
$\tau_1^2/\tau_0^2=1$ and the intercept--slope correlation are fixed, this is
a coherent scalar signal-to-noise anchor. It should not be confused with the
conditional slope ICC, which is recorded as a manipulation check but is not
the calibration target. This terminology matters when comparing the bridge
with Lai-style conditional-ICC designs.

# Design, completeness, and estimands

The 108 conditions cross:

- two calibration arms;
- $m=3,10,25$ observations per cluster;
- posterior-reliability anchors .25, .50, and .80;
- random intercept--slope correlations $\rho=-.5,0,.5$; and
- standardized focal effects $\beta=0$ and $.4$.

Every condition has 100 clusters and 1,000 Monte Carlo replications. The
random-effects covariance shape is fixed at
$\tau_1^2/\tau_0^2=1$; calibration changes the level-1 residual scale.

```{r completion-check}

tibble(
  Check = c(
    "Expected design conditions",
    "Completed condition summaries",
    "Missing expected condition IDs",
    "Unexpected condition IDs",
    "Missing condition-method summaries",
    "Unexpected condition-method summaries",
    "Methods per condition",
    "Minimum replications per condition-method",
    "Maximum replications per condition-method"
  ),
  Result = c(
    nrow(expected_design),
    n_distinct(cell_metrics$condition_id),
    length(missing_ids),
    length(unexpected_ids),
    nrow(missing_method_cells),
    nrow(unexpected_method_cells),
    n_distinct(cell_metrics$method),
    min(cell_metrics$n_rep),
    max(cell_metrics$n_rep)
  )
) %>%
  kable(caption = "Production-output completeness checks.")
```

All performance statistics are condition-level summaries computed from the
replication outputs. Point-estimate summaries require point eligibility.
Coverage and SE summaries require interval eligibility. Coverage is therefore
conditional on obtaining an eligible interval; `success_and_cover` additionally
counts ineligible replications as unsuccessful. Bias Monte Carlo intervals use
the empirical sampling SD divided by the square root of the eligible count.

For the null conditions, Type I error uses the same eligible-fit denominator
and the same confidence-interval inversion rule as coverage. Therefore,

\[
\widehat\alpha = 1-\widehat C
\]

cell by cell. The largest observed numerical deviation from that identity is
`r format(type1_identity_error, scientific = TRUE)`. Type I error and null
coverage are shown on the Type I scale because that is the decision-relevant
interpretation; plotting both would duplicate the same statistic.

# Manipulation checks: what ICC fails to hold constant

```{r reliability-manipulation-plot, fig.width=11, fig.height=7.5}

manipulation %>%
  ggplot(aes(
    x = mean_clus_size,
    y = achieved_reliability,
    colour = arm_label,
    shape = arm_label,
    group = arm_label
  )) +
  geom_hline(
    aes(yintercept = posterior_reliability_anchor),
    linetype = 2,
    colour = "grey45"
  ) +
  geom_line(linewidth = .55) +
  geom_point(size = 2.1) +
  facet_grid(rho_label ~ target_label) +
  scale_x_continuous(breaks = c(3, 10, 25)) +
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  scale_colour_manual(values = arm_palette) +
  labs(
    x = "Observations per cluster (m)",
    y = "Achieved marginal posterior slope reliability",
    colour = NULL,
    shape = NULL,
    title = "A fixed ICC does not hold slope-score information constant",
    subtitle = "Dashed lines are the intended posterior-reliability anchors; arms coincide exactly at m = 10"
  ) +
  theme_bridge()
```

```{r manipulation-table}

manipulation %>%
  group_by(arm_label, mean_clus_size, posterior_reliability_anchor) %>%
  summarise(
    achieved_R_range = sprintf(
      "%.3f--%.3f",
      min(achieved_reliability), max(achieved_reliability)
    ),
    partial_R_range = sprintf(
      "%.3f--%.3f",
      min(achieved_partial_reliability),
      max(achieved_partial_reliability)
    ),
    ICC_range = sprintf(
      "%.3f--%.3f",
      min(intercept_icc), max(intercept_icc)
    ),
    sigma_range = sprintf("%.3f--%.3f", min(sigma), max(sigma)),
    .groups = "drop"
  ) %>%
  arrange(posterior_reliability_anchor, mean_clus_size, arm_label) %>%
  kable(
    col.names = c(
      "Calibration arm", "m", "R anchor", "Achieved marginal R",
      "Achieved residualized R", "Intercept ICC", "Residual SD"
    ),
    caption = "Population manipulation checks; ranges are across the three intercept--slope correlations."
  )
```

At $m=10$, both arms use the same residual scale, the same population DGM,
and the same replication seed. Exact agreement at this anchor is stronger than
ordinary similarity: it verifies that the paired comparison is wired correctly.
At $m=3$, ICC anchoring lowers effective slope reliability substantially;
at $m=25$, it raises it. An ICC-based study would therefore silently combine
changes in cluster size with changes in Stage-1 slope information.

# Coverage

## Actual coverage by estimator and information level

```{r nonnull-coverage-plot, fig.width=12, fig.height=10}

focal %>%
  filter(standardized_beta_target == .4) %>%
  ggplot(aes(
    x = mean_clus_size,
    y = coverage,
    colour = arm_label,
    shape = rho_label,
    group = interaction(arm_label, rho_label)
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey35") +
  geom_errorbar(
    aes(ymin = coverage_mc_low, ymax = coverage_mc_high),
    width = .35,
    alpha = .22,
    linewidth = .35
  ) +
  geom_line(alpha = .45, linewidth = .45) +
  geom_point(size = 1.8) +
  facet_grid(method_label ~ target_label) +
  scale_x_continuous(breaks = c(3, 10, 25)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = arm_palette) +
  labs(
    x = "Observations per cluster (m)",
    y = "Conditional 95% CI coverage",
    colour = NULL,
    shape = NULL,
    title = "Non-null coverage follows the information induced by each calibration",
    subtitle = "Points are design cells; vertical bars are 95% Monte Carlo intervals"
  ) +
  theme_bridge()
```

The oracle remains close to nominal and is identical across arms. This rules
out a truth-coding or generic confidence-interval problem. The feasible methods
respond to the Stage-1 information manipulation.

The low-information corner is most revealing. At $m=3$ and the .25 anchor,
the ICC arm achieves slope reliability of only about .07--.08. Average
non-null coverage across $\rho$ is then
`r fmt_pct(metric_mean("icc_anchored_at_m10", "fuller_closed_form", .4, "coverage", 3, .25))`
for Fuller,
`r fmt_pct(metric_mean("icc_anchored_at_m10", "lai_2spa", .4, "coverage", 3, .25))`
for 2S-PA, and
`r fmt_pct(metric_mean("icc_anchored_at_m10", "msem", .4, "coverage", 3, .25))`
for MSEM. Holding posterior reliability at .25 raises those values to
`r fmt_pct(metric_mean("posterior_reliability_targeted", "fuller_closed_form", .4, "coverage", 3, .25))`,
`r fmt_pct(metric_mean("posterior_reliability_targeted", "lai_2spa", .4, "coverage", 3, .25))`,
and
`r fmt_pct(metric_mean("posterior_reliability_targeted", "msem", .4, "coverage", 3, .25))`.
Posterior targeting does not make the .25 conditions easy, but it prevents the
ICC manipulation from turning a nominal .25 condition into an effective
reliability-.07 condition.

At $m=25$, the ICC arm provides more posterior information than the named
anchor, and arm differences are small or sometimes favor ICC anchoring. That
does not contradict the design recommendation. It demonstrates why ICC is a
poor common factor for hypotheses about slope-score information: the meaning
of an “ICC-held-constant” condition changes with the observation schedule.

## Paired arm contrasts

```{r paired-coverage-difference-plot, fig.width=10.5, fig.height=6.5}

pair_metrics %>%
  filter(
    method %in% focal_methods,
    standardized_beta_target == .4
  ) %>%
  ggplot(aes(
    x = factor(mean_clus_size, levels = c(3, 10, 25)),
    y = coverage_difference,
    colour = factor(posterior_reliability_anchor),
    shape = factor(marginal_rho)
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey35") +
  geom_point(size = 2, alpha = .85) +
  facet_wrap(~ method_label, ncol = 2) +
  scale_y_continuous(labels = label_percent(accuracy = 1)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = "Observations per cluster (m)",
    y = "Coverage difference: posterior-R minus ICC arm",
    colour = "R anchor",
    shape = "rho",
    title = "Paired coverage contrasts are exactly zero at the m = 10 anchor",
    subtitle = "Positive values favor posterior-reliability targeting"
  ) +
  theme_bridge()
```

```{r paired-summary-table}

paired_summary %>%
  transmute(
    mean_clus_size,
    method_label,
    paired_cells,
    coverage_difference = fmt_pp(mean_coverage_difference),
    se_difference = fmt_num(mean_se_difference),
    eligibility_difference = fmt_pp(mean_eligibility_difference)
  ) %>%
  kable(
    col.names = c(
      "m", "Method", "Paired cells", "Coverage difference",
      "Mean-SE difference", "Eligibility difference"
    ),
    caption = paste(
      "Posterior-reliability arm minus ICC arm, averaged across reliability,",
      "rho, and effect cells. Common random numbers sharpen the comparison."
    )
  )
```

# Standard errors and interval scale

Coverage can fail because estimates are off center, because SEs do not track
sampling variability, because standardized errors have heavy tails, or because
fit selection changes the analyzed sample. The cell-level ratio below compares
mean model SE with the empirical SD among interval-eligible replications. A
value of one is the first-order calibration target. The logarithmic axis keeps
the genuine 2S-PA and MSEM tail cells visible rather than truncating them.

```{r se-calibration-plot, fig.width=11, fig.height=9}

focal %>%
  ggplot(aes(
    x = achieved_reliability,
    y = mean_se_to_interval_empirical_sd,
    colour = arm_label,
    shape = m_label
  )) +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey35") +
  geom_point(alpha = .82, size = 1.8) +
  facet_grid(method_label ~ effect_label) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_y_log10(breaks = c(.5, .75, 1, 1.5, 2, 3, 4)) +
  scale_colour_manual(values = arm_palette) +
  labs(
    x = "Achieved marginal posterior slope reliability",
    y = "Mean model SE / empirical SD (log scale)",
    colour = NULL,
    shape = NULL,
    title = "SE calibration depends on achieved slope information, not the ICC label alone",
    subtitle = "Empirical SD uses the same interval-eligible subset as the mean model SE"
  ) +
  theme_bridge()
```

The oracle ratios cluster around one. Fuller is generally stable, but some
low-information cells have ratios near two. 2S-PA and MSEM show still larger
ratios in a few low-information cells, alongside lower eligibility. These are
not plotting outliers to discard: they are evidence that typical SE calibration
and tail behavior both matter.

The relationship between bias and SE scale makes the Study 5 distinction more
explicit:

```{r bias-se-mechanism-plot, fig.width=10.5, fig.height=7.5}

focal %>%
  filter(method %in% feasible_methods) %>%
  ggplot(aes(
    x = bias,
    y = mean_se_to_interval_empirical_sd,
    colour = arm_label,
    shape = m_label
  )) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey45") +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey45") +
  geom_point(alpha = .82, size = 1.8) +
  facet_grid(method_label ~ effect_label) +
  scale_y_log10(breaks = c(.5, .75, 1, 1.5, 2, 3, 4)) +
  scale_colour_manual(values = arm_palette) +
  labs(
    x = "Condition-level mean bias",
    y = "Mean model SE / empirical SD (log scale)",
    colour = NULL,
    shape = NULL,
    title = "Centering bias and interval-scale failure are distinct",
    subtitle = "A point near bias = 0 can still have badly scaled or heavy-tailed uncertainty"
  ) +
  theme_bridge()
```

At the null, Fuller and 2S-PA have very small average bias but Type I error
below 5%; their conservatism therefore cannot be attributed primarily to
centering. In the lowest-information non-null cells, bias, SE scale, and tail
behavior all contribute. A mean SE/SD ratio by itself cannot guarantee 95%
coverage when standardized errors are asymmetric or heavy-tailed.

# Type I error

```{r type1-plot, fig.width=12, fig.height=10}

focal %>%
  filter(standardized_beta_target == 0) %>%
  ggplot(aes(
    x = mean_clus_size,
    y = type1_error,
    colour = arm_label,
    shape = rho_label,
    group = interaction(arm_label, rho_label)
  )) +
  geom_hline(yintercept = .05, linetype = 2, colour = "grey35") +
  geom_errorbar(
    aes(ymin = type1_error_mc_low, ymax = type1_error_mc_high),
    width = .35,
    alpha = .22,
    linewidth = .35
  ) +
  geom_line(alpha = .45, linewidth = .45) +
  geom_point(size = 1.8) +
  facet_grid(method_label ~ target_label) +
  scale_x_continuous(breaks = c(3, 10, 25)) +
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = arm_palette) +
  labs(
    x = "Observations per cluster (m)",
    y = "Conditional Type I error",
    colour = NULL,
    shape = NULL,
    title = "Null coverage translated to the Type I error scale",
    subtitle = "Vertical bars are 95% Monte Carlo intervals; dashed line is alpha = .05"
  ) +
  theme_bridge()
```

```{r null-summary-table}

null_summary %>%
  transmute(
    arm_label,
    mean_clus_size,
    method_label,
    eligibility = fmt_pct(eligibility),
    mean_null_bias = fmt_num(mean_null_bias),
    conditional_type1 = fmt_pct(conditional_type1),
    operational_type1 = fmt_pct(operational_type1),
    cell_range = sprintf(
      "%s--%s",
      fmt_pct(minimum_type1), fmt_pct(maximum_type1)
    )
  ) %>%
  kable(
    col.names = c(
      "Calibration arm", "m", "Method", "Eligibility", "Mean null bias",
      "Conditional Type I", "Operational Type I", "Cell range"
    ),
    caption = "Null-condition diagnostics, equally weighted across reliability and rho cells."
  )
```

The oracle averages approximately 5% Type I error and is identical across
arms. MSEM is the closest feasible method to nominal: its average null Type I
error is
`r fmt_pct(metric_mean("posterior_reliability_targeted", "msem", 0, "type1_error"))`
under posterior targeting and
`r fmt_pct(metric_mean("icc_anchored_at_m10", "msem", 0, "type1_error"))`
under ICC anchoring. These averages still require cell-level inspection.

Fuller and 2S-PA are conservative at the null. Under posterior targeting their
average Type I errors are
`r fmt_pct(metric_mean("posterior_reliability_targeted", "fuller_closed_form", 0, "type1_error"))`
and
`r fmt_pct(metric_mean("posterior_reliability_targeted", "lai_2spa", 0, "type1_error"))`,
respectively. For Fuller, holding ICC fixed creates a strong $m$-gradient:
Type I error moves from
`r fmt_pct(metric_mean("icc_anchored_at_m10", "fuller_closed_form", 0, "type1_error", 3))`
at $m=3$ to
`r fmt_pct(metric_mean("icc_anchored_at_m10", "fuller_closed_form", 0, "type1_error", 25))`
at $m=25$. Posterior targeting holds it near 2.5%--3% across $m$. Thus an
arm-average Type I error can hide exactly the information drift the bridge was
designed to expose.

For 2S-PA, conditional and operational Type I error differ because some fits
are ineligible. Low operational rejection can reflect both conservative
intervals and failure to return an interval. Both quantities therefore appear
in the table.

# Bias and Monte Carlo uncertainty

```{r bias-plot, fig.width=12, fig.height=10}

focal %>%
  ggplot(aes(
    x = achieved_reliability,
    y = bias,
    ymin = bias_mc_low,
    ymax = bias_mc_high,
    colour = arm_label,
    shape = m_label
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey35") +
  geom_errorbar(width = .008, alpha = .28, linewidth = .35) +
  geom_point(size = 1.8, alpha = .85) +
  facet_grid(method_label ~ effect_label, scales = "free_y") +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = arm_palette) +
  labs(
    x = "Achieved marginal posterior slope reliability",
    y = "Mean bias with 95% Monte Carlo interval",
    colour = NULL,
    shape = NULL,
    title = "Bias is indexed more coherently by achieved slope reliability than by ICC arm",
    subtitle = "Intervals quantify Monte Carlo uncertainty in each condition-level mean bias"
  ) +
  theme_bridge()
```

```{r bias-summary-table}

arm_effect_summary %>%
  transmute(
    arm_label,
    method_label,
    effect = standardized_beta_target,
    cells,
    eligibility = fmt_pct(eligibility),
    mean_bias = fmt_num(mean_bias),
    bias_mc_interval = sprintf(
      "[%s, %s]",
      fmt_num(bias_mc_low), fmt_num(bias_mc_high)
    ),
    mean_absolute_bias = fmt_num(mean_absolute_bias),
    mean_coverage = fmt_pct(mean_coverage),
    success_and_cover = fmt_pct(success_and_cover)
  ) %>%
  kable(
    col.names = c(
      "Calibration arm", "Method", "Beta", "Cells", "Eligibility",
      "Mean bias", "95% MC interval", "Mean |bias|", "Coverage",
      "Success and cover"
    ),
    caption = "Effect-specific performance summaries, equally weighted across design cells."
  )
```

Null bias is close to zero for the focal methods, including Fuller and 2S-PA.
Their low null rejection rates are therefore mainly interval-scale, tail, and
selection phenomena rather than directional centering failures.

For the non-null effect, Fuller is the least biased feasible method. Under
posterior targeting its equal-cell mean bias is
`r fmt_num(metric_mean("posterior_reliability_targeted", "fuller_closed_form", .4, "bias"))`,
compared with
`r fmt_num(metric_mean("posterior_reliability_targeted", "lai_2spa", .4, "bias"))`
for 2S-PA and
`r fmt_num(metric_mean("posterior_reliability_targeted", "msem", .4, "bias"))`
for MSEM. The apparent overall advantage of the ICC arm for 2S-PA and MSEM is
not uniform. ICC anchoring gives much lower-than-target reliability at $m=3$
and higher-than-target reliability at $m=25$, so biases of different signs
and information levels partially cancel.

The .25-anchor conditions make this especially clear. At $m=3$, ICC
anchoring yields a Fuller non-null bias of
`r fmt_num(metric_mean("icc_anchored_at_m10", "fuller_closed_form", .4, "bias", 3, .25))`,
whereas posterior targeting yields
`r fmt_num(metric_mean("posterior_reliability_targeted", "fuller_closed_form", .4, "bias", 3, .25))`.
Holding posterior reliability fixed produces similar small Fuller bias at
$m=10$ and $m=25$. This is strong evidence for the proposed Fuller-focused
paper: posterior reliability makes the intended low-information comparison
coherent across observation schedules, even though Fuller's interval procedure
still needs qualification.

# Estimator-specific conclusions

## Oracle

The oracle is the calibration benchmark. It has complete eligibility, is
centered, and remains near 95% coverage and 5% Type I error. Its arm contrasts
are zero because measurement calibration cannot affect a regression that uses
the true random effects. This validates the structural truth and paired-random-
number implementation.

## Fuller

Fuller is the strongest feasible point estimator in this bridge. It is always
eligible and has much smaller non-null bias than 2S-PA or MSEM when posterior
reliability is held fixed. ICC anchoring creates severe small-$m$, low-anchor
attenuation because the actual slope reliability falls far below the label.

Its intervals are not uniformly calibrated. Null Type I error is about 3%
rather than 5% under posterior targeting, consistent with overcoverage. At the
non-null effect, posterior-targeted coverage averages
`r fmt_pct(metric_mean("posterior_reliability_targeted", "fuller_closed_form", .4, "coverage"))`,
but the .25, $m=3$ cells still under-cover. Thus Study 5's fixed-shape repair
solves the covariance-geometry confound, and this bridge solves the ICC-versus-
information confound, but neither repair proves Fuller's Wald intervals are
nominal in the hardest cells.

## Lai 2S-PA

2S-PA broadly follows the same information gradient as Fuller, but conditional
coverage must be interpreted together with eligibility. At $m=3$, mean
interval eligibility rises by
`r fmt_pp(paired_mean(3, "lai_2spa", "eligibility_difference"))`
when posterior reliability is held at its intended value rather than allowed
to fall under ICC anchoring. At $m=25$, the sign reverses because ICC
anchoring supplies more posterior information. This is fit selection, not only
interval calibration. The probability of success and coverage is therefore a
more meaningful operational quantity than conditional coverage alone.

## MSEM

MSEM gives the best Type I-error balance among the feasible methods and has
near-complete eligibility. Its non-null coverage is less conservative than
Fuller's and 2S-PA's, but positive bias remains in the lower reliability
conditions. At the .25 anchor, holding posterior reliability fixed does not
remove that bias, and some cells remain below 95% coverage. MSEM is therefore a
valuable inferential comparator, not an oracle substitute.

## Naive BLUP negative control

Naive dual-BLUP regression is omitted from the focal plots because rare ratio
failures yield condition-level mean biases and SEs many orders of magnitude
larger than the corrected methods. Including those values on a linear scale
would make the focal comparisons unreadable. It remains an important negative
control: its low coverage, elevated null rejection, and numerical instability
show why Stage-1 uncertainty cannot be ignored.

# Conclusions for Studies 1--4

The bridge supports a simple, coherent explanation for a psychological methods
audience:

1. **Hold the random-effects covariance shape fixed.** Study 5 showed that
   manipulating only $\tau_1^2$ creates a geometry confound. The v2 calibrator
   correctly fixes $\tau_1^2/\tau_0^2$ and the intercept--slope correlation.

2. **Manipulate marginal posterior slope reliability as the shared primary
   information factor.** It directly describes the precision of the latent
   slope score under the actual time design. Unlike a fixed ICC, it does not
   silently change when $m$ changes.

3. **Report ICCs as a crosswalk, not as the primary experimental factor.** The
   intercept ICC, marginal slope ICC, and conditional slope ICC remain useful
   population descriptors. Their relation to posterior reliability should be
   reported for every condition, but no single ICC is a design-invariant
   substitute for posterior reliability.

4. **Report residualized-slope reliability whenever the random intercept and
   slope correlate.** Study 5 showed that marginal and residualized reliability
   can diverge. The bridge does not overturn that conclusion; it addresses a
   different comparison.

5. **Judge estimators with a multidimensional criterion.** For null effects,
   report conditional and operational Type I error, directional rejection,
   bias, and eligibility. For non-null effects, report coverage, bias, empirical
   SD, model SE, standardized-error behavior, eligibility, and success-and-cover.
   An arm-average coverage near 95% is insufficient because conservative null
   cells can cancel liberal non-null cells.

The balanced substantive conclusion is that posterior reliability is the more
defensible factor for testing the paper's key hypotheses, while ICC remains an
important reporting and translation device. Posterior reliability is not a
magic scalar: even when it is held fixed, $m$, the time design, correlation,
and estimator behavior can affect coverage. Its advantage is conceptual and
design-based—it holds constant the information quantity most closely connected
to the Stage-1 slope estimates used downstream.

# Appendix: Fuller variants and robust diagnostics

```{r fuller-variant-table}

fuller_variant_labels <- c(
  fuller_closed_form = "Stabilized",
  fuller_book_preliminary_closed_form = "Book preliminary",
  fuller_book_closed_form = "Book equations",
  fuller_alpha_stepdown_closed_form = "Alpha-stepdown"
)

cell_metrics %>%
  filter(method %in% names(fuller_variant_labels)) %>%
  mutate(variant = unname(fuller_variant_labels[method])) %>%
  group_by(arm_label, variant) %>%
  summarise(
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    null_type1 = mean(
      .data$type1_error[standardized_beta_target == 0],
      na.rm = TRUE
    ),
    nonnull_coverage = mean(
      .data$coverage[standardized_beta_target == .4],
      na.rm = TRUE
    ),
    nonnull_bias = mean(
      .data$bias[standardized_beta_target == .4],
      na.rm = TRUE
    ),
    median_cell_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    maximum_cell_p99_se = max(.data$p99_se, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    eligibility = fmt_pct(eligibility),
    null_type1 = fmt_pct(null_type1),
    nonnull_coverage = fmt_pct(nonnull_coverage),
    nonnull_bias = fmt_num(nonnull_bias),
    median_cell_se_ratio = fmt_num(median_cell_se_ratio),
    maximum_cell_p99_se = fmt_num(maximum_cell_p99_se)
  ) %>%
  kable(
    col.names = c(
      "Calibration arm", "Fuller variant", "Eligibility", "Null Type I",
      "Non-null coverage", "Non-null bias", "Median cell SE/SD",
      "Largest cell p99 SE"
    ),
    caption = paste(
      "Fuller variants retained to show which algebraic choices regularize",
      "the SE tail. The stabilized variant is primary in the report."
    )
  )
```

The variant comparison is diagnostic rather than a license to choose the
best-looking result after observing coverage. The stabilized Fuller path is the
predeclared focal variant because it uses the corrected predictor--outcome
measurement-error covariance and the regularized variance construction. The
book-equation and alpha-stepdown variants document which modification controls
the SE tail and should remain labeled as sensitivity analyses.

```{r session-info}

metadata_files <- list.files(
  output_dir,
  pattern = "_run_metadata[.]csv$",
  full.names = TRUE
)
metadata <- data.table::rbindlist(
  lapply(metadata_files, data.table::fread),
  fill = TRUE,
  use.names = TRUE
) %>%
  tibble::as_tibble()

metadata %>%
  distinct(
    pipeline_version, requested_n_sim, r_version, mplus_version,
    package_openmx_version, package_mplusautomation_version
  ) %>%
  kable(caption = "Simulation software provenance recorded by the production run.")
```
