---
title: "VH Study 4v2: Heterogeneous Cluster Information"
author: "Michael Hallquist and Zach Vig"
date: "26 August 2026"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    embed-resources: true
    html-math-method: katex
params:
  output_dir: "outputs/vig_hallquist_v2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 120, out.width = "100%"
)

suppressPackageStartupMessages({
  library(data.table)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
  library(scales)
})

input_path <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) "")
analysis_dir <- if (nzchar(input_path)) {
  dirname(normalizePath(input_path, mustWork = FALSE))
} else {
  getwd()
}
repo_root <- normalizePath(file.path(analysis_dir, "..", ".."), mustWork = TRUE)
resolve_repo_path <- function(path) {
  if (grepl("^/", path)) path else file.path(repo_root, path)
}
source(
  file.path(repo_root, "vig_hallquist_2026", "vh_designs.R"),
  local = knitr::knit_global()
)

output_dir <- resolve_repo_path(params$output_dir)
condition_dir <- file.path(output_dir, "conditions")
canonical_counts <- study_condition_counts()
canonical_offsets <- c(0L, cumsum(canonical_counts)[-length(canonical_counts)])
names(canonical_offsets) <- names(canonical_counts)
expected_ids <- canonical_offsets[["study4v2"]] +
  seq_len(canonical_counts[["study4v2"]])
summary_paths <- file.path(
  condition_dir,
  sprintf("condition_%04d_summary.csv", expected_ids)
)
missing_summary_paths <- summary_paths[!file.exists(summary_paths)]
if (length(missing_summary_paths) > 0L) {
  stop(
    "Study 4v2 is incomplete: missing ", length(missing_summary_paths),
    " condition summaries. First missing file: ", missing_summary_paths[[1]]
  )
}

cell_metrics <- data.table::rbindlist(
  lapply(summary_paths, data.table::fread),
  fill = TRUE,
  use.names = TRUE
) %>%
  tibble::as_tibble() %>%
  filter(.data$study == "study4v2") %>%
  mutate(
    is_falsification_control = dplyr::coalesce(
      .data$is_falsification_control,
      FALSE
    ),
    sd_standardized_error = suppressWarnings(
      as.numeric(.data$sd_standardized_error)
    )
  )

expected_method_names <- c(
  "oracle_dual", "naive_dual_blup", "closed_form_dual",
  "fuller_closed_form", "fuller_book_preliminary_closed_form",
  "fuller_book_closed_form", "fuller_average_measurement",
  "fuller_alpha_stepdown_closed_form", "lai_2spa", "lai_2spaa", "msem"
)
expected_method_cells <- tidyr::crossing(
  condition_id = expected_ids,
  method = expected_method_names
)
observed_method_cells <- cell_metrics %>%
  count(.data$condition_id, .data$method, name = "summary_rows")
missing_method_cells <- dplyr::anti_join(
  expected_method_cells,
  observed_method_cells,
  by = c("condition_id", "method")
)
unexpected_method_cells <- dplyr::anti_join(
  observed_method_cells,
  expected_method_cells,
  by = c("condition_id", "method")
)
v2_contract_ok <-
  n_distinct(cell_metrics$condition_id) == length(expected_ids) &&
  nrow(missing_method_cells) == 0L &&
  nrow(unexpected_method_cells) == 0L &&
  all(observed_method_cells$summary_rows == 1L) &&
  all(cell_metrics$n_rep == 1000L) &&
  all(cell_metrics$study_version == "v2") &&
  all(cell_metrics$calibration_version == "shape_preserving_v2") &&
  all(cell_metrics$covariance_shape_fixed %in% TRUE) &&
  all(cell_metrics$slope_intercept_variance_ratio == 1)
if (!v2_contract_ok) {
  stop(
    "Refusing to render: Study 4v2 summaries violate the v2-only ",
    "completion, method, replication, or calibration contract."
  )
}
summary_file_info <- file.info(summary_paths)
summary_artifact_oldest <- min(summary_file_info$mtime)
summary_artifact_newest <- max(summary_file_info$mtime)

method_labels <- c(
  oracle_dual = "Oracle",
  naive_dual_blup = "Naive dual BLUP",
  closed_form_dual = "Unshrunk-score OLS",
  fuller_closed_form = "Fuller (row-specific)",
  fuller_book_preliminary_closed_form = "Fuller (book preliminary)",
  fuller_book_closed_form = "Fuller (book equations)",
  fuller_average_measurement = "Fuller (average measurement)",
  fuller_alpha_stepdown_closed_form = "Fuller (alpha-stepdown)",
  lai_2spa = "Lai 2S-PA (row-specific)",
  lai_2spaa = "Lai 2S-PAA (average)",
  msem = "MSEM"
)
stopifnot(setequal(names(method_labels), expected_method_names))

profile_labels <- c(
  homogeneous = "Homogeneous",
  moderate = "Moderate heterogeneity",
  severe = "Severe heterogeneity",
  severe_information_matched = "Slope-information control"
)
profile_palette <- c(
  "Homogeneous" = "#4E79A7",
  "Moderate heterogeneity" = "#F28E2B",
  "Severe heterogeneity" = "#E15759",
  "Slope-information control" = "#59A14F"
)
primary_methods <- c(
  "oracle_dual", "fuller_closed_form", "fuller_average_measurement",
  "lai_2spa", "lai_2spaa", "msem"
)
diagnostic_methods <- c("naive_dual_blup", "closed_form_dual")
fuller_methods <- c(
  "fuller_closed_form", "fuller_alpha_stepdown_closed_form",
  "fuller_book_preliminary_closed_form", "fuller_book_closed_form"
)

cell_metrics <- cell_metrics %>%
  mutate(
    method_label = factor(
      unname(method_labels[.data$method]),
      levels = unname(method_labels)
    ),
    profile_label = factor(
      unname(profile_labels[.data$information_profile]),
      levels = unname(profile_labels)
    ),
    reliability_label = factor(
      paste0("R = ", sprintf("%.2f", .data$target_reliability)),
      levels = paste0("R = ", sprintf("%.2f", c(.25, .5, .8)))
    )
  )

pooled_mc_se <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) == 0L) return(NA_real_)
  sqrt(sum(x^2)) / length(x)
}
fmt_pct <- function(x, accuracy = .1) percent(x, accuracy = accuracy)
fmt_num <- function(x, digits = 3) formatC(x, digits = digits, format = "f")
theme_vh <- function() {
  theme_bw(base_size = 10) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      strip.background = element_rect(fill = "grey95", colour = "grey75")
    )
}

factorial_cells <- cell_metrics %>%
  filter(!.data$is_falsification_control)
control_cells <- cell_metrics %>%
  filter(.data$is_falsification_control)

performance_grid <- factorial_cells %>%
  filter(.data$standardized_beta_target > 0) %>%
  group_by(
    .data$method, .data$method_label, .data$information_profile,
    .data$profile_label, .data$num_clus, .data$target_reliability,
    .data$reliability_label
  ) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    bias_mc_se = pooled_mc_se(.data$mc_se_bias),
    bias_mc_low = .data$bias - qnorm(.975) * .data$bias_mc_se,
    bias_mc_high = .data$bias + qnorm(.975) * .data$bias_mc_se,
    coverage = mean(.data$coverage, na.rm = TRUE),
    coverage_mc_se = pooled_mc_se(.data$coverage_mc_se),
    coverage_mc_low = pmax(0, .data$coverage - qnorm(.975) * .data$coverage_mc_se),
    coverage_mc_high = pmin(1, .data$coverage + qnorm(.975) * .data$coverage_mc_se),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

null_grid <- factorial_cells %>%
  filter(.data$standardized_beta_target == 0) %>%
  group_by(
    .data$method, .data$method_label, .data$information_profile,
    .data$profile_label, .data$num_clus, .data$target_reliability,
    .data$reliability_label
  ) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    type1_error = mean(.data$type1_error, na.rm = TRUE),
    type1_mc_se = pooled_mc_se(.data$type1_error_mc_se),
    type1_mc_low = pmax(0, .data$type1_error - qnorm(.975) * .data$type1_mc_se),
    type1_mc_high = pmin(1, .data$type1_error + qnorm(.975) * .data$type1_mc_se),
    operational_type1_error = mean(
      .data$operational_type1_error,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

method_summary <- factorial_cells %>%
  filter(.data$standardized_beta_target > 0) %>%
  group_by(.data$method, .data$method_label) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    coverage = mean(.data$coverage, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    cells_below_nominal = sum(.data$coverage_mc_high < .95, na.rm = TRUE),
    cells_above_nominal = sum(.data$coverage_mc_low > .95, na.rm = TRUE),
    .groups = "drop"
  )

null_summary <- factorial_cells %>%
  filter(.data$standardized_beta_target == 0) %>%
  group_by(.data$method, .data$method_label) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    type1_error = mean(.data$type1_error, na.rm = TRUE),
    operational_type1_error = mean(
      .data$operational_type1_error,
      na.rm = TRUE
    ),
    coverage = mean(.data$coverage, na.rm = TRUE),
    .groups = "drop"
  )

metric_for <- function(summary, method, metric) {
  value <- summary[summary$method == method, metric, drop = TRUE]
  if (length(value) != 1L) NA_real_ else value
}

profile_calibration <- factorial_cells %>%
  filter(.data$method == "oracle_dual") %>%
  distinct(
    .data$condition_id, .data$information_profile, .data$profile_label,
    .data$target_reliability, .data$reliability_label,
    .data$achieved_reliability,
    .data$reliability_min, .data$reliability_max,
    .data$reliability_sd, .data$num_clus
  )

pair_cells <- factorial_cells %>%
  filter(
    .data$standardized_beta_target > 0,
    .data$method %in% c(
      "fuller_closed_form", "fuller_average_measurement",
      "lai_2spa", "lai_2spaa"
    )
  ) %>%
  select(
    .data$condition_id, .data$information_profile, .data$profile_label,
    .data$target_reliability, .data$num_clus, .data$method,
    .data$bias, .data$coverage, .data$interval_eligibility
  ) %>%
  pivot_wider(
    names_from = method,
    values_from = c(bias, coverage, interval_eligibility)
  ) %>%
  mutate(
    fuller_bias_difference =
      .data$bias_fuller_average_measurement - .data$bias_fuller_closed_form,
    fuller_coverage_difference =
      .data$coverage_fuller_average_measurement - .data$coverage_fuller_closed_form,
    lai_bias_difference = .data$bias_lai_2spaa - .data$bias_lai_2spa,
    lai_coverage_difference = .data$coverage_lai_2spaa - .data$coverage_lai_2spa
  )
```

# Executive summary

Study 4v2 asks whether methods that use row-specific Stage-1 uncertainty retain
their advantage when clusters differ sharply in measurement information. The
amended design uses the same shape-preserving posterior-reliability calibrator
as Studies 1v2--3v2, then varies the distribution of cluster sizes while
holding the target mean reliability interpretable.

All 111 planned conditions are complete: 108 factorial cells plus three
slope-information controls. Each condition contains 1,000 replications and all
11 methods. This report has a strict v2 guard and cannot silently read the
legacy Study 4 cache.

The main results are:

1. Row-specific stabilized Fuller is nearly centered (mean bias
   `r fmt_num(metric_for(method_summary, "fuller_closed_form", "bias"))`)
   with `r fmt_pct(metric_for(method_summary, "fuller_closed_form", "coverage"))`
   non-null conditional coverage and complete interval eligibility.
2. Fuller is mildly conservative under the null: average Type I error is
   `r fmt_pct(metric_for(null_summary, "fuller_closed_form", "type1_error"))`.
   Lai 2S-PA is similarly conservative, while MSEM is approximately nominal.
3. Averaging measurement information is not harmless. Average-measurement
   Fuller has more negative bias and lower coverage than row-specific Fuller,
   especially as information becomes heterogeneous. Its success-and-cover
   probability is also reduced by incomplete eligibility.
4. Lai's averaging comparison is more nuanced. 2S-PAA reduces the positive
   bias of row-specific 2S-PA on average, but has lower coverage. Both have
   incomplete fit eligibility, so conditional coverage overstates operational
   success.
5. The severe slope-information control behaves more like homogeneous
   measurement than the ordinary severe profile. Unequal cluster counts alone
   are therefore not the operative mechanism; heterogeneity in slope
   information and the full posterior-error matrices matters.

# Scientific motivation

Real repeated-measures data rarely give every person the same information.
Participants miss assessments, enter at different times, or contribute
measurements at time points with different leverage for estimating a slope.
Two people can therefore have the same kind of latent slope but very different
posterior uncertainty. Study 4 asks whether Stage 2 must retain that
person-specific uncertainty or whether replacing it by one average matrix is
adequate.

The structural model is the Study 2 predictor model,

$$
y_{it}=\gamma_0+\gamma_1x_{it}+u_{0i}+u_{1i}x_{it}+e_{it},
\qquad
z_i=\beta_0+\beta_{0u}u_{0i}+\beta u_{1i}+\varepsilon_i,
$$

but the cluster-specific observation schedules vary. This isolates a practical
question that a homogeneous design cannot answer: whether the *distribution*
of slope information matters even when the mean reliability target is held
fixed.

# Design and v2-only completion contract

The factorial portion crosses three posterior-reliability targets, two
intercept--slope correlations, three information profiles, three cluster counts,
and null versus non-null structural effects. The three additional control
conditions use severely unequal cluster sizes but match slope information.

The method set is designed around the row-specific-versus-average contrast:

- **Oracle**, **naive dual-BLUP OLS**, **unshrunk-score OLS**, and **MSEM** have
  the same benchmark or negative-control roles as in Study 2.
- **Fuller (row-specific)** carries each cluster's Stage-1 error matrix into the
  correction; **Fuller (average measurement)** replaces those matrices by one
  average. The latter is a deliberate approximation, not the proposed method.
- **Lai 2S-PA (row-specific)** and **Lai 2S-PAA (average)** provide the analogous
  contrast in the latent path-analysis framework.
- Fuller book-equation and alpha-stepdown variants remain variance-tail and
  algebra sensitivity checks.

```{r completion-table}
tibble::tibble(
  item = c(
    "Expected conditions", "Observed conditions", "Factorial conditions",
    "Slope-information controls", "Methods per condition",
    "Replications per condition", "Study version", "Calibration version",
    "Covariance shape fixed", "Slope/intercept variance ratio",
    "Summary artifacts read", "Oldest summary artifact (local time)",
    "Newest summary artifact (local time)"
  ),
  value = c(
    length(expected_ids), n_distinct(cell_metrics$condition_id),
    n_distinct(factorial_cells$condition_id),
    n_distinct(control_cells$condition_id), length(expected_method_names),
    unique(cell_metrics$n_rep), unique(cell_metrics$study_version),
    unique(cell_metrics$calibration_version),
    all(cell_metrics$covariance_shape_fixed),
    unique(cell_metrics$slope_intercept_variance_ratio),
    length(summary_paths),
    format(summary_artifact_oldest, tz = "America/New_York"),
    format(summary_artifact_newest, tz = "America/New_York")
  )
) %>%
  kable(align = c("l", "r"))
```

The report reads these exact condition summaries directly, with no legacy or
consolidated-cache fallback.

{{< include _vh_v2_metric_guide.qmd >}}

# Reliability profiles and manipulation checks

```{r reliability-profile-plot, fig.width=11, fig.height=7}
profile_calibration %>%
  ggplot(aes(
    x = factor(.data$num_clus),
    y = .data$achieved_reliability,
    ymin = .data$reliability_min,
    ymax = .data$reliability_max,
    colour = .data$profile_label
  )) +
  geom_hline(
    aes(yintercept = .data$target_reliability),
    linetype = 2,
    colour = "grey55"
  ) +
  geom_linerange(position = position_dodge(width = .45), alpha = .6) +
  geom_point(position = position_dodge(width = .45), size = 2) +
  facet_wrap(~ reliability_label, nrow = 1) +
  scale_y_continuous(limits = c(0, 1), labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = profile_palette) +
  labs(
    x = "Number of clusters",
    y = "Mean and range of cluster-specific reliability",
    colour = NULL,
    title = "Profiles share a mean target but differ in reliability dispersion",
    subtitle = "Points are achieved means; vertical ranges show the least and most informative clusters"
  ) +
  theme_vh()
```

This is the key Study 4 manipulation. The mean target is comparable across
profiles, while the severe condition contains clusters with very different
posterior uncertainties. The row-specific methods use those differences; the
average-measurement variants deliberately discard them.

# Bias

```{r primary-bias-plot, fig.width=12, fig.height=9}
performance_grid %>%
  filter(.data$method %in% primary_methods) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$bias,
    ymin = .data$bias_mc_low,
    ymax = .data$bias_mc_high,
    colour = .data$profile_label,
    linetype = factor(.data$num_clus),
    group = interaction(.data$profile_label, .data$num_clus)
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey40") +
  geom_errorbar(width = .015, alpha = .25, linewidth = .35) +
  geom_line(alpha = .6) +
  geom_point(size = 1.8) +
  facet_wrap(~ method_label, scales = "free_y", ncol = 2) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_colour_manual(values = profile_palette) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Mean bias with 95% Monte Carlo CI",
    colour = NULL,
    linetype = "Clusters",
    title = "Row-specific Fuller remains centered under heterogeneous information",
    subtitle = "Each point averages three rho cells; free y-scales preserve method-specific patterns"
  ) +
  theme_vh()
```

Row-specific Fuller changes little as reliability dispersion increases.
Average-measurement Fuller is more negatively biased, consistent with replacing
heterogeneous measurement-error matrices by a single average correction. 2S-PA
has positive bias, whereas averaging in 2S-PAA shifts the mean bias toward zero
but does not uniformly improve interval behavior. MSEM also has modest positive
bias in difficult cells.

## Negative-control estimators

```{r diagnostic-bias-plot, fig.width=10, fig.height=6.5}
factorial_cells %>%
  filter(
    .data$method %in% diagnostic_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$target_reliability,
    .data$reliability_label, .data$profile_label
  ) %>%
  summarise(
    median_cell_bias = median(.data$bias, na.rm = TRUE),
    q10 = quantile(.data$bias, .10, na.rm = TRUE),
    q90 = quantile(.data$bias, .90, na.rm = TRUE),
    max_absolute_cell_bias = max(abs(.data$bias), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$median_cell_bias,
    ymin = .data$q10,
    ymax = .data$q90,
    colour = .data$profile_label
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey40") +
  geom_linerange(alpha = .5, position = position_dodge(width = .035)) +
  geom_point(size = 1.8, position = position_dodge(width = .035)) +
  facet_wrap(~ method_label, scales = "free_y") +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_colour_manual(values = profile_palette) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Median cell bias (10th--90th percentile)",
    colour = NULL,
    title = "Unshrinking alone remains attenuated; dual BLUP regression has extreme tails",
    subtitle = "Robust summaries prevent rare coefficient explosions from compressing the other methods"
  ) +
  theme_vh()
```

As in Study 2v2, the naive dual-BLUP estimator has rare enormous coefficients
from near-collinearity. Those failures remain in the summaries but are not
allowed to set a common plot scale. Unshrunk-score OLS shows the expected
attenuation because score correction without a Stage-2 measurement-error
correction is insufficient for a predictor.

# Coverage and SE calibration

```{r primary-coverage-plot, fig.width=12, fig.height=9}
performance_grid %>%
  filter(.data$method %in% primary_methods) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$coverage,
    ymin = .data$coverage_mc_low,
    ymax = .data$coverage_mc_high,
    colour = .data$profile_label,
    linetype = factor(.data$num_clus),
    group = interaction(.data$profile_label, .data$num_clus)
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey35") +
  geom_errorbar(width = .015, alpha = .25, linewidth = .35) +
  geom_line(alpha = .6) +
  geom_point(size = 1.8) +
  facet_wrap(~ method_label, ncol = 2) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(limits = c(.75, 1), labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = profile_palette) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Conditional 95% CI coverage",
    colour = NULL,
    linetype = "Clusters",
    title = "Row-specific Fuller is more stable than average-measurement Fuller",
    subtitle = "Conditional 2S-PA coverage is often conservative but excludes ineligible fits"
  ) +
  theme_vh()
```

Row-specific Fuller averages
`r fmt_pct(metric_for(method_summary, "fuller_closed_form", "coverage"))`
coverage, compared with
`r fmt_pct(metric_for(method_summary, "fuller_average_measurement", "coverage"))`
for average-measurement Fuller. The difference is not merely cosmetic: the
average method is also less often eligible. Row-specific 2S-PA has high
conditional coverage, but only
`r fmt_pct(metric_for(method_summary, "lai_2spa", "success_and_cover"))`
of all non-null replications both yield an eligible interval and cover.

```{r se-calibration-plot, fig.width=12, fig.height=8}
performance_grid %>%
  filter(.data$method %in% primary_methods) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$se_ratio,
    colour = .data$profile_label,
    linetype = factor(.data$num_clus),
    group = interaction(.data$profile_label, .data$num_clus)
  )) +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey35") +
  geom_line(alpha = .6) +
  geom_point(size = 1.8) +
  facet_wrap(~ method_label, ncol = 2) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(limits = c(.7, 1.35)) +
  scale_colour_manual(values = profile_palette) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Median condition-level SE ratio",
    colour = NULL,
    linetype = "Clusters",
    title = "Typical SE scale is close to one for the main methods",
    subtitle = paste(
      "Each ratio is mean reported SE / empirical sampling SD;",
      "lines connect reliability levels within a profile and N"
    )
  ) +
  theme_vh()
```

```{r bias-se-mechanism-plot, fig.width=12, fig.height=8}
factorial_cells %>%
  filter(
    .data$method %in% primary_methods,
    .data$standardized_beta_target > 0
  ) %>%
  ggplot(aes(
    x = .data$bias,
    y = .data$mean_se_to_interval_empirical_sd,
    colour = .data$coverage,
    shape = .data$profile_label
  )) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey45") +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey45") +
  geom_point(alpha = .6, size = 1.35) +
  facet_wrap(~ method_label, scales = "free", ncol = 2) +
  scale_colour_viridis_c(
    option = "C", limits = c(0, 1),
    labels = percent_format(accuracy = 1)
  ) +
  labs(
    x = "Condition-level mean bias",
    y = "Condition-level mean-SE / empirical-SD ratio",
    colour = "Conditional coverage",
    shape = NULL,
    title = "Heterogeneity can affect centering without changing typical SE scale",
    subtitle = "Free scales separate method-specific mechanisms and retain difficult cells"
  ) +
  theme_vh()
```

# Row-specific versus average measurement information

```{r row-average-difference-plot, fig.width=11, fig.height=7.5}
pair_cells %>%
  group_by(
    .data$profile_label, .data$target_reliability,
    .data$num_clus
  ) %>%
  summarise(
    `Fuller: average - row-specific` = mean(
      .data$fuller_coverage_difference,
      na.rm = TRUE
    ),
    `Lai: average - row-specific` = mean(
      .data$lai_coverage_difference,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(
      `Fuller: average - row-specific`,
      `Lai: average - row-specific`
    ),
    names_to = "comparison",
    values_to = "coverage_difference"
  ) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$coverage_difference,
    colour = .data$profile_label,
    linetype = factor(.data$num_clus),
    group = interaction(.data$profile_label, .data$num_clus)
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey40") +
  geom_line(alpha = .6) +
  geom_point(size = 1.8) +
  facet_wrap(~ comparison, nrow = 1) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(labels = percent_format(accuracy = .1)) +
  scale_colour_manual(values = profile_palette) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Conditional coverage difference",
    colour = NULL,
    linetype = "Clusters",
    title = "Averaging measurement information changes the two methods differently",
    subtitle = "Negative values favor the row-specific estimator"
  ) +
  theme_vh()
```

The Fuller comparison gives the cleanest support for the row-specific
principle: its average-measurement approximation generally loses coverage and
introduces negative bias as heterogeneity grows. Lai's comparison is not a
simple dominance result. Averaging can reduce 2S-PA's positive bias, but it also
changes coverage and does not repair fit eligibility. This distinction should
be described as an estimator-specific consequence of averaging, not as a
universal theorem that row-specific corrections improve every metric.

# Type I error and operational performance

```{r type1-plot, fig.width=12, fig.height=8}
null_grid %>%
  filter(.data$method %in% primary_methods) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$type1_error,
    ymin = .data$type1_mc_low,
    ymax = .data$type1_mc_high,
    colour = .data$profile_label,
    linetype = factor(.data$num_clus),
    group = interaction(.data$profile_label, .data$num_clus)
  )) +
  geom_hline(yintercept = .05, linetype = 2, colour = "grey35") +
  geom_errorbar(width = .015, alpha = .25, linewidth = .35) +
  geom_line(alpha = .6) +
  geom_point(size = 1.8) +
  facet_wrap(~ method_label, ncol = 2) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(limits = c(0, .12), labels = percent_format(accuracy = .1)) +
  scale_colour_manual(values = profile_palette) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Conditional Type I error",
    colour = NULL,
    linetype = "Clusters",
    title = "Fuller and both 2S-PA variants overcover under the null",
    subtitle = "MSEM and the oracle remain near the nominal 5% rejection rate"
  ) +
  theme_vh()
```

Null rejection makes the overcoverage evidence explicit. Row-specific Fuller,
average Fuller, 2S-PA, and 2S-PAA all reject less than 5% on average. MSEM's
average Type I error is
`r fmt_pct(metric_for(null_summary, "msem", "type1_error"))`, compared with
`r fmt_pct(metric_for(null_summary, "fuller_closed_form", "type1_error"))`
for row-specific Fuller. Thus, Fuller protects more strongly against false
positives, but its coverage advantage partly reflects conservative intervals.

```{r conditional-operational-plot, fig.width=12, fig.height=8}
performance_grid %>%
  filter(.data$method %in% primary_methods) %>%
  group_by(
    .data$method, .data$method_label, .data$target_reliability,
    .data$reliability_label
  ) %>%
  summarise(
    conditional_coverage = mean(.data$coverage, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    eligibility = mean(.data$eligibility, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(conditional_coverage, success_and_cover),
    names_to = "performance",
    values_to = "rate"
  ) %>%
  mutate(
    performance = factor(
      .data$performance,
      levels = c("conditional_coverage", "success_and_cover"),
      labels = c("Conditional coverage", "Success-and-cover")
    )
  ) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$rate,
    colour = .data$performance,
    group = .data$performance
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey50") +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ method_label, ncol = 2) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(limits = c(.65, 1), labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = c("#4E79A7", "#E15759")) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Probability",
    colour = NULL,
    title = "Conditional coverage is insufficient when fit eligibility differs",
    subtitle = "The operational gap is largest for the 2S-PA and average-measurement methods"
  ) +
  theme_vh()
```

# Direct comparison: row-specific stabilized Fuller, Lai 2S-PA, and MSEM

The primary publication comparison uses the three methods that propagate
row-specific information without deliberately averaging it away. Row-specific
Fuller corrects the Stage-2 moments using each cluster's Stage-1 error matrix;
Lai 2S-PA represents the cluster-specific scores and uncertainties in an OpenMx
latent-variable model; and MSEM estimates the repeated-measures and structural
models jointly in Mplus. Because all three methods see the same simulated
sample in each condition, matched-cell differences isolate estimator behavior
from changes in reliability, information heterogeneity, cluster count, and
intercept--slope correlation.

```{r direct-comparison-setup, include=FALSE}
comparison_methods <- c("fuller_closed_form", "lai_2spa", "msem")
comparison_labels <- c(
  fuller_closed_form = "Fuller (row-specific)",
  lai_2spa = "Lai 2S-PA (row-specific)",
  msem = "MSEM"
)

comparison_nonnull <- factorial_cells %>%
  filter(
    .data$method %in% comparison_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$information_profile, .data$profile_label
  ) %>%
  summarise(
    cells = n(),
    mean_bias = mean(.data$bias, na.rm = TRUE),
    mean_absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    conditional_coverage = mean(.data$coverage, na.rm = TRUE),
    interval_eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

comparison_null <- factorial_cells %>%
  filter(
    .data$method %in% comparison_methods,
    .data$standardized_beta_target == 0
  ) %>%
  group_by(
    .data$method, .data$information_profile, .data$profile_label
  ) %>%
  summarise(
    conditional_type1 = mean(.data$type1_error, na.rm = TRUE),
    operational_type1 = mean(.data$operational_type1_error, na.rm = TRUE),
    .groups = "drop"
  )

comparison_summary <- comparison_nonnull %>%
  left_join(
    comparison_null,
    by = c("method", "information_profile", "profile_label")
  ) %>%
  mutate(
    method_label = factor(
      unname(comparison_labels[.data$method]),
      levels = unname(comparison_labels)
    )
  )

comparison_metric <- function(method, profile, metric) {
  value <- comparison_summary[
    comparison_summary$method == method &
      comparison_summary$information_profile == profile,
    metric,
    drop = TRUE
  ]
  if (length(value) != 1L) NA_real_ else value
}

comparison_wide <- factorial_cells %>%
  filter(
    .data$method %in% comparison_methods,
    .data$standardized_beta_target > 0
  ) %>%
  select(
    .data$condition_id, .data$information_profile, .data$profile_label,
    .data$target_reliability, .data$method, .data$bias, .data$coverage,
    .data$interval_eligibility, .data$mean_se_to_interval_empirical_sd
  ) %>%
  pivot_wider(
    names_from = .data$method,
    values_from = c(
      bias, coverage, interval_eligibility,
      mean_se_to_interval_empirical_sd
    )
  )

make_paired_comparison <- function(wide, comparator_key, comparator_label) {
  tibble::tibble(
    comparator = comparator_label,
    information_profile = wide$information_profile,
    profile_label = wide$profile_label,
    target_reliability = wide$target_reliability,
    absolute_bias_difference =
      abs(wide$bias_fuller_closed_form) -
      abs(wide[[paste0("bias_", comparator_key)]]),
    absolute_coverage_error_difference =
      abs(wide$coverage_fuller_closed_form - .95) -
      abs(wide[[paste0("coverage_", comparator_key)]] - .95),
    interval_failure_difference =
      (1 - wide$interval_eligibility_fuller_closed_form) -
      (1 - wide[[paste0("interval_eligibility_", comparator_key)]]),
    absolute_log_se_error_difference =
      abs(log(wide$mean_se_to_interval_empirical_sd_fuller_closed_form)) -
      abs(log(wide[[paste0(
        "mean_se_to_interval_empirical_sd_", comparator_key
      )]]))
  )
}

paired_comparisons <- bind_rows(
  make_paired_comparison(
    comparison_wide,
    "lai_2spa",
    "Lai 2S-PA (row-specific)"
  ),
  make_paired_comparison(comparison_wide, "msem", "MSEM")
) %>%
  pivot_longer(
    cols = c(
      absolute_bias_difference,
      absolute_coverage_error_difference,
      interval_failure_difference,
      absolute_log_se_error_difference
    ),
    names_to = "criterion",
    values_to = "fuller_minus_comparator"
  ) %>%
  mutate(
    criterion = factor(
      .data$criterion,
      levels = c(
        "absolute_bias_difference",
        "absolute_coverage_error_difference",
        "interval_failure_difference",
        "absolute_log_se_error_difference"
      ),
      labels = c(
        "Absolute bias",
        "Absolute conditional-coverage error",
        "Interval-failure probability",
        "Absolute log(SE / empirical SD)"
      )
    )
  ) %>%
  filter(is.finite(.data$fuller_minus_comparator))
```

```{r direct-comparison-plot, fig.width=12, fig.height=9}
paired_comparisons %>%
  ggplot(aes(
    x = .data$profile_label,
    y = .data$fuller_minus_comparator,
    fill = .data$comparator
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey35") +
  geom_boxplot(outlier.alpha = .18, width = .65) +
  facet_grid(.data$criterion ~ .data$comparator, scales = "free_y") +
  scale_fill_manual(values = c(
    "Lai 2S-PA (row-specific)" = "#4E79A7",
    "MSEM" = "#59A14F"
  )) +
  labs(
    x = "Cluster-information profile",
    y = "Fuller minus comparator (negative favors Fuller)",
    fill = NULL,
    title = "Fuller's clearest advantages are centering and interval availability",
    subtitle = paste(
      "Differences are computed within the same non-null design cell;",
      "zero denotes equal performance"
    )
  ) +
  theme_vh() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 20, hjust = 1)
  )
```

The matched-cell comparison clarifies the tradeoff. Row-specific Fuller has
complete interval eligibility in every primary condition and lower absolute
bias than the two principal comparators in most cells. Lai 2S-PA often has
competitive conditional coverage, but its probability of producing an
eligible interval and covering is lower because difficult fits are excluded.
MSEM is nearly always available and has null rejection closer to 5%, but its
non-null bias and coverage are not uniformly better than Fuller's.

```{r direct-comparison-table}
comparison_summary %>%
  transmute(
    Method = .data$method_label,
    Profile = .data$profile_label,
    `Mean bias` = round(.data$mean_bias, 3),
    `Mean |bias|` = round(.data$mean_absolute_bias, 3),
    `Conditional coverage` = percent(
      .data$conditional_coverage,
      accuracy = .1
    ),
    `Interval eligible` = percent(.data$interval_eligibility, accuracy = .1),
    `Success-and-cover` = percent(.data$success_and_cover, accuracy = .1),
    `Median condition-level SE ratio` = round(.data$median_se_ratio, 3),
    `Null Type I error` = percent(.data$conditional_type1, accuracy = .1),
    `Operational null rejection` = percent(
      .data$operational_type1,
      accuracy = .1
    )
  ) %>%
  arrange(.data$Profile, .data$Method) %>%
  kable(align = c("l", "l", rep("r", 8)))
```

At severe heterogeneity, Fuller's success-and-cover probability is
`r fmt_pct(comparison_metric("fuller_closed_form", "severe", "success_and_cover"))`,
compared with
`r fmt_pct(comparison_metric("lai_2spa", "severe", "success_and_cover"))`
for row-specific Lai 2S-PA and
`r fmt_pct(comparison_metric("msem", "severe", "success_and_cover"))`
for MSEM. This operational contrast is the strongest direct evidence for
propagating row-specific uncertainty with a correction that remains available.
It does not imply uniform dominance: MSEM provides more nearly nominal null
calibration, and the separate average-information analysis below addresses a
different question about what is lost by replacing row-specific uncertainty
with a common matrix.

# Slope-information controls

The three control conditions retain severe cluster-size imbalance but match
slope information. They test whether unequal counts alone produce the
row-specific-versus-average differences.

```{r control-summary-table}
control_cells %>%
  filter(.data$method %in% primary_methods) %>%
  transmute(
    `Number of clusters` = .data$num_clus,
    Method = .data$method_label,
    Bias = round(.data$bias, 3),
    `Conditional coverage` = percent(.data$coverage, accuracy = .1),
    `Interval eligible` = percent(.data$interval_eligibility, accuracy = .1),
    `Condition-level mean-SE / empirical-SD ratio` = round(
      .data$mean_se_to_interval_empirical_sd,
      3
    )
  ) %>%
  arrange(.data$`Number of clusters`, .data$Method) %>%
  kable(align = c("r", "l", "r", "r", "r", "r"))
```

The controls substantially narrow the row-versus-average discrepancy. This
supports the intended mechanism: unequal cluster counts matter because they
create heterogeneous posterior slope information, not merely because the data
set contains small and large clusters.

# Fuller variants

```{r fuller-variant-plot, fig.width=11, fig.height=7}
factorial_cells %>%
  filter(
    .data$method %in% fuller_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$target_reliability,
    .data$reliability_label
  ) %>%
  summarise(
    coverage = mean(.data$coverage, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = .data$median_se_ratio,
    y = .data$coverage,
    colour = .data$reliability_label
  )) +
  geom_vline(xintercept = 1, linetype = 2, colour = "grey45") +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey45") +
  geom_point(size = 2) +
  facet_wrap(~ method_label, scales = "free_x", ncol = 2) +
  scale_x_log10() +
  scale_y_continuous(limits = c(.82, 1), labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = c(
    "R = 0.25" = "#E15759", "R = 0.50" = "#F28E2B",
    "R = 0.80" = "#4E79A7"
  )) +
  labs(
    x = "Median condition-level SE ratio (log scale)",
    y = "Conditional coverage",
    colour = NULL,
    title = "Stabilized Fuller controls the SE-tail behavior of the book variants",
    subtitle = "Equation-faithful variants are retained as sensitivity analyses, not primary estimators"
  ) +
  theme_vh()
```

# Conclusions

Study 4v2 supports a coherent message for the VH series. Posterior reliability
can be manipulated while preserving latent covariance shape, and cluster-level
heterogeneity can then be studied as a separate design feature. This is much
easier to explain than changing only the slope variance and hoping that the
resulting ICC, reliability, and covariance geometry remain comparable.

The substantive conclusions are balanced:

- Row-specific stabilized Fuller performs well under heterogeneous information.
  It is nearly unbiased, has complete eligibility, and keeps typical SE scale
  near the empirical sampling SD. Its null intervals are conservative, so the
  evidence supports dependable false-positive control rather than exact
  nominality.
- Averaging Stage-1 error information can be consequential. For Fuller it
  generally introduces attenuation and lower coverage as heterogeneity grows.
  This directly supports carrying cluster-specific posterior uncertainty into
  Stage 2.
- The Lai comparison shows that row-specific information is not guaranteed to
  improve every metric. 2S-PAA reduces average positive bias but has lower
  coverage, and both Lai variants lose operational performance through fit
  ineligibility.
- MSEM remains a valuable benchmark with almost complete eligibility and
  approximately nominal Type I error. Its modest positive bias and non-null
  undercoverage in some difficult cells caution against treating it as an
  infallible reference.
- The slope-information controls show that the operative problem is not
  unequal cluster size itself. It is heterogeneity in the measurement
  information attached to the latent slope.

For the paper, Study 4 should therefore be framed as evidence for preserving
and propagating the full row-specific uncertainty structure. The primary
Fuller estimator should be the corrected, stabilized implementation; the book
and alpha-stepdown variants remain informative sensitivity analyses for how
regularization changes SE tails.

# Appendix

## Compact method summaries

```{r method-summary-table}
method_summary %>%
  transmute(
    Method = .data$method_label,
    Cells = .data$cells,
    `Interval eligible` = percent(.data$eligibility, accuracy = .1),
    `Mean bias` = round(.data$bias, 3),
    `Mean |bias|` = round(.data$absolute_bias, 3),
    `Conditional coverage` = percent(.data$coverage, accuracy = .1),
    `Success-and-cover` = percent(.data$success_and_cover, accuracy = .1),
    `Median condition-level SE ratio` = round(.data$median_se_ratio, 3),
    `MC interval below .95` = .data$cells_below_nominal,
    `MC interval above .95` = .data$cells_above_nominal
  ) %>%
  kable(
    align = c("l", rep("r", 9)),
    caption = paste(
      "Primary factorial cells only. Cell counts require the condition-level",
      "95% Monte Carlo interval to lie wholly below or above .95."
    )
  )
```

```{r null-summary-table}
null_summary %>%
  transmute(
    Method = .data$method_label,
    Cells = .data$cells,
    `Interval eligible` = percent(.data$eligibility, accuracy = .1),
    `Mean bias` = round(.data$bias, 3),
    `Conditional Type I error` = percent(.data$type1_error, accuracy = .1),
    `Operational Type I error` = percent(
      .data$operational_type1_error,
      accuracy = .1
    ),
    `Conditional coverage` = percent(.data$coverage, accuracy = .1)
  ) %>%
  kable(align = c("l", rep("r", 6)))
```

{{< include _vh_v2_provenance.qmd >}}

## Session information

```{r session-info}
sessionInfo()
```
