---
title: "VH Study 2v2: Random Slope as a Structural Predictor"
author: "Michael Hallquist and Zach Vig"
date: "26 August 2026"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    embed-resources: true
    html-math-method: katex
params:
  output_dir: "outputs/vig_hallquist_v2"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  fig.align = "center", dpi = 120, out.width = "100%"
)

suppressPackageStartupMessages({
  library(data.table)
  library(dplyr)
  library(tidyr)
  library(ggplot2)
  library(knitr)
  library(scales)
})

input_path <- tryCatch(knitr::current_input(dir = TRUE), error = function(e) "")
analysis_dir <- if (nzchar(input_path)) {
  dirname(normalizePath(input_path, mustWork = FALSE))
} else {
  getwd()
}
repo_root <- normalizePath(file.path(analysis_dir, "..", ".."), mustWork = TRUE)
resolve_repo_path <- function(path) {
  if (grepl("^/", path)) path else file.path(repo_root, path)
}
source(
  file.path(repo_root, "vig_hallquist_2026", "vh_designs.R"),
  local = knitr::knit_global()
)

output_dir <- resolve_repo_path(params$output_dir)
condition_dir <- file.path(output_dir, "conditions")
canonical_counts <- study_condition_counts()
canonical_offsets <- c(0L, cumsum(canonical_counts)[-length(canonical_counts)])
names(canonical_offsets) <- names(canonical_counts)
expected_ids <- canonical_offsets[["study2v2"]] +
  seq_len(canonical_counts[["study2v2"]])
summary_paths <- file.path(
  condition_dir,
  sprintf("condition_%04d_summary.csv", expected_ids)
)
missing_summary_paths <- summary_paths[!file.exists(summary_paths)]
if (length(missing_summary_paths) > 0L) {
  stop(
    "Study 2v2 is incomplete: missing ", length(missing_summary_paths),
    " condition summaries. First missing file: ", missing_summary_paths[[1]]
  )
}

cell_metrics <- data.table::rbindlist(
  lapply(summary_paths, data.table::fread),
  fill = TRUE,
  use.names = TRUE
) %>%
  tibble::as_tibble() %>%
  filter(.data$study == "study2v2") %>%
  mutate(
    sd_standardized_error = suppressWarnings(
      as.numeric(.data$sd_standardized_error)
    )
  )

expected_method_names <- c(
  "oracle_dual", "naive_dual_blup", "closed_form_dual",
  "fuller_closed_form", "fuller_book_preliminary_closed_form",
  "fuller_book_closed_form", "fuller_alpha_stepdown_closed_form",
  "lai_2spa", "msem"
)
expected_method_cells <- tidyr::crossing(
  condition_id = expected_ids,
  method = expected_method_names
)
observed_method_cells <- cell_metrics %>%
  count(.data$condition_id, .data$method, name = "summary_rows")
missing_method_cells <- dplyr::anti_join(
  expected_method_cells,
  observed_method_cells,
  by = c("condition_id", "method")
)
unexpected_method_cells <- dplyr::anti_join(
  observed_method_cells,
  expected_method_cells,
  by = c("condition_id", "method")
)
v2_contract_ok <-
  n_distinct(cell_metrics$condition_id) == length(expected_ids) &&
  nrow(missing_method_cells) == 0L &&
  nrow(unexpected_method_cells) == 0L &&
  all(observed_method_cells$summary_rows == 1L) &&
  all(cell_metrics$n_rep == 1000L) &&
  all(cell_metrics$study_version == "v2") &&
  all(cell_metrics$calibration_version == "shape_preserving_v2") &&
  all(cell_metrics$covariance_shape_fixed %in% TRUE) &&
  all(cell_metrics$slope_intercept_variance_ratio == 1)
if (!v2_contract_ok) {
  stop(
    "Refusing to render: Study 2v2 summaries violate the v2-only ",
    "completion, method, replication, or calibration contract."
  )
}
summary_file_info <- file.info(summary_paths)
summary_artifact_oldest <- min(summary_file_info$mtime)
summary_artifact_newest <- max(summary_file_info$mtime)

method_labels <- c(
  oracle_dual = "Oracle",
  naive_dual_blup = "Naive dual BLUP",
  closed_form_dual = "Unshrunk-score OLS",
  fuller_closed_form = "Fuller (stabilized)",
  fuller_book_preliminary_closed_form = "Fuller (book preliminary)",
  fuller_book_closed_form = "Fuller (book equations)",
  fuller_alpha_stepdown_closed_form = "Fuller (alpha-stepdown)",
  lai_2spa = "Lai 2S-PA",
  msem = "MSEM"
)
stopifnot(setequal(names(method_labels), expected_method_names))

primary_methods <- c(
  "oracle_dual", "fuller_closed_form", "lai_2spa", "msem"
)
diagnostic_methods <- c("naive_dual_blup", "closed_form_dual")
fuller_methods <- c(
  "fuller_closed_form", "fuller_alpha_stepdown_closed_form",
  "fuller_book_preliminary_closed_form", "fuller_book_closed_form"
)

cell_metrics <- cell_metrics %>%
  mutate(
    method_label = factor(
      unname(method_labels[.data$method]),
      levels = unname(method_labels)
    ),
    effect_label = factor(
      paste0("beta = ", .data$standardized_beta_target),
      levels = paste0("beta = ", c(0, .2, .4, .6))
    ),
    reliability_label = factor(
      paste0("R = ", sprintf("%.2f", .data$target_reliability)),
      levels = paste0("R = ", sprintf("%.2f", c(.25, .5, .8)))
    )
  )

pooled_mc_se <- function(x) {
  x <- x[is.finite(x)]
  if (length(x) == 0L) return(NA_real_)
  sqrt(sum(x^2)) / length(x)
}
fmt_pct <- function(x, accuracy = .1) percent(x, accuracy = accuracy)
fmt_num <- function(x, digits = 3) formatC(x, digits = digits, format = "f")
theme_vh <- function() {
  theme_bw(base_size = 10) +
    theme(
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      strip.background = element_rect(fill = "grey95", colour = "grey75")
    )
}
reliability_palette <- c(
  "R = 0.25" = "#E15759",
  "R = 0.50" = "#F28E2B",
  "R = 0.80" = "#4E79A7"
)

performance_grid <- cell_metrics %>%
  filter(.data$standardized_beta_target > 0) %>%
  group_by(
    .data$method, .data$method_label, .data$effect_label,
    .data$standardized_beta_target, .data$num_clus,
    .data$target_reliability, .data$reliability_label
  ) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    bias_mc_se = pooled_mc_se(.data$mc_se_bias),
    bias_mc_low = .data$bias - qnorm(.975) * .data$bias_mc_se,
    bias_mc_high = .data$bias + qnorm(.975) * .data$bias_mc_se,
    coverage = mean(.data$coverage, na.rm = TRUE),
    coverage_mc_se = pooled_mc_se(.data$coverage_mc_se),
    coverage_mc_low = pmax(0, .data$coverage - qnorm(.975) * .data$coverage_mc_se),
    coverage_mc_high = pmin(1, .data$coverage + qnorm(.975) * .data$coverage_mc_se),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

# Collapse over structural effect before drawing reliability trajectories.
# Without this step, several effect-specific points share each x value and a
# line grouped only by N draws meaningless vertical segments between them.
se_calibration_grid <- cell_metrics %>%
  filter(
    .data$method %in% primary_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$num_clus,
    .data$target_reliability
  ) %>%
  summarise(
    se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

null_grid <- cell_metrics %>%
  filter(.data$standardized_beta_target == 0) %>%
  group_by(
    .data$method, .data$method_label, .data$num_clus,
    .data$target_reliability, .data$reliability_label
  ) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    type1_error = mean(.data$type1_error, na.rm = TRUE),
    type1_mc_se = pooled_mc_se(.data$type1_error_mc_se),
    type1_mc_low = pmax(0, .data$type1_error - qnorm(.975) * .data$type1_mc_se),
    type1_mc_high = pmin(1, .data$type1_error + qnorm(.975) * .data$type1_mc_se),
    operational_type1_error = mean(
      .data$operational_type1_error,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

method_summary <- cell_metrics %>%
  filter(.data$standardized_beta_target > 0) %>%
  group_by(.data$method, .data$method_label) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    coverage = mean(.data$coverage, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    cells_below_nominal = sum(.data$coverage_mc_high < .95, na.rm = TRUE),
    cells_above_nominal = sum(.data$coverage_mc_low > .95, na.rm = TRUE),
    .groups = "drop"
  )

null_summary <- cell_metrics %>%
  filter(.data$standardized_beta_target == 0) %>%
  group_by(.data$method, .data$method_label) %>%
  summarise(
    cells = n(),
    eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    bias = mean(.data$bias, na.rm = TRUE),
    type1_error = mean(.data$type1_error, na.rm = TRUE),
    operational_type1_error = mean(
      .data$operational_type1_error,
      na.rm = TRUE
    ),
    coverage = mean(.data$coverage, na.rm = TRUE),
    .groups = "drop"
  )

metric_for <- function(summary, method, metric) {
  value <- summary[summary$method == method, metric, drop = TRUE]
  if (length(value) != 1L) NA_real_ else value
}

calibration_grid <- cell_metrics %>%
  filter(.data$method == "oracle_dual") %>%
  group_by(
    .data$target_reliability, .data$mean_clus_size,
    .data$marginal_rho
  ) %>%
  summarise(
    achieved_population = mean(.data$achieved_reliability, na.rm = TRUE),
    fitted_marginal = mean(
      .data$mean_stage1_fitted_marginal_slope_posterior_reliability_mean,
      na.rm = TRUE
    ),
    fitted_residualized = mean(
      .data$mean_stage1_fitted_residualized_slope_posterior_reliability_mean,
      na.rm = TRUE
    ),
    .groups = "drop"
  )
```

# Executive summary

Study 2v2 is the amended random-slope-as-predictor experiment. It uses the
shared shape-preserving calibration: the intercept variance, slope variance,
and covariance move together while the target marginal posterior slope
reliability is manipulated. This removes the covariance-shape confounding
identified by the preliminary slope-calibration investigation while preserving
posterior reliability as the quantity most directly tied to Stage-1 score
uncertainty. That investigation was an internal design diagnostic, not a study
intended for publication.

The production run contains all 720 conditions, all nine planned methods, and
1,000 replications per condition. The report reads only rows marked
`study2v2`, `study_version = "v2"`, and
`calibration_version = "shape_preserving_v2"`; it stops rather than falling
back to a legacy cache.

The main findings are:

1. The broad low-reliability coverage collapse seen under the old one-coordinate
   calibration does not recur. Stabilized Fuller averages
   `r fmt_pct(metric_for(method_summary, "fuller_closed_form", "coverage"))`
   non-null conditional coverage with mean bias
   `r fmt_num(metric_for(method_summary, "fuller_closed_form", "bias"))`.
2. This is not exact nominality. Fuller has both under- and overcovering cells,
   and its null rejection rate is
   `r fmt_pct(metric_for(null_summary, "fuller_closed_form", "type1_error"))`,
   showing net conservatism. Lai 2S-PA is similarly conservative under the
   null, whereas MSEM is close to 5%.
3. Conditional and operational performance differ materially for 2S-PA.
   Its conditional coverage is
   `r fmt_pct(metric_for(method_summary, "lai_2spa", "coverage"))`, but its
   success-and-cover probability is only
   `r fmt_pct(metric_for(method_summary, "lai_2spa", "success_and_cover"))`
   because interval eligibility averages
   `r fmt_pct(metric_for(method_summary, "lai_2spa", "eligibility"))`.
4. Unshrinking alone is not a predictor correction. Unshrunk-score OLS remains
   attenuated. Naive regression on the dual BLUP vector also has rare, enormous
   coefficient failures; it is retained as a negative-control diagnostic, not
   as a viable estimator.
5. The Fuller book-equation variants confirm why the stabilized variance path
   is primary. Their typical and tail SE behavior is much more conservative,
   even when their average coverage looks attractive.

# Scientific motivation, the calibration repair, and Study 1v2

Study 2 represents the common applied practice of estimating a person-specific
within-person feature and then using that feature to predict a separate
person-level outcome. Examples include using an estimated affective-reactivity
slope, learning rate, or treatment-response slope to predict later symptoms or
functioning. The scientific target is the relation involving the latent slope,
not the regression on a noisy score substituted for it.

The data-generating model can be summarized as

$$
y_{it}=\gamma_0+\gamma_1x_{it}+u_{0i}+u_{1i}x_{it}+e_{it},
\qquad
z_i=\beta_0+\beta_{0u}u_{0i}+\beta u_{1i}+\varepsilon_i.
$$

The focal estimand is the standardized $\beta$ for the random slope, with the
random intercept included as a nuisance predictor. Including both is important
when intercept and slope correlate: omitting or mismeasuring the intercept can
change the apparent slope effect.

The preliminary slope-calibration investigation showed that fixing only the
slope variance changes the entire latent covariance geometry and can make a
nominal reliability contrast carry changes in ICC, latent correlation, and
numerical conditioning. Study 2v2 instead holds the intercept-to-slope variance
ratio and intercept--slope correlation fixed. Thus, reliability is changed by
scaling a common covariance shape.

Study 1v2 established that this repair largely restores bias and interval
calibration when the estimated random slope is an outcome. Study 2v2 is a
harder test: the noisy estimated slope is a predictor, so score error enters
the structural moment matrix. Fuller, 2S-PA, and MSEM explicitly address that
problem in different ways.

# Design and completeness

The factorial design crosses five cluster counts, four mean cluster sizes,
three reliability targets, three intercept--slope correlations, and four
standardized effects. The structural regression includes both the random
intercept and random slope; the focal coefficient is the standardized slope
effect.

The method set separates benchmarks, proposed corrections, and deliberate
negative controls:

- **Oracle** regresses $z$ on the simulated true intercept and slope.
- **Naive dual-BLUP OLS** substitutes both shrunken empirical Bayes scores;
  **unshrunk-score OLS** reverses shrinkage but ignores the scores' remaining
  measurement error. Neither is a candidate correction.
- **Fuller (stabilized)** corrects the predictor moment matrix with each
  cluster's Stage-1 score-error covariance. The retained book-equation and
  alpha-stepdown variants audit the algebra and variance stabilization.
- **Lai 2S-PA** uses an OpenMx latent-variable Stage 2 supplied with Stage-1
  scores and uncertainties.
- **MSEM** estimates the repeated-measures and structural portions jointly in
  Mplus and is the one-stage benchmark.

```{r completion-table}
tibble::tibble(
  item = c(
    "Expected conditions", "Observed conditions", "Methods per condition",
    "Replications per condition", "Study version", "Calibration version",
    "Covariance shape fixed", "Slope/intercept variance ratio",
    "Summary artifacts read", "Oldest summary artifact (local time)",
    "Newest summary artifact (local time)"
  ),
  value = c(
    length(expected_ids), n_distinct(cell_metrics$condition_id),
    length(expected_method_names), unique(cell_metrics$n_rep),
    unique(cell_metrics$study_version),
    unique(cell_metrics$calibration_version),
    all(cell_metrics$covariance_shape_fixed),
    unique(cell_metrics$slope_intercept_variance_ratio),
    length(summary_paths),
    format(summary_artifact_oldest, tz = "America/New_York"),
    format(summary_artifact_newest, tz = "America/New_York")
  )
) %>%
  kable(align = c("l", "r"))
```

The report reads these exact condition summaries directly. It does not use a
consolidated cache and stops if the v2 contract is violated.

{{< include _vh_v2_metric_guide.qmd >}}

# Calibration and Stage-1 diagnostics

```{r reliability-calibration-plot, fig.width=11, fig.height=6.5}
calibration_grid %>%
  pivot_longer(
    cols = c(
      achieved_population, fitted_marginal, fitted_residualized
    ),
    names_to = "quantity",
    values_to = "reliability"
  ) %>%
  mutate(
    quantity = factor(
      .data$quantity,
      levels = c(
        "achieved_population", "fitted_marginal", "fitted_residualized"
      ),
      labels = c(
        "Population marginal target", "Fitted marginal reliability",
        "Fitted residualized reliability"
      )
    )
  ) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$reliability,
    colour = factor(.data$mean_clus_size),
    group = interaction(.data$mean_clus_size, .data$marginal_rho)
  )) +
  geom_abline(slope = 1, intercept = 0, linetype = 2, colour = "grey35") +
  geom_line(alpha = .35) +
  geom_point(alpha = .65, size = 1.6) +
  facet_wrap(~ quantity, nrow = 1) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(limits = c(.1, .9), breaks = c(.25, .5, .8)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Reliability",
    colour = "Mean m",
    title = "The population target is exact; fitted reliability preserves the ordering",
    subtitle = "Lines distinguish m and rho cells; residualized reliability is diagnostic, not the manipulated target"
  ) +
  theme_vh()
```

The population calibration lands on the requested marginal reliability while
preserving the latent covariance shape. Fitted marginal reliability is noisier
in small-cluster cells but remains ordered. Residualized reliability differs
when the intercept and slope correlate; reporting it prevents the marginal
target from being mistaken for reliability of the unique slope contrast.

# Bias

## Primary corrected methods and benchmarks

```{r corrected-bias-plot, fig.width=12, fig.height=9}
performance_grid %>%
  filter(.data$method %in% primary_methods) %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$bias,
    ymin = .data$bias_mc_low,
    ymax = .data$bias_mc_high,
    colour = .data$reliability_label,
    group = .data$reliability_label
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey40") +
  geom_errorbar(width = 5, alpha = .28, linewidth = .35) +
  geom_line(alpha = .55) +
  geom_point(size = 1.7) +
  facet_grid(method_label ~ effect_label, scales = "free_y") +
  scale_x_continuous(breaks = c(30, 50, 100, 150, 300)) +
  scale_colour_manual(values = reliability_palette) +
  labs(
    x = "Number of clusters",
    y = "Mean bias with 95% Monte Carlo CI",
    colour = NULL,
    title = "Fuller is nearly centered; 2S-PA and MSEM retain positive bias",
    subtitle = "Each point averages 12 m-by-rho cells; free y-scales preserve method-specific structure"
  ) +
  theme_vh()
```

Stabilized Fuller is close to the oracle on average, although condition-level
bias is not uniformly zero. Lai 2S-PA and MSEM have modest positive bias that
is concentrated in difficult low-reliability and smaller-$N$ conditions. The
2S-PA pattern must also be read as conditional on an eligible fit.

## Negative controls without outlier compression

```{r diagnostic-bias-plot, fig.width=10, fig.height=6.5}
cell_metrics %>%
  filter(
    .data$method %in% diagnostic_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$target_reliability,
    .data$reliability_label
  ) %>%
  summarise(
    median_cell_bias = median(.data$bias, na.rm = TRUE),
    q10 = quantile(.data$bias, .10, na.rm = TRUE),
    q90 = quantile(.data$bias, .90, na.rm = TRUE),
    max_absolute_cell_bias = max(abs(.data$bias), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$median_cell_bias,
    ymin = .data$q10,
    ymax = .data$q90,
    colour = .data$reliability_label
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey40") +
  geom_linerange(alpha = .55) +
  geom_point(size = 2) +
  facet_wrap(~ method_label, scales = "free_y") +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_colour_manual(values = reliability_palette, guide = "none") +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Median cell bias (10th--90th percentile)",
    title = "Robust summaries keep rare dual-BLUP explosions from hiding typical behavior",
    subtitle = "Unshrinking alone remains attenuated; naive dual-BLUP OLS also has extreme coefficient tails"
  ) +
  theme_vh()
```

The robust display is intentional. A few naive dual-BLUP cells have enormous
mean coefficients because the two shrunken predictors become nearly collinear.
Allowing those values to determine a common y-axis would erase the scientifically
relevant patterns for every other method. They remain in the audit tables and
are treated as method failure, not silently deleted.

# Coverage and standard errors

```{r corrected-coverage-plot, fig.width=12, fig.height=9}
performance_grid %>%
  filter(.data$method %in% primary_methods) %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$coverage,
    ymin = .data$coverage_mc_low,
    ymax = .data$coverage_mc_high,
    colour = .data$reliability_label,
    group = .data$reliability_label
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey35") +
  geom_errorbar(width = 5, alpha = .25, linewidth = .35) +
  geom_line(alpha = .55) +
  geom_point(size = 1.7) +
  facet_grid(method_label ~ effect_label) +
  scale_x_continuous(breaks = c(30, 50, 100, 150, 300)) +
  scale_y_continuous(
    limits = c(.75, 1),
    labels = percent_format(accuracy = 1)
  ) +
  scale_colour_manual(values = reliability_palette) +
  labs(
    x = "Number of clusters",
    y = "Conditional 95% CI coverage",
    colour = NULL,
    title = "The repaired design avoids a general low-reliability coverage collapse",
    subtitle = "Coverage is conditional on an eligible interval; bars are pooled Monte Carlo CIs"
  ) +
  theme_vh()
```

The repaired design changes the interpretation of the earlier coverage
problem. There is no universal failure below $R=.50$. Fuller is somewhat low
on average in non-null cells, but the oracle also shows finite-$N$ departures.
2S-PA often conditionally overcovers, while MSEM has modest undercoverage in
some non-null cells. These are method-specific patterns rather than a common
artifact of the reliability manipulation.

```{r se-calibration-plot, fig.width=11, fig.height=7}
se_calibration_grid %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$se_ratio,
    colour = factor(.data$num_clus),
    group = .data$num_clus
  )) +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey35") +
  geom_line(alpha = .55) +
  geom_point(size = 1.7) +
  facet_wrap(~ method_label, ncol = 2) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(limits = c(.7, 1.35)) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Median condition-level SE ratio",
    colour = "Clusters",
    title = "Typical SE scale is near one for the primary methods",
    subtitle = paste(
      "Each ratio is mean reported SE / empirical sampling SD;",
      "lines connect reliability levels within a fixed N"
    )
  ) +
  theme_vh()
```

```{r bias-se-mechanism-plot, fig.width=11, fig.height=7.5}
cell_metrics %>%
  filter(
    .data$method %in% primary_methods,
    .data$standardized_beta_target > 0
  ) %>%
  ggplot(aes(
    x = .data$bias,
    y = .data$mean_se_to_interval_empirical_sd,
    colour = .data$coverage,
    shape = .data$reliability_label
  )) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey45") +
  geom_hline(yintercept = 1, linetype = 2, colour = "grey45") +
  geom_point(alpha = .5, size = 1.2) +
  facet_wrap(~ method_label, scales = "free", ncol = 2) +
  scale_colour_viridis_c(
    option = "C", limits = c(0, 1),
    labels = percent_format(accuracy = 1)
  ) +
  labs(
    x = "Condition-level mean bias",
    y = "Condition-level mean-SE / empirical-SD ratio",
    colour = "Conditional coverage",
    shape = NULL,
    title = "Bias, interval scale, and coverage diagnose different mechanisms",
    subtitle = "Free scales prevent difficult cells from compressing well-behaved methods"
  ) +
  theme_vh()
```

# Type I error and operational inference

At the null, CI noncoverage and two-sided rejection are the same event among
eligible intervals. Type I error is therefore the clearest inferential reading
of null-condition coverage.

```{r type1-plot, fig.width=11, fig.height=7.5}
null_grid %>%
  filter(.data$method %in% primary_methods) %>%
  ggplot(aes(
    x = .data$num_clus,
    y = .data$type1_error,
    ymin = .data$type1_mc_low,
    ymax = .data$type1_mc_high,
    colour = .data$reliability_label,
    group = .data$reliability_label
  )) +
  geom_hline(yintercept = .05, linetype = 2, colour = "grey35") +
  geom_errorbar(width = 5, alpha = .28, linewidth = .35) +
  geom_line(alpha = .55) +
  geom_point(size = 1.7) +
  facet_wrap(~ method_label, ncol = 2) +
  scale_x_continuous(breaks = c(30, 50, 100, 150, 300)) +
  scale_y_continuous(
    limits = c(0, .12),
    labels = percent_format(accuracy = .1)
  ) +
  scale_colour_manual(values = reliability_palette) +
  labs(
    x = "Number of clusters",
    y = "Conditional Type I error",
    colour = NULL,
    title = "Fuller and 2S-PA are conservative under the null",
    subtitle = "MSEM and the oracle are close to 5% on average"
  ) +
  theme_vh()
```

The null results provide direct evidence of overcoverage for Fuller and 2S-PA,
not merely wide intervals in non-null cells. Their average Type I errors are
`r fmt_pct(metric_for(null_summary, "fuller_closed_form", "type1_error"))`
and `r fmt_pct(metric_for(null_summary, "lai_2spa", "type1_error"))`,
respectively, compared with
`r fmt_pct(metric_for(null_summary, "msem", "type1_error"))` for MSEM.
This is a meaningful tradeoff: the two-stage corrections protect more strongly
against false positives, but can pay with reduced power and unnecessarily wide
intervals.

```{r conditional-operational-plot, fig.width=11, fig.height=7}
performance_grid %>%
  filter(.data$method %in% primary_methods) %>%
  group_by(
    .data$method, .data$method_label, .data$target_reliability,
    .data$reliability_label
  ) %>%
  summarise(
    conditional_coverage = mean(.data$coverage, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    eligibility = mean(.data$eligibility, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols = c(conditional_coverage, success_and_cover),
    names_to = "performance",
    values_to = "rate"
  ) %>%
  mutate(
    performance = factor(
      .data$performance,
      levels = c("conditional_coverage", "success_and_cover"),
      labels = c("Conditional coverage", "Success-and-cover")
    )
  ) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$rate,
    colour = .data$performance,
    group = .data$performance
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey50") +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~ method_label, ncol = 2) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(limits = c(.55, 1), labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = c("#4E79A7", "#E15759")) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Probability",
    colour = NULL,
    title = "2S-PA's conditional coverage masks substantial fit selection",
    subtitle = "Success-and-cover counts a failed interval as an inferential failure"
  ) +
  theme_vh()
```

# Direct comparison: stabilized Fuller, Lai 2S-PA, and MSEM

The preceding sections diagnose each method separately. Because the proposed
contribution is stabilized Fuller, this section makes the comparison more
explicit. The comparison is deliberately condition matched: Fuller, 2S-PA,
and MSEM are evaluated on the same simulated samples within every design cell.
No method is declared superior from conditional coverage alone.

```{r direct-comparison-setup, include=FALSE}
comparison_methods <- c("fuller_closed_form", "lai_2spa", "msem")
comparison_labels <- c(
  fuller_closed_form = "Fuller (stabilized)",
  lai_2spa = "Lai 2S-PA",
  msem = "MSEM"
)

comparison_nonnull <- cell_metrics %>%
  filter(
    .data$method %in% comparison_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(.data$method, .data$target_reliability) %>%
  summarise(
    cells = n(),
    mean_bias = mean(.data$bias, na.rm = TRUE),
    mean_absolute_bias = mean(abs(.data$bias), na.rm = TRUE),
    conditional_coverage = mean(.data$coverage, na.rm = TRUE),
    absolute_coverage_error = mean(abs(.data$coverage - .95), na.rm = TRUE),
    interval_eligibility = mean(.data$interval_eligibility, na.rm = TRUE),
    success_and_cover = mean(.data$success_and_cover, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    absolute_log_se_error = mean(
      abs(log(.data$mean_se_to_interval_empirical_sd)),
      na.rm = TRUE
    ),
    .groups = "drop"
  )

comparison_null <- cell_metrics %>%
  filter(
    .data$method %in% comparison_methods,
    .data$standardized_beta_target == 0
  ) %>%
  group_by(.data$method, .data$target_reliability) %>%
  summarise(
    conditional_type1 = mean(.data$type1_error, na.rm = TRUE),
    operational_type1 = mean(.data$operational_type1_error, na.rm = TRUE),
    null_interval_eligibility = mean(
      .data$interval_eligibility,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

comparison_summary <- comparison_nonnull %>%
  left_join(
    comparison_null,
    by = c("method", "target_reliability")
  ) %>%
  mutate(
    method_label = factor(
      unname(comparison_labels[.data$method]),
      levels = unname(comparison_labels)
    ),
    reliability_label = paste0(
      "R = ", sprintf("%.2f", .data$target_reliability)
    )
  )

comparison_metric <- function(method, reliability, metric) {
  value <- comparison_summary[
    comparison_summary$method == method &
      comparison_summary$target_reliability == reliability,
    metric,
    drop = TRUE
  ]
  if (length(value) != 1L) NA_real_ else value
}

comparison_wide <- cell_metrics %>%
  filter(
    .data$method %in% comparison_methods,
    .data$standardized_beta_target > 0
  ) %>%
  select(
    .data$condition_id, .data$target_reliability, .data$num_clus,
    .data$mean_clus_size, .data$marginal_rho,
    .data$standardized_beta_target, .data$method, .data$bias,
    .data$coverage, .data$interval_eligibility,
    .data$success_and_cover, .data$mean_se_to_interval_empirical_sd
  ) %>%
  pivot_wider(
    names_from = method,
    values_from = c(
      bias, coverage, interval_eligibility, success_and_cover,
      mean_se_to_interval_empirical_sd
    )
  )

make_paired_comparison <- function(wide, comparator, comparator_label) {
  comparator_key <- comparator
  tibble::tibble(
    condition_id = wide$condition_id,
    target_reliability = wide$target_reliability,
    num_clus = wide$num_clus,
    comparator = comparator_label,
    absolute_bias_difference =
      abs(wide$bias_fuller_closed_form) -
      abs(wide[[paste0("bias_", comparator_key)]]),
    absolute_coverage_error_difference =
      abs(wide$coverage_fuller_closed_form - .95) -
      abs(wide[[paste0("coverage_", comparator_key)]] - .95),
    interval_failure_difference =
      (1 - wide$interval_eligibility_fuller_closed_form) -
      (1 - wide[[paste0("interval_eligibility_", comparator_key)]]),
    operational_calibration_error_difference =
      abs(wide$success_and_cover_fuller_closed_form - .95) -
      abs(wide[[paste0("success_and_cover_", comparator_key)]] - .95),
    se_calibration_error_difference =
      abs(log(wide$mean_se_to_interval_empirical_sd_fuller_closed_form)) -
      abs(log(wide[[paste0(
        "mean_se_to_interval_empirical_sd_",
        comparator_key
      )]]))
  )
}

paired_comparisons <- bind_rows(
  make_paired_comparison(comparison_wide, "lai_2spa", "Lai 2S-PA"),
  make_paired_comparison(comparison_wide, "msem", "MSEM")
) %>%
  pivot_longer(
    cols = ends_with("_difference"),
    names_to = "criterion",
    values_to = "fuller_minus_comparator"
  ) %>%
  mutate(
    criterion = factor(
      .data$criterion,
      levels = c(
        "absolute_bias_difference",
        "absolute_coverage_error_difference",
        "se_calibration_error_difference",
        "interval_failure_difference",
        "operational_calibration_error_difference"
      ),
      labels = c(
        "Absolute bias",
        "Absolute conditional-coverage error",
        "Absolute log(SE / empirical SD)",
        "Interval failure probability",
        "Absolute operational-coverage error"
      )
    )
  ) %>%
  filter(is.finite(.data$fuller_minus_comparator))

paired_difference_summary <- paired_comparisons %>%
  group_by(.data$comparator, .data$target_reliability, .data$criterion) %>%
  summarise(
    median_difference = median(.data$fuller_minus_comparator),
    q25_difference = quantile(.data$fuller_minus_comparator, .25),
    q75_difference = quantile(.data$fuller_minus_comparator, .75),
    fuller_better = mean(.data$fuller_minus_comparator < 0),
    ties = mean(abs(.data$fuller_minus_comparator) < 1e-12),
    .groups = "drop"
  )
```

## The central tradeoff: conditional calibration versus usable inference

```{r direct-coverage-eligibility-tradeoff, fig.width=10.5, fig.height=7}
comparison_summary %>%
  ggplot(aes(
    x = .data$interval_eligibility,
    y = .data$conditional_coverage,
    colour = .data$method_label,
    shape = factor(.data$target_reliability),
    size = .data$success_and_cover,
    group = .data$method_label
  )) +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey45") +
  geom_vline(xintercept = 1, linetype = 2, colour = "grey65") +
  geom_path(alpha = .45, linewidth = .7) +
  geom_point(alpha = .9) +
  scale_x_continuous(
    limits = c(.60, 1.005),
    labels = percent_format(accuracy = 1)
  ) +
  scale_y_continuous(
    limits = c(.88, .98),
    labels = percent_format(accuracy = 1)
  ) +
  scale_size_continuous(
    range = c(3, 7), limits = c(.60, .96),
    labels = percent_format(accuracy = 1)
  ) +
  scale_colour_manual(values = c(
    "Fuller (stabilized)" = "#E15759",
    "Lai 2S-PA" = "#4E79A7",
    "MSEM" = "#59A14F"
  )) +
  labs(
    x = "Interval eligibility",
    y = "Conditional 95% CI coverage",
    colour = NULL,
    shape = "Reliability",
    size = "Success-and-cover",
    title = "Nominal conditional coverage can conceal very different operational performance",
    subtitle = "Each point averages 180 non-null design cells at one reliability target"
  ) +
  theme_vh()
```

At $R=.25$, 2S-PA has the best conditional coverage of the three methods
(`r fmt_pct(comparison_metric("lai_2spa", .25, "conditional_coverage"))`),
but only `r fmt_pct(comparison_metric("lai_2spa", .25, "interval_eligibility"))`
of intervals are eligible. Consequently, its probability of both producing an
eligible interval and covering is
`r fmt_pct(comparison_metric("lai_2spa", .25, "success_and_cover"))`.
Fuller covers less often conditionally
(`r fmt_pct(comparison_metric("fuller_closed_form", .25, "conditional_coverage"))`),
but has complete eligibility and therefore the much higher operational rate of
`r fmt_pct(comparison_metric("fuller_closed_form", .25, "success_and_cover"))`.
MSEM lies between them operationally: eligibility is high, but conditional
coverage is similar to Fuller at the lowest reliability.

At $R=.50$, Fuller combines complete eligibility, the smallest bias, and
`r fmt_pct(comparison_metric("fuller_closed_form", .50, "conditional_coverage"))`
coverage. At $R=.80$, all three methods are much more similar; the substantive
choice is no longer driven primarily by measurement quality.

## Matched-cell advantages and disadvantages

```{r paired-method-difference-plot, fig.width=12, fig.height=8.5}
paired_difference_summary %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$median_difference,
    ymin = .data$q25_difference,
    ymax = .data$q75_difference,
    colour = .data$comparator,
    group = .data$comparator
  )) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey40") +
  geom_errorbar(width = .015, alpha = .45, linewidth = .45) +
  geom_line(alpha = .65) +
  geom_point(size = 2) +
  facet_wrap(~ criterion, scales = "free_y", ncol = 2) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_colour_manual(values = c("Lai 2S-PA" = "#4E79A7", "MSEM" = "#59A14F")) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Fuller minus comparator",
    colour = NULL,
    title = "Fuller's comparative advantage is strongest for bias and fit availability",
    subtitle = "Negative values favor Fuller; points are matched-cell medians and bars are interquartile ranges"
  ) +
  theme_vh()
```

This plot uses differences within the same design condition, so nuisance
variation in $N$, $m$, effect size, and intercept--slope correlation is removed
before comparison. Fuller usually has lower absolute bias and fewer interval
failures than either comparator. Its conditional-coverage advantage is less
uniform. Relative to 2S-PA, Fuller often sacrifices conditional coverage in
the hardest cells in exchange for actually returning an eligible interval.
Relative to MSEM, Fuller has substantially less bias but similar low-reliability
coverage error. Typical SE calibration also does not uniformly favor Fuller at
$R=.25$; all three methods have problematic studentization or tail behavior in
that region.

## Null rejection: Fuller's protection is conservative

```{r direct-type1-comparison, fig.width=12, fig.height=7.5}
null_grid %>%
  filter(.data$method %in% comparison_methods) %>%
  mutate(
    comparison_method = factor(
      unname(comparison_labels[.data$method]),
      levels = unname(comparison_labels)
    )
  ) %>%
  ggplot(aes(
    x = .data$target_reliability,
    y = .data$type1_error,
    ymin = .data$type1_mc_low,
    ymax = .data$type1_mc_high,
    colour = .data$comparison_method,
    group = .data$comparison_method
  )) +
  geom_hline(yintercept = .05, linetype = 2, colour = "grey35") +
  geom_errorbar(width = .015, alpha = .3, linewidth = .4) +
  geom_line(alpha = .65) +
  geom_point(size = 1.8) +
  facet_wrap(~ num_clus, nrow = 1, labeller = label_both) +
  scale_x_continuous(breaks = c(.25, .5, .8)) +
  scale_y_continuous(
    limits = c(0, .12),
    labels = percent_format(accuracy = .1)
  ) +
  scale_colour_manual(values = c(
    "Fuller (stabilized)" = "#E15759",
    "Lai 2S-PA" = "#4E79A7",
    "MSEM" = "#59A14F"
  )) +
  labs(
    x = "Target marginal posterior slope reliability",
    y = "Conditional Type I error",
    colour = NULL,
    title = "Fuller and 2S-PA strongly protect against false positives at low reliability",
    subtitle = "That protection is substantial overcoverage, not exact nominal calibration"
  ) +
  theme_vh()
```

Fuller and 2S-PA are nearly indistinguishable under the null: at $R=.25$ their
Type I errors are approximately
`r fmt_pct(comparison_metric("fuller_closed_form", .25, "conditional_type1"))`
and `r fmt_pct(comparison_metric("lai_2spa", .25, "conditional_type1"))`.
MSEM is much closer to the nominal target at that reliability
(`r fmt_pct(comparison_metric("msem", .25, "conditional_type1"))`).
Thus, Fuller's false-positive protection is a genuine advantage when avoiding
liberal inference is paramount, but it should not be advertised as more
accurate calibration. It is conservative calibration, and the cost may appear
as reduced power.

## Side-by-side numerical summary

```{r direct-comparison-table}
comparison_summary %>%
  transmute(
    Method = .data$method_label,
    Reliability = sprintf("%.2f", .data$target_reliability),
    `Mean bias` = round(.data$mean_bias, 3),
    `Mean |bias|` = round(.data$mean_absolute_bias, 3),
    `Conditional coverage` = percent(
      .data$conditional_coverage,
      accuracy = .1
    ),
    `Interval eligible` = percent(.data$interval_eligibility, accuracy = .1),
    `Success-and-cover` = percent(.data$success_and_cover, accuracy = .1),
    `Median condition-level SE ratio` = round(.data$median_se_ratio, 3),
    `Null Type I error` = percent(.data$conditional_type1, accuracy = .1),
    `Operational null rejection` = percent(
      .data$operational_type1,
      accuracy = .1
    )
  ) %>%
  arrange(.data$Reliability, .data$Method) %>%
  kable(align = c("l", "r", rep("r", 8)))
```

## Comparative conclusion for the proposed method

The results support describing stabilized Fuller as an attractive option, but
not as a uniformly dominant one.

**Why Fuller is attractive.** It has the smallest absolute bias of the three
methods at every reliability target, complete interval eligibility, and the
best or nearly best operational coverage once failed intervals are counted.
These advantages are especially important for a two-stage workflow in which a
method must return a usable result for every fitted Stage-1 model. At moderate
and high reliability, Fuller combines these properties with conditional
coverage and typical SE scale close to their nominal targets.

**Where 2S-PA is competitive.** Lai 2S-PA often has conditional coverage closer
to 95%, including at low reliability. However, it has larger positive bias and
substantial fit selection in the same difficult cells. Its conditional
coverage therefore answers a narrower question: how well intervals cover
among the subset of replications that produce an eligible OpenMx solution.
When all replications count, Fuller is markedly more dependable.

**Where MSEM is competitive.** MSEM has high eligibility and the best null
calibration at low reliability. It is therefore the strongest comparator when
the primary concern is exact Type I error rather than bias. Its disadvantages
here are larger positive bias than Fuller and lower non-null coverage in many
cells. MSEM and Fuller have similarly imperfect non-null coverage at $R=.25$,
so the simulation does not support claiming that Fuller solves extremely weak
measurement conditions.

**The appropriate claim.** Stabilized Fuller offers an appealing balance of
low bias, complete availability, and broadly reasonable interval calibration
without fitting a joint multilevel structural model. The strongest evidence is
for $R\geq .50$. At $R=.25$, all three methods exhibit an important weakness:
Fuller and MSEM undercover non-null effects, while 2S-PA loses many eligible
fits and strongly overcovers under the null. Those cells should motivate a
measurement-design warning and sensitivity analysis rather than a winner.

# Fuller variants and SE tails

```{r fuller-variant-plot, fig.width=11, fig.height=7}
cell_metrics %>%
  filter(
    .data$method %in% fuller_methods,
    .data$standardized_beta_target > 0
  ) %>%
  group_by(
    .data$method, .data$method_label, .data$target_reliability,
    .data$reliability_label
  ) %>%
  summarise(
    coverage = mean(.data$coverage, na.rm = TRUE),
    median_se_ratio = median(
      .data$mean_se_to_interval_empirical_sd,
      na.rm = TRUE
    ),
    p95_cell_p99_se = quantile(.data$p99_se, .95, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(
    x = .data$median_se_ratio,
    y = .data$coverage,
    colour = .data$reliability_label
  )) +
  geom_vline(xintercept = 1, linetype = 2, colour = "grey45") +
  geom_hline(yintercept = .95, linetype = 2, colour = "grey45") +
  geom_point(size = 2) +
  facet_wrap(~ method_label, scales = "free_x", ncol = 2) +
  scale_x_log10() +
  scale_y_continuous(limits = c(.82, 1), labels = percent_format(accuracy = 1)) +
  scale_colour_manual(values = reliability_palette) +
  labs(
    x = "Median condition-level SE ratio (log scale)",
    y = "Conditional coverage",
    colour = NULL,
    title = "The Fuller book paths reveal the SE-tail regularization tradeoff",
    subtitle = "Stabilized Fuller stays near an SE ratio of one; equation-faithful variants can become extremely conservative"
  ) +
  theme_vh()
```

The comparison supports retaining the variants as diagnostics while using the
stabilized implementation as the proposed method. Similar average point bias
does not imply similar uncertainty behavior. In particular, the book-equation
variance bread can generate very large SE ratios and overcoverage in the same
cells where the stabilized estimator remains usable.

# Conclusions

Study 2v2 resolves the central question raised by the preliminary
slope-calibration investigation. Correcting the covariance shape was necessary:
it removes the appearance that low posterior reliability by itself causes
catastrophic coverage. It was not sufficient to make all estimators
interchangeable.

For a psychological-methods audience, the clearest interpretation is:

- Posterior reliability remains the preferred design target because it indexes
  the Stage-1 uncertainty actually propagated into the structural analysis.
  The fixed covariance-shape rule makes that target interpretable and keeps ICC
  and posterior reliability from drifting for avoidable geometric reasons.
- Stabilized Fuller is the strongest two-stage performer in this predictor
  setting: bias is small and typical SE calibration is close to one. Its
  intervals are mildly conservative under the null and somewhat below nominal
  in a subset of non-null cells, so the claim should be good finite-sample
  calibration rather than exact 95% behavior.
- Lai 2S-PA often has nominal or conservative conditional coverage, but positive
  bias and incomplete interval eligibility make unconditional performance less
  favorable. Conditional coverage alone overstates its practical reliability.
- MSEM is less conservative under the null and has nearly complete eligibility,
  but modest positive bias and some non-null undercoverage remain. It is a
  useful direct-model benchmark rather than an automatic gold standard in every
  finite sample.
- Unshrunk-score OLS and naive dual-BLUP OLS are negative controls. The former
  demonstrates attenuation from additive score error; the latter additionally
  exposes near-collinearity and coefficient-tail failures.

The paper should therefore report bias, conditional coverage, operational
success-and-cover, Type I error, SE calibration, and fit eligibility together.
No single column adequately represents inferential performance in this study.

# Appendix

## Compact method summaries

```{r method-summary-table}
method_summary %>%
  transmute(
    Method = .data$method_label,
    Cells = .data$cells,
    `Interval eligible` = percent(.data$eligibility, accuracy = .1),
    `Mean bias` = round(.data$bias, 3),
    `Mean |bias|` = round(.data$absolute_bias, 3),
    `Conditional coverage` = percent(.data$coverage, accuracy = .1),
    `Success-and-cover` = percent(.data$success_and_cover, accuracy = .1),
    `Median condition-level SE ratio` = round(.data$median_se_ratio, 3),
    `MC interval below .95` = .data$cells_below_nominal,
    `MC interval above .95` = .data$cells_above_nominal
  ) %>%
  kable(
    align = c("l", rep("r", 9)),
    caption = paste(
      "Cell counts require the condition-level 95% Monte Carlo interval",
      "to lie wholly below or above .95."
    )
  )
```

```{r null-summary-table}
null_summary %>%
  transmute(
    Method = .data$method_label,
    Cells = .data$cells,
    `Interval eligible` = percent(.data$eligibility, accuracy = .1),
    `Mean bias` = round(.data$bias, 3),
    `Conditional Type I error` = percent(.data$type1_error, accuracy = .1),
    `Operational Type I error` = percent(
      .data$operational_type1_error,
      accuracy = .1
    ),
    `Conditional coverage` = percent(.data$coverage, accuracy = .1)
  ) %>%
  kable(align = c("l", rep("r", 6)))
```

{{< include _vh_v2_provenance.qmd >}}

## Session information

```{r session-info}
sessionInfo()
```
